### 2013-02-28 doc/report/intro.tex
added docs and LICENSE

\section{Introduction}
\label{sec:introduction}

So far, our proposal Looks fairly similar to \cite{Merrill:2007:ALP:1758156.1758158}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2013-02-28 doc/report/main.tex
added docs and LICENSE

\documentclass{sig-alternate}

\usepackage{times}
\usepackage{graphics}

\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{tabularx}
\usepackage{color}
\usepackage{xspace}
\usepackage{hyperref}    % Creates hyperlinks from ref/cite
\hypersetup{pdfstartview=FitH}
\usepackage{graphicx}    % For importing graphics
\usepackage{url}         %

\hypersetup{%
pdftitle={Your Title}, pdfauthor={Your Authors}, pdfkeywords={your
keywords}, bookmarksnumbered, pdfstartview={FitH}, colorlinks,
citecolor=black, filecolor=black, linkcolor=black, urlcolor=black,
breaklinks=true,}

\renewcommand{\arraystretch}{1.2} % Space out rows in tables

\newcommand{\ml}[1]{{\color{green} {\it ML: #1}}}

% No space between bibliography items:
\let\oldthebibliography=\thebibliography
  \let\endoldthebibliography=\endthebibliography
  \renewenvironment{thebibliography}[1]{%
    \begin{oldthebibliography}{#1}%
      \setlength{\parskip}{0ex}%
      \setlength{\itemsep}{0ex}%
  }%
  {%
    \end{oldthebibliography}%
  }

%\pagenumbering{arabic}  % Arabic page numbers for submission.  Remove this line to eliminate page numbers for the camera ready copy

\begin{document}

% use this command to override the default ACM copyright statement
% (e.g. for preprints). Remove for camera ready copy.
%\toappear{Submitted for review to IPSN 2012.}
% \conferenceinfo{ConfName} {Date, Location}
% \CopyrightYear{Year} 
% \crdata{978-1-4503-1227-1/12/04} 
% \clubpenalty=10000 
% \widowpenalty = 10000

% to make various LaTeX processors do the right thing with page size
\special{papersize=8.5in,11in}
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\title{Attention Tracker}

\author{
{Ben Zhang, Sean Chen}\\
\affaddr{University of California, Berkeley}\\
%\affaddr{}\\
\email{benzh@eecs.berkeley.edu, sean.yhc@berkeley.edu}
}

\maketitle

% \begin{abstract}

% \end{abstract}

%\category{C.2.1}{Computer-Communication Networks}{Network Architecture and Design}[Wireless communication]
%\keywords{Collection, CTP, Sensor Network, Routing}

% \category{B.0}{Hardware}{General}
% \category{B.4}{Hardware}{Input/Output \& Data Communication} 
% \category{H.4.m}{Information Systems Applications}{Miscellaneous}

% \terms{Design, Experimentation, Measurement} 
% \keywords{Localization, Magneto-Inductive, Tracking, Virtual Zone}

\input{intro.tex}

% \section{Acknowledgments}

\bibliographystyle{abbrv}
\bibliography{reference}

\end{document}

### 2013-04-04 server/wireless.py
added code for wireless communication. now it only reads from the serial port and print anything it saw

@@ -1,0 +1,33 @@
+import serial, glob
+
+# we know for mac it will show as /dev/usb.tty*, so list all of them and ask user to choose
+
+availables = glob.glob('/dev/tty.usb*')
+
+print "All available ports:" 
+
+index = 1
+for port in availables:
+	print '  ', index, port
+	index += 1
+
+try:
+	num = int(raw_input('Enter the number:'))
+	if num > 0 and num < len(availables):
+		selected = availables[num-1]
+except:
+	print 'invalid input!'
+
+try:
+	baud = int(raw_input('Baud rate (9600):'))
+except:
+	baud = 9600
+
+
+ser = serial.Serial(selected, baud)
+while True:
+	print ser.read()
+
+
+
+	

### 2013-04-05 doc/report/intro.tex
some old check in stuff, just cleaned it

@@
 \label{sec:introduction}
+
+Traditional ways of interacting with applications on phones or tablets are browsing the available applications and click on them.
+You may have the freedom to set shortcuts on home screen for easy access.
+While this is unavoidable in cyber world, for interaction with physical devices, there is a more natural way to select by exploiting spatial information - proximity and attention.
 

### 2013-04-05 server/wireless.py
added arguments for script

@@ -1,8 +1,23 @@
-import serial, glob
+import serial, glob, argparse
+
+parser = argparse.ArgumentParser(description='Pyserial for XBee')
+parser.add_argument('--debug', default=False, action='store_true', help='Print More Errors and Launch interactive console on exception or forced exit')
+parser.add_argument('--baud', type=int, action='store', default=9600, help='Specify the baud rate')
+parser.add_argument('--timeout', type=float, action='store', default=1, help='Timeout parameter for serial connection')
 
 # we know for mac it will show as /dev/usb.tty*, so list all of them and ask user to choose
 
 availables = glob.glob('/dev/tty.usb*')
+
+arguments = parser.parse_args()
+try:
+	baud = arguments.baud
+	timeout = arguments.timeout
+	print "  baud rate:", baud
+	print "  serial timeout:", timeout
+except Exception as e:
+	print e
+	exit(1)
 
 print "All available ports:" 
 
@@ -13,20 +28,21 @@ for port in availables:
 
 try:
 	num = int(raw_input('Enter the number:'))
-	if num > 0 and num < len(availables):
+	if num > 0 and num <= len(availables):
 		selected = availables[num-1]
-except:
-	print 'invalid input!'
+except Exception as e:
+	print "invalid input!"
+	exit(1)
+
 
 try:
-	baud = int(raw_input('Baud rate (9600):'))
-except:
-	baud = 9600
-
-
-ser = serial.Serial(selected, baud)
+	ser = serial.Serial(selected, baud, timeout=timeout)
+except Exception as e:
+  print "failed to connect the serial port", e
+	
 while True:
-	print ser.read()
+	ser.write("testing")
+	print ser.readline()
 
 
 

### 2013-04-06 server/wireless.py
changed to wait for input

@@ -41,7 +41,8 @@ except Exception as e:
   print "failed to connect the serial port", e
 	
 while True:
-	ser.write("testing")
+	text = raw_input("text message to send:")
+	ser.write(text)
 	print ser.readline()
 
 

### 2013-04-06 server/wireless.py
added glass code, which is not complete yet

@@ -44,7 +44,4 @@ while True:
 	text = raw_input("text message to send:")
 	ser.write(text)
 	print ser.readline()
-
-
-
 	

### 2013-05-03 slider_test/slider.h
add testing slider codes, TBC

@@ -1,0 +1,29 @@
+#ifndef slider_h
+#define slider_h
+
+#define gsIDLE 0
+#define gsPRESS 1
+#define gsRELEASE 2
+
+#define SOFTPOT_THREASHOLD 950
+#define SOFTPOT_DELTA_THREASHOLD 80
+#define DELAY_IN_WAIT 1000000
+
+
+
+#ifdef DEBUG
+  #define DEBUG_PRINT(x)  Serial.print(x)
+  #define DEBUG_PRINTLN(x)  Serial.println(x)
+#else
+  #define DEBUG_PRINT(x)
+  #define DEBUG_PRINTLN(x)
+#endif
+
+typedef enum {
+  gNONE,
+  gHOVER,
+  gTAP,
+  gD_TAP,
+} gesture_t;
+
+#endif

### 2013-05-03 slider_test/slider.h
new slider code v1.0
able to return none, hover, tap, d_tap, release

@@ -24,6 +24,7 @@ typedef enum {
   gHOVER,
   gTAP,
   gD_TAP,
+  gRelease
 } gesture_t;
 
 #endif

### 2013-05-03 slider_test/slider.h
slider v1.1: add hover change detection
TBC: take out tap if it might be a double tap

@@ -24,7 +24,9 @@ typedef enum {
   gHOVER,
   gTAP,
   gD_TAP,
-  gRelease
+  gRELEASE,
+  gHOVERCHANGE,
+  gPRESS
 } gesture_t;
 
 #endif

### 2013-05-03 glass/slider.h
merge slider code to glass (switch state needs to be modified)
currently return right after getting the guesture

@@ -1,16 +1,11 @@
-
-#define DEBUG
-#include "slider.h"
+#ifndef slider_h
+#define slider_h
 
 #define gsIDLE 0
 #define gsPRESS 1
 #define gsRELEASE 2
 
-int softpotReading = 0;
-int softpotInitV = 0;
-int softpotPin = A0; //analog pin 0
-int ledPin = 13;
-
+#define SOFTPOT_THREASHOLD 950
 
 unsigned long start_time = 0;
 unsigned long release_time = 0;
@@ -18,40 +13,26 @@ unsigned long last_release_time = 0;
 int duration_threshold = 300;
 int dtap_duration_threshold = 600;
 int delta_threshold = 20;
-gesture_t g = gNONE;
 int sliderState = gsIDLE;
 
-int sliderVal = 0;
-int sliderDelta = 0;
+int softpotReading = 0;
+int softpotPin = A0; //analog pin 0
 
+typedef enum {
+  gNONE,
+  gHOVER,
+  gTAP,
+  gD_TAP,
+  gRELEASE,
+  gHOVERCHANGE,
+  gPRESS
+} gesture_t;
 
-
-void setup()
-{
-  Serial.begin(9600);
-  digitalWrite(softpotPin, HIGH); //enable pullup resistor
-  randomSeed(analogRead(5));
-  digitalWrite(ledPin, LOW);
-  // XBee.begin(9600);
-  // XBeePacketCounter = 0;
-
-  
-  delay(300);
-  Serial.println("system begins");
-}
-
-void loop() {
-  
-  g = sliderEvent(&sliderDelta, &sliderVal);
-  if (g!= 0) {
-    printGesture(g);
-  }
-  
-  delay(10);
-}
+// gesture_t g = gNONE;
 
 gesture_t sliderEvent(int* sDelta, int* sVal) {
   softpotReading = analogRead(softpotPin);
+  // int softpotReading = *pot;
   // DEBUG_PRINTLN(softpotReading);
   if(softpotReading > SOFTPOT_THREASHOLD ) { // currently not pressed
     if(sliderState == gsPRESS) {
@@ -130,7 +111,7 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
   }
 }
 
-void printGesture(int g){
+void printGesture(int g, int *delta, int *val){
   // DEBUG_PRINT("---");
   switch(g){
     case 0:
[…]

### 2013-05-03 glass/utils.h

@@ -145,12 +145,4 @@ void digitalToggle(int pin){
 
 
 
-typedef enum {
-  gNONE,
-  gHOVER,
-  gTAP,
-  gD_TAP,
-  gCLICK,
-} gesture_t;
-
 #endif

### 2013-05-03 doc/report/design.tex
adding files to the report

\section{Design}
\label{sec:design}

Talks about the system design space.
What might be available to achieve such a function.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2013-05-03 doc/report/intro.tex
adding files to the report

--- 2013-02-28 doc/report/intro.tex
@@
 
-So far, our proposal Looks fairly similar to \cite{Merrill:2007:ALP:1758156.1758158}.
+There has been the vision of smart-home and smart-office where everything around is intelligent enough to assist human in everyday life.
+The development of radio technology, sensors, and actuators are making this dream come true.
+You can buy those basic components online \cite{SmartHome, NinjaBlocks} and many hobbyists have already created their own smart environment (see \cite{BRAD} as an example).
+Recently, with the the prevalence of smartphones, many systems \cite{SmartThings, Lockitron} have been integrated into smartphone applications to lower the barrier for ordinary users.
+
+Admittedly, people have been used to accessing the Internet, conversing with friends, and even dealing with online business through their smartphones.
+These handheld devices serve as an important portal for users to be connected to the Internet.
+And now many systems are making smartphones the portal to interact with physicalworld.
+Given all the benefits we can get by using smartphones -- networking,  display, gesture input, and increasing computability,  we argue that traditional GUI on mobile platform still imposes difficulty in interacting with physical objects.
+Take iPhone for example, users are required to browse a list of possible icons and then select the right one.
+Even on Android platform, you can create widget to simplify the browsing overhead.
+We hold the opinion that squeezing physically distributed objects onto a-few-inch screen is not the right approach.
+And a more direct approach is desirable.
+
+In this project, we aim to explore the ability of direct interaction by using glass-based attention tracking to pin-point the target.
+
+% explain why we need glass form factor
+
+The rest of this paper is structured as follows.
+In Sec.\,\ref{sec:related-work}, we will give an overview of some related projects.
+Then we present our approach in detail about the system design and implementation in Sec.\,\ref{sec:design} and Sec.\,\ref{sec:implementation} respectively.
+Then Sec.\,\ref{sec:evaluation} shows our evaluation results and we provide an in-depth discussion (Sec.\,\ref{sec:discussion}) about different kinds of interaction since not all are suitable for our system.
+Sec.\,\ref{sec:conclusion} concludes this paper.
 

### 2013-05-03 doc/report/main.tex
adding files to the report

@@
 \hypersetup{%
-pdftitle={Your Title}, pdfauthor={Your Authors}, pdfkeywords={your
-keywords}, bookmarksnumbered, pdfstartview={FitH}, colorlinks,
+pdftitle={Project Glass 2.0}, pdfauthor={Ben Zhang, Sean Chen}, pdfkeywords={HCI}, bookmarksnumbered, pdfstartview={FitH}, colorlinks,
 citecolor=black, filecolor=black, linkcolor=black, urlcolor=black,
@@
 
-\title{Attention Tracker}
+\title{Project Glass 2.0 -- Interaction with Physical Devices through Attention}
 
@@
 
-% \begin{abstract}
-
-% \end{abstract}
+\begin{abstract}
+With the increasing interest of assigning intelligence to everyday object, many approaches of interacting with them fall back to smartphone based applications.
+We argue about the inconvenience, overhead, and distraction in such an interaction paradigm, and propose a more natural solution.
+A glass-based controller which enables user to express their interest in interaction directly simplifies the process of selection in a smart-home or smart-office senario.
+We present our thoughts, design and evaluation of such a system in this paper to motivate researchers in considering ``interaction in attention''.
+In the meantime, we noticed that Google Glass might serve as the perfect enabling technology, though in our prototype we've adopted simple infrared approach for proof-of-concept.
+\end{abstract}
 
@@
 % \terms{Design, Experimentation, Measurement} 
-% \keywords{Localization, Magneto-Inductive, Tracking, Virtual Zone}
+\keywords{HCI, Interaction, Attention, Infrared, Glass}
 
 \input{intro.tex}
+\input{related.tex}
+\input{design.tex}
+\input{implementation.tex}
+\input{evaluation.tex}
+\input{discussion.tex}
+\input{conclusion.tex}
 

### 2013-05-03 doc/report/related.tex
some old check in stuff, just cleaned it
adding files to the report

--- 2013-02-28 doc/report/intro.tex
@@
-\section{Introduction}
-\label{sec:introduction}
+\section{Related Work}
+\label{sec:related-work}
 
-So far, our proposal Looks fairly similar to \cite{Merrill:2007:ALP:1758156.1758158}.
+In this section, we will first summarize some prior work and how it relates to our project in three folds: Tangible User Interface (TUI), Glass-based device, and Attention Aware System (AAS), discussing each in turn.
+
+\subsection{Tangible User Interface}
+\label{sec:tang-user-interf}
+
+The concept of Tangible User Interface was proposed in \cite{Ishii:1997:TBT:258549.258715}, and according to the original paper, such system would enable user to ``grasp \& manipulate bits in the center of users' attention by coupling the bits with everyday physical objects and architectural surfaces''.
+Thereafter, numerous TUI systems (such as \cite{Nanayakkara:2012:EEF:2212776.2212382, Patel:2006:IPA:2094945.2094962}) are developed.
+\cite{Merrill:2007:ALP:1758156.1758158} is one that is closest to our system, which base users' interaction with physical objects on people's natural looking, pointing and reaching metaphor.
+However, these systems have limitations that they focus too much on sensing the physical world, while the actuation is mostly done in cyber world.
+Recent years, many startups \cite{SmartThings, NinjaBlocks, Lockitron} tries to ``hack'' the physical world to achieve the envisioned Internet of Things by enabling actuating devices from phone/web portal.
+
+Our system combines some merits from prior work.
+We also user people's natural looking as attention tracking as in \cite{Merrill:2007:ALP:1758156.1758158}, but we are not limited to browsing or searching physical objects.
+We provide such capability of interacting directly with physical device, in addition to control with a mobile phone or web portal \cite{SmartThings, NinjaBlocks}.
+
+\subsection{Glass Form-factor}
+\label{sec:glass-form-factor}
+
+\cite{mann2004continuous}, as the earliest wearable device, uses the form factor of glass to serve as a logging machine that records user's’ daily life.
+According to one review paper \cite{morris2010emerging}, there are many glass-based device developed for the past decades.
+Those who use glasses as input have primarily focused on gaze-tracking \cite{Selker:2001:EGE:634067.634176, Nagamatsu:2010:MDG:1753846.1753983}, and the projects that uses glasses for available output device works hard to achieve virtual reality \cite{Lumus, GoogleGlass}.
+
+The most recent and potentially most impactful project \cite{GoogleGlass} integrate many crucial components -- camera, projection, microphone, speaker, IMU and networking -- into a single glass that can achieve virtual reality to an unimaginable extent.
+We do believe this might be a powerful platform for future interaction development.
+However, so far the vision is still being limited to Cyber world.
+Our project explores the potential of using glasses for direct interacting with physical devices, rather than doing ``digital'' tasks.
+
+\subsection{Attention Aware System}
+\label{sec:attent-aware-syst}
+
+Microsoft Research's project ``Attentional User Interface'' \cite{horvitz2003models} explores how attention can be employed to enhance human-computer interaction, from the lesson learned in human-human interaction.
+Attention detection can also assist interface design to avoid context shifting overhead, e.g.
+the peripheral displays and the notification level should be adjusted according to user's attention \cite{parkdesigning}.
+
+Our work also borrows ideas from these project in that we seek to identify users' attention to select the right physical object to interact.
 

### 2013-05-03 doc/report/main.tex
some minor change

@@
 We present our thoughts, design and evaluation of such a system in this paper to motivate researchers in considering ``interaction in attention''.
-In the meantime, we noticed that Google Glass might serve as the perfect enabling technology, though in our prototype we've adopted simple infrared approach for proof-of-concept.
+In the meantime, we noticed that Google Glass might serve as the perfect enabling technology, though in our prototype we've adopted infrared approach for proof-of-concept.
+The evaluation and discussion of such system might be valuable for future work to explore on more mature hardware platform.
 \end{abstract}

### 2013-05-03 glass/slider.h
further merged the slider code and went thorugh each connection states

@@ -6,13 +6,14 @@
 #define gsRELEASE 2
 
 #define SOFTPOT_THREASHOLD 950
+#define SOFTPOT_DELTA_THRESHOLD 20
 
 unsigned long start_time = 0;
 unsigned long release_time = 0;
 unsigned long last_release_time = 0;
 int duration_threshold = 300;
 int dtap_duration_threshold = 600;
-int delta_threshold = 20;
+
 int sliderState = gsIDLE;
 
 int softpotReading = 0;
@@ -89,7 +90,7 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
     if(sliderState == gsPRESS) {
       // continuously pressing
       *sDelta = softpotReading - *sVal;
-      if(abs(*sDelta) > delta_threshold){ //avoid noise
+      if(abs(*sDelta) > SOFTPOT_DELTA_THRESHOLD){ //avoid noise
         *sVal = softpotReading;
         
         return gHOVERCHANGE;

### 2013-05-05 doc/report/design.tex
added latex docs for the report, some graphs

@@
 
-Talks about the system design space.
-What might be available to achieve such a function.
+% ok, I have to admit that when I start writing, I haven't figured out what to write.
+% so kinda messy.
+
+% this section is used to discuss the design considerations of the system
+% starting from a motivating example, and then extract the key requireements, then about system design
+
+In this section, we will talk about our thoughts on enabling direct interaction with physical device, driven by users' attention.
+We will first discuss one example application which motivates the work in this project.
+We also seek to summarize the key design requirements, and use them for the system design.
+Then several possible approaches are presented as an exploration o f the design space.
+To build a proof-of-concept system, we've simplified several tasks and implement a working system.
+The detailed implementation details will be covered in the next section.
+
+\subsection{Example Application}
+\label{sec:example-application}
+
+Our work is primarily motivated by the observation that in CS294-84 course at Berkeley, the instructor has to go to the switch to turn off the light during presentation.
+The separation of device and their controller makes it possible for people control things that are not reachable (such as the lights on the ceiling).
+However, such indirection introduces additional overhead since people still need to locate the controller and then interact.
+Remote control might mitigate such problem, but again a controller for a device is not a scalable solution.
+Squeezing controllers into the smartphone display is also not the right approach.
+And these points will be more clear when we show our example application -- Smart Home.
+
+% {\bf Smart Home}:
+% Sean, maybe you can elaborate on this application more to motivate.
+Imagine in a smart-home envinroment that most device are controllable remotely.
+When the users are watching TV shows, they want to turn off the lights and turn up the volume so they can enjoy this memorable time.
+Such ordinary-life tasks have been simplified so that they are able to issue ``turn off'' commands to the lights by simply staring at the light plus a specific gesture.
+Then they turn their head back to the TV, and again simply ``ask'' the TV to turn up the volume.
+When transitioning, the disconnection with lights and connection with TV happens seamlessly, with the correct feedback.
+Of course all the magic has to be resolved by an always-available device, such as glasses, rings, watches, or any other proper form factor.
+We will hold the discussion to the design space exploration.
+
+Occasionally, the user needs to perform some advanced configuration to the TV, such as adjusting the white balance, or the brightness.
+Since the need of such complicated tasks are not frequent, and it will be less efficient to embed these features into the always-available device (which are designed to be small and easy-to-access, rather than too flexible).
+The smart-home falls back to a more generic approach which relies on touch screen or other rich input device.
+In this example case, since the user's attention is still on TV, his smartphone pops up the control menu for advanced usage automatically.
+Such menu could also just pop up on walls or other large interactive devices \cite{unPad:eWallpaper, MSVision}, but the attention-based interaction simplifies the selection on those flexible devices.
+
+\subsection{System Requirements}
+\label{sec:system-requirements}
+
+From what we have described in the previous section, to achieve such direct interaction in attention, we have identified several key requirements.
+
+% leads to the adoption of glass
+The first and foremost is the ability of always-available sensing and acutation. Since our target is to simplify everyday task in smart-home/office scenario. Any cubersome solution is not desired.
+
+Secondly, users are directly interacting with physical device. Since a screen is not always available, there should be some feedback indicator to explicit tell users what is going on.
+
+As discussed before, we envision the simple always-available device would not be able to cover all tasks, especially some advanced ones. We will have to design the system to enable falling back.
+
+For any human-centric system, additional requirements like responsiveness would aslo be required.
+
+\subsection{System Design}
+\label{sec:system-design}
+
+In this section, we will present several design consideration that fulfills the requirements.
+From \cite{Merrill:2007:ALP:1758156.1758158}, they tried the solution of putting IR into earphones, rings to enable pointing metaphor.
+Other than that direction, recently published Google Glass project motivates us to consider glasses form factor.
+Glasses are natural in align with human's vision, and can reflect users' attention in a more precise way.
+But given the form factor of glasses, what might be a good approach to achieve sensing and actuation remains a challenge for us.
+
+% kinda feel lost in describing
+
+In our discussion, we've come up with three possible approaches.
+The first is a purely glass solution.
+The second is augmented by some computer vision detection.
+The third  will be a combination of glasses and smart watch.
+
+% Lessons learned from \cite{Bellotti:2002:MSS:503376.503450} have also guided us in thinking the Five A's when designing novel interaction system.
 

### 2013-05-05 doc/report/implementation.tex
adding files to the report
added latex docs for the report, some graphs

--- 2013-05-03 doc/report/design.tex
@@
-\section{Design}
-\label{sec:design}
+\section{Implementation}
+\label{sec:implementation}
 
-Talks about the system design space.
-What might be available to achieve such a function.
+In this section, we details our implementation on Arduino \cite{Arduino} platform.
+The whole system is comprised of a glass module, numbers of clients, and a gateway.
+We will start by the clients, for which we only implement the simpliest functionalities.
+We push the burden of coordination to the glass.
+
+\subsection{Clients}
+The clients are comprised of a main Arduino board, an IR receiver, an XBee radio, and various actuators.
+In our current system, we have a relay which can control the AC power plug, and we have a USB connected computer to control video playing.
+
+The main function of client is to respond IR signal, and communicate with glass to receive corresponding commands.
+When there is no glass initiating connection, the client simply lives in {\it IDLE} mode; while when they receive IR signal, indicating the user is expressing interest in interacting with it, the client responds with an XBee acknowledgement and goes into {\it PENDING} state.
+In this case, since the client doesn't know how many other clients have also responded to the glass, it will wait until a connection signal.
+If there are multiple clients waiting to be verified, the glass will coordinate all of them by sending the right verifying message.
+The active client that is being verified will blink faster, as visual feedback to users.
+When it receives the connection message, it goes to {\it CONNECTED} state and is ready to take commands.
+To summarize the behavior, we have a state machine in Fig.\,\ref{fig:clientFSM} for illustration.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{../figs/clientFSM.pdf}
+  \caption{FSM model for client}
+  \label{fig:clientFSM}
+\end{figure}
+
+Within {\it CONNECTED} state, the clients' behaviors are different from each other according to the device it is connected to.
+In our prototype, we have two different types of devices.
+The first is a relay, which can then control the whole AC power supply.
+To control the relay, we only need to set a digital pin to {\it LOW} or {\it HIGH}, and each action is encoded in the command set from the glass.
+The second is a computer that is used to play video.
+And the direct interaction includes adjusting the volume and pause/play.
+These commands are sent to the computer through USB serial data.
+
+\subsection{Glass}
+\label{sec:glass}
+
+The Glass is complicated in two folds. First, we move the complexity of coordinating all clients to the Glass. Second, we have to handle user gesture on Glass.
+
+% add graphs and FSM for Glass
+
+\subsection{Gateway}
+\label{sec:gateway}
+
+The gateway provides access to the personal area network for computers, and presumably this can be further open to the Internet.
+Since there have already been many products \cite{NinjaBlocks, Lockitron} that essentially function in this way, we do not spend much time on this aspect in our project.
 

### 2013-05-05 doc/report/intro.tex
added latex docs for the report, some graphs

@@
 
+In comparison, taking out the phones (or even simply locating them) is relatively time-consuming.
+According to a study in \cite{Ashbrook:2008:QIM:1357054.1357092}, 4 seconds are required on average just to get a phone out of the pocket or hip holster, not to mention the time to choose the target and initiate interaction.
+
 The rest of this paper is structured as follows.

### 2013-05-05 doc/report/conclusion.tex
adding files to the report
added some texts to the report, and a glass FSM figure

--- 2013-02-28 doc/report/intro.tex
@@
-\section{Introduction}
-\label{sec:introduction}
+\section{Conclusion}
+\label{sec:conclusion}
 
-So far, our proposal Looks fairly similar to \cite{Merrill:2007:ALP:1758156.1758158}.
+In this project, we explored the novel way of interacting with physical device by expressing users' attention.
+The prototype we have built demonstrate the possibility of identifying user's attention and control various devices.
+
+We will leave X, Y, Z to future work.
 

### 2013-05-05 doc/report/implementation.tex
minor change to figure title
added some texts to the report, and a glass FSM figure

@@
 The whole system is comprised of a glass module, numbers of clients, and a gateway.
-We will start by the clients, for which we only implement the simpliest functionalities.
-We push the burden of coordination to the glass.
+We will start by the clients, which responds to different message sent by others.
+The complexity of coordination is pushed to the glass, and only minimal function are provided in the gateway.
 
@@
   \includegraphics[width=\linewidth]{../figs/clientFSM.pdf}
-  \caption{FSM model for client}
+  \caption{FSM model for client. For each transition, ``e'' stands for event, and ``a'' stands for action}
   \label{fig:clientFSM}
@@
 
-The Glass is complicated in two folds. First, we move the complexity of coordinating all clients to the Glass. Second, we have to handle user gesture on Glass.
+The Glass is complicated in two folds.
+First, we move the complexity of coordinating all clients to the Glass.
+Second, we have to handle user gesture on Glass.
+For a clean design, we separate the gesture detection and XBee transmission into two customized library, and only expose interesting information to the main program.
 
-% add graphs and FSM for Glass
+The whole work flow is as follows.
+When the user hasn't shown any interest of interaction, the system lives in {\it IDLE} state.
+When ``tap'' is detected, the Glass will broadcast IR signal to inform the devices which is in front of the user (the one that can receive IR signal).
+Then the Glass waits for a specific period of time for acknowledge message sent by clients.
+The timer in {\it WAIT} state now is set to 1 second.
+When timer expires, Glass will go to {\it IDLE} if no client has responded, or {\it CONFIRM} if it receives just one message.
+However, the complicated case is when there are multiple clients that have received IR signal and responded.
+Though we expect such scenario is rare, there is still chance that the physical devices are put together and it is necessary to differentiate them.
+In this case, the system goes to {\it VERIFY} state, and send out corresponding verifying message based on user's gesture.
+Once the user confirmed the selection, it goes to {\it CONNECTED} state where it's ready to send out commands.
+Again, to illustrate the flows, we have the FSM in Fig.\,\ref{fig:glassFSM}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{../figs/glassFSM.pdf}
+  \caption{FSM model for Glass. For each transition, ``e'' stands for event, and ``a'' stands for action}
+  \label{fig:glassFSM}
+\end{figure}
 
@@
 
+Just for completeness, we wrote a python script that controls a USB XBee adaptor, so that we can remotely control those devices without the Glass and IR initiation.
+
 %%% Local Variables: 

### 2013-05-06 doc/report/intro.tex
added some texts to the report, and a glass FSM figure
update intro, refine the language and some description, good with me

--- 2013-04-05 doc/report/intro.tex
@@
 
-Traditional ways of interacting with applications on phones or tablets are browsing the available applications and click on them.
-You may have the freedom to set shortcuts on home screen for easy access.
-While this is unavoidable in cyber world, for interaction with physical devices, there is a more natural way to select by exploiting spatial information - proximity and attention.
+The vision of smart-home and smart office is being realized recently with the development of radio technology, sensors, and actuators.
+Nowadays, you can buy the basic components online \cite{SmartHome, NinjaBlocks} and many hobbyists have already created their own smart environment (see \cite{BRAD} as an example).
+Though various challenges are ahead in the field to build a more robust, scalable, energy-efficient, and user-friendly systems, the general trend of having smart devices around us is undoubted.
+Those smart devices each can be controlled separately (like what is existing now for television and heater), however, with the prevalence of smartphones, many systems \cite{SmartThings, Lockitron} are being written as an application on mobile platforms so that the smartphone act as the bridge of connecting digital world and physical world.
 
-So far, our proposal Looks fairly similar to \cite{Merrill:2007:ALP:1758156.1758158}.
+Admittedly, people have been used to accessing the Internet, conversing with friends, and even dealing with online business through their smartphones.
+And putting the control function of physical device into smartphones can significantly lowers the barrier for ordinary users.
+Also users can get other benefits by using smartphones -- networking,  display, gesture input, and the increasing computability.
+
+However, we argue that traditional graphical user interface (GUI) on mobile platforms still imposes difficulty in interacting with physical objects.
+Imagine when we have all the control panel integrated into iPhone, users then need to browse a list of possible icons and select the right one.
+Even if you can create widget to simplify the browsing overhead, such as on Android platform, the problem still holds since squeezing physically distributed objects onto a-few-inch screen is not the right approach.
+In an ideal system, the user should be able to express their interest of interaction in a more direct way.
+On the other hand, taking out the phones (or even simply locating them when at home) is relatively time-consuming.
+According to a study in \cite{Ashbrook:2008:QIM:1357054.1357092}, 4 seconds are required on average just to get a phone out of the pocket or hip holster, not to mention the time to choose the target and initiate interaction.
+This project is trying to propose a new way of interacting with physical objects without losing attention, essentially solving the ``address'' problem in interacting with physical devices around.
+
+Attention Aware System (AAS) \cite{horvitz2003models} has been studied for a while in understanding how users' attention affects their interaction, and how system should affect or manage users' attention.
+This comes from some observations in human-human interaction.
+For people to direct interact with physical objects, their attention can be a strong indicator of their interest.
+When staring at the television, the user is more likely to change channels or do other TV related controls rather than turn off the heater.
+If they do want to control the heater, they will need to express this explicitly.
+Discussion of understanding human's thoughts through brain-computer interface is beyond the scope of this work.
+We utilize users' ``gesture'' to know their interest for interaction.
+
+Some direct gesture include looking, pointing and reaching, and there is one interesting paper \cite{Merrill:2007:ALP:1758156.1758158} which put sensors to users' ears and fingers to detect their intention.
+And this information is fed back to the system to assist user in browsing physical objects.
+Our approach borrows idea from this by putting sensors on users' body, however, we enable the direct interaction rather than just passive sensing.
+The other major difference is that we choose to use glass form factor for such attention expression.
+Firstly, line of sight is easier to detect on glasses, rather than earpiece.
+And this is also more intuitive to users that when they are looking at something, they might be interested in that, and this information is captured by a on-glass sensor.
+Second, with the release of Google Glass, we believe that it could serve as the enabling hardware platform.
+Though in this paper, we will still present some of our thoughts on ``what if we have an advanced glass and various sensors'', we focus most of the discussion on our prototype which uses IR to express users' interst, XBee radio for device communication, and a slider for users' gesture.
+We admit that improvements are huge to make the device more compact and compelling, the system more robust, and the size smaller and more convenient.
+But for this semester project, our goal (for us) is to explore the possibility of enabling direct interaction with physical objects, learn how to prototype novel interaction device on our own, and apply what we have learned and discussed in class about human computer interaction into a real project.
+
+The contribution in this work is to propose a new way of interacting with physical device and simplify every task.
+Thoughts and discussion on this topic might also be beneficial to researchers on similar work.
+
+The rest of this paper is structured as follows.
+In Sec.\,\ref{sec:related-work}, we will give an overview of some related projects.
+Then we present our approach in detail about the system design and implementation in Sec.\,\ref{sec:design} and Sec.\,\ref{sec:implementation} respectively.
+Then Sec.\,\ref{sec:evaluation} shows our evaluation results and we provide an in-depth discussion (Sec.\,\ref{sec:discussion}) about different kinds of interaction since not all are suitable for our system.
+Sec.\,\ref{sec:conclusion} concludes this paper.
 

### 2013-05-06 doc/report/main.tex
changed to sigchi style “

@@
-\documentclass{sig-alternate}
+\documentclass{sigchi}
 
@@
 \usepackage{xspace}
-\usepackage{hyperref}    % Creates hyperlinks from ref/cite
-\hypersetup{pdfstartview=FitH}
+
 \usepackage{graphicx}    % For importing graphics
 \usepackage{url}         %
-
-\hypersetup{%
-pdftitle={Project Glass 2.0}, pdfauthor={Ben Zhang, Sean Chen}, pdfkeywords={HCI}, bookmarksnumbered, pdfstartview={FitH}, colorlinks,
-citecolor=black, filecolor=black, linkcolor=black, urlcolor=black,
-breaklinks=true,}
 
@@
 
-%\pagenumbering{arabic}  % Arabic page numbers for submission.  Remove this line to eliminate page numbers for the camera ready copy
+\pagenumbering{arabic}  % Arabic page numbers for submission.  Remove this line to eliminate page numbers for the camera ready copy
 
-\begin{document}
+% Use this command to override the default ACM copyright statement (e.g. for preprints). 
+% Consult the conference website for the camera-ready copyright statement.
+\toappear{
+Permission to make digital or hard copies of all or part of this
+work for personal or classroom use is granted without fee provided that 
+copies are not made or distributed for profit or commercial advantage and
+that copies bear this notice and the full citation on the first page. To
+copy otherwise, or republish, to post on servers or to redistribute to 
+lists, requires prior specific permission and/or a fee.\\
+Copyright 20XX ACM X-XXXXX-XX-X/XX/XX ...\$10.00.
+}
 
-% use this command to override the default ACM copyright statement
-% (e.g. for preprints). Remove for camera ready copy.
-%\toappear{Submitted for review to IPSN 2012.}
-% \conferenceinfo{ConfName} {Date, Location}
-% \CopyrightYear{Year} 
-% \crdata{978-1-4503-1227-1/12/04} 
-% \clubpenalty=10000 
-% \widowpenalty = 10000
+% Load basic packages
+\usepackage{balance}  % to better equalize the last page
+\usepackage{graphics} % for EPS, load graphicx instead
+\usepackage{times}    % comment if you want LaTeX's default font
+\usepackage{url}      % llt: nicely formatted URLs
+
+% llt: Define a global style for URLs, rather that the default one
+\makeatletter
+\def\url@leostyle{%
+  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
+\makeatother
+\urlstyle{leo}
 
@@
 
+% Make sure hyperref comes last of your loaded packages, 
+% to give it a fighting chance of not being over-written, 
+% since its job is to redefine many LaTeX commands.
+\usepackage[pdftex]{hyperref}
+\hypersetup{
+pdftitle={Project Glass 2.0}, 
+pdfauthor={Ben Zhang, Sean Chen},
+pdfkeywords={HCI},
+bookmarksnumbered,
+pdfstartview={FitH},
+colorlinks,
+citecolor=black,
+filecolor=black,
+linkcolor=black,
+urlcolor=black,
+breaklinks=true,
+}
+
+% create a shortcut to typeset table headings
+\newcommand\tabhead[1]{\small\textbf{#1}}
+
+% End of preamble. Here it comes the document.
+\begin{document}
+
 \title{Project Glass 2.0 -- Interaction with Physical Devices through Attention}
 
+\numberofauthors{2}
 \author{

### 2013-05-06 doc/report/related.tex
changed to sigchi style “

@@
 
-In this section, we will first summarize some prior work and how it relates to our project in three folds: Tangible User Interface (TUI), Glass-based device, and Attention Aware System (AAS), discussing each in turn.
+In this section, we will summarize some prior work and how it relates to our project in three folds: Tangible User Interface (TUI), Glass-based device, and Attention Aware System (AAS), discussing each in turn.
 
@@
 Our system combines some merits from prior work.
-We also user people's natural looking as attention tracking as in \cite{Merrill:2007:ALP:1758156.1758158}, but we are not limited to browsing or searching physical objects.
-We provide such capability of interacting directly with physical device, in addition to control with a mobile phone or web portal \cite{SmartThings, NinjaBlocks}.
+As has been pointed in Sec.\,\ref{sec:introduction}, we are also using people's natural looking as attention tracking as in \cite{Merrill:2007:ALP:1758156.1758158}, but we are not limited to browsing or searching physical objects.
+We provide such capability of interacting directly with physical devices.
+This also improves some of existing work that is based on smartphones or web portal \cite{SmartThings, NinjaBlocks}.
 
@@
 The most recent and potentially most impactful project \cite{GoogleGlass} integrate many crucial components -- camera, projection, microphone, speaker, IMU and networking -- into a single glass that can achieve virtual reality to an unimaginable extent.
-We do believe this might be a powerful platform for future interaction development.
-However, so far the vision is still being limited to Cyber world.
-Our project explores the potential of using glasses for direct interacting with physical devices, rather than doing ``digital'' tasks.
+We do believe this would become a powerful platform for future interaction development.
+However, so far the vision is still being limited to digital world.
+In contrast, our project explores the potential of using glasses for direct interaction with physical devices, rather than performing ``digital'' tasks.
 
@@
 
-Our work also borrows ideas from these project in that we seek to identify users' attention to select the right physical object to interact.
+These work researched more on how the system should manage or adapt to users' attention.
+While for us, we want user to express their attention directly and use this to address the target for interaction.
 

### 2013-05-06 glass/slider.h
handles accidental release

@@ -4,20 +4,27 @@
 #define gsIDLE 0
 #define gsPRESS 1
 #define gsRELEASE 2
+#define gsPENDING 3
 
 #define SOFTPOT_THREASHOLD 950
 #define SOFTPOT_DELTA_THRESHOLD 20
 
+#define HOLD_THRESHOLD 300   //determine tap vs hold
+#define DTAP_THRESHOLD 600  //two taps within this duration -> dtap
+#define ACCIDENT_THRESHOLD 300
+
 unsigned long start_time = 0;
 unsigned long release_time = 0;
 unsigned long last_release_time = 0;
-int duration_threshold = 300;
-int dtap_duration_threshold = 600;
 
 int sliderState = gsIDLE;
 
 int softpotReading = 0;
 int softpotPin = A0; //analog pin 0
+
+//error handling
+boolean checkDTap = false;
+boolean checkRelease = false;
 
 typedef enum {
   gNONE,
@@ -35,27 +42,63 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
   softpotReading = analogRead(softpotPin);
   // int softpotReading = *pot;
   // DEBUG_PRINTLN(softpotReading);
+
+  if(checkRelease) {
+    //if pressed again in short duration -> accident
+    //else (timeout) -> return release
+    if(millis() - release_time > ACCIDENT_THRESHOLD) {
+      //it's a real release
+      last_release_time = release_time;
+      start_time = 0;
+      release_time = 0;
+      *sVal = 0;
+      *sDelta = 0;
+      checkRelease = false;
+      sliderState = gsRELEASE;
+      return gRELEASE;
+
+    } else {
+      if(softpotReading <= SOFTPOT_THREASHOLD) {
+        //withn duration but have been pressed -> accident
+        DEBUG_PRINTLN("catch accidental release");
+        checkRelease = false;
+        sliderState = gsPRESS;
+        release_time = last_release_time;
+        return gHOVER;
+
+      } else {
+        //within duration but still unpressed -> keep waiting
+        
+        return gHOVER;
+      }
+    }
+
+    
+  }
+
+
   if(softpotReading > SOFTPOT_THREASHOLD ) { // currently not pressed
     if(sliderState == gsPRESS) {
       // previously pressed -> look at duration
-      sliderState = gsRELEASE;
+      // sliderState = gsRELEASE;
       release_time = millis();
       DEBUG_PRINT("pressed duration: ");
[…]

### 2013-05-06 glass/slider.h
handles dtap (hold on to sending tap until confirmed not dtap)

@@ -10,7 +10,7 @@
 #define SOFTPOT_DELTA_THRESHOLD 20
 
 #define HOLD_THRESHOLD 300   //determine tap vs hold
-#define DTAP_THRESHOLD 600  //two taps within this duration -> dtap
+#define DTAP_THRESHOLD 300  //two taps within this duration -> dtap
 #define ACCIDENT_THRESHOLD 300
 
 unsigned long start_time = 0;
@@ -76,6 +76,25 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
     
   }
 
+  if(checkDTap) {
+    //if a tap is sensed, wait till a duration
+    //if no press sensed within threshold -> declare tap
+    //if second release sensed before timeout -> declare dtap
+    if(millis() - release_time > DTAP_THRESHOLD ) {
+      //checkDTap haven't been canceled -> single tap
+      start_time = 0;
+      release_time = 0;
+      *sVal = 0;
+      *sDelta = 0;
+      // sliderState = gsRELEASE;
+      checkDTap = false;
+      return gTAP;
+    } else {
+      //before timeout expire
+      //
+    }
+  }
+
 
   if(softpotReading > SOFTPOT_THREASHOLD ) { // currently not pressed
     if(sliderState == gsPRESS) {
@@ -99,15 +118,23 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
           *sVal = 0;
           *sDelta = 0;
           sliderState = gsRELEASE;
+          checkDTap = false;
           return gD_TAP;
         } else {
+
+          //wait to see if it's a D_TAP
           last_release_time = release_time;
-          start_time = 0;
-          release_time = 0;
-          *sVal = 0;
-          *sDelta = 0;
           sliderState = gsRELEASE;
-          return gTAP;
+          checkDTap = true;
+          return gHOVER;
+
+          // last_release_time = release_time;
+          // start_time = 0;
+          // release_time = 0;
+          // *sVal = 0;
+          // *sDelta = 0;
+          // sliderState = gsRELEASE;
+          // return gTAP;
         }
 
         

### 2013-05-06 server/adapter.py
start on hoverChange -> set volume, TBC

@@ -53,10 +53,10 @@ while True:
 
 	if char == 'u':
 		sa = OSAX()
-		sa.set_volume(100)
+		sa.set_volume(sa.get_volume_settings()+5)
 	if char == 'd':
 		sa = OSAX()
-		sa.set_volume(100)
+		sa.set_volume(sa.get_volume_settings()-5)
 	elif char == 'p':
 		app('System Events').keystroke(' ')
 	

### 2013-05-06 doc/report/design.tex
added system architecture graph
updating texts for design section. changed intro due to new template reference problem. A grammar error in implementation.tex

--- 2013-05-03 doc/report/design.tex
@@
 
-Talks about the system design space.
-What might be available to achieve such a function.
+% this section is used to discuss the design considerations of the system
+% starting from a motivating example, and then extract the key requireements, then about system design
+
+In this section, we will talk about our design considerations in solving the ``addressing'' problem in the smart environment.
+We start by describing the example application, where in an ideal case, how the user should interact with physical devices.
+From that, we draw some design requirements.
+To explore the design space, we propose several possible solutions.
+Some of them are difficult to achieve within a shrot time.
+In next section, we will build a proof-of-concept system, where we have simplified several tasks.
+
+\subsection{Example Application}
+\label{sec:example-application}
+
+Our work is primarily motivated by the observation that in CS294-84 course at Berkeley, the instructor has to go to the switch to turn off the light during presentation.
+The separation of device and their controller makes it possible for people control things that are not reachable (such as the lights on the ceiling).
+However, such indirection introduces additional overhead since people still need to locate the controller and then interact.
+Remote control might mitigate such a problem, but one controller for one device is not a scalable solution.
+And squeezing controllers onto the smartphone display is also not the right approach.
+We will demonstrate this by giving the following Smart Home example.
+
+% {\bf Smart Home}:
+% Sean, maybe you can elaborate on this application more to motivate.
+Imagine in a smart-home environment that most devices are remotely controllable.
+When the users are watching TV shows, they want to turn off the lights and turn up the volume so they can enjoy this memorable time.
+Such ordinary-life tasks are happening everyday, and any simplification will result in significant time savings.
+One ideal approach is that the user expresses his interest of turning off the lights by simply staring at the light, plus a specific gesture.
+Then he turns his head back to the TV, and again simply ``asks'' the TV to turn up the volume.
+When transitioning, the disconnection with the light and the connection with TV happens seamlessly, with the correct feedback for the user.
+And all the magic are resolved by an always-available device; this might be the glasses, rings or watches.
+
+Occasionally, the user needs to perform some advanced configurations to the TV, such as adjusting the white balance, or the brightness.
+Since the need of these complicated tasks are not frequent, and it will be less efficient to embed these features into the always-available device (which is designed to be small and easy-to-access, rather than too flexible).
+The smart-home falls back to a more generic approach which relies on touch screen or other rich input device.
+In this example case, since the user's attention is still on TV, his smartphone may pop up the control menu for advanced usage automatically.
+Such menu could also just pop up on walls or other large interactive displays \cite{unPad:eWallpaper, MSVision}.
+Even in this case, the attention system knows users' intention and saves time for user to address the device to interact.
+
+\subsection{System Requirements}
+\label{sec:system-requirements}
+
+From what we have described in the previous section, the example application is enabled by a few necessary requirements. We will discuss some important ones in this section.
+
+% leads to the adoption of glass
+The first and foremost is the ability to sense and control in an always-available fashion.
+Since the goal is to simplify everyday task in a smart-home or smart-office scenario, any cubersome solution is not desired.
+The control of the device would also have to be simple, not to confuse the user of using it.
+
+Secondly, since screens might not always be available when users are interacting directly with physical device.
+Adequate feedback should assist user understanding what is going on.
+Even if screens are everywhere to assist in the future, the requirements of having proper feedback still hold.
+
+As discussed before, we envision the simple always-available device would not be capable to cover all tasks.
+For those which are really complicated and less likely being executed, we will have to enable the system to fall back.
+
+Besides the previous three, for any human-centric system, additional requirements like responsiveness would also be required, though not explicit discussed through this paper.
+
+\subsection{System Design}
+\label{sec:system-design}
+
+In this section, we will present several design consideration that fulfills the requirements.
+From \cite{Merrill:2007:ALP:1758156.1758158}, they tried the solution of putting IR into earphones, rings to enable looking, pointing and reaching metaphor.
+When we were thinking about how to capture user's attention, we are actually leaning towards the glasses solution.
+Part of the motivation is from Google Glass.
+Generally, glasses are more natural in align with human's light of sight.
+Therefore, sensors on glasses can reflect the users' attention in a more precise way.
+Then given the glass form factor, what might be the right sensor or component to use to obtain users' attention remains a major question for us.
+
+The first possible approach is a pure Google Glass solution.
+Since the glasses come with a camera, sophiscated computer vision algorithms might be able to detect what is in front of the user's sight.
+Following this thoughts, the camera would also be able to tell users' gestures.
+In fact, there are more features which would make this interaction cooler -- the voice recognition is another important source of user input, the projection onto user's eyeballs can provide timely feedback to users' action, the networking capability provide seamless connection with other peripheral devices such as displays.
+And initially, we were also thinking about building a prototype that resembles Google Glass, but the complexity behind of achieving acceptable image processing is huge.
+Not to mention we don't have the Google Glass yet.
+This approach is more an exploration of design space.
+
+Since our target is to validate the thoughts of interacting directly with physical objects through users' attention.
+We might break the initial proposal to several achievable tasks and build the proof-of-concept.
+Our second approach still adopts the glass form factor, but we've turned to other sensors to detect users' line of sight.
+
+Specifically, we choose to use IR.
+When surrounded by tubes, the IR is pretty directional.
+And from our test, it can differentiate objects in {\color{red} 20 (do some experiments here)} centimeters.
+User input can be captured by a slider on glass frame, where it's easy to access (though not hand-free).
+The networking capability is enabled by having an XBee module on the main board.
+We propose the interaction architecture in Fig.\,\ref{fig:sysarch}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{../figs/sysarch.pdf}
+  \caption{System Architecture}
+  \label{fig:sysarch}
+\end{figure}
+
+The dotted lines show some existing approaches that seek to provide access to physical objects through gateways from the broader Internet.
+In our project, we focus on the solid lines where the users can interact with them directly.
+Also as has been pointed out in the requirements, the fall-back scheme for complicated tasks are enabled by the communication between the glasses and the gateway.
+Though simple enough, we find it essentially useful to have this architectural figure in mind when considering users' interaction.
+
+% Lessons learned from \cite{Bellotti:2002:MSS:503376.503450} have also guided us in thinking the Five A's when designing novel interaction system.
 

### 2013-05-06 doc/report/intro.tex
updating texts for design section. changed intro due to new template reference problem. A grammar error in implementation.tex

@@
 According to a study in \cite{Ashbrook:2008:QIM:1357054.1357092}, 4 seconds are required on average just to get a phone out of the pocket or hip holster, not to mention the time to choose the target and initiate interaction.
-This project is trying to propose a new way of interacting with physical objects without losing attention, essentially solving the ``address'' problem in interacting with physical devices around.
+This project is trying to propose a new way of interacting with physical objects without losing attention, essentially solving the ``addressing'' problem in interacting with physical devices around.
 
@@
 The rest of this paper is structured as follows.
+We will first give an overview of some related projects, and discuss the similarities and different of our work from theirs.
+Then we present our approach in detail about the system design and implementation in the following two sections respectively.
+Evaluations and in-depth discussions will be provided before we concludes this paper and talk about potential future work.
+
+% ref didn't work in SIGCHI template
+% The rest of this paper is structured as follows.
 In Sec.\,\ref{sec:related-work}, we will give an overview of some related projects.

### 2013-05-06 server/adapter.py
correct the function used in adaptor to change volume

@@ -14,49 +14,57 @@ availables = glob.glob('/dev/tty.usb*')
 
 arguments = parser.parse_args()
 try:
-	baud = arguments.baud
-	timeout = arguments.timeout
-	print "  baud rate:", baud
-	print "  serial timeout:", timeout
+  baud = arguments.baud
+  timeout = arguments.timeout
+  print "  baud rate:", baud
+  print "  serial timeout:", timeout
 except Exception as e:
-	print e
-	exit(1)
+  print e
+  exit(1)
 
 print "All available ports:" 
 
 index = 1
 for port in availables:
-	print '  ', index, port
-	index += 1
+  print '  ', index, port
+  index += 1
 
 try:
-	num = int(raw_input('Enter the number:'))
-	if num > 0 and num <= len(availables):
-		selected = availables[num-1]
+  num = int(raw_input('Enter the number:'))
+  if num > 0 and num <= len(availables):
+    selected = availables[num-1]
 except Exception as e:
-	print "invalid input!"
-	exit(1)
+  print "invalid input!"
+  exit(1)
 
 
 try:
-	ser = serial.Serial(selected, baud, timeout=timeout)
+  ser = serial.Serial(selected, baud, timeout=timeout)
 except Exception as e:
   print "failed to connect the serial port", e
 
-# not complete yet
-# sa.get_volume_settings() will return the current volumn settings
-while True:
-	char = ser.read()
-	# debugging print out
-	if char != '':
-		print "char read: ", char
 
-	if char == 'u':
-		sa = OSAX()
-		sa.set_volume(sa.get_volume_settings()+5)
-	if char == 'd':
-		sa = OSAX()
-		sa.set_volume(sa.get_volume_settings()-5)
-	elif char == 'p':
-		app('System Events').keystroke(' ')
-	
+def change_volume(sa, amount):
+  current_settings = sa.get_volume_settings()
+  original = -1
+  for key in current_settings:
+    if str(key) == 'k.output_volume':
+      original = current_settings[key]
+  if original == -1:
+    raise Exception('Problem in parsing get_volume_settings output')
+  sa.set_volume(original+amount)
+  
+sa = OSAX()
+
+while True:
+  char = ser.read()
+  # debugging print out
+  if char != '':
+    print "char read: ", char
+  if char == 'u':
+    change_volumne(sa, 5)
+  if char == 'd':
+    change_volumne(sa, -5)
+  elif char == 'p':
+    app('System Events').keystroke(' ')
+  

### 2013-05-06 doc/report/design.tex
update texts in some section, now turn back to coding and presentation preparation

@@
 
-The dotted lines show some existing approaches that seek to provide access to physical objects through gateways from the broader Internet.
+The dotted lines show some existing (even gradually commercialized) approaches that seek to provide access to physical objects through gateways from the broader Internet.
 In our project, we focus on the solid lines where the users can interact with them directly.

### 2013-05-06 doc/report/implementation.tex
updating texts for design section. changed intro due to new template reference problem. A grammar error in implementation.tex
update texts in some section, now turn back to coding and presentation preparation

@@
 
-In this section, we details our implementation on Arduino \cite{Arduino} platform.
-The whole system is comprised of a glass module, numbers of clients, and a gateway.
-We will start by the clients, which responds to different message sent by others.
-The complexity of coordination is pushed to the glass, and only minimal function are provided in the gateway.
+In this section, we detail our implementation on Arduino \cite{Arduino} platform.
+The whole system is comprised of a glass module, numbers of clients, and a gateway, as shown in the architecture figure (Fig.\,\ref{fig:sysarch}) in last section.
+We will start from the clients, whose primary tasks are to interface between the Glass and the devices that are attached to it.
+Then we move on to the implementation of the Glass, who manges users' gesture recognition and communication with clients.
+The functionality in gateway so far is only minimum to communicate with clients, since it's not the primary focus of this project.
 
@@
 The clients are comprised of a main Arduino board, an IR receiver, an XBee radio, and various actuators.
-In our current system, we have a relay which can control the AC power plug, and we have a USB connected computer to control video playing.
+We used Arduino Uno, which has an ATmega328 microcontroller and suffice our purpose in this project.
+We use off-the-shell IR receivers ({\color{red} confirm it is PNA4602}), which outputs 0V (low) on detection of 38 KHz carrier or 5V (high) otherwise.
+The XBee radio, accompanied with an XBee modems, takes charge of the communication with the Glass.
+Additional components are subject to the device being attached to.
+In our current system, we have a relay ({\color{red} need to know relay model}) which can control the AC power plug, and we have a USB connected computer to control video playing.
 
-The main function of client is to respond IR signal, and communicate with glass to receive corresponding commands.
-When there is no glass initiating connection, the client simply lives in {\it IDLE} mode; while when they receive IR signal, indicating the user is expressing interest in interacting with it, the client responds with an XBee acknowledgement and goes into {\it PENDING} state.
+The main function of the client is to respond IR signal, and communicate with the Glass to receive corresponding commands.
+Normally, the client lives in {\it IDLE} mode.
+When they receive IR signals, indicating the user is expressing interest in interacting with it, the client responds with an XBee acknowledgement and goes into {\it PENDING} state.
 In this case, since the client doesn't know how many other clients have also responded to the glass, it will wait until a connection signal.
-If there are multiple clients waiting to be verified, the glass will coordinate all of them by sending the right verifying message.
-The active client that is being verified will blink faster, as visual feedback to users.
-When it receives the connection message, it goes to {\it CONNECTED} state and is ready to take commands.
-To summarize the behavior, we have a state machine in Fig.\,\ref{fig:clientFSM} for illustration.
+If there are multiple clients waiting to be verified by the Glass, the coordination is done at the Glass through sending the right verifying message.
+To provide visual feedback to the user, the active client that is being verified will blink faster.
+When it receives the connecting message, it goes to {\it CONNECTED} state and is now ready to take commands.
+To summarize the clients' behavior, we have a state machine in Fig.\,\ref{fig:clientFSM} for illustration.
 
@@
 
-Within {\it CONNECTED} state, the clients' behaviors are different from each other according to the device it is connected to.
+Within {\it CONNECTED} state, the clients will react differently according to the device it is connected to.
 In our prototype, we have two different types of devices.
-The first is a relay, which can then control the whole AC power supply.
-To control the relay, we only need to set a digital pin to {\it LOW} or {\it HIGH}, and each action is encoded in the command set from the glass.
-The second is a computer that is used to play video.
-And the direct interaction includes adjusting the volume and pause/play.
-These commands are sent to the computer through USB serial data.
+The first is a relay, which can control the whole AC power supply.
+The interface to relay is simply an digital pin.
+We only need to set it to {\it LOW} or {\it HIGH}.
+Such command is transmitted remotely from the Glass through XBee using encoded packet.
+The second tyep is a computer that is used to play video.
+To send command to the computer, we have a USB adaptor and a python script taking charge of the translation.
+We use {\it pyserial} and {\it appscript} in the software.
+And this enables controlling the play and pause of an video, and system-wide volume control.
+
+{\color{red} I arrived here for the second pass - Ben}
 

### 2013-05-06 server/adapter.py

@@ -47,6 +47,7 @@ except Exception as e:
 def change_volume(sa, amount):
   current_settings = sa.get_volume_settings()
   original = -1
+  # hacky way of finding out current volume
   for key in current_settings:
     if str(key) == 'k.output_volume':
       original = current_settings[key]

### 2013-05-07 glass/utils.h
add debug method

@@ -116,10 +116,16 @@ struct XBeePacket readXBeePacket (SoftwareSerial *XBee) {
   /* } */
   while ((*XBee).available()) {
     strArray[i] = (*XBee).read();
+    DEBUG_PRINT(strArray[i]);
     i++;
+    if(i>20) {
+      //abandon the packet
+      return p;
+    }
   }
   strArray[i] = '\0';
   // we appended an '\0', plus an '\n' since it was XBee.println
+  DEBUG_PRINTLN();
   DEBUG_PRINT("got packet, now i=");
   DEBUG_PRINT(i);
   DEBUG_PRINT(", and packet: ");

### 2013-05-07 glass/slider.h
increase threshold so vol control doesn't send as frequent

@@ -7,7 +7,7 @@
 #define gsPENDING 3
 
 #define SOFTPOT_THREASHOLD 950
-#define SOFTPOT_DELTA_THRESHOLD 20
+#define SOFTPOT_DELTA_THRESHOLD 40
 
 #define HOLD_THRESHOLD 300   //determine tap vs hold
 #define DTAP_THRESHOLD 300  //two taps within this duration -> dtap

### 2013-05-07 server/adapter.py
added reset, and update adaptor

@@ -53,7 +53,10 @@ def change_volume(sa, amount):
       original = current_settings[key]
   if original == -1:
     raise Exception('Problem in parsing get_volume_settings output')
-  sa.set_volume(original+amount)
+  print original, amount
+  # a number between 0 to 7
+  # original + amount is 0 to 100
+  sa.set_volume((original+amount) / 100.0 * 7)
   
 sa = OSAX()
 
@@ -63,9 +66,9 @@ while True:
   if char != '':
     print "char read: ", char
   if char == 'u':
-    change_volumne(sa, 5)
+    change_volume(sa, 5)
   if char == 'd':
-    change_volumne(sa, -5)
+    change_volume(sa, -5)
   elif char == 'p':
     app('System Events').keystroke(' ')
   

### 2013-05-07 server/wireless.py

@@ -42,8 +42,9 @@ except Exception as e:
 
 print "All available commands:"
 print "  1, print -- print all id(s)"
-print "  2, on id -- turn on according to id"
-print "  3, off id -- turn off according to id"
+print "  2, reset -- reset all from any state"
+print "  3, on id -- turn on according to id"
+print "  4, off id -- turn off according to id"
 
 
 while True:
@@ -52,6 +53,11 @@ while True:
   if cmd[0] == "print":
     # print all
     ser.write("FFFF")
+    time.sleep(2)
+    print ser.readlines()
+  elif cmd[0] == "reset":
+    # print all
+    ser.write("FFrFFFF0\n")
     time.sleep(2)
     print ser.readlines()
   elif cmd[0] == "on" or cmd[0] == "off":

### 2013-05-07 doc/report/discussion.tex
adding files to the report
added some texts to the report, and a glass FSM figure
finished implementation, and start discussion

--- 2013-05-03 doc/report/design.tex
@@
-\section{Design}
-\label{sec:design}
+\section{Discussion}
+\label{sec:discussion}
 
-Talks about the system design space.
-What might be available to achieve such a function.
+\textcolor{red}{Discuss the problems we have encountered and other random stuffs in this section.}
+
+\subsection{Interaction in Attention}
+\label{sec:inter-attent}
+
+During our brainstorming and discussion of this idea, we have identified some tasks where user might not pay their attention to the device they are controlling.
+One such example is giving presentation.
+The attention of presenter is to interact with the audience, but this is achieved through controlling the slides and projection.
 

### 2013-05-07 doc/report/implementation.tex
finished implementation, and start discussion

@@
 The XBee radio, accompanied with an XBee modems, takes charge of the communication with the Glass.
+LEDs are added as an indicator of the state of clients and visual feedback for users.
 Additional components are subject to the device being attached to.
 In our current system, we have a relay ({\color{red} need to know relay model}) which can control the AC power plug, and we have a USB connected computer to control video playing.
+The prototype is shown in Fig.\,\ref{fig:client}.
+
+{\color{red} add clients photos here}
+\begin{figure}
+  \centering
+%  \includegraphics[width=\linewidth]{../figs/client.png}
+  \caption{Prototype of the clients}
+  \label{fig:client}
+\end{figure}
 
@@
 
-{\color{red} I arrived here for the second pass - Ben}
-
 \subsection{Glass}
@@
 Second, we have to handle user gesture on Glass.
-For a clean design, we separate the gesture detection and XBee transmission into two customized library, and only expose interesting information to the main program.
+From an engineer perspective, we separate the gesture detection and XBee transmission into two customized library files in Arduino.
+And only interesting information is exposed to the main program.
+We use the same Arduino Uno and XBee module as the client.
+In addition, we have a membrane potentiometer (100 mm long) as the slider to take user's input, and the most important IR emitter to broadcast IR signals.
+After we've put the components onto the Glass, we have the prototype shown in Fig.\,\ref{fig:glass}.
 
-The whole work flow is as follows.
-When the user hasn't shown any interest of interaction, the system lives in {\it IDLE} state.
-When ``tap'' is detected, the Glass will broadcast IR signal to inform the devices which is in front of the user (the one that can receive IR signal).
-Then the Glass waits for a specific period of time for acknowledge message sent by clients.
-The timer in {\it WAIT} state now is set to 1 second.
-When timer expires, Glass will go to {\it IDLE} if no client has responded, or {\it CONFIRM} if it receives just one message.
-However, the complicated case is when there are multiple clients that have received IR signal and responded.
-Though we expect such scenario is rare, there is still chance that the physical devices are put together and it is necessary to differentiate them.
-In this case, the system goes to {\it VERIFY} state, and send out corresponding verifying message based on user's gesture.
-Once the user confirmed the selection, it goes to {\it CONNECTED} state where it's ready to send out commands.
-Again, to illustrate the flows, we have the FSM in Fig.\,\ref{fig:glassFSM}.
+{\color{red} add glass photos here}
+\begin{figure}
+  \centering
+%  \includegraphics[width=\linewidth]{../figs/glass.png}
+  \caption{Prototype of the Glass}
+  \label{fig:glass}
+\end{figure}
+
+The whole work flow on the Glass is as follows.
+When the user hasn't shown any interest for interaction, the system stays in {\it IDLE} state.
+When ``tapping'' is detected, the Glass will broadcast IR signals to inform the devices which are in front of the user.
+Since we have the tube around the IR transmitter, the resulting signal coverage is a pretty straight line.
+After the Glass has sent out IR message, at this time, it doesn't know how many devices might respond.
+In this case, the Glass would wait for a specific period of time, and count the number of acknowledgement messages sent by clients.
+To ensure responsiveness to end user, we set the timer in {\it WAIT} state to be 1 second.
+When the timer expires, Glass will go to {\it IDLE} if no client has responded, or {\it CONFIRM} if it receives just one message.
+However, the complicated case is when there are multiple clients that have received the IR signal (because they are near to each other) and responded.
+Though we expect this scenario rarely happens, there is still chance when the physical devices are put together and it is necessary to differentiate them.
+In this case, the system goes to {\it VERIFY} state, and sends out corresponding verifying message based on user's gesture.
+We adopt the slider on glass to change active selection.
+Once the user confirmed the selection, it goes to {\it CONNECTED} state and now the Glass is ready to send out commands.
+We have the FSM in Fig.\,\ref{fig:glassFSM}.
+
+We've configured the XBee modules such that the Glass can talk in both direction with all clients, while clients cannot talk with each other.
+In this way, the Glass would serve as master of this personal area network (PAN), whose ID is 4321 in our scenario.
 
@@
 
-The gateway provides access to the personal area network for computers, and presumably this can be further open to the Internet.
+The gateway provides access to this PAN for computers, and presumably this can be further open to the Internet.
 Since there have already been many products \cite{NinjaBlocks, Lockitron} that essentially function in this way, we do not spend much time on this aspect in our project.
 
-Just for completeness, we wrote a python script that controls a USB XBee adaptor, so that we can remotely control those devices without the Glass and IR initiation.
+For completeness, we wrote a python script that controls a USB XBee adaptor.
+And this XBee module has the same configuration as the Glass, so it serves as the master of the network.
+This enables the remote control of the clients.
 

### 2013-05-07 doc/report/discussion.tex
added some discussion

@@
 
-\textcolor{red}{Discuss the problems we have encountered and other random stuffs in this section.}
+% Discuss the problems we have encountered and other random stuffs in this section
+In this section, we will touch something that is related to our project, or pops up during our thinking that might be beneficial to people who consider working on similar topics.
+We will first talk about the attention-based interaction and its limitation.
+Then specific to our project implementation, we would like to express some of our experience.
 
@@
 
+{\color{red} a very short brainstorming would be able to finish this part.}
+
+\subsection{Experience learned in this project}
+\label{sec:exper-learn-this}
+
+XBee setting up.
+
 %%% Local Variables: 

### 2013-05-18 doc/report/conclusion.tex
update conclusion

--- 2013-02-28 doc/report/intro.tex
@@
-\section{Introduction}
-\label{sec:introduction}
+\section{Conclusion}
+\label{sec:conclusion}
 
-So far, our proposal Looks fairly similar to \cite{Merrill:2007:ALP:1758156.1758158}.
+In this project, we explored a novel way of interacting with physical devices by capturing user's attention.
+The prototype we built demonstrates the possibility of identifying user's line of sight to select and control target appliances.
+Furthermore, we designed a mechanism to intuitively choose from multiple targets when they are close-by.
+
+There are still a few more components that require enhancement: (A) Thorough verifications of the IR strength, angles, and reflecting effects.
+(B) A more robust communication protocol with better error handling.
+(C) A more stable and richer input method.
+(D) User study.
 

### 2013-05-18 doc/report/intro.tex
update intro (TBC)

@@
 Nowadays, you can buy the basic components online \cite{SmartHome, NinjaBlocks} and many hobbyists have already created their own smart environment (see \cite{BRAD} as an example).
-Though various challenges are ahead in the field to build a more robust, scalable, energy-efficient, and user-friendly systems, the general trend of having smart devices around us is undoubted.
-Those smart devices each can be controlled separately (like what is existing now for television and heater), however, with the prevalence of smartphones, many systems \cite{SmartThings, Lockitron} are being written as an application on mobile platforms so that the smartphone act as the bridge of connecting digital world and physical world.
+Though various challenges are ahead in the field to build a more robust, scalable, energy-efficient, and user-friendly systems, the general trend of having smart appliances around us is foreseeable.
+Those smart appliances each can be controlled separately (like what is existing now for television and heater), however, with the prevalence of smartphones, many systems \cite{SmartThings, Lockitron} are being designed as applications on mobile platforms so that the smartphone act as an universal bridge between human and appliances.
 

### 2013-05-18 doc/report/implementation.tex
added glass photo
resolve conflict

@@
 In this section, we detail our implementation on Arduino \cite{Arduino} platform.
-The whole system is comprised of a glass module, numbers of clients, and a gateway, as shown in the architecture figure (Fig.\,\ref{fig:sysarch}) in last section.
+The whole system is comprised of a glass module, numbers of clients, and a gateway, as shown in the architecture figure (Fig.\,\ref{fig:sysarch}) in the last section.
 We will start from the clients, whose primary tasks are to interface between the Glass and the devices that are attached to it.
-Then we move on to the implementation of the Glass, who manges users' gesture recognition and communication with clients.
-The functionality in gateway so far is only minimum to communicate with clients, since it's not the primary focus of this project.
+Then we move on to the implementation of the Glass, which manages user's gesture recognition and communication with the clients.
+The functionality of the gateway so far is confined to only communicate with the clients, since it's not the primary focus of this project.
 
@@
   \centering
-%  \includegraphics[width=\linewidth]{../figs/client.png}
+  \includegraphics[width=\linewidth]{../figs/client.jpg}
   \caption{Prototype of the clients}
@@
 
-The main function of the client is to respond IR signal, and communicate with the Glass to receive corresponding commands.
-Normally, the client lives in {\it IDLE} mode.
-When they receive IR signals, indicating the user is expressing interest in interacting with it, the client responds with an XBee acknowledgement and goes into {\it PENDING} state.
-In this case, since the client doesn't know how many other clients have also responded to the glass, it will wait until a connection signal.
-If there are multiple clients waiting to be verified by the Glass, the coordination is done at the Glass through sending the right verifying message.
-To provide visual feedback to the user, the active client that is being verified will blink faster.
-When it receives the connecting message, it goes to {\it CONNECTED} state and is now ready to take commands.
+The main function of the client is to respond IR signal and communicate with the Glass to receive commands from the user.
+Normally, the client stays in {\it IDLE} mode.
+When it receives an IR signal, suggesting that the user could be interested in interacting with it, the client responds with an XBee acknowledgement and goes into {\it PENDING} state.
+The glass, after seding the IR, will wait for a certain amoount of time.
+If there is only one client responding during the wait, the Glass will directly send a connecting message to the only client.
+When there are multiple clients responding, the Glass will broadcast a verify message with a client ID.
+The client with a matching ID is called the active client and its LED will blink faster then the others as a visual feedback to the user.
+After the user decide which one to connect with, the Glass will send out a connecting message.
+Client turned into {\it CONNECTED} state will be ready to take commands.
 To summarize the clients' behavior, we have a state machine in Fig.\,\ref{fig:clientFSM} for illustration.
@@
 To send command to the computer, we have a USB adaptor and a python script taking charge of the translation.
-We use {\it pyserial} and {\it appscript} in the software.
-And this enables controlling the play and pause of an video, and system-wide volume control.
+We use the {\it pyserial} and {\it appscript} library to trigger the play, pause command of an video, and system-wide volume control.
 
@@
 From an engineer perspective, we separate the gesture detection and XBee transmission into two customized library files in Arduino.
-And only interesting information is exposed to the main program.
+And only relevant information is sent to the main program.
 We use the same Arduino Uno and XBee module as the client.
-In addition, we have a membrane potentiometer (100 mm long) as the slider to take user's input, and the most important IR emitter to broadcast IR signals.
+An IR emitter is placed within a black tube in order to make the signal a straight beam of light.
+Then the the tube is attached to the Glass and adjusted in a proper direction.
+In addition, we have a membrane potentiometer (100 mm long) as the slider to take user's input.
 After we've put the components onto the Glass, we have the prototype shown in Fig.\,\ref{fig:glass}.
@@
   \centering
-%  \includegraphics[width=\linewidth]{../figs/glass.png}
+  \includegraphics[width=0.85\linewidth]{../figs/glass.png}
   \caption{Prototype of the Glass}
@@
 The whole work flow on the Glass is as follows.
-When the user hasn't shown any interest for interaction, the system stays in {\it IDLE} state.
-When ``tapping'' is detected, the Glass will broadcast IR signals to inform the devices which are in front of the user.
-Since we have the tube around the IR transmitter, the resulting signal coverage is a pretty straight line.
-After the Glass has sent out IR message, at this time, it doesn't know how many devices might respond.
-In this case, the Glass would wait for a specific period of time, and count the number of acknowledgement messages sent by clients.
+When the user hasn't shown any interest for interaction, the Glass stays in {\it IDLE} state.
+When ``tapping'' is detected, the Glass will send a beam of IR signal to inform the devices which are in front of the user.
+The Glass waited for a specific period of time and count the number of acknowledgement messages responded by the clients.
 To ensure responsiveness to end user, we set the timer in {\it WAIT} state to be 1 second.
@@
 In this case, the system goes to {\it VERIFY} state, and sends out corresponding verifying message based on user's gesture.
-We adopt the slider on glass to change active selection.
-Once the user confirmed the selection, it goes to {\it CONNECTED} state and now the Glass is ready to send out commands.
+The user can slide back and forth to select the desired target.
+Once the user confirmed the selection and release his finger, it goes to {\it CONNECTED} state and now the Glass is ready to send out commands.
 We have the FSM in Fig.\,\ref{fig:glassFSM}.

### 2013-05-18 doc/report/intro.tex
update intro

@@
 
-However, we argue that traditional graphical user interface (GUI) on mobile platforms still imposes difficulty in interacting with physical objects.
+However, we argue that traditional graphical user interface (GUI) on mobile platforms still imposes difficulty in interacting with real-world objects.
 Imagine when we have all the control panel integrated into iPhone, users then need to browse a list of possible icons and select the right one.
-Even if you can create widget to simplify the browsing overhead, such as on Android platform, the problem still holds since squeezing physically distributed objects onto a-few-inch screen is not the right approach.
+Even if you can create widget to simplify the browsing overhead, such as on Android platform, the problem still holds since squeezing physically distributed objects onto a few-inch screen is not an intuitive mapping.
 In an ideal system, the user should be able to express their interest of interaction in a more direct way.
-On the other hand, taking out the phones (or even simply locating them when at home) is relatively time-consuming.
+In addition, taking out the phones (including locating them, turn on the screen, perhaps entering the passcode, and open the application) is relatively time-consuming.
 According to a study in \cite{Ashbrook:2008:QIM:1357054.1357092}, 4 seconds are required on average just to get a phone out of the pocket or hip holster, not to mention the time to choose the target and initiate interaction.
@@
 And this information is fed back to the system to assist user in browsing physical objects.
-Our approach borrows idea from this by putting sensors on users' body, however, we enable the direct interaction rather than just passive sensing.
+Our approach was inspired by this paper.
+We utilize sensors on users' body, however, we enable the direct interaction rather than just passive sensing.
 The other major difference is that we choose to use glass form factor for such attention expression.
@@
 Though in this paper, we will still present some of our thoughts on ``what if we have an advanced glass and various sensors'', we focus most of the discussion on our prototype which uses IR to express users' interst, XBee radio for device communication, and a slider for users' gesture.
-We admit that improvements are huge to make the device more compact and compelling, the system more robust, and the size smaller and more convenient.
+We recognize that there is a great room of improvement for making the system more robust and the device more compact.
 But for this semester project, our goal (for us) is to explore the possibility of enabling direct interaction with physical objects, learn how to prototype novel interaction device on our own, and apply what we have learned and discussed in class about human computer interaction into a real project.

### 2013-05-18 doc/report/design.tex
update design

@@
 When transitioning, the disconnection with the light and the connection with TV happens seamlessly, with the correct feedback for the user.
-And all the magic are resolved by an always-available device; this might be the glasses, rings or watches.
+And all the magics are resolved by an always-available device; this might be the glasses, rings or watches.
 
@@
 The smart-home falls back to a more generic approach which relies on touch screen or other rich input device.
-In this example case, since the user's attention is still on TV, his smartphone may pop up the control menu for advanced usage automatically.
+In this example, since the user's attention is still on TV, his smartphone may pop up the control menu for advanced usage automatically.
 Such menu could also just pop up on walls or other large interactive displays \cite{unPad:eWallpaper, MSVision}.
-Even in this case, the attention system knows users' intention and saves time for user to address the device to interact.
+The system knows the user's intention and still saves time even if a smartphone is required.
 
@@
 Since the goal is to simplify everyday task in a smart-home or smart-office scenario, any cubersome solution is not desired.
-The control of the device would also have to be simple, not to confuse the user of using it.
+The control of the device would also have to be simple, not to add too much burden to the user.
 
-Secondly, since screens might not always be available when users are interacting directly with physical device.
+Secondly, since screens might not always be available when users are interacting directly with physical appliances.
 Adequate feedback should assist user understanding what is going on.
@@
 
-As discussed before, we envision the simple always-available device would not be capable to cover all tasks.
-For those which are really complicated and less likely being executed, we will have to enable the system to fall back.
+As discussed earlier, the always-available device we envisioned would only focus on solving simple tasks.
+For those tasks which are complicated and less frequently executed, we would still utilize the smartphones as a richer medium for interaction.
 
@@
 
-In this section, we will present several design consideration that fulfills the requirements.
+In this section, we present several design considerations that fulfill the requirements.
 From \cite{Merrill:2007:ALP:1758156.1758158}, they tried the solution of putting IR into earphones, rings to enable looking, pointing and reaching metaphor.
@@
 Part of the motivation is from Google Glass.
-Generally, glasses are more natural in align with human's light of sight.
+Generally, glasses are more natural in align with human's line of sight.
 Therefore, sensors on glasses can reflect the users' attention in a more precise way.
@@
 In fact, there are more features which would make this interaction cooler -- the voice recognition is another important source of user input, the projection onto user's eyeballs can provide timely feedback to users' action, the networking capability provide seamless connection with other peripheral devices such as displays.
-And initially, we were also thinking about building a prototype that resembles Google Glass, but the complexity behind of achieving acceptable image processing is huge.
-Not to mention we don't have the Google Glass yet.
-This approach is more an exploration of design space.
+And initially, we were also thought about building a prototype that resembles Google Glass, but the complexity behind of achieving acceptable image processing is huge, and wiill also cause power concerns.
 
@@
 Specifically, we choose to use IR.
-When surrounded by tubes, the IR is pretty directional.
+When surrounded by dark tubes, the IR is pretty directional.
 And from our test, it can differentiate objects in {\color{red} 20 (do some experiments here)} centimeters.

### 2013-05-19 doc/report/discussion.tex
update discussion

--- 2013-05-03 doc/report/design.tex
@@
-\section{Design}
-\label{sec:design}
+\section{Discussion}
+\label{sec:discussion}
 
-Talks about the system design space.
-What might be available to achieve such a function.
+% Discuss the problems we have encountered and other random stuffs in this section
+% In this section, we will touch something that is related to our project, or pops up during our thinking that might be beneficial to people who consider working on similar topics.
+We will first talk about the attention-based interaction and its limitation.
+Then specific to our project implementation, we would like to express some of our experience.
+
+\textbf{IR sending and receiving:} We found that it is an intuitive and fairly accurate method for detecting the user's attention.
+However, there are two limitations under current design: (1) We want the user be able to control the appliances as long as they can see it.
+However, in some cases where the user is blocked by other objects or in a particular angle, being able to see part of the appliance does not necessarily mean the user can see the IR receiver.
+(2) When the wall is somewhat reflective or has some reflective materials such as white board hanging on it, the IR might bounce back and trigger the appliances in back of the user.
+We are discussing ways to mitigate the problem.
+
+\textbf{XBee:} The communication between XBees are complicated.
+For one, the signals are transmitted one by one instead of in a packet.
+Hence, we need to handle with extra care.
+In addition, sometimes multiple client can be sending signals simultaneously.
+We added a delay to each client that is proportional to their client IDs.
+
+\textbf{Slider:} The slider we ordered is actually a soft potentiometer that returns a value between 0 to 1023 when pressure is applied to it.
+And the returned value reflects the position being pressed along the x-axis.
+To detect tap is easy.
+To detect hold, we need to handle accidentally release since the soft potentiometer requires certain amount of strength and the glass could move when the user is sliding along the temple.
+We also added a bounce handling so that the main program is only notified of slide change if a certain threshold of value is changed on the slider.
+Finally, we added the double tap gesture and it was a bit tricky to tell it from accidentally release.
+We believe that we could purchase a better piece in our next prototype so that we can concentrate more on the interaction design rather than the basic components.
+
+% \subsection{Interaction in Attention}
+% \label{sec:inter-attent}
+
+% During our brainstorming and discussion of this idea, we have identified some tasks where user might not pay their attention to the device they are controlling.
+One such example is giving presentation.
+The attention of presenter is to interact with the audience, but this is achieved through controlling the slides and projection.
+
+% {\color{red} a very short brainstorming would be able to finish this part.}
+
+% \subsection{Experience learned in this project}
+% \label{sec:exper-learn-this}
+
+% XBee setting up.
 

### 2013-05-19 doc/report/evaluation.tex
adding files to the report
added some texts to the report, and a glass FSM figure
update evaluation

--- 2013-05-03 doc/report/design.tex
@@
-\section{Design}
-\label{sec:design}
+\section{Evaluation}
+\label{sec:evaluation}
 
-Talks about the system design space.
-What might be available to achieve such a function.
+\textcolor{red}{Will fill this evaluation when we get to it.}
+Some of the ideas as to how to work on evaluation include: (A) Make a mobile (web) app that can remotely control the appliances through the gateway.
+Give a sequence of tasks to the tester and have him complete the tasks first on smartphone then use the Glass.
+(B) Observe the differences between using LED, Recon screen, and other feedbacks such as audio, and how that affects the user experience.
 

### 2013-05-19 doc/report/discussion.tex
typos, and removed some red texts, also removed .pdf file..., compile the src when you need, don't include pdf file in it

@@
 Hence, we need to handle with extra care.
-In addition, sometimes multiple client can be sending signals simultaneously.
-We added a delay to each client that is proportional to their client IDs.
+In addition, sometimes multiple clients can be sending signals simultaneously.
+We added a delay to each client that is proportional to their client IDs to avoid conflicts.
 
@@
 Finally, we added the double tap gesture and it was a bit tricky to tell it from accidentally release.
-We believe that we could purchase a better piece in our next prototype so that we can concentrate more on the interaction design rather than the basic components.
+We believe that we could use a better piece
+%purchase a better piece 
+in our next prototype so that we can concentrate more on the interaction design rather than the basic components.
 

### 2013-05-19 doc/report/evaluation.tex
typos, and removed some red texts, also removed .pdf file..., compile the src when you need, don't include pdf file in it

@@
 
-\textcolor{red}{Will fill this evaluation when we get to it.}
-Some of the ideas as to how to work on evaluation include: (A) Make a mobile (web) app that can remotely control the appliances through the gateway.
+%\textcolor{red}{Will fill this evaluation when we get to it.}
+Constrained by the time within a semester, we haven't conducted a formal user study.
+Based on our discussions with the instructors, we have some initial plans as to how to work on evaluation -- (A) Make a mobile (web) app that can remotely control the appliances through the gateway.
 Give a sequence of tasks to the tester and have him complete the tasks first on smartphone then use the Glass.
@@
 
+We are planning to have evaluation done within the summer.
 %%% Local Variables: 

### 2013-05-19 doc/report/implementation.tex
update client image
remove some reminder texts
typos, and removed some red texts, also removed .pdf file..., compile the src when you need, don't include pdf file in it

@@
 We used Arduino Uno, which has an ATmega328 microcontroller and suffice our purpose in this project.
-We use off-the-shell IR receivers ({\color{red} confirm it is PNA4602}), which outputs 0V (low) on detection of 38 KHz carrier or 5V (high) otherwise.
+We use off-the-shell IR receivers % ({\color{red} confirm it is PNA4602})
+, which outputs 0V (low) on detection of 38 KHz carrier or 5V (high) otherwise.
 The XBee radio, accompanied with an XBee modems, takes charge of the communication with the Glass.
@@
 Additional components are subject to the device being attached to.
-In our current system, we have a relay ({\color{red} need to know relay model}) which can control the AC power plug, and we have a USB connected computer to control video playing.
-The prototype is shown in Fig.\,\ref{fig:client}.
+In our current system, we have a relay %({\color{red} need to know relay model})
+which can control the AC power plug, and we have a USB connected computer to control video playing. The prototype is shown in Fig.\,\ref{fig:client}.
 
-{\color{red} add clients photos here}
 \begin{figure}
   \centering
-  \includegraphics[width=\linewidth]{../figs/client.jpg}
+  \includegraphics[width=\linewidth]{../figs/client.png}
   \caption{Prototype of the clients}
@@
 
-{\color{red} add glass photos here}
 \begin{figure}

### 2013-08-23 glass/utils.h
refine debug msg

@@ -4,7 +4,7 @@
 #include <SoftwareSerial.h>
 #include "Arduino.h"
 
-// #define DEBUG
+#define DEBUG
 
 #ifdef DEBUG
   #define DEBUG_PRINT(x)  Serial.print(x)
@@ -133,11 +133,9 @@ struct XBeePacket readXBeePacket (SoftwareSerial *XBee) {
     i--;
   }
   // we appended an '\0', plus an '\n' since it was XBee.println
-  DEBUG_PRINTLN();
-  DEBUG_PRINT("got packet, now i=");
   DEBUG_PRINT(i);
-  DEBUG_PRINT(", and packet: ");
-  DEBUG_PRINT(strArray);
+  DEBUG_PRINT(", ");
+  DEBUG_PRINTLN(strArray);
   
   if (i == 10 || i == 9) {
     string_copy(p.id, strArray, 0, 1);

### 2013-08-23 server/target_exp.py

@@ -83,7 +83,7 @@ while True:
       ser.write(out)
     elif line[0:3] == "end":
       print 'writing results'
-      f = open("result"+user_id+".txt", 'w')
+      f = open("result_"+user_id+".txt", 'w')
       for t in result_log:
         print t
         f.write(' '.join(str(s) for s in t) + '\n') 
@@ -123,7 +123,7 @@ while True:
         logResult("wrong_mul", in_msg[0:2])
     elif in_msg[3:9] == "SELCAN":
       logResult("cancel_mul", '')
-      
+
 
 
 

### 2013-08-23 glass/utils.h
refine debug msg

@@ -4,7 +4,7 @@
 #include <SoftwareSerial.h>
 #include "Arduino.h"
 
-#define DEBUG
+// #define DEBUG
 
 #ifdef DEBUG
   #define DEBUG_PRINT(x)  Serial.print(x)
@@ -15,7 +15,7 @@
 #endif
 
 #ifdef DEBUG_TAG
-  #define DEBUG_TAGGING(x, y)  Serial.print(x); Serial.println(y)
+  #define DEBUG_TAGGING(x, y)  Serial.print(x); Serial.print(y)
 #else
   #define DEBUG_TAGGING(x, y) 
 #endif
@@ -133,9 +133,10 @@ struct XBeePacket readXBeePacket (SoftwareSerial *XBee) {
     i--;
   }
   // we appended an '\0', plus an '\n' since it was XBee.println
-  DEBUG_PRINT(i);
-  DEBUG_PRINT(", ");
-  DEBUG_PRINTLN(strArray);
+  // DEBUG_PRINT(i);
+  // DEBUG_PRINT(", ");
+  // DEBUG_PRINTLN(strArray);
+
   
   if (i == 10 || i == 9) {
     string_copy(p.id, strArray, 0, 1);

### 2013-08-23 server/target_exp.py
refine log file name

@@ -83,7 +83,7 @@ while True:
       ser.write(out)
     elif line[0:3] == "end":
       print 'writing results'
-      f = open("result_"+user_id+".txt", 'w')
+      f = open("result_"+user_id[:-1]+".txt", 'w')
       for t in result_log:
         print t
         f.write(' '.join(str(s) for s in t) + '\n') 

### 2013-08-24 server/adapter.py
new interface for serial, add argument for --serial

@@ -12,6 +12,7 @@ parser = argparse.ArgumentParser(description='Adapter that controls laptop throu
 parser.add_argument('--debug', default=False, action='store_true', help='Print More Errors and Launch interactive console on exception or forced exit')
 parser.add_argument('--baud', type=int, action='store', default=9600, help='Specify the baud rate')
 parser.add_argument('--timeout', type=float, action='store', default=1, help='Timeout parameter for serial connection')
+parser.add_argument('--serial', action='store', default=None, help='serial port')
 
 # we know for mac it will show as /dev/usb.tty*, so list all of them and ask user to choose
 
@@ -21,26 +22,30 @@ arguments = parser.parse_args()
 try:
   baud = arguments.baud
   timeout = arguments.timeout
+  serial_port = arguments.serial
   print "  baud rate:", baud
   print "  serial timeout:", timeout
+  print "  serial port:", serial_port
 except Exception as e:
   print e
   exit(1)
 
-print "All available ports:" 
+if serial_port:
+  selected = serial_port
+else:
+  print "All available ports:" 
+  index = 1
+  for port in availables:
+    print '  ', index, port
+    index += 1
 
-index = 1
-for port in availables:
-  print '  ', index, port
-  index += 1
-
-try:
-  num = int(raw_input('Enter the number:'))
-  if num > 0 and num <= len(availables):
-    selected = availables[num-1]
-except Exception as e:
-  print "invalid input!"
-  exit(1)
+  try:
+    num = int(raw_input('Enter the number:'))
+    if num > 0 and num <= len(availables):
+      selected = availables[num-1]
+  except Exception as e:
+    print "invalid input!"
+    exit(1)
 
 try:
   ser = serial.Serial(selected, baud, timeout=timeout)

### 2013-08-26 doc/sigchi14/sigchi14.tex
initialized sigchi14 folder, with some modifications to the sample latex files

\documentclass{sigchi}

% Use this command to override the default ACM copyright statement (e.g. for preprints). 
% Consult the conference website for the camera-ready copyright statement.

%% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is 	granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from permissions@acm.org. \\
% {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
% Copyright \copyright~2014 ACM ISBN/14/04...\$15.00. \\
% DOI string from ACM form confirmation}
%% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)

% Arabic page numbers for submission. 
% Remove this line to eliminate page numbers for the camera ready copy
\pagenumbering{arabic}

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={SIGCHI Conference Proceedings Format},
pdfauthor={LaTeX},
pdfkeywords={SIGCHI, proceedings, archival format},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

% create a shortcut to typeset table headings
\newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{SIGCHI Conference Proceedings Format}

\numberofauthors{3}
\author{
  \alignauthor 1st Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}
  \alignauthor 2nd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}    
  \alignauthor 3rd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}
}

\maketitle

\begin{abstract}
In this paper we describe the formatting requirements for
SIGCHI Conference Proceedings, and this sample file
offers recommendations on writing for the worldwide
SIGCHI readership. Please review this document even if
you have submitted to SIGCHI conferences before, some
format details have changed relative to previous years.
\end{abstract}

\keywords{
	Guides; instructions; author's kit; conference publications;
	keywords should be separated by a semi-colon.
	\textcolor{red}{Mandatory section to be included in your final version.}
}

\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}

See: \url{http://www.acm.org/about/class/1998/}
for more information and the full list of ACM classifiers
and descriptors. 
\textcolor{red}{Mandatory section to be included in your
final version. On the submission page only the classifiers'
letter-number combination will need to be entered.}

\section{Introduction}

This format is to be used for submissions that are
published in the conference proceedings.  We wish to give
this volume a consistent, high-quality appearance. We
therefore ask that authors follow some simple
guidelines. In essence, you should format your paper
exactly like this document. The easiest way to do this is
simply to download a template from the conference web
site, and replace the content with your own material.

\section{Page Size and Columns}

On each page your material (not including the page number) should fit
within a rectangle of 18 x 23.5 cm (7 x 9.25 in.), centered on a US
letter page, beginning 1.9 cm (.75 in.) from the top of the page, with
a .85 cm (.33 in.) space between two 8.4 cm (3.3 in.) columns.  Right
margins should be justified, not ragged. Beware, especially when using
this template on a Macintosh, Word can change these dimensions in
unexpected ways. Please be sure that your PDF is US letter and not
A4. If your PDF or paper are formatted for A4, the submission will be
returned to you to fix.

\section{Typeset Text}

Prepare your submissions on a word processor or typesetter.  Please
note that page layout may change slightly depending upon the printer
you have specified.  \LaTeX\ sometimes will create overfull lines
that extend into columns.  To attempt to combat this, the .cls
file has a command, {\textbackslash}sloppy, that essentially asks
\LaTeX\ to prefer underfull lines with extra whitespace.  For more
details on this, and info on how to control it more finely, check out
{\url{http://www.economics.utoronto.ca/osborne/latex/PMAKEUP.HTM}}.

\subsection{Title and Authors}

Your paper's title, authors and affiliations should run across the
full width of the page in a single column 17.8 cm (7 in.) wide.  The
title should be in Helvetica 18-point bold; use Arial if Helvetica is
not available.  Authors' names should be in Times Roman 12-point bold,
and affiliations in Times Roman 12-point.  For more than three authors,
you may have to place some address information in a footnote, or in a named
section at the end of your paper. Please use full international addresses and
telephone dialing prefixes.  Leave one 10-pt line of white space below the last
line of affiliations.

\subsection{Abstract and Keywords}

Every submission should begin with an abstract of about 150 words,
followed by a set of keywords. The abstract and keywords should be
placed in the left column of the first page under the left half of the
title. The abstract should be a concise statement of the problem,
approach and conclusions of the work described.  It should clearly
state the paper's contribution to the field of HCI.

The first set of keywords will be used to index the paper in the
proceedings. The second set are used to catalogue the paper in the ACM
Digital Library. The latter are entries from the ACM Classification
System~\cite{acm_categories}.  In general, it should only be necessary
to pick one or more of the H5 subcategories, see
\url{http://www.acm.org/class/1998/ccs98.html}

\subsection{Normal or Body Text}

Please use a 10-point Times Roman font or, if this is unavailable,
another proportional font with serifs, as close as possible in
appearance to Times Roman 10-point. The Press 10-point font available
to users of Script is a good substitute for Times Roman. If Times
Roman is not available, try the font named Computer Modern Roman. On a
Macintosh, use the font named Times and not Times New Roman. Please
use sans-serif or non-proportional fonts only for special purposes,
such as headings or source code text.

\subsection{First Page Copyright Notice}

Leave 3 cm (1.25 in.) of blank space for the copyright notice at the
bottom of the left column of the first page. In this template a
floating text box will automatically generate the required space. Note
however that the text box is anchored to the \textbf{ABSTRACT}
heading, so if that heading is deleted the text box will disappear as
well.  You can replace the default copyright notice by uncommenting
the {\textbackslash}toappear block at the beginning of the document
and inserting your own text, for example, for versions under review.

\subsection{Subsequent Pages}

On pages beyond the first, start at the top of the page and continue
in double-column format.  The two columns on the last page should be
of equal length.

\begin{figure}[!h]
\centering
\includegraphics[width=0.9\columnwidth]{Figure1}
\caption{With Caption Below, be sure to have a good resolution image
  (see item D within the preparation instructions).}
\label{fig:figure1}
\end{figure}

\subsection{References and Citations}

Use a numbered list of references at the end of the article, ordered
alphabetically by first author, and referenced by numbers in brackets
\cite{ethics,
  Klemmer:2002:WSC:503376.503378,
  Mather:2000:MUT,
  Zellweger:2001:FAO:504216.504224}. For
papers from conference proceedings, include the title of the paper and
an abbreviated name of the conference (e.g., for Interact 2003
proceedings, use \textit{Proc. Interact 2003}). Do not include the
location of the conference or the exact date; do include the page
numbers if available. See the examples of citations at the end of this
document. Within this template file, use the \texttt{References} style
for the text of your citation.

Your references should be published materials accessible to the
public.  Internal technical reports may be cited only if they are
easily accessible (i.e., you provide the address for obtaining the
report within your citation) and may be obtained by any reader for a
nominal fee.  Proprietary information may not be cited. Private
communications should be acknowledged in the main text, not referenced
(e.g., ``[Robertson, personal communication]'').

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    \tabhead{Objects} &
    \multicolumn{1}{|p{0.3\columnwidth}|}{\centering\tabhead{Caption --- pre-2002}} &
    \multicolumn{1}{|p{0.4\columnwidth}|}{\centering\tabhead{Caption --- 2003 and afterwards}} \\
    \hline
    Tables & Above & Below \\
    \hline
    Figures & Below & Below \\
    \hline
  \end{tabular}
  \caption{Table captions should be placed below the table.}
  \label{tab:table1}
\end{table}

\section{Sections}

The heading of a section should be in Helvetica 9-point bold, all in
capitals. Use Arial if Helvetica is not available. Sections should
not be numbered.

\subsection{Subsections}

Headings of subsections should be in Helvetica 9-point bold with
initial letters capitalized.  For
sub-sections and sub-subsections, a word like \emph{the} or \emph{of}
is not capitalized unless it is the first word of the heading.)

\subsubsection{Sub-subsections}

Headings for sub-subsections should be in Helvetica 9-point italic
with initial letters capitalized.  Standard {\textbackslash}section,
{\textbackslash}subsection, and {\textbackslash}subsubsection commands
will work fine.

\section{Figures/Captions}

Place figures and tables at the top or bottom of the appropriate
column or columns, on the same page as the relevant text
(see Figure~\ref{fig:figure1}). A figure or table may extend across both
columns to a maximum width of 17.78 cm (7 in.).

Captions should be Times New Roman 9-point bold.  They should be numbered (e.g.,
``Table~\ref{tab:table1}'' or ``Figure~\ref{fig:figure2}''), centered
and placed beneath the figure or table.  Please note that the words
``Figure'' and ``Table'' should be spelled out (e.g., ``Figure''
rather than ``Fig.'') wherever they occur.

Papers and notes may use color figures, which are included in the page
limit; the figures must be usable when printed in black and white in
the proceedings.  The paper may be accompanied by a short video figure
up to five minutes in length.  However, the paper should stand on its
own without the video figure, as the video may not be available to
everyone who reads the paper.

\section{Language, Style and Content}

The written and spoken language of SIGCHI is English. Spelling and
punctuation may use any dialect of English (e.g., British, Canadian,
US, etc.) provided this is done consistently. Hyphenation is
optional. To ensure suitability for an international audience, please
pay attention to the following:

\begin{itemize}
\item Write in a straightforward style.
\item Try to avoid long or complex sentence structures.
\item Briefly define or explain all technical terms that may be
  unfamiliar to readers.
\item Explain all acronyms the first time they are used in your text---e.g.,
``Digital Signal Processing (DSP)''.
\item Explain local references (e.g., not everyone knows all city
  names in a particular country).
\item Explain ``insider'' comments. Ensure that your whole audience
  understands any reference whose meaning you do not describe (e.g.,
  do not assume that everyone has used a Macintosh or a particular
  application).
\item Explain colloquial language and puns. Understanding phrases like
  ``red herring'' may require a local knowledge of English.  Humor and
  irony are difficult to translate.
\item Use unambiguous forms for culturally localized concepts, such as
  times, dates, currencies and numbers (e.g., ``1-5-97'' or ``5/1/97''
  may mean 5 January or 1 May, and ``seven o'clock'' may mean 7:00 am or
  19:00).  For currencies, indicate equivalences---e.g., ``Participants
  were paid 10,000 lire, or roughly \$5.''
\item Be careful with the use of gender-specific pronouns (he, she)
  and other gendered words (chairman, manpower, man-months). Use
  inclusive language that is gender-neutral (e.g., she or he, they,
  s/he, chair, staff, staff-hours,
  person-years). See~\cite{Schwartz:1995:GBF} for further advice and
  examples regarding gender and other personal attributes.
\item If possible, use the full (extended) alphabetic character set
  for names of persons, institutions, and places (e.g.,
  Gr{\o}nb{\ae}k, Lafreni\'ere, S\'anchez, Universit{\"a}t,
  Wei{\ss}enbach, Z{\"u}llighoven, \r{A}rhus, etc.).  These characters
  are already included in most versions of Times, Helvetica, and Arial
  fonts.
\end{itemize}

\section{Accessibility}
The Executive Council of SIGCHI has committed to making SIGCHI conferences more inclusive for researchers, practitioners, and educators with disabilities. As a part of this goal, the all authors are asked to work on improving the accessibility of their submissions. Specifically, we encourage authors to carry out the following five steps:
\begin{enumerate}
	\item Add alternative text to all figures
	\item Mark table headings
	\item Add tags to the PDF
	\item Verify the default language
	\item Set the tab order to ``Use Document Structure''
\end{enumerate}
Unfortunately good tools do not yet exist to create tagged PDF files from Latex. LaTeX users will need to carry out all of the above steps in the PDF directly using Adobe Acrobat, after the PDF has been generated.

For more information and links to instructions and resources, please see:
{\url{http://chi2014.acm.org/authors/guide-to-an-accessible-submission}}.

\section{Page Numbering, Headers and Footers}

Please submit your anonymous version for reviewing with page numbers
centered in the footer.  These must be removed in the final version of
accepted papers, as page numbers, headers, and footers will be added
by the conference printers.  Comment out the {\textbackslash}pagenumbering
command at the top of the document to remove page numbers.

\section{Producing and Testing PDF Files}

We recommend that you produce a PDF version of your submission well
before the final deadline.  Your PDF file must be ACM DL
Compliant. The requirements for an ACM Compliant PDF are available at:
{\url{http://www.sheridanprinting.com/typedept/ACM-distilling-settings.htm}}.

Test your PDF file by viewing or printing it with the same software we
will use when we receive it, Adobe Acrobat Reader Version 7. This is
widely available at no cost from~\cite{acrobat}.  Note that most
reviewers will use a North American/European version of Acrobat
reader, which cannot handle documents containing non-North American or
non-European fonts (e.g. Asian fonts).  Please therefore do not use
Asian fonts, and verify this by testing with a North American/European
Acrobat reader (obtainable as above). Something as minor as including
a space or punctuation character in a two-byte font can render a file
unreadable.

\section{Blind Review}

For archival submissions, CHI requires a ``blind review.'' To prepare
your submission for blind review, remove author and institutional
identities in the title and header areas of the paper. You may also
need to remove part or all of the Acknowledgments text.  Further
suppression of identity in the body of the paper and references is
left to the authors' discretion. For more details, see the submission
guidelines and checklist for your submission category.

\section{Conclusion}

It is important that you write for the SIGCHI audience.  Please read
previous years' Proceedings to understand the writing style and
conventions that successful authors have used.  It is particularly
important that you state clearly what you have done, not merely what
you plan to do, and explain how your work is different from previously
published work, i.e., what is the unique contribution that your work
makes to the field?  Please consider what the reader will learn from
your submission, and how they will find your work useful.  If you
write with these questions in mind, your work is more likely to be
successful, both in being accepted into the Conference, and in
influencing the work of our field.

\section{Acknowledgments}

We thank CHI, PDC and CSCW volunteers, and all publications support
and staff, who wrote and provided helpful comments on previous
versions of this document.  Some of the references cited in this paper
are included for illustrative purposes only.  \textbf{Don't forget
to acknowledge funding sources as well}, so you don't wind up
having to correct it later.

% Balancing columns in a ref list is a bit of a pain because you
% either use a hack like flushend or balance, or manually insert
% a column break.  http://www.tex.ac.uk/cgi-bin/texfaq2html?label=balance
% multicols doesn't work because we're already in two-column mode,
% and flushend isn't awesome, so I choose balance.  See this
% for more info: http://cs.brown.edu/system/software/latex/doc/balance.pdf
%
% Note that in a perfect world balance wants to be in the first
% column of the last page.
%
% If balance doesn't work for you, you can remove that and
% hard-code a column break into the bbl file right before you
% submit:
%
% http://stackoverflow.com/questions/2149854/how-to-manually-equalize-columns-
% in-an-ieee-paper-if-using-bibtex
%
% Or, just remove \balance and give up on balancing the last page.
%
\balance

\section{References format}
References must be the same font size as other body text.
% REFERENCES FORMAT
% References must be the same font size as other body text.

\bibliographystyle{acm-sigchi}
\bibliography{sigchi14}
\end{document}

### 2013-09-02 doc/sigchi14/device.tex
bjoern's pass at overall narrative

\section{Hardware Device}

Our prototype consists of a Google Glass Explorer Edition head-worn computing device, augmented with an infrared emitter that is mounted on the frame, pointing out in the direction of the wearer's view (Figure~\ref{fig:glass}). The IR emitter LED is mounted in an opaque hollow tube, that restricts the outgoing angle of illumination. The angle can be adjusted by moving the LED in or out of the tube.

Glass communicates a device id over the IR channel analogous to Patel's approach~\cite{patel_2-way_2003}. Target appliances have IR receivers and offer immediate visual feedback by toggling an LED whenever a valid id is received.

When users confirm that, further communication switches over to a ZigBee wireless network so that line of sight to the target is no longer needed.

In our prototype, Glass communicates over Bluetooth to an additional microcontroller PCB the user has to wear. This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame. This architecture was mostly chosen for reasons of expediency. Future headmounted devices could clearly integrate IR emitters; the choice of local area wireless technology could also change . \bjoern{why not all bluetooth? There's probably a pairing/latency argument to be made here.}

\subsection{Device Characterization}
We characterized the usable range and selection accuracy... (see Figure~\ref{fig:coverage} and Table~\ref{table:measurements}).
\bjoern{insert text from google doc here}

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/glass-with-ir}
\caption{Our augmented Glass prototype has a frame-mounted infrared emitter.}
\label{fig:glass}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/glass-ir-coverage}
\caption{Our augmented Glass prototype has a frame-mounted infrared emitter.}
\label{fig:coverage}
\end{figure}

\begin{table}
    \begin{tabular}{l|lllll}
    distance/ depth & \ft2       & \ft4       & \ft8   & \ft{12}  & \ft{16}  \\ \hline
    \inch{0}                     & $74\degree$ & $78\degree$ & N/A  & N/A  & N/A  \\
  \inch{0.5}                   & $60\degree$ & $48\degree$     & $28\degree$ & $22\degree$ & $16\degree$ \\
    \inch{1.0}                     & $46\degree$     & $36\degree$     & $26\degree$ & $18\degree$ & $10\degree$ \\
    \inch{1.5}                   & $36\degree$     & $32\degree$     & $18\degree$ & $14\degree$ & $6\degree$  \\
    \end{tabular}
    \caption{Measured IR coverage angles $\Theta$ at different target distances and different depths of IR emitter inside shielding tube.}
    \label{table:measurements}

\end{table}

### 2013-09-02 doc/sigchi14/introduction.tex
bjoern's pass at overall narrative

\section{Introduction}
\bjoern{What's the overall motivation for the project? What problem does it address?}
Increasingly, devices and services in our built environment such as intelligent lighting, televisions, HVAC systems, laundry machines, blinds or kitchen appliances are networked and can be controlled remotely.

The proliferation of smart, controllable devices raises the question of how to best interact with them. Prior work has introduced several approaches that use handheld mobile devices as {\em universal remote controls} to select and control appliances in the proximate environment- e.g., through pointing at such devices to select them, and through touch-screen or button interfaces to send control commands~\cite{}. Commercial universal control applications for smart phones are also available~\cite{}, though they usually do not offer spatial selection of target devices, forcing users to browse through lists of configured devices instead.

\bjoern{what are the drawbacks of this existing work?}
Some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is split between looking down at a screen and out at the device to-be-controlled. \bjoern{these all sound kind of weak.} These general limitations have led researchers to investigate ``always-available'' physiological input modalities on one hand~\cite{saponas}, and wearable computing devices~\cite{} on the other hand.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/teaser}
\caption{Teaser figure. Needs a caption.}
\label{fig:teaser}
\end{figure}

In this paper, we introduce a novel method for selecting and controlling smart appliances in physical spaces though the use of a head-worn computing device with near-eye display and wireless communication. We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose. Users first look in the direction of the device they wish to control to acquire it (e.g., at a lamp for controlling lighting, or at a speaker to change music playback volume). If multiple devices fall within communication range, an on-screen disambiguation dialog lets users select a single target. Once acquired, a device specific control UI enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).

Our hardware relies on infrared communication between Glass and target devices to establish a connection; and on wireless ZigBee radio communication to exchange control messages.  Target devices similarly have IR receivers and ZigBee radios. These choices provide an orientation-specific  step selection appropriate for the limitations of head orientation; once acquired, users do not have to look at the target device for control commands.

To understand the system performance, usability, and user experience of head-orientation targeting and head-mounted displays for device control, we first report measurements of the target range and accuracy of our device. We then conduct a comparative study of device acquisition time and error of different interface variants. We find that target acquisition through head orientation is faster than selecting items from a list, given the constraints of linear input using a head-worn touch controller. Finally, we report high-level feedback from $n$ users who use our system for home automation tasks.

### 2013-09-02 doc/sigchi14/relatedwork.tex
bjoern's pass at overall narrative

\section{Related Work}
Relevant prior work exists in the areas of remote control of physical appliances and evaluations of pointing in physical space. We discuss each in turn.

\subsection{Remote Control of Physical Appliances}
A number of researchers propose to use different handheld devices to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012} or exchange other information with smart infrastructure like reading sensor netowrk data~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. Key design decisions are the method by which a target device is selected; and the method by which it is then later controlled or configured.

Several techniques use laser pointers to directly point at objects of interest. This has the advantage of also providing immediate visual feedback to the user in the shape of the laser dot. Beigl's early AIDA handheld combines laser pointing for selection with IR communication to exchange commands~\cite{beigl_point_1999}. Patel extends this technique by modulating the laser light to communicate the controllers identity~\cite{patel_2-way_2003}. Kemp et al. use a laser pointer to indicate to robots which item to pick up in a room~\cite{kemp_point-and-click_2008}.

The XWand~\cite{wilson_xwand:_2003} determines its absolute position and orientation and uses a virtual room model to select target devices. Position is determined through two ceiling-mounted cameras; orientation is determined using a built-in IMU. Users can employ physical gestures or utter speech commands to control selected devices. This technique requires room instrumentation and an up-to-date virtual model of device locations. The Tricorder~\cite{lifton_tricorder:_2007} uses IMU orientation coupled with room-localization based on received signal strength indicators (RSSI) to estimate what a user is pointing at.

Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in PiControl~\cite{schmidt_picontrol:_2012}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Tags like QR codes can also be affixed to devices and read by cameras. Common tagging systems are optimized to be read from a close distance, though it is possible to redesign codes that can be read further away (by encoding less information)~\cite{lowcostsensing-uist13}. 
%However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.

Standard infrared remote controls for televisions and AV equipment tend to use wide-angle LEDs - since they are only meant to control a single device, there is no ambiguity  and it is unnecessary to have the user point precisely.

Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing - the user looks at the target device. Selection techniques with very small selectors such as laser pointing are less appropriate for head-mounted applications. Head orientation only indicates a general area of visual interest. It does not necessarily match gaze orientation as the eyes can be positioned relative to the head using the extraocular muscles. We therefore select an infrared LED - a source with a wider angle of illumination, but restrict its angle to be narrower than in general purpose IR applications.

\bjoern{Other citations to cover here: Raskar et al's radio frequency identity and geometry~\cite{raskar_rfig_2004}.}

\subsection{Evaluation of room-scale selection}
Pausch et al.'s early investigation of head-mounted displays compared orientation tracking of the head to handheld orientation control for a target acquisition task in a virtual reality room shown on a head-mounted display~\cite{pausch_user_1993}. They found a clear performance benefit for head-mounted tracking.

On the other hand, Card et al. experimentally determined that the bandwidth of neck muscles is much lower than that of arm, wrist or finger muscle groups~\cite{card_morphological_1991}, which limits the performance of any head orientation-based interaction scheme. However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques. Compared to screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower.

Myers et al compared different methods of interacting with displays at a distance~\cite{myers_interacting_2002} and quantified selection time and jitter or position error when using laser pointers. Various other techniques outperformed laser pointers.

Rukzio found that users strongly preferred either touching a mobile device to a target appliance to initiate interacting or pointing at a distance to ``scanning'' (initiating a Bluetooth scan and then selecting a target appliance from a list of scan results on the phone )~\cite{rukzio_experimental_2006}.

\subsection{Augmented Reality Interfaces}
\bjoern{most of the prior headmounted work is in the AR space. I don't know much about it yet.}

### 2013-09-02 doc/sigchi14/sigchi14.tex
bjoern's pass at overall narrative

@@
 \usepackage{url}      % llt: nicely formatted URLs
+\usepackage{gensymb}
+% see http://tex.stackexchange.com/questions/46055/typesetting-with-inch-symbols-and-sizes-in-inches
+\usepackage{mathpazo,amsmath}
+\def\inch#1{#1''}
+\def\ft#1{#1'\thinspace}
 
@@
 
+%use these commands while writing
+\newcommand {\bjoern}[1]{{\color{red}\bf{BH: #1}\normalfont}}
+\newcommand {\ben}[1]{{\color{blue}\bf{BZ: #1}\normalfont}}
+\newcommand {\sean}[1]{{\color{blue}\bf{SC: #1}\normalfont}}
+\newcommand {\claire}[1]{{\color{blue}\bf{CT: #1}\normalfont}}
+
+%comment out the above and uncomment these for final submit
+%\newcommand {\bjoern}[1]{}
+%\newcommand {\ben}[1]{}
+%\newcommand {\sean}[1]{}
+%\newcommand {\claire}[1]{}
+
 % End of preamble. Here it comes the document.
@@
 
-\title{SIGCHI Conference Proceedings Format}
+\title{Selecting and Interacting with Smart Devices Through a Head-Mounted Display with Infrared Targeting}
 
-\numberofauthors{3}
+\numberofauthors{1}
 \author{
-  \alignauthor 1st Author Name\\
-    \affaddr{Affiliation}\\
-    \affaddr{Address}\\
-    \email{e-mail address}\\
-    \affaddr{Optional phone number}
-  \alignauthor 2nd Author Name\\
-    \affaddr{Affiliation}\\
-    \affaddr{Address}\\
-    \email{e-mail address}\\
-    \affaddr{Optional phone number}    
-  \alignauthor 3rd Author Name\\
-    \affaddr{Affiliation}\\
-    \affaddr{Address}\\
-    \email{e-mail address}\\
-    \affaddr{Optional phone number}
+  \alignauthor Anonymous for submission
+%  \alignauthor 3rd Author Name\\
+%    \affaddr{Affiliation}\\
+%    \affaddr{Address}\\
+%    \email{e-mail address}\\
+%    \affaddr{Optional phone number}
 }
@@
 \begin{abstract}
-In this paper we describe the formatting requirements for
-SIGCHI Conference Proceedings, and this sample file
-offers recommendations on writing for the worldwide
-SIGCHI readership. Please review this document even if
-you have submitted to SIGCHI conferences before, some
-format details have changed relative to previous years.
+Insert abstract
 \end{abstract}
@@
 \keywords{
-	Guides; instructions; author's kit; conference publications;
-	keywords should be separated by a semi-colon.
-	\textcolor{red}{Mandatory section to be included in your final version.}
+	smart devices; universal remote control; wearable computing; glass
 }
@@
 
-See: \url{http://www.acm.org/about/class/1998/}
-for more information and the full list of ACM classifiers
-and descriptors. 
-\textcolor{red}{Mandatory section to be included in your
-final version. On the submission page only the classifiers'
-letter-number combination will need to be entered.}
+%$See: \url{http://www.acm.org/about/class/1998/}
+%for more information and the full list of ACM classifiers
+%and descriptors. 
+%\textcolor{red}{Mandatory section to be included in your
+%final version. On the submission page only the classifiers'
+%letter-number combination will need to be entered.}
 
-\section{Introduction}
+\input{introduction}
+\input{relatedwork}
+\input{device}
+\input{targeting}
+\input{control}
+\input{conclusion}
 
-This format is to be used for submissions that are
-published in the conference proceedings.  We wish to give
-this volume a consistent, high-quality appearance. We
-therefore ask that authors follow some simple
-guidelines. In essence, you should format your paper
-exactly like this document. The easiest way to do this is
-simply to download a template from the conference web
-site, and replace the content with your own material.
-
-\section{Page Size and Columns}
-
-On each page your material (not including the page number) should fit
-within a rectangle of 18 x 23.5 cm (7 x 9.25 in.), centered on a US
-letter page, beginning 1.9 cm (.75 in.) from the top of the page, with
-a .85 cm (.33 in.) space between two 8.4 cm (3.3 in.) columns.  Right
-margins should be justified, not ragged. Beware, especially when using
-this template on a Macintosh, Word can change these dimensions in
-unexpected ways. Please be sure that your PDF is US letter and not
-A4. If your PDF or paper are formatted for A4, the submission will be
-returned to you to fix.
-
-\section{Typeset Text}
-
-Prepare your submissions on a word processor or typesetter.  Please
-note that page layout may change slightly depending upon the printer
-you have specified.  \LaTeX\ sometimes will create overfull lines
-that extend into columns.  To attempt to combat this, the .cls
-file has a command, {\textbackslash}sloppy, that essentially asks
-\LaTeX\ to prefer underfull lines with extra whitespace.  For more
-details on this, and info on how to control it more finely, check out
-{\url{http://www.economics.utoronto.ca/osborne/latex/PMAKEUP.HTM}}.
-
-\subsection{Title and Authors}
-
-Your paper's title, authors and affiliations should run across the
-full width of the page in a single column 17.8 cm (7 in.) wide.  The
-title should be in Helvetica 18-point bold; use Arial if Helvetica is
-not available.  Authors' names should be in Times Roman 12-point bold,
-and affiliations in Times Roman 12-point.  For more than three authors,
-you may have to place some address information in a footnote, or in a named
-section at the end of your paper. Please use full international addresses and
-telephone dialing prefixes.  Leave one 10-pt line of white space below the last
-line of affiliations.
-
-\subsection{Abstract and Keywords}
-
-Every submission should begin with an abstract of about 150 words,
-followed by a set of keywords. The abstract and keywords should be
-placed in the left column of the first page under the left half of the
-title. The abstract should be a concise statement of the problem,
-approach and conclusions of the work described.  It should clearly
-state the paper's contribution to the field of HCI.
-
-The first set of keywords will be used to index the paper in the
-proceedings. The second set are used to catalogue the paper in the ACM
-Digital Library. The latter are entries from the ACM Classification
-System~\cite{acm_categories}.  In general, it should only be necessary
-to pick one or more of the H5 subcategories, see
-\url{http://www.acm.org/class/1998/ccs98.html}
-
-\subsection{Normal or Body Text}
-
-Please use a 10-point Times Roman font or, if this is unavailable,
-another proportional font with serifs, as close as possible in
-appearance to Times Roman 10-point. The Press 10-point font available
-to users of Script is a good substitute for Times Roman. If Times
-Roman is not available, try the font named Computer Modern Roman. On a
-Macintosh, use the font named Times and not Times New Roman. Please
-use sans-serif or non-proportional fonts only for special purposes,
-such as headings or source code text.
-
-\subsection{First Page Copyright Notice}
-
-Leave 3 cm (1.25 in.) of blank space for the copyright notice at the
-bottom of the left column of the first page. In this template a
-floating text box will automatically generate the required space. Note
-however that the text box is anchored to the \textbf{ABSTRACT}
-heading, so if that heading is deleted the text box will disappear as
-well.  You can replace the default copyright notice by uncommenting
-the {\textbackslash}toappear block at the beginning of the document
-and inserting your own text, for example, for versions under review.
-
-\subsection{Subsequent Pages}
-
-On pages beyond the first, start at the top of the page and continue
-in double-column format.  The two columns on the last page should be
-of equal length.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=0.9\columnwidth]{Figure1}
-\caption{With Caption Below, be sure to have a good resolution image
-  (see item D within the preparation instructions).}
-\label{fig:figure1}
-\end{figure}
-
-\subsection{References and Citations}
-
-Use a numbered list of references at the end of the article, ordered
-alphabetically by first author, and referenced by numbers in brackets
-\cite{ethics,
-  Klemmer:2002:WSC:503376.503378,
-  Mather:2000:MUT,
-  Zellweger:2001:FAO:504216.504224}. For
-papers from conference proceedings, include the title of the paper and
-an abbreviated name of the conference (e.g., for Interact 2003
-proceedings, use \textit{Proc. Interact 2003}). Do not include the
-location of the conference or the exact date; do include the page
-numbers if available. See the examples of citations at the end of this
-document. Within this template file, use the \texttt{References} style
-for the text of your citation.
-
-Your references should be published materials accessible to the
-public.  Internal technical reports may be cited only if they are
-easily accessible (i.e., you provide the address for obtaining the
-report within your citation) and may be obtained by any reader for a
-nominal fee.  Proprietary information may not be cited. Private
-communications should be acknowledged in the main text, not referenced
-(e.g., ``[Robertson, personal communication]'').
-
-\begin{table}
-  \centering
-  \begin{tabular}{|c|c|c|}
-    \hline
-    \tabhead{Objects} &
-    \multicolumn{1}{|p{0.3\columnwidth}|}{\centering\tabhead{Caption --- pre-2002}} &
-    \multicolumn{1}{|p{0.4\columnwidth}|}{\centering\tabhead{Caption --- 2003 and afterwards}} \\
-    \hline
-    Tables & Above & Below \\
-    \hline
-    Figures & Below & Below \\
-    \hline
-  \end{tabular}
-  \caption{Table captions should be placed below the table.}
-  \label{tab:table1}
-\end{table}
-
-\section{Sections}
-
-The heading of a section should be in Helvetica 9-point bold, all in
-capitals. Use Arial if Helvetica is not available. Sections should
-not be numbered.
-
-\subsection{Subsections}
-
-Headings of subsections should be in Helvetica 9-point bold with
-initial letters capitalized.  For
-sub-sections and sub-subsections, a word like \emph{the} or \emph{of}
-is not capitalized unless it is the first word of the heading.)
-
-\subsubsection{Sub-subsections}
-
-Headings for sub-subsections should be in Helvetica 9-point italic
-with initial letters capitalized.  Standard {\textbackslash}section,
-{\textbackslash}subsection, and {\textbackslash}subsubsection commands
-will work fine.
-
-\section{Figures/Captions}
-
-Place figures and tables at the top or bottom of the appropriate
-column or columns, on the same page as the relevant text
-(see Figure~\ref{fig:figure1}). A figure or table may extend across both
-columns to a maximum width of 17.78 cm (7 in.).
-
-Captions should be Times New Roman 9-point bold.  They should be numbered (e.g.,
-``Table~\ref{tab:table1}'' or ``Figure~\ref{fig:figure2}''), centered
-and placed beneath the figure or table.  Please note that the words
-``Figure'' and ``Table'' should be spelled out (e.g., ``Figure''
-rather than ``Fig.'') wherever they occur.
-
-Papers and notes may use color figures, which are included in the page
-limit; the figures must be usable when printed in black and white in
-the proceedings.  The paper may be accompanied by a short video figure
-up to five minutes in length.  However, the paper should stand on its
-own without the video figure, as the video may not be available to
-everyone who reads the paper.
-
-\section{Language, Style and Content}
-
-The written and spoken language of SIGCHI is English. Spelling and
-punctuation may use any dialect of English (e.g., British, Canadian,
-US, etc.) provided this is done consistently. Hyphenation is
-optional. To ensure suitability for an international audience, please
-pay attention to the following:
-
-\begin{itemize}
-\item Write in a straightforward style.
-\item Try to avoid long or complex sentence structures.
-\item Briefly define or explain all technical terms that may be
-  unfamiliar to readers.
-\item Explain all acronyms the first time they are used in your text---e.g.,
-``Digital Signal Processing (DSP)''.
-\item Explain local references (e.g., not everyone knows all city
-  names in a particular country).
-\item Explain ``insider'' comments. Ensure that your whole audience
-  understands any reference whose meaning you do not describe (e.g.,
-  do not assume that everyone has used a Macintosh or a particular
-  application).
-\item Explain colloquial language and puns. Understanding phrases like
-  ``red herring'' may require a local knowledge of English.  Humor and
-  irony are difficult to translate.
-\item Use unambiguous forms for culturally localized concepts, such as
-  times, dates, currencies and numbers (e.g., ``1-5-97'' or ``5/1/97''
-  may mean 5 January or 1 May, and ``seven o'clock'' may mean 7:00 am or
-  19:00).  For currencies, indicate equivalences---e.g., ``Participants
-  were paid 10,000 lire, or roughly \$5.''
-\item Be careful with the use of gender-specific pronouns (he, she)
-  and other gendered words (chairman, manpower, man-months). Use
-  inclusive language that is gender-neutral (e.g., she or he, they,
-  s/he, chair, staff, staff-hours,
-  person-years). See~\cite{Schwartz:1995:GBF} for further advice and
-  examples regarding gender and other personal attributes.
-\item If possible, use the full (extended) alphabetic character set
-  for names of persons, institutions, and places (e.g.,
-  Gr{\o}nb{\ae}k, Lafreni\'ere, S\'anchez, Universit{\"a}t,
-  Wei{\ss}enbach, Z{\"u}llighoven, \r{A}rhus, etc.).  These characters
-  are already included in most versions of Times, Helvetica, and Arial
-  fonts.
-\end{itemize}
-
-\section{Accessibility}
-The Executive Council of SIGCHI has committed to making SIGCHI conferences more inclusive for researchers, practitioners, and educators with disabilities.
-As a part of this goal, the all authors are asked to work on improving the accessibility of their submissions.
-Specifically, we encourage authors to carry out the following five steps:
-\begin{enumerate}
-	\item Add alternative text to all figures
-	\item Mark table headings
-	\item Add tags to the PDF
-	\item Verify the default language
-	\item Set the tab order to ``Use Document Structure''
-\end{enumerate}
-Unfortunately good tools do not yet exist to create tagged PDF files from Latex.
-LaTeX users will need to carry out all of the above steps in the PDF directly using Adobe Acrobat, after the PDF has been generated.
-
-For more information and links to instructions and resources, please see:
-{\url{http://chi2014.acm.org/authors/guide-to-an-accessible-submission}}.
-
-\section{Page Numbering, Headers and Footers}
-
-Please submit your anonymous version for reviewing with page numbers
-centered in the footer.  These must be removed in the final version of
-accepted papers, as page numbers, headers, and footers will be added
-by the conference printers.  Comment out the {\textbackslash}pagenumbering
-command at the top of the document to remove page numbers.
-
-\section{Producing and Testing PDF Files}
-
-We recommend that you produce a PDF version of your submission well
-before the final deadline.  Your PDF file must be ACM DL
-Compliant. The requirements for an ACM Compliant PDF are available at:
-{\url{http://www.sheridanprinting.com/typedept/ACM-distilling-settings.htm}}.
-
-Test your PDF file by viewing or printing it with the same software we
-will use when we receive it, Adobe Acrobat Reader Version 7. This is
-widely available at no cost from~\cite{acrobat}.  Note that most
-reviewers will use a North American/European version of Acrobat
-reader, which cannot handle documents containing non-North American or
-non-European fonts (e.g. Asian fonts).  Please therefore do not use
-Asian fonts, and verify this by testing with a North American/European
-Acrobat reader (obtainable as above). Something as minor as including
-a space or punctuation character in a two-byte font can render a file
-unreadable.
-
-\section{Blind Review}
-
-For archival submissions, CHI requires a ``blind review.'' To prepare
-your submission for blind review, remove author and institutional
-identities in the title and header areas of the paper. You may also
-need to remove part or all of the Acknowledgments text.  Further
-suppression of identity in the body of the paper and references is
-left to the authors' discretion. For more details, see the submission
-guidelines and checklist for your submission category.
-
-\section{Conclusion}
-
-It is important that you write for the SIGCHI audience.  Please read
-previous years' Proceedings to understand the writing style and
-conventions that successful authors have used.  It is particularly
-important that you state clearly what you have done, not merely what
-you plan to do, and explain how your work is different from previously
-published work, i.e., what is the unique contribution that your work
-makes to the field?  Please consider what the reader will learn from
-your submission, and how they will find your work useful.  If you
-write with these questions in mind, your work is more likely to be
-successful, both in being accepted into the Conference, and in
-influencing the work of our field.
-
-\section{Acknowledgments}
-
-We thank CHI, PDC and CSCW volunteers, and all publications support
-and staff, who wrote and provided helpful comments on previous
-versions of this document.  Some of the references cited in this paper
-are included for illustrative purposes only.  \textbf{Don't forget
-to acknowledge funding sources as well}, so you don't wind up
-having to correct it later.
+%figure template:
+%\begin{figure}[!h]
+%\centering
+%\includegraphics[width=1.0\columnwidth]{Figure1}
+%\caption{With Caption Below, be sure to have a good resolution image
+%  (see item D within the preparation instructions).}
+%\label{fig:figure1}
+%\end{figure}
 
@@
 
-\section{References format}
-References must be the same font size as other body text.
 % REFERENCES FORMAT

### 2013-09-02 doc/sigchi14/control.tex
bjoern's pass at overall narrative

\section{Example Scenarios}
\bjoern{Here we should either describe a couple of example scenarios we've built, or get some students to run through realistic control tasks based on those scenarios - e.g.: ``Dim the light to 20\% and then start the movie, turn up the volume to 80\%."}

### 2013-09-02 doc/sigchi14/device.tex
bjoern's pass at overall narrative

@@
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/glass-with-ir}
+\caption{Our augmented Glass prototype has a frame-mounted infrared emitter.}
+\label{fig:glass}
+\end{figure}
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/architecture}
+\caption{System architecture.}
+\label{fig:architecture}
+\end{figure}
+
 \section{Hardware Device}
@@
 The IR emitter LED is mounted in an opaque hollow tube, that restricts the outgoing angle of illumination.
-The angle can be adjusted by moving the LED in or out of the tube.
 
@@
 
-In our prototype, Glass communicates over Bluetooth to an additional microcontroller PCB the user has to wear.
-This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame.
+In our prototype, Glass communicates over Bluetooth to an additional microcontroller PCB the user has to wear (Atmel ATMega256).
+This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame (Figure~\ref{fig:architecture}).
 This architecture was mostly chosen for reasons of expediency.
-Future headmounted devices could clearly integrate IR emitters; the choice of local area wireless technology could also change .
-\bjoern{why not all bluetooth?
-There's probably a pairing/latency argument to be made here.}
+We selected XBee 802.15.4 radios to avoid the latency associated with connecting and disconnecting to Bluetooth devices \bjoern{Is this true or not?} Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
+In particular, one could substitute WiFi modules.
 
 \subsection{Device Characterization}
-We characterized the usable range and selection accuracy... (see Figure~\ref{fig:coverage} and Table~\ref{table:measurements}).
-\bjoern{insert text from google doc here}
+We determined the usable range and accuracy empirically with one IR emitter and two IR receivers.
+The IR emitter constantly sent out an id signal.
+The receivers that correctly received the signal turn their LED on for 300 ms.
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/glass-with-ir}
-\caption{Our augmented Glass prototype has a frame-mounted infrared emitter.}
-\label{fig:glass}
-\end{figure}
+We placed all three devices at the same height with clear line of sight.
+The IR emitter is first places 2 feet away from the receivers.
+The receivers were moved sideways apart from each other until they could no longer receive stable signals.
+We then recorded the distance of the two receivers for the calculation of coverage angles.
+The steps are repeated for IR emitters in different distances (as shown in Table~\ref{table:measurements}).
+We then repeated measurements with the emitter  placed at various depths in the tube (see Figure~\ref{fig:coverage}).
+
+In summary, \bjoern{describe what the measured results mean in practice.}
 

### 2013-09-02 doc/sigchi14/introduction.tex
bjoern's pass at overall narrative

@@
 \bjoern{What's the overall motivation for the project? What problem does it address?}
-Increasingly, devices and services in our built environment such as intelligent lighting, televisions, HVAC systems, laundry machines, blinds or kitchen appliances are networked and can be controlled remotely.
-
-The proliferation of smart, controllable devices raises the question of how to best interact with them.
-Prior work has introduced several approaches that use handheld mobile devices as {\em universal remote controls} to select and control appliances in the proximate environment- e.g., through pointing at such devices to select them, and through touch-screen or button interfaces to send control commands~\cite{}.
-Commercial universal control applications for smart phones are also available~\cite{}, though they usually do not offer spatial selection of target devices, forcing users to browse through lists of configured devices instead.
+Increasingly, devices and services in our built environment are networked and can be controlled remotely.
+The proliferation of smart, controllable devices such as intelligent lighting, AV equipment, HVAC systems, or kitchen appliances raises the question of how to best interact with them.
+Prior work has introduced several approaches that use handheld mobile devices as {\em universal remote controls} to select and control appliances in the proximate environment- e.g., through pointing at such devices to select them, and through touch-screen or button interfaces to send control commands~\cite{beigl_point_1999,patel_2-way_2003}.
 
@@
 Some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is split between looking down at a screen and out at the device to-be-controlled.
-\bjoern{these all sound kind of weak.} These general limitations have led researchers to investigate ``always-available'' physiological input modalities on one hand~\cite{saponas}, and wearable computing devices~\cite{} on the other hand.
+\bjoern{these all sound kind of weak.} These general limitations have led researchers to investigate ``always-available'' physiological input modalities for phones on one hand~\cite{morris_emerging_2011}, and wearable computing devices on the other hand.
 

### 2013-09-02 doc/sigchi14/targeting.tex
bjoern's pass at overall narrative

\section{Target Acquisition Experiment}
\bjoern{text from Google Doc goes here}

### 2013-09-02 doc/sigchi14/relatedwork.tex
bjoern's pass at overall narrative
small tweaks

@@
 
+Commercial universal control applications for smart phones are also available (e.g., Belkin WeMo\footnote{http://www.belkin.com/us/wemo}), though they usually do not offer spatial selection of target devices, forcing users to browse through lists of configured devices instead.
+
 Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing - the user looks at the target device.
@@
 \subsection{Augmented Reality Interfaces}
-\bjoern{most of the prior headmounted work is in the AR space. I don't know much about it yet.}
+\bjoern{most of the prior head-mounted work is in the AR space. I don't know much about it yet, but Kortuem~\cite{kortuem_context-aware_1998} might be a good starting point.}

### 2013-09-02 doc/sigchi14/introduction.tex
tweaks

@@
 Target devices similarly have IR receivers and ZigBee radios.
-These choices provide an orientation-specific  step selection appropriate for the limitations of head orientation; once acquired, users do not have to look at the target device for control commands.
+These choices offer selection-through-head-orientation to initiate interaction; once initiated, users do not have to look at the target device for control commands.
 

### 2013-09-02 doc/sigchi14/introduction.tex
pass over related work and device section

@@
 We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose.
-Users first look in the direction of the device they wish to control to acquire it (e.g., at a lamp for controlling lighting, or at a speaker to change music playback volume).
+Users first look in the direction of the device they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume).
 If multiple devices fall within communication range, an on-screen disambiguation dialog lets users select a single target.
-Once acquired, a device specific control UI enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).
+Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).
 
 Our hardware relies on infrared communication between Glass and target devices to establish a connection; and on wireless ZigBee radio communication to exchange control messages.
+Glass is augmented with a narrow IR emitter and a ZigBee ratio.
 Target devices similarly have IR receivers and ZigBee radios.
-These choices offer selection-through-head-orientation to initiate interaction; once initiated, users do not have to look at the target device for control commands.
+These choices offer selection-through-head-orientation to initiate interaction; once initiated, users can look away from the target device while issuing control commands.
 

### 2013-09-03 server/target_exp.py
refine auto cue target

@@ -70,9 +70,10 @@ def end_exp():
   result_log = []
   exit(0)
 
-def send_cue(id):
-  out = target_id + "CSELTAR"
-  logResult("target", target_id)
+def send_cue(tid):
+  # print "debug:", tid
+  out = tid + "CSELTAR"
+  logResult("target", tid)
   global list_cursor
   print 'round', list_cursor+1
   print 'Target msg:', out
@@ -155,7 +156,7 @@ while True:
       cmd = raw_input('Enter command:')  
       print '[Console]: ',cmd
       # print 'line length: ',len(cmd)
-      if ('t' in cmd or 'T' in cmd) and len(cmd) == 4:
+      if ('t' in cmd or 'T' in cmd) and len(cmd) == 3:
         # target a client to turn on cue led
         target_id = cmd[1:3]
         state_cue = False
@@ -172,7 +173,20 @@ while True:
 
     out = send_cue(target_id)
     ser.write(out)
-    
+  
+  # read line without blocking
+  while sys.stdin in select.select([sys.stdin], [], [], 0)[0]:
+    line = sys.stdin.readline()
+    print '[Console]: ',line
+    print 'line length: ',len(line)
+    if ('t' in line or 'T' in line) and len(line) == 4:
+      # target a client to turn on cue led
+      target_id = line[1:3]
+      state_cue = False
+    elif line[0:3] == "end":
+      end_exp()
+    else:
+      ser.write(line)  
 
   while ser.inWaiting() > 0:
     in_msg = ser.readline()

### 2013-09-03 doc/sigchi14/conclusion.tex

\section{Conclusion and Future Work}
Our prototype does not yet have a general way of communicating device descriptions and capabilities - this means our prototype cannot yet control new, unknown devices - they first have to be modeled. We may extend the PUC work~\cite{nichols} to address this issue.

More importantly though, the main limitation is that extra hardware for IR communication is needed for the HMD and each controllable device. One potential approach would be to combine the growing availability of high-resolution indoor maps with live data from the HMD's camera to determine what a user is looking at without IR data exchange.

\bjoern{add some general summary}

### 2013-09-03 doc/sigchi14/device.tex

@@
 
-\section{Hardware Device}
+\section{Hardware Device and Interaction}
 
@@
 
-Glass communicates a device id over the IR channel analogous to Patel's approach~\cite{patel_2-way_2003}.
+\subsection{Interaction Flow}
+{\bf Look:} Users select a target device by looking in its general direction.
+Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
 Target appliances have IR receivers and offer immediate visual feedback by toggling an LED whenever a valid id is received.
+They also respond by initiating an XBee radio connection and sending their id.
+This allows the Glass application to show which devices are currently targeted on the display.
 
-When users confirm that, further communication switches over to a ZigBee wireless network so that line of sight to the target is no longer needed.
+{\bf Initiate:} Users confirm to initiate interaction by tapping on the Glass touchpad.
+The next section on disambiguation deals with cases in which multiple devices received valid IR signals.
+At this point, all further communication switches over to the XBee wireless network so that line of sight to the target is no longer needed (otherwise restricting head movement can become straining).
 
-In our prototype, Glass communicates over Bluetooth to an additional microcontroller PCB the user has to wear (Atmel ATMega256).
+{\bf Control:} Glass displays a control interface for parameters of the chosen device. \bjoern{Describe the interaction scheme for navigating and adjusting parameters here}.
+Control commands are sent over XBee radios.
+
+{\bf Disengagement:} \bjoern{how do you end an interaction? Is there a timeout if the user forgets to manually back out of the control screen?}
+
+\subsection{Disambiguation}
+\bjoern{describe why disambiguation is necessary - combination of the relative inaccuracy of head orientation and the spread of our IR signal. Then describe how to overcome it.}
+
+\subsection{Prototype Implementation}
+In our prototype, Glass communicates over Bluetooth to an additional microcontroller board the user has to wear (Atmel ATMega256).
 This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame (Figure~\ref{fig:architecture}).
@@
 
-\subsection{Device Characterization}
+\section{Device Characterization}
 We determined the usable range and accuracy empirically with one IR emitter and two IR receivers.

### 2013-09-03 doc/sigchi14/relatedwork.tex

--- 2013-09-02 doc/sigchi14/relatedwork.tex
@@
 \section{Related Work}
-Relevant prior work exists in the areas of remote control of physical appliances and evaluations of pointing in physical space. We discuss each in turn.
+Relevant prior work exists in the areas of remote control of physical appliances, evaluations of pointing in physical space and augmented reality applications. We discuss each in turn.
 
 \subsection{Remote Control of Physical Appliances}
-A number of researchers propose to use different handheld devices to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012} or exchange other information with smart infrastructure like reading sensor netowrk data~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}.
+Universal device control applications for smart phones are commercially available (e.g., Belkin WeMo\footnote{http://www.belkin.com/us/wemo}), though they usually do not offer spatial selection of target devices, forcing users to browse through lists of configured devices instead.
+Rukzio found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.
+
+Several approaches to spatial selection with handheld devices exist to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012} or to exchange information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}.
 Key design decisions are the method by which a target device is selected; and the method by which it is then later controlled or configured.
 
-Several techniques use laser pointers to directly point at objects of interest.
-This has the advantage of also providing immediate visual feedback to the user in the shape of the laser dot.
-Beigl's early AIDA handheld combines laser pointing for selection with IR communication to exchange commands~\cite{beigl_point_1999}.
-Patel extends this technique by modulating the laser light to communicate the controllers identity~\cite{patel_2-way_2003}.
+In several techniques, users select objects of interest with laser pointers.
+The laser dot provides immediate visual feedback to the user what is being selected.
+Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}.
+Patel extends this technique by modulating the laser light to communicate the controllers' identity~\cite{patel_2-way_2003} to initiate radio communication.
+These proofs-of-concept do not include thorough evaluations.
 Kemp et al.
@@
 
-Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in PiControl~\cite{schmidt_picontrol:_2012}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
-Tags like QR codes can also be affixed to devices and read by cameras.
-Common tagging systems are optimized to be read from a close distance, though it is possible to redesign codes that can be read further away (by encoding less information)~\cite{lowcostsensing-uist13}.
+Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
+Printed tags like QR codes can also be affixed to devices and read by cameras.
+Common tagging systems are optimized to be read from a close distance, though it is possible to redesign codes that can be read further away (by encoding less information)~\cite{cross_low-cost_2012}.
 %However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.
 
-Standard infrared remote controls for televisions and AV equipment tend to use wide-angle LEDs - since they are only meant to control a single device, there is no ambiguity  and it is unnecessary to have the user point precisely.
+Standard infrared remote controls for televisions and AV equipment are only meant to control a single device and it is unnecessary to have the user point precisely.
+These controllers tend to use wide-angle infrared LEDs.
 
-Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing - the user looks at the target device.
-Selection techniques with very small selectors such as laser pointing are less appropriate for head-mounted applications.
-Head orientation only indicates a general area of visual interest.
-It does not necessarily match gaze orientation as the eyes can be positioned relative to the head using the extraocular muscles.
-We therefore select an infrared LED - a source with a wider angle of illumination, but restrict its angle to be narrower than in general purpose IR applications.
-
-\bjoern{Other citations to cover here: Raskar et al's radio frequency identity and geometry~\cite{raskar_rfig_2004}.}
+Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing - the user looks at the target device to initiate interaction.
+Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications as head orientation only indicates a general area of visual interest.
+It does not necessarily match gaze orientation as extra-ocular muscles can move the eyes.
+We therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.
 
 \subsection{Evaluation of room-scale selection}
-Pausch et al.'s early investigation of head-mounted displays compared orientation tracking of the head to handheld orientation control for a target acquisition task in a virtual reality room shown on a head-mounted display~\cite{pausch_user_1993}.
-They found a clear performance benefit for head-mounted tracking.
+Pausch et al.'s early investigation of head-mounted displays compared head-tracking to handheld orientation control for a target acquisition task in a virtual reality room shown on a head-mounted display~\cite{pausch_user_1993}.
+They found a clear performance benefit for head-tracking.
 
@@
 However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques.
-Compared to screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower.
+Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy.
 
-Myers et al compared different methods of interacting with displays at a distance~\cite{myers_interacting_2002} and quantified selection time and jitter or position error when using laser pointers.
-Various other techniques outperformed laser pointers.
+Myers et al compared different methods of interacting with displays at a distance~\cite{myers_interacting_2002} and quantified selection time and jitter or position error when using remote handheld pointing.
+Various techniques outperformed laser pointers.
 
-Rukzio found that users strongly preferred either touching a mobile device to a target appliance to initiate interacting or pointing at a distance to ``scanning'' (initiating a Bluetooth scan and then selecting a target appliance from a list of scan results on the phone )~\cite{rukzio_experimental_2006}.
+Our work is complementary as it provides concrete performance data on using head orientation to select targets in a physical environment.
 
 \subsection{Augmented Reality Interfaces}
-\bjoern{most of the prior headmounted work is in the AR space. I don't know much about it yet.}
+Agumented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
+Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world.
+
+%\bjoern{most of the prior head-mounted work is in the AR space. I don't know much about it yet, but Kortuem~\cite{kortuem_context-aware_1998} might be a good starting point.}

### 2013-09-03 doc/sigchi14/introduction.tex
add thoughts in introduction > drawbacks of existing work

@@
 Some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is split between looking down at a screen and out at the device to-be-controlled.
-\bjoern{these all sound kind of weak.} These general limitations have led researchers to investigate ``always-available'' physiological input modalities for phones on one hand~\cite{morris_emerging_2011}, and wearable computing devices on the other hand.
+\bjoern{these all sound kind of weak.} \sean{When the number of smart devices increases, it also takes time to traverse within a list of dozens of items.
+In addition, mapping the name of a device to its real physical presence is not always intuitive.
+For example, it is hard to identify where is ``light in area E'' in an office building.
+Moreover, in a shared space, the person trying to control the device might not be the one that named it.
+} These general limitations have led researchers to investigate ``always-available'' physiological input modalities for phones on one hand~\cite{morris_emerging_2011}, and wearable computing devices on the other hand.
 

### 2013-09-04 server/target_exp.py
refine exp server code

@@ -8,6 +8,7 @@ import serial, glob, argparse, Queue
 import sys, select
 import time
 import random
+from time import sleep
 
 parser = argparse.ArgumentParser(description='Pyserial for XBee')
 parser.add_argument('--debug', default=False, action='store_true', help='Print More Errors and Launch interactive console on exception or forced exit')
@@ -24,12 +25,14 @@ user_id = "NA"
 start_time = time.time()
 
 target_list = []
-predefined_list = ['01', '02', '03']
+predefined_list = ['01', '02', '01']
 total_client_no = 3
 total_task_round = 10
 list_cursor = 0
-task_interval = 3000  # after selecting a correct candidate, wait for couple seconds till next target shows
+task_interval = 3  # after selecting a correct candidate, wait for couple seconds till next target shows
 state_cue = True
+correct_target = False
+
 MODE_PREDEFINE = 1
 MODE_RANDOM = 2
 MODE_MANUAL = 3
@@ -45,7 +48,7 @@ def logResult(action, obj):
   result_log.append(tup)
   print tup
 
-def populate_target(type):
+def prepare_target_list(type):
   print 'populate, type: ', type
   result = []
   if type == 1: # predefined
@@ -128,7 +131,7 @@ try:
   num = int(raw_input('Enter the number:'))
   if num > 0 and num <= 3:
     exp_mode = num
-    target_list = populate_target(exp_mode)
+    target_list = prepare_target_list(exp_mode)
   else:
     print 'out of range'
     exit(1)
@@ -155,9 +158,10 @@ while True:
 
       cmd = raw_input('Enter command:')  
       print '[Console]: ',cmd
-      # print 'line length: ',len(cmd)
+      print 'line length: ',len(cmd)
       if ('t' in cmd or 'T' in cmd) and len(cmd) == 3:
         # target a client to turn on cue led
+        print 'target set to', cmd[1:3]
         target_id = cmd[1:3]
         state_cue = False
       elif cmd[0:3] == "end":
@@ -167,6 +171,7 @@ while True:
     else: # predefined or random list
       if list_cursor > len(target_list) - 1:
         print 'exp done'
+        end_exp()
         exit(0)
       target_id = target_list[list_cursor]
       state_cue = False
@@ -196,7 +201,7 @@ while True:
       logResult("connect", '')
     elif in_msg[2:9] == "ASELTAR":
       # is ack for target cue, record timestamp
-      logResult("cue shown", '')
[…]

### 2013-09-04 server/target_exp.py
add 'reset' cmd to send cue again in case msg corrupted

@@ -179,15 +179,16 @@ while True:
     out = send_cue(target_id)
     ser.write(out)
   
-  # read line without blocking
+  # can reset or end exp within a task phase
   while sys.stdin in select.select([sys.stdin], [], [], 0)[0]:
     line = sys.stdin.readline()
     print '[Console]: ',line
     print 'line length: ',len(line)
-    if ('t' in line or 'T' in line) and len(line) == 4:
-      # target a client to turn on cue led
-      target_id = line[1:3]
-      state_cue = False
+    if line[0:5] == 'reset':
+      # repeat the cue stage (if cue ack isn't received)
+      list_cursor -= 1
+      state_cue = True
+
     elif line[0:3] == "end":
       end_exp()
     else:

### 2013-09-04 server/target_exp.py
turn off all cue leds when program ended

@@ -22,11 +22,12 @@ parser.add_argument('--serial', action='store', default=None, help='serial port'
 target_id = "NA"
 result_log = []
 user_id = "NA"
+ser = None
 start_time = time.time()
 
 target_list = []
-predefined_list = ['01', '02', '01']
-total_client_no = 3
+predefined_list = ['01', '02', '01', '01', '02']
+total_client_no = 2
 total_task_round = 10
 list_cursor = 0
 task_interval = 3  # after selecting a correct candidate, wait for couple seconds till next target shows
@@ -64,13 +65,14 @@ def prepare_target_list(type):
 
 def end_exp():
   print 'writing results'
-  global user_id, result_log
+  global user_id, result_log, ser
   f = open("result_"+user_id[:-1]+".txt", 'w')
   for t in result_log:
     print t
     f.write(' '.join(str(s) for s in t) + '\n') 
   f.close()
   result_log = []
+  ser.write('XXCSELTAR')  # turn off all cue leds
   exit(0)
 
 def send_cue(tid):

### 2013-09-05 doc/sigchi14/device.tex
added comments on paper

@@
 This allows the Glass application to show which devices are currently targeted on the display.
+\sean{Currently the led is implemented but clients don't response with their IDs and display on Glass at this stage.
+We can add that.
+But we felt like the users would use led as the primary indicator and not to distract them with frequently changing the content on the screen?
+}
 
@@
 
-{\bf Disengagement:} \bjoern{how do you end an interaction? Is there a timeout if the user forgets to manually back out of the control screen?}
+{\bf Disengagement:} \bjoern{how do you end an interaction?
+Is there a timeout if the user forgets to manually back out of the control screen?} \sean{currently the only way is to swipe back.
+A timeout mechanism can be added easily.
+I guess the design decision lies in what is an appropriate duration?}
 
@@
 This architecture was mostly chosen for reasons of expediency.
-We selected XBee 802.15.4 radios to avoid the latency associated with connecting and disconnecting to Bluetooth devices \bjoern{Is this true or not?} Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
+We selected XBee 802.15.4 radios to avoid the latency associated with connecting and disconnecting to Bluetooth devices \bjoern{Is this true or not?} \sean{Origainlly we didn't know we'd work on Google Glass, therefore BT didn't have a higher priority.
+We learned that bluetooth can concurrently communicate with up to 7 devices (while up to 255 can be paired but inactive), giving us more complexity when scaling up.
+In addition ZigBee has shorter wakeup delay and lower power consumption.
+(referenced here: http://superuser.com/questions/332767/limit-to-the-number-of-devices-that-can-be-paired-with-a-bluetooth-device)} Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
 In particular, one could substitute WiFi modules.

### 2013-09-05 doc/sigchi14/introduction.tex
added comments on paper

@@
 If multiple devices fall within communication range, an on-screen disambiguation dialog lets users select a single target.
-Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).
+\sean{as well as led indicators on the devices?} Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).
 
 Our hardware relies on infrared communication between Glass and target devices to establish a connection; and on wireless ZigBee radio communication to exchange control messages.
-Glass is augmented with a narrow IR emitter and a ZigBee ratio.
+Glass is augmented with a narrow IR emitter and a ZigBee radio.
 Target devices similarly have IR receivers and ZigBee radios.

### 2013-09-05 doc/sigchi14/relatedwork.tex
added comments on paper

@@
 However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques.
-Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy.
+\sean{Would like to add that we are only using head gesture for connecting, since we believe a user's attention is often drawn to the objects he intend to interact with.
+For further interactions, we still use hand gestures with heads-up-display.} Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy.
 

### 2013-09-12 server/target_exp.py
refine

@@ -26,7 +26,7 @@ ser = None
 start_time = time.time()
 
 target_list = []
-predefined_list = ['01', '02', '03', '04', '01']
+predefined_list = ['01', '02', '03', '04', '05', '06', '07', '08', '09', '10']
 total_client_no = 4
 total_task_round = 6
 list_cursor = 0
@@ -166,9 +166,12 @@ while True:
         print 'target set to', cmd[1:3]
         target_id = cmd[1:3]
         state_cue = False
+        out = send_cue(target_id)
+        ser.write(out)
       elif cmd[0:3] == "end":
         end_exp()
       else:
+        print 'writing cmd to XBee'
         ser.write(cmd)
     else: # predefined or random list
       if list_cursor > len(target_list) - 1:
@@ -177,9 +180,8 @@ while True:
         exit(0)
       target_id = target_list[list_cursor]
       state_cue = False
-
-    out = send_cue(target_id)
-    ser.write(out)
+      out = send_cue(target_id)
+      ser.write(out)
   
   # can reset or end exp within a task phase
   while sys.stdin in select.select([sys.stdin], [], [], 0)[0]:
@@ -194,6 +196,7 @@ while True:
     elif line[0:3] == "end":
       end_exp()
     else:
+      print 'writing cmd to XBee'
       ser.write(line)  
 
   while ser.inWaiting() > 0:

### 2013-09-12 server/target_exp.py
move on only after server receive cue ack

@@ -83,7 +83,6 @@ def send_cue(tid):
   print 'round', list_cursor+1
   print 'Target msg:', out
   print 'out length:', len(out)
-  list_cursor += 1
   return out
 
 availables = glob.glob('/dev/tty.*')
@@ -165,9 +164,10 @@ while True:
         # target a client to turn on cue led
         print 'target set to', cmd[1:3]
         target_id = cmd[1:3]
-        state_cue = False
+        
         out = send_cue(target_id)
         ser.write(out)
+
       elif cmd[0:3] == "end":
         end_exp()
       else:
@@ -208,6 +208,8 @@ while True:
     elif in_msg[2:9] == "ASELTAR":
       # is ack for target cue, record timestamp
       logResult("cue shown", in_msg[0:2])
+      list_cursor += 1
+      state_cue = False
     elif in_msg[3:9] == "SEL1st":
       logResult("multiple", in_msg[0:2])
     elif in_msg[3:9] == "SEL NA":

### 2013-09-12 server/target_exp.py
refine

@@ -26,7 +26,8 @@ ser = None
 start_time = time.time()
 
 target_list = []
-predefined_list = ['01', '02', '03', '04', '05', '06', '07', '08', '09', '10']
+# predefined_list = ['01', '02', '03', '04', '05', '06', '07', '08', '09', '10']
+predefined_list = ['08', '02', '04', '01', '05', '10', '07', '03', '09', '06', '01', '09', '02', '10', '07']
 total_client_no = 4
 total_task_round = 6
 list_cursor = 0
@@ -192,7 +193,6 @@ while True:
       # repeat the cue stage (if cue ack isn't received)
       list_cursor -= 1
       state_cue = True
-
     elif line[0:3] == "end":
       end_exp()
     else:

### 2013-09-12 server/target_exp.py
refine catching success msg out of a cluster

@@ -27,13 +27,14 @@ start_time = time.time()
 
 target_list = []
 # predefined_list = ['01', '02', '03', '04', '05', '06', '07', '08', '09', '10']
-predefined_list = ['08', '02', '04', '01', '05', '10', '07', '03', '09', '06', '01', '09', '02', '10', '07']
-total_client_no = 4
-total_task_round = 6
+predefined_list = ['08', '02', '04', '01', '05', '10', '07', '03', '09', '06', '05', '09', '02', '06', '07']
+trial_list = ['01', '04', '09']
+total_client_no = 10
+total_task_round = 15
 list_cursor = 0
 task_interval = 3  # after selecting a correct candidate, wait for couple seconds till next target shows
 state_cue = True
-correct_target = False
+success = 0;
 
 MODE_PREDEFINE = 1
 MODE_RANDOM = 2
@@ -43,7 +44,7 @@ exp_mode = 0
 def logResult(action, obj):
   ts = time.time()
   global start_time
-  if action == 'target':
+  if action == 'start':
     # reset start time everytime a target is sent
     start_time = ts
   tup = (ts-start_time, action, obj)
@@ -61,13 +62,15 @@ def prepare_target_list(type):
       result.append("%02d" % t)
   elif type == 3:
     restult = []
+  elif type == 4:
+    result = trial_list
   print 'list: ', result
   return result
 
 def end_exp():
   print 'writing results'
   global user_id, result_log, ser
-  f = open("result_"+user_id[:-1]+".txt", 'w')
+  f = open("../data/result_"+user_id[:-1]+".txt", 'w')
   for t in result_log:
     print t
     f.write(' '.join(str(s) for s in t) + '\n') 
@@ -79,11 +82,10 @@ def end_exp():
 def send_cue(tid):
   # print "debug:", tid
   out = tid + "CSELTAR"
-  logResult("target", tid)
+  # logResult("target", tid)
   global list_cursor
-  print 'round', list_cursor+1
-  print 'Target msg:', out
-  print 'out length:', len(out)
+  print '=== round', list_cursor+1, '/', len(target_list), 'success:', success, '==='
+  print 'Target msg:', out, 'length:', len(out)
   return out
 
 availables = glob.glob('/dev/tty.*')
@@ -128,21 +130,23 @@ user_id = sys.stdin.readline()
 print "1. auto from predefined list"
 print "2. auto from random"
 print "3. manually enter (e.g. T01)"
+print "4. trial test"
 
 try:
   num = int(raw_input('Enter the number:'))
[…]

### 2013-09-13 doc/sigchi14/control.tex
progress with text and figures

@@
-\section{Example Scenarios}
+\section{Example SmartHome Scenarios}
+To understand how our device could be used to interact with smart appliances, we also asked all study participants to work through a concrete scenario.
+The main goal was to obtain qualitative feedback on the usability and utility of our device with more realistic tasks.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/smarthome-scenario.jpg}
+\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment.}
+\label{fig:smarthome}
+\end{figure}
+
+\subsection{Methodology}
+We recreated a living room environment that had three (four?) controllable appliances: a fan, a lamp, and two laptops functioning as an audio and video player, respectively (see Figure~\ref{fig:smarthome}).
+The fan and lamp had binary controls: they could be switched on or off.
+The laptop had multiple parameterized functions: participants could start, pause, stop, and rewind  video and audio, and adjust volume.
+
+We then asked users to work through the following script:
+
+\small{
+\begin{enumerate}
+\item Turn off the lights as you want to watch the movie in a darkened room.
+\item You feel a little hot in the room, so you turn on the fan
+\item You connect to the Smart TV and start playing the movie. 
+\item The volume seems too soft to hear over the fan- turn it up a bit. 
+\item After a while, you want to take a break to get a snack. Pause the movie. 
+\item When you come back, you've forgotten what was said last - rewind by ~30 seconds and restart the movie. 
+\item You stop the movie and turn the lights back on.
+\item You want to listen to some music.
+\item After awhile, you stop the music
+you turn off the fan and leave the room.
+\end{enumerate}}
+
+\subsection{Qualitative results}
+
 \bjoern{Here we should either describe a couple of example scenarios we've built, or get some students to run through realistic control tasks based on those scenarios - e.g.: ``Dim the light to 20\% and then start the movie, turn up the volume to 80\%."}

### 2013-09-13 doc/sigchi14/device.tex
progress with text and figures

@@
 Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
-Target appliances have IR receivers and offer immediate visual feedback by toggling an LED whenever a valid id is received.
+Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received.
+This enables {\em scanning} the environment with ones gaze to see which devices can be controlled.
 They also respond by initiating an XBee radio connection and sending their id.

### 2013-09-13 doc/sigchi14/targeting.tex
progress with text and figures

@@
 \section{Target Acquisition Experiment}
-\bjoern{text from Google Doc goes here}
+To understand the accuracy and performance of head-orientation-based  selection through our device, we carried out a comparative target acquisition study, where participants had to connect to wireless nodes with our technique, and a list selection approach.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
+\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment.}
+\label{fig:targeting-study-layout}
+\end{figure}
+
+\subsection{Apparatus}
+In an indoor environment, 10 wireless nodes are spread across a room at various heights and distances (see Figure~\ref{fig:targeting-study-layout}).
+The nodes are stand-ins for potential smart appliances and have all relevant functionality for targeting and selection, but do not control any actual appliances.
+Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs: an orange LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple clients are within IR range (see Figure~\ref{fig:target}).
+Next to each client device, a paper sheet shows a number and letter combination, which are used for uniquely identifying the devices.
+
+\begin{figure}[b]
+\centering
+\includegraphics[width=0.8\columnwidth]{figures/study-node.pdf}
+\caption{An example node from the targeting study --- we constructed 10 such nodes - each mounted in a box.}
+\label{fig:targeting-study-layout}
+\end{figure}
+
+\subsection{Methodology}
+In our within-subjects design, participants performed 15 target acquisition tasks each with two interaction styles.
+In the {\em infrared} condition, participants used our IR targeting approach; in the {\em list} condition, participants had to look up a device's letter code and select that letter code from a list displayed on their Glass device.
+The list was navigated with swipe motions on the Glass touchpad.
+For each task, participants started at a fixed position in the room.
+The experimenter called out a number and simultaneously started a timer.
+Participants then had to find the number on sheets printed and affixed to walls and other surfaces.
+In the {\em infrared} condition, participants then selected and acquired the target by aiming the infrared beam at the target, and confirmed their selection with a touchpad tap.
+If more than one target was within range, participants had to either use the disambiguation dialog or reposition themselves.
+In the {\em list} condition, participants had to read the letter next to the number and then select that letter by browsing a linear list shown in their Glass display.
+While the list was alphabetized, letter arrangement in the room was not.
+This design required participants to find the target in the room before starting a list navigation to keep conditions similar to each other.
+
+Afterwards, participants completed a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
+
+\subsection{Participants}
+We recruited N participants from our institution. X had never used Glass before.
+
+\subsection{Measures}
+The main measure was {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device.
+We also elicited {\bf user preference}: which interface users preferred for the task after completing the study.
+
+\subsection{Results}
+
+\subsubsection{Performance data}
+
+\subsubsection{User preference}
+X of Y users preferred infrared mode over list mode. As slef-report data can skew to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.
+
+\bjoern{Insert summary of survey data on this point here.}
+
+\subsubsection{Qualitative results}

### 2013-09-13 doc/sigchi14/control.tex
more text and figures

--- 2013-09-02 doc/sigchi14/control.tex
@@
-\section{Example Scenarios}
-\bjoern{Here we should either describe a couple of example scenarios we've built, or get some students to run through realistic control tasks based on those scenarios - e.g.: ``Dim the light to 20\% and then start the movie, turn up the volume to 80\%."}
+\section{Example SmartHome Scenario}
+To understand how our device could be used to interact with smart appliances, we also asked all study participants to work through a concrete scenario.
+The main goal was to obtain qualitative feedback on the usability and utility of our device with more realistic tasks.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/smarthome-scenario.jpg}
+\caption{In the smart home scenario, participants completed a series of appliance control tasks in a simulated living room.}
+\label{fig:smarthome}
+\end{figure}
+
+\subsection{Methodology}
+We recreated a living room environment that had four controllable appliances: a fan, a lamp, and two laptops functioning as an audio and video player, respectively (see Figure~\ref{fig:smarthome}).
+The fan and lamp had binary controls: they could be switched on or off.
+The laptops had multiple parameterized functions: participants could start, pause, stop, and rewind  video and audio, and adjust volume.
+
+We then asked users to work through the following script for controlling the room for watching a movie in the evening:
+{\small
+\begin{enumerate}
+\item Turn off the lights as you want to watch the movie in a darkened room.
+\item You feel a little hot in the room, so you turn on the fan.
+\item You connect to the Smart TV and start playing the movie. 
+\item The volume seems too soft to hear over the fan- turn it up a bit. 
+\item After a while, you want to take a break to get a snack. Pause the movie. 
+\item When you come back, you've forgotten what was said last - rewind by ~30 seconds and restart the movie. 
+\item When the credits roll, you stop the movie and turn the lights back on.
+\item You want to listen to some music.
+\item After awhile, you stop the music you turn off the fan and leave the room.
+\end{enumerate}
+}
+
+For this study, we only elicited subjective data in the form of Likert data and 
+\begin{figure}[b]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/scenario-likert.pdf}
+\caption{Likert scale ratings for ease of use of aspects of the smart home control scenario. Error bars show standard error.}
+\label{fig:smarthome-likert}
+\end{figure}
+
+\subsection{Results}
+All participants successfully completed the list of tasks.
+Participants stated it \studyquote{felt easy to interact with the appliances}, but there were distinct differences between appliances.
+Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for more complex TV and music player devices (see Figure~\ref{fig:smarthome-likert}).
+Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
+The screen size and 1D input thus likely put a limit on the complexity of interfaces that can be presented.
+
+\bjoern{explain why music is so low, if this trend holds}
+
+\subsubsection{Visual overlap between Near-Eye Display and Target Appliance}
+When a participant looks at a target, the near-eye display may occlude or overlap the target appliance.
+This may make it difficult to read either the on-screen display or see information displayed on the target device.
+As one participant remarked \studyquote{This was especially annoying with the TV because there were two screens overlapping each other.} While it is possible to look away once a device has been acquired to better see the Glass display, this tension is likely to remain for users who want to rely on feedback from the devices themselves instead of the Glass display.
+
+\subsubsection{Control}
+
+%Overall it felt easy to interact with the appliances.
+I imagine that moving backwards and forwards in the video by swiping could get frustrating if one were trying to move by just a little bit, or to a specific spot."
+%Operating the lamp and fan felt very natural because each only involved turning it on or off, which can be done with tapping Glass once.
+Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad, but only clicking once made use a lot easier.
+The TV was a little more difficult to operate because there were more options on the display.
+Tapping felt more natural than the multi-touch gestures.
+%"Again, the lack of movement is both good and bad.
+It's good in that you can just stay in one place and control everything.
+Also, everything is centralized, so you don't need to use multiple switches.
+I liked the two-finger touch to change the volume as well as to fast-forward/rewind.
+It's clever.
+The only issue with it, though, was that I felt like I didn't have as fine-grained control, but perhaps that's because I have not practiced swiping with Glass much.
+
+%I didn't like that it promotes a sedentary lifestyle.
+Also, I disliked how it took away the tactile feeling that you get from pushing buttons.
+I didn't like how you had to scroll through a bunch of things to access certain items when otherwise I could just have flipped a switch."
+%I think it's cool to control home appliances by using Glass if the home devices are far from the place where the user is.
+It might also be useful for people who need to take care of small children that they can complete all the tasks while keeping an eye on their children at the same time.
+%it was easy to connect, but it was annoying to have to first connect to turn on/off - I kept trying to point and then immediately click to turn on/off, but I had to do the extra step of connecting THEN turning on/off.
+I also tried to point back at the appliance when I was already connected.
+I intuitively want the screen to automatically appear when the IR detects the appliance rather than having to tap to connect.
+%"i liked the feeling i had while doing the selection etc
+
+%i disliked that lag time where i was looking at the thing i wanted to control and it wasn't responding to me - about 500ms or so of interaction discomfort"

### 2013-09-13 doc/sigchi14/discussion.tex
more text and figures

\section{Discussion}

Study-related topics:
\begin{enumerate}
\item Our technique should have a wider margin as the number of targets increases.

\end{enumerate}

Other topics:
\begin{enumerate}
\item Some fundamental differences to earlier laser-pointer based systems: with a laser pointer in hand one can very precisely point at an arbitrary object and know that one is pointing; however, that alone does not provide feedback on whether a device can actually be controlled or not. By shifting the visual feedback to the target device,

\item Battery life considerations: we still need to know when to start emitting the IR signal, otherwise, battery life will suffer, which is especially critical for wireless devices. However, requiring complicated steps to turn IR on goes counter the design goals of seamless interactions. A frame-mounted button may be an appropriate design choice.

\item Where to look? For a light, one might put a received on the light itself, or on the light switch. For volume, it might be on a speaker or on the amplifier. Someone should study preferences.
\end{enumerate}

### 2013-09-13 doc/sigchi14/targeting.tex
more text and figures

--- 2013-09-02 doc/sigchi14/targeting.tex
@@
-\section{Target Acquisition Experiment}
-\bjoern{text from Google Doc goes here}
+\section{Physical Target Acquisition Study}
+To understand the accuracy and performance of head-orientation-based  selection through our device, we carried out a comparative target acquisition study, where participants had to connect to wireless nodes distributed in a room with our technique, and with an alternate list selection approach.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
+\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment.}
+\label{fig:targeting-study-layout}
+\end{figure}
+
+\subsection{Apparatus}
+In an indoor environment, 10 wireless nodes are spread across a room at various heights and distances (see Figure~\ref{fig:targeting-study-layout}).
+The nodes are stand-ins for potential smart appliances and have all relevant functionality for targeting and wireless communication, but do not control any actual appliances.
+Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs (see Figure~\ref{fig:target}).
+An orange LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple clients are within IR range .
+Next to each client device, a paper sheet shows a number and letter combination, which are used for uniquely identifying the devices.
+
+Some trials also used a {\em distractor} configuration in which 15 additional paper sheets with additional identifiers are added to the environment.
+We did not create 15 additional wireless nodes though as each node has to be hand-assembled and soldered.
+
+\begin{figure}[b]
+\centering
+\includegraphics[width=0.8\columnwidth]{figures/study-node.pdf}
+\caption{An example node from the targeting study --- we constructed 10 such nodes - each mounted in a box.}
+\label{fig:targeting-study-layout}
+\end{figure}
+
+\subsection{Methodology}
+In our within-subjects design, participants performed 15 target acquisition tasks each with two interaction styles.
+In the {\em infrared} condition, participants used our IR targeting approach; in the {\em list} condition, participants had to look up a device's letter code on the printed paper next to the device and then select that letter code from a list displayed on their Glass device.
+The list was navigated with swipe motions on the Glass touchpad.
+For each task, participants started at a fixed position in the room.
+The experimenter called out a number and simultaneously started a timer.
+Participants then had to find the corresponding device (by looking for its printed code).
+In the {\em infrared} condition, participants then selected and acquired the target by aiming the infrared beam at the target, and confirmed their selection with a touch pad tap.
+If more than one target was within range, participants had to either use the disambiguation dialog or reposition themselves.
+In the {\em list} condition, participants had to read the letter next to the number and then select that letter by browsing a linear list shown in their Glass display.
+While the list was alphabetized, letter arrangement in the room was not.
+This design required participants to find the target in the room before starting a list navigation to keep conditions similar to each other.
+
+Afterwards, participants completed a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
+
+\subsection{Participants}
+We recruited \bjoern{N} participants from our institution.
+\bjoern{X} had never used Glass before.
+\bjoern{Y} wore prescription glasses, which may have affected their task performance as wearing both Glass and regular glasses is possible, but leads to \bjoern{explain problems}.
+
+\subsection{Measures}
+The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; and{\bf user preference}: which interface users preferred for the task after completing the study.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/selection-times2.pdf}
+\caption{Task completion times. \bjoern{update data}}
+\label{fig:selection-times}
+\end{figure}
+\subsection{Results}
+
+\subsubsection{Performance data}
+The time to complete each task can be broken down into the following pieces:
+
+$t_{infrared}=t_{locate}+t_{reorient}+t_{disambiguate}+t_{tap}$
+
+$t_{list}=t_{locate}+t_{listnav}+t_{tap}$
+
+In both conditions, participants first have to locate the target announced by the experimenter ($t_{locate}$).
+In the {\em infrared} condition, participants may then directly confirm their selection if only a single target was selected ($t_{tap}$).
+However, if they don't immediately receive feedback that their target was selected, or if multiple targets were selected, users either have change their position or head orientation ($t_{reorient}$) or they have to step through the on-screen disambiguation dialog ($t_{disambiguate}$).
+In the {\em list} condition, participants must scroll through the list to find the desired target identifier ($t_{listnav}$).
+This analysis makes it clear that any performance benefit depends on whether $t_{reorient}+t_{disambiguate}<t_{listnav}$.
+This depends on the number of total devices in the environment (increasing $t_{listnav}$), and their density (which will increase $t_{disambiguate}$).
+
+At 10 devices, total selection times are comparable:
+\bjoern{Hopefully, when adding distractors, the list navigation will become very cumbersome.}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/target-likert.pdf}
+\caption{Likert scale ratings for ease of use of aspects of the targeting task. Error bars show standard error. \bjoern{update data}}
+\label{fig:target-likert}
+\end{figure}
+
+\subsubsection{Preference}
+\bjoern{update: 6 of 7} users preferred infrared mode over list mode.
+While both interfaces were judged similarly on overall ease of connecting, list navigation received the lowest likert scores (see Figure~\ref{fig:target-likert}).
+As self-report data can easily skew positive as participants try to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.
+
+List mode had certain advantages: It was judged to be more accurate and predictable as there was always exactly one device selected in the list (\studyquote{With the list you never have to worry about accidentally picking up two targets}).
+Also, it did not require a clear line of sight to the target device so participants did not have to move from their starting position (\studyquote{The shortcoming of the IR mode was that you had to be a certain distance away in order for it to detect the appliance}).
+
+On the other hand, list mode was judged to be more \studyquote{annoying} and tedious.
+The temple-based touchpad for selection was difficult to use for a participant with long hair: \studyquote{List mode was physically difficult for me to navigate, since my long hair wasn't tied back and it kept interfering with my swiping.} Another participant also commented on the ergonomic challenge of touchpad use on Glass: \studyquote{The strength of the IR mode was that I didn't have to use my fingers as much to control.
+If the items were spaced relatively far apart, it was easy to select a specific appliance.}
+
+\bjoern{add more from survey answers}.
+
+%IR mode has semantic meaning (like I said).
+I'm likely to want visual feedback from any distant device I'm trying to control, and since I'm already looking at it, this is very easy to get.
+List mode was physically difficult for me to navigate, since my long hair wasn't tied back and it kept interfering with my swiping.
+Another comment about list mode was that swiping faster didn't seem to navigate the list faster, which would have been convenient for accessing items like W and Y very far from the beginning of the list.
+This could also be partially overcome by using a circular linked list (where the list wraps from Y to A) instead of a one-way list.
+%With the list you never have to worry about accidentally picking up two targets, but since you always have to swipe through a (longer) list that's not much of an advantage.
+%List mode felt like I could be more confident in what client I was choosing because I couldn't select more than one client.
+The IR mode was a little more natural because it involved less clicking and navigating, but it felt easier to select the wrong client which made the list mode more comfortable.
+%"The shortcoming of the List mode was that the list didn't wrap around. So for example if I wanted to go to Y, I couldn't just go left from A—I had to scroll through all the letters in between.
+
+%The strength of the list mode was that it's pretty accurate, so as long as I swiped the correct amount, I could activate and power off the appliances.
+I also didn't have to move at all, although perhaps that is a shortcoming because otherwise we might just all turn into Wall-E people.
+
+%The strength of the IR mode was that I didn't have to use my fingers as much to control. If the items were spaced relatively far apart, it was easy to select a specific appliance.
+
+%The shortcoming of the IR mode was that you had to be a certain distance away in order for it to detect the appliance. "
+%The List Mode is easier to choose the desired item when the item list is short.
+It became a bit frustrated when I had to swipe many times in order to get to the targeted item.
+The IR mode works more intuitive in this case, but need a longer time to learn and get used to it.
+%list mode is easier to select what you want, but may take a while to get to since you can only swipe through one at a time; IR is quicker and more intuitive, but unstable
+%"list mode is annoying. i find myself counting how many swipes i have to do.
+
+%IR mode is fun and pretty convenient"

### 2013-09-14 server/target_exp.py
more results

@@ -209,6 +209,8 @@ while True:
       state_cue = True
       list_cursor += 1
       logResult("skip", "")
+    elif line[0:3] == "dis":
+      logResult("***discard", "")
     else:
       print 'writing cmd to XBee'
       ser.write(line)  
@@ -261,6 +263,8 @@ while True:
         # sleep(task_interval)  # wait 3s to cue next target
       else:
         logResult("disconnected_wrong", target_id)
+    else:
+      logResult("msg", in_msg)
 
 
 

### 2013-09-14 doc/sigchi14/device.tex
more text and figures
various improvements

@@
 
-\section{Hardware Device and Interaction}
+\section{Interaction and Hardware Device}
 
-Our prototype consists of a Google Glass Explorer Edition head-worn computing device, augmented with an infrared emitter that is mounted on the frame, pointing out in the direction of the wearer's view (Figure~\ref{fig:glass}).
-The IR emitter LED is mounted in an opaque hollow tube, that restricts the outgoing angle of illumination.
+\subsection{Design Goals}
+{\bf Leverage visual attention:} Take advantage of the fact that visual attention can express intention - trigger interaction based on where a user is already looking
+
+{\bf Provide feedback in the environment:} While a near-eye display can push information to the user, locate visual feedback about selection targets in the environment, to prevent distraction and interruption.
+
+{\bf Deal with dense device arrangements and the limit of head positioning}
+
+\bjoern{Finish and expand this section.}
 
@@
 \subsection{Prototype Implementation}
+Our prototype consists of a Google Glass Explorer Edition head-worn computing device, augmented with an infrared emitter that is mounted on the frame, pointing out in the direction of the wearer's view (Figure~\ref{fig:glass}).
+The IR emitter LED is mounted in an opaque hollow tube, that restricts the outgoing angle of illumination.
+
 In our prototype, Glass communicates over Bluetooth to an additional microcontroller board the user has to wear (Atmel ATMega256).
@@
 \includegraphics[width=1.0\columnwidth]{figures/glass-ir-coverage}
-\caption{Our augmented Glass prototype has a frame-mounted infrared emitter.}
+\caption{Different IR configurations suggest that usable beam widths of 2 to 4 feet and distances up to 16 feet }
 \label{fig:coverage}

### 2013-09-14 doc/sigchi14/introduction.tex
various improvements

--- 2013-09-02 doc/sigchi14/introduction.tex
@@
 \section{Introduction}
-\bjoern{What's the overall motivation for the project? What problem does it address?}
 Increasingly, devices and services in our built environment are networked and can be controlled remotely.
 The proliferation of smart, controllable devices such as intelligent lighting, AV equipment, HVAC systems, or kitchen appliances raises the question of how to best interact with them.
-Prior work has introduced several approaches that use handheld mobile devices as {\em universal remote controls} to select and control appliances in the proximate environment- e.g., through pointing at such devices to select them, and through touch-screen or button interfaces to send control commands~\cite{beigl_point_1999,patel_2-way_2003}.
 
-\bjoern{what are the drawbacks of this existing work?}
-Some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is split between looking down at a screen and out at the device to-be-controlled.
-\bjoern{these all sound kind of weak.} These general limitations have led researchers to investigate ``always-available'' physiological input modalities for phones on one hand~\cite{morris_emerging_2011}, and wearable computing devices on the other hand.
+Today, commercial solutions use handheld mobile devices as {\em universal remote controls} to control such appliances~\bjoern{refs!}.
+Commonly, users first browse a list of all available devices, and then call up a device-specific user interface.
+This method faces two challenges: {\em naming} and {\em scoping}.
+Assigning clear names is non-trivial.
+In shared spaces, the person trying to control the device might not be the one that named it - e.g., while an office building manager may know what ``Light 4 in area E'' corresponds to, an occupant may not.
+Second, without a method of scoping selection to automatically filter non-relevant devices, paging though long lists of names or navigating hierarchies becomes potentially more cumbersome than the physical action the ``convenient'' software solution was meant to replace.
+
+To address these challenges, research has introduced techniques of augmenting mobile devices, e.g., with laser pointers, to enable direct aiming at target devices~\cite{beigl_point_1999,patel_2-way_2003}.
+While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at the device to-be-controlled.
 
@@
 Users first look in the direction of the device they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume).
-If multiple devices fall within communication range, an on-screen disambiguation dialog lets users select a single target.
+If multiple devices fall within communication range, an disambiguation technique that combines on-screen information as well as visual feedback on the target devices lets users select their desired target.
 Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).
 
-Our hardware relies on infrared communication between Glass and target devices to establish a connection; and on wireless ZigBee radio communication to exchange control messages.
-Glass is augmented with a narrow IR emitter and a ZigBee ratio.
-Target devices similarly have IR receivers and ZigBee radios.
-These choices offer selection-through-head-orientation to initiate interaction; once initiated, users can look away from the target device while issuing control commands.
+Our hardware relies on infrared communication between Glass and target devices to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.
+Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio.
+Target devices similarly have IR receivers and radios.
+This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target device while issuing control commands.
 
-To understand the system performance, usability, and user experience of head-orientation targeting and head-mounted displays for device control, we first report measurements of the target range and accuracy of our device.
-We then conduct a comparative study of device acquisition time and error of different interface variants.
-We find that target acquisition through head orientation is faster than selecting items from a list, given the constraints of linear input using a head-worn touch controller.
-Finally, we report high-level feedback from $n$ users who use our system for home automation tasks.
+While prior work has tended to focus on proofs-of-concept, we also contribute empirical data on the system performance, usability, and user experience of  head-orientation targeting and device control.
+We first report measurements of range and beam characteristics of our controller.
+We then conduct a study that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
+We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.
+We also report qualitative results from $n$ participants who use our system for home automation tasks.

### 2013-09-14 doc/sigchi14/relatedwork.tex
added some refs
various improvements

--- 2013-09-03 doc/sigchi14/relatedwork.tex
@@
 \subsection{Remote Control of Physical Appliances}
-Universal device control applications for smart phones are commercially available (e.g., Belkin WeMo\footnote{http://www.belkin.com/us/wemo}), though they usually do not offer spatial selection of target devices, forcing users to browse through lists of configured devices instead.
+Standard infrared remote controls for televisions and AV equipment are only meant to control a single device.
+These controllers tend to use wide-angle infrared LEDs.
+Universal remote controls are available as dedicated devices (e.g, Crestron\footnote{\url{http://www.crestron.com}}) or applications for smart phones (e.g., Belkin WeMo\footnote{http://www.belkin.com/us/wemo}).
+They do not offer spatial selection of target devices, forcing users to browse through lists of pre-configured devices instead.
 Rukzio found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.
@@
 In several techniques, users select objects of interest with laser pointers.
-The laser dot provides immediate visual feedback to the user what is being selected.
+The laser dot provides immediate visual feedback to the user what is being selected (though it does not indicate whether the pointed-at object can indeed be controlled).
 Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}.
@@
 
-Standard infrared remote controls for televisions and AV equipment are only meant to control a single device and it is unnecessary to have the user point precisely.
-These controllers tend to use wide-angle infrared LEDs.
-
-Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing - the user looks at the target device to initiate interaction.
+Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing --- the user looks at the target device to initiate interaction.
 Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications as head orientation only indicates a general area of visual interest.
@@
 Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy.
+Our work only uses head orientation for the initial selection step; since we believe a user's attention is often drawn to the objects they intend to interact with.
+Further interactions do not use head orientation.
 
@@
 Agumented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
-Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world.
+Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world, though our techniques could be extended to such configurations.
 
+\bjoern{add something about the assembly instruction work here.}
 %\bjoern{most of the prior head-mounted work is in the AR space. I don't know much about it yet, but Kortuem~\cite{kortuem_context-aware_1998} might be a good starting point.}

### 2013-09-14 doc/sigchi14/device.tex
placeholder
pass through first couple of pages

@@
+\section{Hardware Device}
 \begin{figure}[t]
@@
 \end{figure}
-
-\section{Interaction and Hardware Device}
-
-\subsection{Design Goals}
-{\bf Leverage visual attention:} Take advantage of the fact that visual attention can express intention - trigger interaction based on where a user is already looking
-
-{\bf Provide feedback in the environment:} While a near-eye display can push information to the user, locate visual feedback about selection targets in the environment, to prevent distraction and interruption.
-
-{\bf Deal with dense device arrangements and the limit of head positioning}
-
-\bjoern{Finish and expand this section.}
-
-\subsection{Interaction Flow}
-{\bf Look:} Users select a target device by looking in its general direction.
-Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
-Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received.
-This enables {\em scanning} the environment with ones gaze to see which devices can be controlled.
-They also respond by initiating an XBee radio connection and sending their id.
-This allows the Glass application to show which devices are currently targeted on the display.
-\sean{Currently the led is implemented but clients don't response with their IDs and display on Glass at this stage.
-We can add that.
-But we felt like the users would use led as the primary indicator and not to distract them with frequently changing the content on the screen?
-}
-
-{\bf Initiate:} Users confirm to initiate interaction by tapping on the Glass touchpad.
-The next section on disambiguation deals with cases in which multiple devices received valid IR signals.
-At this point, all further communication switches over to the XBee wireless network so that line of sight to the target is no longer needed (otherwise restricting head movement can become straining).
-
-{\bf Control:} Glass displays a control interface for parameters of the chosen device. \bjoern{Describe the interaction scheme for navigating and adjusting parameters here}.
-Control commands are sent over XBee radios.
-
-{\bf Disengagement:} \bjoern{how do you end an interaction?
-Is there a timeout if the user forgets to manually back out of the control screen?} \sean{currently the only way is to swipe back.
-A timeout mechanism can be added easily.
-I guess the design decision lies in what is an appropriate duration?}
-
-\subsection{Disambiguation}
-\bjoern{describe why disambiguation is necessary - combination of the relative inaccuracy of head orientation and the spread of our IR signal. Then describe how to overcome it.}
 
@@
 
-\section{Device Characterization}
+\subsection{Device Characterization}
 We determined the usable range and accuracy empirically with one IR emitter and two IR receivers.
@@
 In summary, \bjoern{describe what the measured results mean in practice.}
-
-\begin{figure}[t]
+\afterpage{
+\begin{figure}[t!]
 \centering
@@
 
-\begin{table}
+\begin{table}[h!]
     \begin{tabular}{l|lllll}
@@
     \label{table:measurements}
-
 \end{table}
+} %end afterpage

### 2013-09-14 doc/sigchi14/relatedwork.tex
pass through first couple of pages

@@
 Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing --- the user looks at the target device to initiate interaction.
-Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications as head orientation only indicates a general area of visual interest.
-It does not necessarily match gaze orientation as extra-ocular muscles can move the eyes.
-We therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.
+Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications, we therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.
 
@@
 Our work only uses head orientation for the initial selection step; since we believe a user's attention is often drawn to the objects they intend to interact with.
-Further interactions do not use head orientation.
 
@@
 
-\bjoern{add something about the assembly instruction work here.}
+%\bjoern{add something about the assembly instruction work here.}
 %\bjoern{most of the prior head-mounted work is in the AR space. I don't know much about it yet, but Kortuem~\cite{kortuem_context-aware_1998} might be a good starting point.}

### 2013-09-14 data/script/aggregate.py
I've add an aggregate script to combine all our dataset to a single csv file, which is data.csv, you may check it out. Next step is to use R to analyze and visualize

@@ -1,0 +1,67 @@
+from glob import glob
+## schema would be 
+## participant, mode, target, correct_time, correct_type, first_miss_time, miss_count, switch_count, switch_target
+class log:
+  def __init__(self, participant, mode, target):
+    self.participant = participant
+    self.mode = mode
+    self.target = target
+    self.correct_type = ""
+    self.correct_time = ""
+    self.first_miss_time = ""
+    self.miss_count = 0
+    self.switch_count = 0
+    self.switch_target = []
+    self.wrong_count = 0
+    self.wrong_target = []
+    self.fate = []
+  def __repr__(self):
+    return self.participant + ',\t' +\
+           self.mode + ',\t' +\
+           self.target + ',\t' +\
+           self.correct_time + ',\t' +\
+           self.correct_type + ',\t' +\
+           self.first_miss_time + ',\t' +\
+           str(self.miss_count) + ',\t' +\
+           str(self.switch_count) + ',\t' +\
+           '[' + ';'.join(map(str, self.switch_target)) + '],\t' +\
+           str(self.wrong_count) + ',\t' +\
+           '[' + ';'.join(map(str, self.wrong_target)) + '],\t' +\
+           '[' + ';'.join(map(str, self.fate)) + ']'
+
+filelist = glob("../*.txt")
+f1=open('./data.csv', 'w+')
+print >>f1, "participant, mode, target, correct_time, correct_type, first_miss_time, miss_count, switch_count, switch_target, wrong_count, wrong_target, fate"
+        
+for f in filelist:
+  logs = []
+  new_log = None
+  lines = [line.strip().split() for line in open(f)]
+  participant = f.split('_')[1]
+  mode = f.split('_')[2][:-4]
+  for l in lines:
+    if not l or l[1] == "msg":
+      continue
+    if l[1] == "start":
+      if new_log is not None:
+        print >>f1, str(new_log).expandtabs(4)
+      new_log = log(participant, mode, l[2])
+    if any(item in l[1] for item in ["reset", "disconnect", "next", "jump"]) and new_log is not None:
+      new_log.fate.append(l[1])      
+    if new_log is not None:
+      ## then it's good to start monitoring keywords here
+      if l[1].split('_')[0] == "correct":
+        new_log.correct_type = l[1].split('_')[1]
+        new_log.correct_time = l[0]
+      if l[1].split('_')[0] == "miss":
+        new_log.miss_count += 1
[…]

### 2013-09-15 doc/sigchi14/control.tex
update Likert Ratings: Home Control Scenario

@@
 \subsection{Methodology}
-We recreated a living room environment that had four controllable appliances: a fan, a lamp, and two laptops functioning as an audio and video player, respectively (see Figure~\ref{fig:smarthome}).
+We recreated a living room environment that had three controllable appliances: a fan, a lamp, and one laptop functioning as a video player, respectively (see Figure~\ref{fig:smarthome}).
 The fan and lamp had binary controls: they could be switched on or off.
-The laptops had multiple parameterized functions: participants could start, pause, stop, and rewind  video and audio, and adjust volume.
+The laptop had multiple parameterized functions: participants could start, pause, fastforward, rewind, and adjust volume.
 
@@
 \item When the credits roll, you stop the movie and turn the lights back on.
-\item You want to listen to some music.
-\item After awhile, you stop the music you turn off the fan and leave the room.
+\item After awhile, you turn off the fan and leave the room.
 \end{enumerate}

### 2013-09-15 data/script/aggregate.py
a few new figures, temporarily just in the data/script folder

@@ -1,6 +1,6 @@
 from glob import glob
 ## schema would be 
-## participant, mode, target, correct_time, correct_type, first_miss_time, miss_count, switch_count, switch_target
+## participant, mode, target, correct_time, correct_type, first_miss_time, miss_count, switch_count, switch_target, wrong_count, wrong_target, multiple_time, fate
 class log:
   def __init__(self, participant, mode, target):
     self.participant = participant
@@ -14,24 +14,26 @@ class log:
     self.switch_target = []
     self.wrong_count = 0
     self.wrong_target = []
+    self.multiple_time = ""
     self.fate = []
   def __repr__(self):
-    return self.participant + ',\t' +\
-           self.mode + ',\t' +\
-           self.target + ',\t' +\
-           self.correct_time + ',\t' +\
-           self.correct_type + ',\t' +\
-           self.first_miss_time + ',\t' +\
-           str(self.miss_count) + ',\t' +\
-           str(self.switch_count) + ',\t' +\
+    return self.participant + ',' +\
+           self.mode + ',' +\
+           self.target + ',' +\
+           self.correct_time + ',' +\
+           self.correct_type + ',' +\
+           self.first_miss_time + ',' +\
+           str(self.miss_count) + ',' +\
+           str(self.switch_count) + ',' +\
            '[' + ';'.join(map(str, self.switch_target)) + '],\t' +\
-           str(self.wrong_count) + ',\t' +\
+           str(self.wrong_count) + ',' +\
            '[' + ';'.join(map(str, self.wrong_target)) + '],\t' +\
+           self.multiple_time + ',' +\
            '[' + ';'.join(map(str, self.fate)) + ']'
 
 filelist = glob("../*.txt")
 f1=open('./data.csv', 'w+')
-print >>f1, "participant, mode, target, correct_time, correct_type, first_miss_time, miss_count, switch_count, switch_target, wrong_count, wrong_target, fate"
+print >>f1, "participant, mode, target, correct_time, correct_type, first_miss_time, miss_count, switch_count, switch_target, wrong_count, wrong_target, multiple_time, fate"
         
 for f in filelist:
   logs = []
@@ -63,5 +65,9 @@ for f in filelist:
       if l[1].split('_')[0] == "wrong":
         new_log.wrong_count += 1
         new_log.wrong_target.append(l[2])
+      if l[1].split('_')[0] == "multiple":
+        new_log.multiple_time = l[0]
+
+
       
         

### 2013-09-15 data/script/visualize.R
R code

@@ -1,0 +1,49 @@
+require('ggplot2')
+
+## read data in first
+df <- read.csv("data.csv")
+
+## discard data frmo valkerie and Micheal
+df <- df[!is.na(df$correct_time),]
+df <- df[!(grepl("valkerie", df$participant) | grepl("Micheal", df$participant)), ]
+
+df.IR.single <- df[grepl("IR", df$mode) & grepl("single", df$correct_type), ]
+df.IR.mul <- df[grepl("IR", df$mode) & grepl("mul", df$correct_type), ]
+
+
+#### draw the list mode, time vs. position
+pdf("List_by_Target.pdf")
+
+df.List <- df[grepl("List", df$mode), ]
+# mapping from id to name
+name <- "GOWFJSYRAM"
+df.List$target <- sapply(df.List$target, function(x) substr(name, x, x))
+
+label.x <- sapply(names(table(df.List$target)), function(x) regexpr(x, name))
+box <- ggplot(df.List, aes(factor(target), correct_time, fill = df.List$target))
+box <- box + geom_boxplot(lwd=0.5) + ylim(2.5,15) + xlab("target by name")
+box + theme(legend.position = "none", axis.title.y=element_blank()) +
+
+## I don't know how to do annotation
+## geom_text(aes(x=label.x, y=3, label=table(df.List$target))) 
+dev.off()
+
+
+
+#### draw the list mode, time vs. position
+mode.bind.type <- cbind(df, paste(df$mode, df$correct_type))
+colnames(mode.bind.type)[14] <- "mode_type"
+
+pdf("time_by_Category.pdf")
+toPlot <- mode.bind.type
+box <- ggplot(toPlot, aes(factor(mode_type), correct_time, fill = toPlot$mode_type))
+box <- box + geom_boxplot(lwd=0.5) + ylim(2.5,25) + xlab("different mode")
+       box + theme(legend.position = "none", axis.title.y=element_blank())
+dev.off()
+
+##    IR mul IR single  List mul 
+##       19       168       205 
+pdf("mul_vs_single.pdf")
+pie(c(19, 168), c("IR multiple", "IR single"))
+dev.off()
+

### 2013-09-15 doc/sigchi14/interaction.tex
added interaction.tex

\begin{figure*}[t!]
\centering
\includegraphics[width=7in]{figures/stepbystep}
\caption{Step by step walkthrough of interaction goes here.}
\label{fig:interaction}
\end{figure*}

\section{Initiating Interaction through Attention}
This section describes the design goals of our systems and their realization in particular interaction techniques.

\subsection{Design Goals}
Our work is motivated by the following design goals that leverage opportunities of head-worn computing, but also acknowledge potential challenged:

{\bf Leverage visual attention:} Take advantage of the fact that visual attention can express intention - initiate interaction based on where a user is already looking.

{\bf Provide immediate feedback about selection targets in the environment:} While a near-eye display can push information to the user, users don't always want to control an object simply because they are looking at it (a problem known in gaze-based interaction as the Midas touch). A calmer~\cite{weiser_coming_1997} approach is to locate visual feedback about selection targets in the environment, to prevent distraction and interruption. Such feedback should be delivered instantaneously, while users look around a room.

{\bf Offer flexible orientation after initiating interaction:} After initiating attention through head orientation, enable the user to reorient their head or body position during the remaining interaction to prevent neck strain.

{\bf Offer efficient ways to disambiguate orientation input:} It may not always be possible to identify a unique target device based on a user's attention and orientation. Offer ways to supplement orientation-based interaction with screen-based interaction to provide disambiguation information.

These design goals find their expression in the following interaction model:

\subsection{Interaction Flow}
{\bf Look:} Users select a target device by looking in its general direction.
Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}. Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}A). This enables {\em scanning} the environment with ones gaze to see which devices can be controlled.

{\bf Initiate:} Users confirm their desire to connect to a device by tapping on the Glass touchpad. The next section on disambiguation deals with cases in which multiple devices received valid IR signals. At this point, all further communication switches over to the 802.15.4 wireless network so that line of sight to the target is no longer needed.

{\bf Control:} Glass displays a user interface for parameters of the chosen device (Figure~\ref{fig:interaction}B). The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters; single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters. This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the saggital plane (up and down). 
Control commands are sent over XBee radios.

{\bf Disengagement:} Users stay connected to the last selected device up to a timeout period. During that period, users can disengage through down swipes.

\subsection{Disambiguation}
Head orientation only indicates a general area of visual interest. It does not necessarily match gaze orientation as extra-ocular muscles can move the eyes. The IR beam of our device also has a certain spread (see next section). In an environment dense with potential targets, multiple targets could be within range. Users can tell when multiple feedback LEDs in the environment illuminate. To disambiguate, users can either move to adjust their head position, or, alternatively, call up a disambiguation dialog on the Glass display. The dialog presents a list filtered to only those devices that are within IR range. Users navigate the list using the touchpad, and then continue their interaction as described above.

### 2013-09-15 doc/sigchi14/targeting.tex
add comments

@@
 Next to each client device, a paper sheet shows a number and letter combination, which are used for uniquely identifying the devices.
-
-Some trials also used a {\em distractor} configuration in which 15 additional paper sheets with additional identifiers are added to the environment.
-We did not create 15 additional wireless nodes though as each node has to be hand-assembled and soldered.
+\sean{The numbers are the primary identifiers while the letters are in raondom order.
+The nodes are somewhat sorted by their numbers, making it easy to find in order to simulate a place the users are more familiar with.
+Therefore, the time spent on finding the target was minimized and the time measured was mainly the acqusition time.
+}
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/target-likert.pdf}
-\caption{Likert scale ratings for ease of use of aspects of the targeting task. Error bars show standard error. \bjoern{update data}}
+\caption{Likert scale ratings for ease of use of aspects of the targeting task. Error bars show standard error. }
 \label{fig:target-likert}

### 2013-09-16 doc/sigchi14/targeting.tex
add time-vs-list-order figure to the text

@@
 \subsection{Measures}
-The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; and{\bf user preference}: which interface users preferred for the task after completing the study.
+The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; and {\bf user preference}: which interface users preferred for the task after completing the study.
 
@@
 \bjoern{Hopefully, when adding distractors, the list navigation will become very cumbersome.}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
+\caption{Average time taken vs. the order of nodes on the list }
+\label{fig:time-vs-list-order}
+\end{figure}
 

### 2013-09-16 data/script/visualize.R
using new data set, and changed R code to directly create figure under paper folder

@@ -1,5 +1,52 @@
 require('ggplot2')
 
+# Multiple plot function
+#
+# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
+# - cols:   Number of columns in layout
+# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
+#
+# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
+# then plot 1 will go in the upper left, 2 will go in the upper right, and
+# 3 will go all the way across the bottom.
+#
+multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  require(grid)
+
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+
+  numPlots = length(plots)
+
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                    ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+
+ if (numPlots==1) {
+    print(plots[[1]])
+
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}
+
+
 ## read data in first
 df <- read.csv("data.csv")
 
@@ -12,38 +59,72 @@ df.IR.mul <- df[grepl("IR", df$mode) & grepl("mul", df$correct_type), ]
 
 
 #### draw the list mode, time vs. position
-pdf("List_by_Target.pdf")
-
+pdf("../../doc/sigchi14/figures/R_List_by_Target.pdf", width=5, height=3)
 df.List <- df[grepl("List", df$mode), ]
[…]

### 2013-09-16 data/script/visualize.R
update figure height

@@ -59,7 +59,7 @@ df.IR.mul <- df[grepl("IR", df$mode) & grepl("mul", df$correct_type), ]
 
 
 #### draw the list mode, time vs. position
-pdf("../../doc/sigchi14/figures/R_List_by_Target.pdf", width=5, height=3)
+pdf("../../doc/sigchi14/figures/R_List_by_Target.pdf", width=7, height=4)
 df.List <- df[grepl("List", df$mode), ]
 # mapping from id to name
 name <- "GOWFJSYRAM"
@@ -98,7 +98,7 @@ dev.off()
 #### draw the list mode, time vs. position
 mode.bind.type <- cbind(df, paste(df$mode, df$correct_type))
 colnames(mode.bind.type)[14] <- "mode_type"
-pdf("../../doc/sigchi14/figures/R_time_by_Category.pdf", width=5, height=4)
+pdf("../../doc/sigchi14/figures/R_time_by_Category.pdf", width=7, height=4)
 box1 <- ggplot(mode.bind.type, aes(factor(mode), correct_time, fill = mode.bind.type$mode))
 box1 <- box1 + geom_boxplot(lwd=0.5) + ylim(2.5,25) + xlab("different mode") +
     theme(legend.position = "none") +

### 2013-09-16 doc/sigchi14/targeting.tex
added numbers to participants

@@
 \subsection{Participants}
-We recruited \bjoern{N} participants from our institution.
-\bjoern{X} had never used Glass before.
-\bjoern{Y} wore prescription glasses, which may have affected their task performance as wearing both Glass and regular glasses is possible, but leads to \bjoern{explain problems}.
+We recruited 15 participants from our institution.
+13 had never used Glass before.
+4 wore prescription glasses, which may have affected their task performance \claire as the glasses beneath Glass make it more cumbersome to secure the position of the Glass and to adjust the screen to the optimal angle.
 

### 2013-09-16 doc/sigchi14/targeting.tex
add descriptions to charts
added strengths/shortcomings of IR

@@
 
-\bjoern{add more from survey answers}.
+\claire One noted benefit of IR mode was a feeling that it was \studyquote{more direct}.
+One subject appreciated the ability to \studyquote{interact with things just by looking at them}.
+Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}.
+
+\claire A perceived shortcoming of IR mode was the necessity to move the head in order to control a device.
+One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}.
+Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}.
 

### 2013-09-16 doc/sigchi14/targeting.tex
modified targeting
fixed CT attributions

@@
 13 had never used Glass before.
-4 wore prescription glasses, which may have affected their task performance \claire as the glasses beneath Glass make it more cumbersome to secure the position of the Glass and to adjust the screen to the optimal angle.
+4 wore prescription glasses, which may have affected their task performance \claire{ as the glasses beneath Glass make it more cumbersome to secure the position of the Glass and to adjust the screen to the optimal angle.}
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/selection-times2.pdf}
-\caption{Task completion times. \bjoern{update data}}
-\label{fig:selection-times}
-\end{figure}
 \subsection{Results}
-
 \subsubsection{Performance data}
@@
 \bjoern{Hopefully, when adding distractors, the list navigation will become very cumbersome.}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
+\caption{Task completion times.} \sean{label mean on the chart}
+\label{fig:selection-times}
+\end{figure}
+
+As a result, infrared mode out performs list mode by 2.186 seconds (6.672 vs 8.858), as shown in figure~\ref{fig:selection-times}. \sean{TBC}
 
@@
 
-\claire One noted benefit of IR mode was a feeling that it was \studyquote{more direct}.
+\claire{ One noted benefit of IR mode was a feeling that it was \studyquote{more direct}.
 One subject appreciated the ability to \studyquote{interact with things just by looking at them}.
@@
 
-\claire A perceived shortcoming of IR mode was the necessity to move the head in order to control a device.
+ A perceived shortcoming of IR mode was the necessity to move the head in order to control a device.
 One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}.
-Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}.
+Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}.}
 

### 2013-09-16 doc/sigchi14/targeting.tex
add result discussions, to be refined

@@
 
+\subsubsection{Discussion}
+\sean{to be refined}
+
+\begin{itemize}
+\item During the experiment, a lot of wireless messages were transferred for communication and logging purposes.
+Some messages got lost or delayed and caused the users had to redo that particular action.
+Infrared mode requires a lot of more handshake communications than the list mode.
+This means that in a more robust infrastructure, the performance of infrared mode can improve more than that of the list mode.
+
+\item Some of the infrared performance were affected by the strength and stableness of the infrared emitter.
+By getting a stronger emitter that would be able to reach to a longer distance, connecting to nodes further away such as 4 and 10 would become easier.
+
+\item As shown in figure X, the average time taken to connect to a certain client is correlated to the order of the node within the list.
+We therefore can assume that as the system scales, the average time required in list mode would have a linear growth.
+As for the infrared mode, it is more about the density of nodes in a fix space.
+There’s an upper bound how many smart appliance one can put on a single ta
+
+\item In list mode, users’ eyes focuses more on the display to traverse between the list while in infrared mode, users’ eyes remain in the physical world more.
+
+\item In our prototype, the IR emitter is nearly fixed.
+Due to different sizes and shapes of the users’ heads, the emitter wouldn’t necessarily point to the center of their eye sights.
+An adjustable emitter would even further improve the performance of our design.
+
+\item The list mode could also be improved by making the swiping gestures more responsive, e.g. swiping speed affects traversing speed.
+
+\item In our experiment, after users locates the nodes by ID, they can instantly figure out their names by the label provided.
+In realistic environments, users have to recall the names of the appliances, which would take a lot more time.
+Furthermore, finding names such as “Living room corner stand lamp” would be more difficult than an alphabet.
+On the other hand, infrared mode will most likely to be unaffected since it does not require name-recalling or text matching.
+
+ \end{itemize}
+
 %IR mode has semantic meaning (like I said).

### 2013-09-17 doc/sigchi14/targeting.tex
draft texts for analysis results

@@
 
-At 10 devices, total selection times are comparable:
-\bjoern{Hopefully, when adding distractors, the list navigation will become very cumbersome.}
-
 \begin{figure}[t]
@@
 \includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
-\caption{Task completion times.} \sean{label mean on the chart}
+\caption{Task completion times.} \ben{label mean, change xtick, change color}
 \label{fig:selection-times}
@@
 
-As a result, infrared mode out performs list mode by 2.186 seconds (6.672 vs 8.858), as shown in figure~\ref{fig:selection-times}. \sean{TBC}
+We will first show our results at 10 devices and then provide discussions about extrapolation.
+The $t_{infrared}$ and $t_{list}$ are shown in Figure~\ref{fig:selection-times} {\em (left)}.
+The average time in two conditions are 6.67 seconds and 8.86 seconds, respectively.
+Student's t-test result (p=0.00017) shows that they are significantly different.
+
+To further understand the performance gain in {\em infrared} condition, especially the factor $t_{disambiguate}$, we compare the time it takes in condition of whether there are multiple devices involved (Figure~\ref{fig:selection-times} {\em (right)}).
+When there is a single device, $t_{disambiguate}$ is 0; it takes 6.40 seconds (on average) to complete one connection.
+When multiple in range, the time increases to 9.16 seconds, indicating 2.76 seconds required to disambiguate.
+Though it takes significantly longer (p=0.012 using t-test) in the {\em multiple} case, it's worth pointing out that it happens less frequent (10\%), even we intentionally placed a few devices nearer to each other.
+% 19 / (19+168 )
+
+% \bjoern{Hopefully, when adding distractors, the list navigation will become very cumbersome.}
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
-\caption{Average time taken vs. the order of nodes on the list }
+\caption{Average time taken vs. the order of nodes on the list  } \ben{trend line}
 \label{fig:time-vs-list-order}
 \end{figure}
+
+For each device, $t_{listnav}$ would depends on their relative position in the list.
+And the average time for all targets would increase if the number of total devices increases.
+To verify this conjecture, we visualize the time it takes to connect to each device in Figure~\ref{fig:time-vs-list-order}, ordered by their relative position in the list.
+The order would also reflect the number of swipes to reach each device.
+A linear trend is visible, and the regression returns with coefficient of determination ($R^2=0.032$).
 

### 2013-09-17 data/script/visualize.R

@@ -67,32 +67,52 @@ df.List$target <- sapply(df.List$target, function(x) substr(name, x, x))
 label.x <- sapply(names(table(df.List$target)), function(x) regexpr(x, name))
 box <- ggplot(df.List, aes(factor(target), correct_time, fill = df.List$target))
 box <- box + geom_boxplot(lwd=0.5) + ylim(2.5,15) + xlab("order in list") + scale_x_discrete(labels=seq(0, 9)) + ylab("time (seconds)")
-# linear fit
+## linear fit
 sorted.name <- sort(unique(df.List$target))
 target1 <- sapply(df.List$target, function(x) which(sorted.name == x) - 1)
 fit <- lm(df.List$correct_time ~ target1)
-target1 <- sapply(df.List$target, function(x) which(sorted.name == x))
-box <- box + geom_line(aes(target1, predict(fit)))
-box + theme(legend.position = "none")
-abline(fit)
-dev.off()  
+# geom_line(x = c(0,9), y = c(6.740350,10.85737), colour = "red")
+box + theme(legend.position = "none") 
+dev.off()
+
+## > summary(fit)
+
+## Call:
+## lm(formula = df.List$correct_time ~ target1)
+
+## Residuals:
+##    Min     1Q Median     3Q    Max 
+## -3.904 -2.106 -1.066  0.260 95.364 
+
+## Coefficients:
+##             Estimate Std. Error t value Pr(>|t|)    
+## (Intercept)   6.7403     0.9685   6.959 4.62e-11 ***
+## target1       0.4574     0.1768   2.588   0.0104 *  
+## ---
+## Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1 
+
+## Residual standard error: 7.416 on 203 degrees of freedom
+## Multiple R-squared: 0.03193,	Adjusted R-squared: 0.02716 
+## F-statistic: 6.696 on 1 and 203 DF,  p-value: 0.01036 
 
 
+pdf("../../doc/sigchi14/figures/R_List_by_Target.pdf", width=7, height=4)
+ggplot(data = data.frame(x = c(0,9), y = c(6.740350,10.85737)), aes = aes(x = x, y = y, color="red")) + geom_line()
+dev.off()
+
+## fix the color
 #### draw the list mode, time vs. position
 pdf("../../doc/sigchi14/figures/R_IR_by_Target.pdf")
[…]

### 2013-09-17 data/script/visualize.R
update figures

@@ -65,13 +65,14 @@ df.List <- df[grepl("List", df$mode), ]
 name <- "GOWFJSYRAM"
 df.List$target <- sapply(df.List$target, function(x) substr(name, x, x))
 label.x <- sapply(names(table(df.List$target)), function(x) regexpr(x, name))
-box <- ggplot(df.List, aes(factor(target), correct_time, fill = df.List$target))
-box <- box + geom_boxplot(lwd=0.5) + ylim(2.5,15) + xlab("order in list") + scale_x_discrete(labels=seq(0, 9)) + ylab("time (seconds)")
+box <- ggplot(df.List, aes(factor(target), correct_time))
+box <- box + geom_boxplot(lwd=0.5, fill = '#9FB987') + ylim(2.5,15) + xlab("order in list") + ylab("time (seconds)") # + scale_x_discrete(labels=seq(0, 9))
+box <- box + geom_smooth(lwd=1.5, linetype=2, colour = "#477843", aes(group = 1), method="lm", se = F)
+#box <- box + geom_line(data.frame(x = c(0,9), y = c(6.740350,10.85737)), aes(x, y), colour = "red")
 ## linear fit
 sorted.name <- sort(unique(df.List$target))
 target1 <- sapply(df.List$target, function(x) which(sorted.name == x) - 1)
 fit <- lm(df.List$correct_time ~ target1)
-# geom_line(x = c(0,9), y = c(6.740350,10.85737), colour = "red")
 box + theme(legend.position = "none") 
 dev.off()
 
@@ -122,13 +123,14 @@ pdf("../../doc/sigchi14/figures/R_time_by_Category.pdf", width=7, height=4)
 box1 <- ggplot(mode.bind.type, aes(factor(mode), correct_time, fill = mode.bind.type$mode))
 box1 <- box1 + geom_boxplot(lwd=0.5) + ylim(2.5,25) + xlab("different mode") +
     theme(legend.position = "none") +
+    scale_fill_manual(values = c("#A4C3D9", "#9FB987")) + #477843
     stat_summary(fun.y = "mean", geom = "text", label="---", size= 8, color= "white") +
     ylab("time (seconds)")
 toPlot <- mode.bind.type[grepl("IR", mode.bind.type$mode_type), ]
 box2 <- ggplot(toPlot, aes(factor(mode_type), correct_time, fill = toPlot$mode_type))
[…]

### 2013-09-17 doc/sigchi14/targeting.tex
added some to advan/disadvan

@@
 
-\claire{ One noted benefit of IR mode was a feeling that it was \studyquote{more direct}.
-One subject appreciated the ability to \studyquote{interact with things just by looking at them}.
+\claire{ One noted benefit of IR mode was a feeling that it was \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
+One subject called it \studyquote{natural to interact with things just by looking at them}.
 Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}.
 
- A perceived shortcoming of IR mode was the necessity to move the head in order to control a device.
+A few perceived weaknesses of IR mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target.
 One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}.
-Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}.}
+Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}.
+Users had to learn how the angle of the IR before they became successful at controlling the devices:\studyquote{I had to compensate by tilting my head up a little bit.}}
 

### 2013-09-18 doc/sigchi14/control.tex
figure optimizations
update smarthome-scenario figure
bjoern's pass

@@
-\section{Example SmartHome Scenario}
+\section{Smart Home Control Scenario}
 To understand how our device could be used to interact with smart appliances, we also asked all study participants to work through a concrete scenario.
@@
 \subsection{Methodology}
-We recreated a living room environment that had three controllable appliances: a fan, a lamp, and one laptop functioning as a video player, respectively (see Figure~\ref{fig:smarthome}).
+We recreated a living room environment that had three controllable appliances: a fan, a lamp, and one laptop functioning as a video player (see Figure~\ref{fig:smarthome}).
 The fan and lamp had binary controls: they could be switched on or off.
-The laptop had multiple parameterized functions: participants could start, pause, fastforward, rewind, and adjust volume.
+The laptop had multiple parameterized functions: participants could start, pause, fast forward, rewind, and adjust volume.
 
@@
 
-For this study, we only elicited subjective data in the form of Likert data and 
+For this study, we only elicited subjective data in the form of Likert data and open-ended responses.
 \begin{figure}[b]
@@
 All participants successfully completed the list of tasks.
-Participants stated it \studyquote{felt easy to interact with the appliances}, but there were distinct differences between appliances.
-Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for more complex TV and music player devices (see Figure~\ref{fig:smarthome-likert}).
+Participants stated it was easy to interact with the appliances, but there were distinct differences between appliances.
+Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for the more complex movie player (see Figure~\ref{fig:smarthome-likert}).
 Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
-The screen size and 1D input thus likely put a limit on the complexity of interfaces that can be presented.
-
-\bjoern{explain why music is so low, if this trend holds}
+The screen size and 1D input put a limit on the complexity of interfaces that can be presented.
 
@@
 As one participant remarked \studyquote{This was especially annoying with the TV because there were two screens overlapping each other.} While it is possible to look away once a device has been acquired to better see the Glass display, this tension is likely to remain for users who want to rely on feedback from the devices themselves instead of the Glass display.
-
-\subsubsection{Control}
 

### 2013-09-18 doc/sigchi14/device.tex
bjoern's pass

@@
 \includegraphics[width=1.0\columnwidth]{figures/architecture}
-\caption{System architecture.}
+\caption{In our system architecture, selection is initiated through infrared but confirmed over 802.15.4.
+This permits wearers to move their head freely after connecting to an appliance.
+In the research prototype, users have to carry an additional microcontroller board that marshalls messages between Glass' Bluetooth radio an IR/802.15.4, but our custom hardware could also be integrated into the wearable device.}
 \label{fig:architecture}
@@
 This architecture was mostly chosen for reasons of expediency.
-We selected XBee 802.15.4 radios to avoid the latency associated with connecting and disconnecting to Bluetooth devices \bjoern{Is this true or not?} \sean{Origainlly we didn't know we'd work on Google Glass, therefore BT didn't have a higher priority.
-We learned that bluetooth can concurrently communicate with up to 7 devices (while up to 255 can be paired but inactive), giving us more complexity when scaling up.
-In addition ZigBee has shorter wakeup delay and lower power consumption.
-(referenced here: http://superuser.com/questions/332767/limit-to-the-number-of-devices-that-can-be-paired-with-a-bluetooth-device)} Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
-In particular, one could substitute WiFi modules.
+We selected XBee 802.15.4 radios to avoid Bluetooth  wake up latencies but we do not claim optimality for our design decisions.
+Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
+In particular, one could substitute WiFi modules or design an all-Bluetooth network.
 
@@
 
-In summary, \bjoern{describe what the measured results mean in practice.}
+In summary, our measurements suggest that IR communication can be targeted to an area about 2--4' in diameter, up to 16' in front of the user.
+These values are a reasonable match for selecting appliances in a room-size environment.
+A wider beam would lead to an increased chance of multiple appliances receiving IR signals simultaneously.
+A narrower beam will make targeting more challenging, given the precision constraints of human head movement.
+
 \afterpage{

### 2013-09-18 doc/sigchi14/discussion.tex
various improvements
bjoern's pass

@@
 \begin{enumerate}
+\item Midas Touch problem? don't want to permanently flash info on screen based on gaze direction.
 \item Some fundamental differences to earlier laser-pointer based systems: with a laser pointer in hand one can very precisely point at an arbitrary object and know that one is pointing; however, that alone does not provide feedback on whether a device can actually be controlled or not.
@@
 \end{enumerate}
+
+\subsubsection{Discussion}
+\sean{to be refined}
+
+\begin{itemize}
+\item During the experiment, a lot of wireless messages were transferred for communication and logging purposes.
+Some messages got lost or delayed and caused the users had to redo that particular action.
+Infrared mode requires a lot of more handshake communications than the list mode.
+This means that in a more robust infrastructure, the performance of infrared mode can improve more than that of the list mode.
+
+\item Some of the infrared performance were affected by the strength and stableness of the infrared emitter.
+By getting a stronger emitter that would be able to reach to a longer distance, connecting to nodes further away such as 4 and 10 would become easier.
+
+\item As shown in figure X, the average time taken to connect to a certain client is correlated to the order of the node within the list.
+We therefore can assume that as the system scales, the average time required in list mode would have a linear growth.
+As for the infrared mode, it is more about the density of nodes in a fix space.
+There’s an upper bound how many smart appliance one can put on a single ta
+
+\item In list mode, users’ eyes focuses more on the display to traverse between the list while in infrared mode, users’ eyes remain in the physical world more.
+
+\item In our prototype, the IR emitter is nearly fixed.
+Due to different sizes and shapes of the users’ heads, the emitter wouldn’t necessarily point to the center of their eye sights.
+An adjustable emitter would even further improve the performance of our design.
+
+\item The list mode could also be improved by making the swiping gestures more responsive, e.g. swiping speed affects traversing speed.
+
+\item In our experiment, after users locates the nodes by ID, they can instantly figure out their names by the label provided.
+In realistic environments, users have to recall the names of the appliances, which would take a lot more time.
+Furthermore, finding names such as “Living room corner stand lamp” would be more difficult than an alphabet.
+On the other hand, infrared mode will most likely to be unaffected since it does not require name-recalling or text matching.
+
+ \end{itemize}

### 2013-09-18 doc/sigchi14/interaction.tex
updated illustration and home scenario figures
figure optimizations
bjoern's pass

@@
-\begin{figure*}[t!]
+\begin{figure}[t!]
 \centering
-\includegraphics[width=7in]{figures/stepbystep}
-\caption{Step by step walkthrough of interaction goes here.}
+\includegraphics[width=\columnwidth]{figures/stepbystep-edited-bh.png}
+\caption{Targeting interaction: when users turn towards a controllable appliance (A->B), the appliance shows immediate visual feedback (B).
+Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback on the appliance.}
 \label{fig:interaction}
-\end{figure*}
+\end{figure}
 

### 2013-09-18 doc/sigchi14/introduction.tex
update teaser figuire
figure optimizations
bjoern's pass

@@
 
-Today, commercial solutions use handheld mobile devices as {\em universal remote controls} to control such appliances~\bjoern{refs!}.
+Today, commercial solutions use handheld mobile devices as {\em universal remote controls} to control such appliances.
 Commonly, users first browse a list of all available devices, and then call up a device-specific user interface.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/teaser}
-\caption{Teaser figure. Needs a caption.}
+\includegraphics[width=1.0\columnwidth]{figures/teaser.jpg}
+\caption{Using an augmented head-worn device (1), users can control smart home appliances (2) using head orientation targeting.
+A near-eye display then shows an appliance control UI, which users navigate through multitouch gestures.}
 \label{fig:teaser}
@@
 We first report measurements of range and beam characteristics of our controller.
-We then conduct a study that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
+We then conduct a study with $15$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
 We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.
-We also report qualitative results from $n$ participants who use our system for home automation tasks.
+We also report qualitative results from participants who use our system for home automation tasks.

### 2013-09-18 doc/sigchi14/targeting.tex
bjoern's pass

--- 2013-09-16 doc/sigchi14/targeting.tex
@@
 \includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
-\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment.}
+\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment.
+Targets were called out by number; for the list condition, participants had to match numbers to letters.}
 \label{fig:targeting-study-layout}
@@
 Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs (see Figure~\ref{fig:target}).
-An orange LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple clients are within IR range .
+An orange LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple clients are within IR range.
 Next to each client device, a paper sheet shows a number and letter combination, which are used for uniquely identifying the devices.
-\sean{The numbers are the primary identifiers while the letters are in raondom order.
-The nodes are somewhat sorted by their numbers, making it easy to find in order to simulate a place the users are more familiar with.
-Therefore, the time spent on finding the target was minimized and the time measured was mainly the acqusition time.
-}
+The numbers are the primary identifiers, ordered from left to right in the room.
+This ordering makes it easy to locate them, which simulates looking towards an appliance with a well-known location in a room, and minimizes visual search time.
 
@@
 While the list was alphabetized, letter arrangement in the room was not.
-This design required participants to find the target in the room before starting a list navigation to keep conditions similar to each other.
+This design required participants to find the target in the room before starting a list navigation to keep visual search times similar in each condition.
 
@@
 13 had never used Glass before.
-4 wore prescription glasses, which may have affected their task performance \claire{ as the glasses beneath Glass make it more cumbersome to secure the position of the Glass and to adjust the screen to the optimal angle.}
+4 wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
 
 \subsection{Measures}
-The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; and {\bf user preference}: which interface users preferred for the task after completing the study.
+The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; {\bf error rate} and {\bf user preference}: which interface users preferred for the task after completing the study.
 
@@
 
-$t_{infrared}=t_{locate}+t_{reorient}+t_{disambiguate}+t_{tap}$
+$t_{infrared}=t_{locate}[+t_{reorient}][+t_{disambiguate}]+t_{tap}$
 
@@
 
-In both conditions, participants first have to locate the target announced by the experimenter ($t_{locate}$).
+In both conditions, participants first have to locate the target announced by the experimenter through visual search ($t_{locate}$).
 In the {\em infrared} condition, participants may then directly confirm their selection if only a single target was selected ($t_{tap}$).
@@
 In the {\em list} condition, participants must scroll through the list to find the desired target identifier ($t_{listnav}$).
-This analysis makes it clear that any performance benefit depends on whether $t_{reorient}+t_{disambiguate}<t_{listnav}$.
+Thus, {\em infrared} will show a performance benefit if $(t_{reorient}+t_{disambiguate})<t_{listnav}$.
 This depends on the number of total devices in the environment (increasing $t_{listnav}$), and their density (which will increase $t_{disambiguate}$).
-
-At 10 devices, total selection times are comparable:
-\bjoern{Hopefully, when adding distractors, the list navigation will become very cumbersome.}
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
-\caption{Task completion times.} \sean{label mean on the chart}
+\caption{Task completion times.} \ben{label mean, change xtick}
 \label{fig:selection-times}
@@
 
-As a result, infrared mode out performs list mode by 2.186 seconds (6.672 vs 8.858), as shown in figure~\ref{fig:selection-times}. \sean{TBC}
+We first show results for 10 targets and then discuss extrapolations of these results.
+Average target acquisition time $t_{infrared}$ was 6.67 seconds, while $t_{list}$ was 8.86 seconds (Figure~\ref{fig:selection-times} {\em (left)}).
+This difference is significant (Student's t-test, p=0.00017).
+\bjoern{you need to report the t statistic and degrees of freedom as well - t(df)= n,p=m.}
+
+To further understand the performance gain in {\em infrared} condition, especially the factor $t_{disambiguate}$, we compare selection times when multiple devices are targeted (and disambiguation or is required) to single-device selection times (Figure~\ref{fig:selection-times} {\em (right)}).
+When there is a single device, $t_{disambiguate}$ is 0 and it takes 6.40 seconds (on average) to complete the connection.
+When multiple devices are in range, the time increases to 9.16 seconds, indicating 2.76 seconds required to disambiguate.
+Though it takes significantly longer (p=0.012 using t-test) in the {\em multiple} case, these cases made up only 10\% of total infrared trials.
+%, even we intentionally placed a few devices nearer to each other. % 19 / (19+168 )
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
-\caption{Average time taken vs. the order of nodes on the list }
+\caption{Average time taken vs. the order of nodes on the list  } \bjoern{Add horizontal lines at $t_IRsingle$ and $t_IRmultiple$.}
 \label{fig:time-vs-list-order}
 \end{figure}
+
+For each device, $t_{listnav}$ depends on their relative position in the list.
+Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device as a function of its list position - the trend line enables extrapolation to estimate at what number of devices different interaction techniques will outperform list mode.
+\bjoern{A linear trend is visible, and the regression returns with coefficient of determination ($R^2=0.032$).
+--- This R value is basically no correlation at all!
+try to run a correlation only on the median times.
+}
 
@@
 
+Error rates were low in both conditions (\bjoern{enter number of user errors in each condition}), precluding us from running a more detailed analysis.
+
 \subsubsection{Preference}
-\bjoern{update: 6 of 7} users preferred infrared mode over list mode.
-While both interfaces were judged similarly on overall ease of connecting, list navigation received the lowest likert scores (see Figure~\ref{fig:target-likert}).
+Eleven of 15 users preferred infrared mode over list mode (three preferred list, one was undecided).
+While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (see Figure~\ref{fig:target-likert}).
 As self-report data can easily skew positive as participants try to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.
@@
 
-\claire{ One noted benefit of IR mode was a feeling that it was \studyquote{more direct}.
-One subject appreciated the ability to \studyquote{interact with things just by looking at them}.
+One noted benefit of infrared mode was a feeling that it was \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
+One subject called it \studyquote{natural to interact with things just by looking at them}.
 Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}.
 
- A perceived shortcoming of IR mode was the necessity to move the head in order to control a device.
+A few perceived weaknesses of IR mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target.
 One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}.
-Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}.}
-
-\subsubsection{Discussion}
-\sean{to be refined}
-
-\begin{itemize}
-\item During the experiment, a lot of wireless messages were transferred for communication and logging purposes.
-Some messages got lost or delayed and caused the users had to redo that particular action.
-Infrared mode requires a lot of more handshake communications than the list mode.
-This means that in a more robust infrastructure, the performance of infrared mode can improve more than that of the list mode.
-
-\item Some of the infrared performance were affected by the strength and stableness of the infrared emitter.
-By getting a stronger emitter that would be able to reach to a longer distance, connecting to nodes further away such as 4 and 10 would become easier.
-
-\item As shown in figure X, the average time taken to connect to a certain client is correlated to the order of the node within the list.
-We therefore can assume that as the system scales, the average time required in list mode would have a linear growth.
-As for the infrared mode, it is more about the density of nodes in a fix space.
-There’s an upper bound how many smart appliance one can put on a single ta
-
-\item In list mode, users’ eyes focuses more on the display to traverse between the list while in infrared mode, users’ eyes remain in the physical world more.
-
-\item In our prototype, the IR emitter is nearly fixed.
-Due to different sizes and shapes of the users’ heads, the emitter wouldn’t necessarily point to the center of their eye sights.
-An adjustable emitter would even further improve the performance of our design.
-
-\item The list mode could also be improved by making the swiping gestures more responsive, e.g. swiping speed affects traversing speed.
-
-\item In our experiment, after users locates the nodes by ID, they can instantly figure out their names by the label provided.
-In realistic environments, users have to recall the names of the appliances, which would take a lot more time.
-Furthermore, finding names such as “Living room corner stand lamp” would be more difficult than an alphabet.
-On the other hand, infrared mode will most likely to be unaffected since it does not require name-recalling or text matching.
-
- \end{itemize}
+Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}.
+Users had to learn the usable angle of the IR emitter before they became successful at controlling the devices: \studyquote{I had to compensate by tilting my head up a little bit.}
 

### 2013-09-18 doc/sigchi14/abstract.tex
abstract

We introduce a novel method for selecting and controlling smart appliances in physical spaces through a head-worn computing device with near-eye display and wireless communication. We augment a commercial wearable computing device, Google Glass, with a narrow-beam IR emitter for this purpose. This configuration yields a usable beam width of 2 to 4 feet (60 to 120cm) for targeting at room scale. We  describe a disambiguation technique if infrared targeting hits multiple targets simultaneously. A target acquisition study with 15 participants shows that selection using head orientation with our device outperforms list selection on a wearable device. We also report qualitative data from using our device to control multiple appliances in a smart home scenario.

### 2013-09-18 doc/sigchi14/targeting.tex
change participant number to 14 (it is actually 14)

--- 2013-09-17 doc/sigchi14/targeting.tex
@@
 \subsection{Participants}
-We recruited 15 participants from our institution.
+We recruited 14 participants from our institution.
 13 had never used Glass before.
-4 wore prescription glasses, which may have affected their task performance \claire{ as the glasses beneath Glass make it more cumbersome to secure the position of the Glass and to adjust the screen to the optimal angle.}
+4 wore prescription glasses, which may have affected their task performance \claire{ as the glasses beneath Glass make it more cumbersome to secure the position of the Glass and to adjust the screen to the optimal angle.} Half of them performed {\em infrared} mode first and the other half did {/em list} mode first.
 

### 2013-09-18 doc/sigchi14/interaction.tex
update captions and descriptions for figure interaction

--- 2013-09-15 doc/sigchi14/interaction.tex
@@
-\begin{figure*}[t!]
+\begin{figure}[t!]
 \centering
-\includegraphics[width=7in]{figures/stepbystep}
-\caption{Step by step walkthrough of interaction goes here.}
+\includegraphics[width=\columnwidth]{figures/stepbystep.png}
+\caption{(A) Glass periodically sends out infrared signal aligned with the direction user is facing.
+(B) Device toggles red led as visual feedback when receiving valid signals.
+(C) User taps to connect with responding client}
 \label{fig:interaction}
-\end{figure*}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\columnwidth]{figures/stepbystep_multi.png}
+\caption{(A) Multiple clients are within range. (B) All responding clients lights up blue led while the current hovered one blinks. (C) swipe the touchpad to traverse between responding devices. }
+\label{fig:interaction_multi}
+\end{figure}
 
@@
 Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
-Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}A).
+Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}B).
 This enables {\em scanning} the environment with ones gaze to see which devices can be controlled.
@@
 
-{\bf Control:} Glass displays a user interface for parameters of the chosen device (Figure~\ref{fig:interaction}B).
+{\bf Control:} Glass displays a user interface for parameters of the chosen device.
 The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters; single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters.
@@
 In an environment dense with potential targets, multiple targets could be within range.
-Users can tell when multiple feedback LEDs in the environment illuminate.
+Users can tell when multiple feedback LEDs in the environment illuminate (Figure~\ref{fig:interaction_multi}A).
 To disambiguate, users can either move to adjust their head position, or, alternatively, call up a disambiguation dialog on the Glass display.
-The dialog presents a list filtered to only those devices that are within IR range.
-Users navigate the list using the touchpad, and then continue their interaction as described above.
+The dialog presents a list filtered to only those devices that are within IR range, whlie devices also lights up blue led as visual cues.
+Users navigate the list using the touchpad (Figure~\ref{fig:interaction_multi}B), and then continue their interaction as described above.

### 2013-09-18 doc/sigchi14/introduction.tex
update teaser caption
finished pass

@@
 
-In this paper, we introduce a novel method for selecting and controlling smart appliances in physical spaces though the use of a head-worn computing device with near-eye display and wireless communication.
+In this paper, we introduce a novel method for selecting and controlling smart appliances in physical spaces through the use of a head-worn computing device with near-eye display and wireless communication.
 We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose.

### 2013-09-18 doc/sigchi14/interaction.tex

@@
 \includegraphics[width=\columnwidth]{figures/stepbystep.png}
-\caption{(A) Glass periodically sends out infrared signal aligned with the direction user is facing.
-(B) Device toggles red led as visual feedback when receiving valid signals.
-(C) User taps to connect with responding client}
+\caption{Targeting interaction: when users turn towards a controllable appliance (A->B), the appliance shows immediate visual feedback (B).
+Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback on the appliance.}
 \label{fig:interaction}

### 2013-09-18 doc/sigchi14/conclusion.tex

@@
-\section{Conclusion and Future Work}
-Our prototype does not yet have a general way of communicating device descriptions and capabilities - this means our prototype cannot yet control new, unknown devices - they first have to be modeled.
-We may extend the PUC work~\cite{nichols} to address this issue.
-
-More importantly though, the main limitation is that extra hardware for IR communication is needed for the HMD and each controllable device.
-One potential approach would be to combine the growing availability of high-resolution indoor maps with live data from the HMD's camera to determine what a user is looking at without IR data exchange.
-
-\bjoern{add some general summary}
+\section{Conclusion}
+We introduced a novel method for selecting and controlling smart appliances in physical spaces through  infrared targeting based on head orientation.
+While we present a prototype approach that requires that the user carry additional hardware, all parts can readily be miniaturized and integrated into future head-worn hardware.
+We also introduced a disambiguation technique in case head orientation is not sufficient to determine a unique target.
+We characterized our devices performance, arguing that it is matched well to the amount of head movement people can control without strain.
+A target acquisition study showed that the technique is efficient; a home control scenario showed promise but also limitations when trying to control complex appliances.
+As our environment continues to be populated by a swarm of sensing and actuation devices, methods to interrogate and control our smart environments may become increasingly important.
+%\section{Acknowledgments}
+%We thank our user study participants and the a

### 2013-09-18 doc/sigchi14/abstract.tex
update participants to 14

@@
 We  describe a disambiguation technique if infrared targeting hits multiple targets simultaneously.
-A target acquisition study with 15 participants shows that selection using head orientation with our device outperforms list selection on a wearable device.
+A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device.
 We also report qualitative data from using our device to control multiple appliances in a smart home scenario.

### 2013-09-18 doc/sigchi14/interaction.tex
add ui control screenshots

@@
 \subsection{Design Goals}
-Our work is motivated by the following design goals that leverage opportunities of head-worn computing, but also acknowledge potential challenged:
+Our work is motivated by the following design goals that leverage opportunities of head-worn computing, but also acknowledge potential challenges:
 
@@
 
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=0.5\columnwidth]{figures/screenshot_LAMP_CONTROL.png}
+% \caption{ }
+% \label{fig:control_lamp}
+% \end{figure}
+
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=0.5\columnwidth]{figures/screenshot_TV_CONTROLS.png}
+% \caption{( devices. }
+% \label{fig:control_tv}
+% \end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}{.5\columnwidth}
+  \centering
+  \includegraphics[width=0.9\columnwidth]{figures/screenshot_LAMP_CONTROL.png}
+  \caption{Lamp control}
+  \label{fig:control_lamp}
+\end{subfigure}%
+\begin{subfigure}{.5\columnwidth}
+  \centering
+  \includegraphics[width=0.9\columnwidth]{figures/screenshot_TV_CONTROLS.png}
+  \caption{Video player control}
+  \label{fig:control_tv}
+\end{subfigure}
+\caption{Screenshots of appliance controls}
+\label{fig:test}
+\end{figure}
+
 \subsection{Interaction Flow}
@@
 Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
-Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}B).
+Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}A~\ref{fig:interaction}B).
 This enables {\em scanning} the environment with ones gaze to see which devices can be controlled.
 
-{\bf Initiate:} Users confirm their desire to connect to a device by tapping on the Glass touchpad.
+{\bf Initiate:} Users confirm their desire to connect to a device by tapping on the Glass touchpad (Figure~\ref{fig:interaction}C).
 The next section on disambiguation deals with cases in which multiple devices received valid IR signals.
@@
 {\bf Control:} Glass displays a user interface for parameters of the chosen device.
-The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters; single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters.
+The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:control_lamp}); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:control_tv}).
 This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the saggital plane (up and down).

### 2013-09-18 doc/sigchi14/interaction.tex
fixed a few typos

--- 2013-09-18 doc/sigchi14/interaction.tex
@@
-\begin{figure}[t!]
-\centering
-\includegraphics[width=\columnwidth]{figures/stepbystep.png}
-\caption{Targeting interaction: when users turn towards a controllable appliance (A->B), the appliance shows immediate visual feedback (B).
-Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback on the appliance.}
-\label{fig:interaction}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width=\columnwidth]{figures/stepbystep_multi.png}
-\caption{(A) Multiple clients are within range. (B) All responding clients lights up blue led while the current hovered one blinks. (C) swipe the touchpad to traverse between responding devices. }
-\label{fig:interaction_multi}
-\end{figure}
-
 \section{Initiating Interaction through Attention}
@@
 \subsection{Interaction Flow}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\columnwidth]{figures/stepbystep.png}
+\caption{Targeting interaction: when users turn towards a controllable appliance (A$\rightarrow$B), the appliance shows immediate visual feedback (B).
+Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback on the appliance.}
+\label{fig:interaction}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\columnwidth]{figures/stepbystep_multi.png}
+\caption{When multiple clients are within range, they would all have red LEDs illuminated for feedback (A).
+Users call up the disambiguation where all responding clients lights up blue LED while the current hovered one blinks (B).
+Swiping the touchpad is to traverse among responding devices (C).}
+\label{fig:interaction_multi}
+\end{figure}
+
 {\bf Look:} Users select a target device by looking in its general direction.
@@
 Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}B).
-This enables {\em scanning} the environment with ones gaze to see which devices can be controlled.
+This enables {\em scanning} the environment with one's gaze to see which devices can be controlled.
 
@@
 To disambiguate, users can either move to adjust their head position, or, alternatively, call up a disambiguation dialog on the Glass display.
-The dialog presents a list filtered to only those devices that are within IR range, whlie devices also lights up blue led as visual cues.
-Users navigate the list using the touchpad (Figure~\ref{fig:interaction_multi}B), and then continue their interaction as described above.
+The dialog presents a list filtered to only those devices that are within IR range, whlie devices also lights up blue LED as visual cues (Figure~\ref{fig:interaction_multi}B).
+Users navigate the list using the touchpad (Figure~\ref{fig:interaction_multi}C), and then continue their interaction as described above.

### 2013-09-18 doc/sigchi14/targeting.tex
fixed a few typos

--- 2013-09-18 doc/sigchi14/targeting.tex
@@
 \subsection{Apparatus}
-In an indoor environment, 10 wireless nodes are spread across a room at various heights and distances (see Figure~\ref{fig:targeting-study-layout}).
+In an indoor environment, 10 wireless nodes are spread across a room at various heights and distances (Figure~\ref{fig:targeting-study-layout}).
 The nodes are stand-ins for potential smart appliances and have all relevant functionality for targeting and wireless communication, but do not control any actual appliances.
-Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs (see Figure~\ref{fig:target}).
-An orange LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple clients are within IR range.
-Next to each client device, a paper sheet shows a number and letter combination, which are used for uniquely identifying the devices.
+Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs (Figure~\ref{fig:target}).
+An yellow LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple clients are within IR range.
+Next to each client device, a paper sheet shows a number and letter combination, which is used for uniquely identifying the device.
 The numbers are the primary identifiers, ordered from left to right in the room.
@@
 \caption{An example node from the targeting study --- we constructed 10 such nodes - each mounted in a box.}
-\label{fig:targeting-study-layout}
+\label{fig:target}
 \end{figure}
@@
 \subsection{Participants}
-We recruited 15 participants from our institution.
+We recruited 14 participants from our institution.
 13 had never used Glass before.
 4 wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
+Half of them performed {\em infrared} condition first and the other half did {\em list} condition first.
 
@@
 We first show results for 10 targets and then discuss extrapolations of these results.
-Average target acquisition time $t_{infrared}$ was 6.67 seconds, while $t_{list}$ was 8.86 seconds (Figure~\ref{fig:selection-times} {\em (left)}).
-This difference is significant (Student's t-test, p=0.00017).
-\bjoern{you need to report the t statistic and degrees of freedom as well - t(df)= n,p=m.}
+Average target acquisition time $t_{infrared}$ was 6.67 seconds, while $t_{list}$ was 8.86 seconds (Figure~\ref{fig:selection-times}A).
+This difference is significant (Student's t-test, $t(279)=-3.81, p=0.00017$).
+% \bjoern{you need to report the t statistic and degrees of freedom as well - t(df)= n,p=m.}
 
-To further understand the performance gain in {\em infrared} condition, especially the factor $t_{disambiguate}$, we compare selection times when multiple devices are targeted (and disambiguation or is required) to single-device selection times (Figure~\ref{fig:selection-times} {\em (right)}).
+To further understand the performance gain in {\em infrared} condition, especially the factor $t_{disambiguate}$, we compare selection times when multiple devices are targeted (and disambiguation is required) to single-device selection times (Figure~\ref{fig:selection-times}B).
 When there is a single device, $t_{disambiguate}$ is 0 and it takes 6.40 seconds (on average) to complete the connection.
@@
 For each device, $t_{listnav}$ depends on their relative position in the list.
-Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device as a function of its list position - the trend line enables extrapolation to estimate at what number of devices different interaction techniques will outperform list mode.
-\bjoern{A linear trend is visible, and the regression returns with coefficient of determination ($R^2=0.032$).
+Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device as a function of its list position - the trend line enables extrapolation to estimate at what number of devices different interaction techniques will outperform {\em list} condition.
+We see a linear trend of the median values, and the fit returns $R^2=0.952$, indicating a strong linear prediction.
+% \bjoern{A linear trend is visible, and the regression returns with coefficient of determination ($R^2=0.032$).
 --- This R value is basically no correlation at all!
@@
 Eleven of 15 users preferred infrared mode over list mode (three preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (see Figure~\ref{fig:target-likert}).
+While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (Figure~\ref{fig:target-likert}).
 As self-report data can easily skew positive as participants try to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.

### 2013-09-18 doc/sigchi14/interaction.tex
fixed conflict, I prefer to say only figure B

@@
 These design goals find their expression in the following interaction model:
+
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=0.5\columnwidth]{figures/screenshot_LAMP_CONTROL.png}
+% \caption{ }
+% \label{fig:control_lamp}
+% \end{figure}
+
+% \begin{figure}[t!]
+% \centering
+% \includegraphics[width=0.5\columnwidth]{figures/screenshot_TV_CONTROLS.png}
+% \caption{( devices. }
+% \label{fig:control_tv}
+% \end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}{.5\columnwidth}
+  \centering
+  \includegraphics[width=0.9\columnwidth]{figures/screenshot_LAMP_CONTROL.png}
+  \caption{Lamp control}
+  \label{fig:control_lamp}
+\end{subfigure}%
+\begin{subfigure}{.5\columnwidth}
+  \centering
+  \includegraphics[width=0.9\columnwidth]{figures/screenshot_TV_CONTROLS.png}
+  \caption{Video player control}
+  \label{fig:control_tv}
+\end{subfigure}
+\caption{Screenshots of appliance controls}
+\label{fig:test}
+\end{figure}
 
@@
 
-{\bf Initiate:} Users confirm their desire to connect to a device by tapping on the Glass touchpad.
+{\bf Initiate:} Users confirm their desire to connect to a device by tapping on the Glass touchpad (Figure~\ref{fig:interaction}C).
 The next section on disambiguation deals with cases in which multiple devices received valid IR signals.

### 2013-09-18 doc/sigchi14/targeting.tex
more refinements

--- 2013-09-18 doc/sigchi14/targeting.tex
@@
 \subsection{Participants}
-We recruited 15 participants from our institution.
+We recruited 14 participants from our institution.
 13 had never used Glass before.
 4 wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
+Half of them performed {\em infrared} mode first and the other half did {/em list} mode first.
 
@@
 \subsubsection{Preference}
-Eleven of 15 users preferred infrared mode over list mode (three preferred list, one was undecided).
+Eleven of 14 users preferred infrared mode over list mode (three preferred list, one was undecided).
 While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (see Figure~\ref{fig:target-likert}).

### 2013-09-18 doc/sigchi14/introduction.tex

@@
 \caption{Using an augmented head-worn device (1), users can control smart home appliances (2) using head orientation targeting.
-A near-eye display then shows an appliance control UI, which users navigate through multitouch gestures.}
+A near-eye display (3) then shows an appliance control UI, which users navigate through multitouch gestures.}
 \label{fig:teaser}
@@
 Users first look in the direction of the device they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume).
-If multiple devices fall within communication range, an disambiguation technique that combines on-screen information as well as visual feedback on the target devices lets users select their desired target.
+If multiple devices fall within communication range, a disambiguation technique that combines on-screen information as well as visual feedback on the target devices lets users select their desired target.
 Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).
 
-Our hardware relies on infrared communication between Glass and target devices to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.
+Our hardware relies on infrared (IR) communication between Glass and target devices to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.
 Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio.

### 2013-09-18 doc/sigchi14/targeting.tex

--- 2013-09-18 doc/sigchi14/targeting.tex
@@
 \subsubsection{Preference}
-Eleven of 15 users preferred infrared mode over list mode (three preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (Figure~\ref{fig:target-likert}).
+
+Eleven of 14 users preferred infrared mode over list mode (three preferred list, one was undecided).
+While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (see Figure~\ref{fig:target-likert}).
 As self-report data can easily skew positive as participants try to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.

### 2013-09-18 doc/sigchi14/interaction.tex
remove old figure syntax
moved the figure
current status of the appliance is retreived by Glass and shown on the UI
1. update figure and caption 2. note that system retrieves appliance status

@@
 
-These design goals find their expression in the following interaction model:
+These design goals find their expression in the following interaction model.
 
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=0.5\columnwidth]{figures/screenshot_LAMP_CONTROL.png}
-% \caption{ }
-% \label{fig:control_lamp}
-% \end{figure}
+\subsection{Interaction Flow}
 
-% \begin{figure}[t!]
-% \centering
-% \includegraphics[width=0.5\columnwidth]{figures/screenshot_TV_CONTROLS.png}
-% \caption{( devices. }
-% \label{fig:control_tv}
-% \end{figure}
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\columnwidth]{figures/stepbystep.png}
+\caption{Targeting interaction: when users turn towards a controllable appliance (A$\rightarrow$B), the appliance shows immediate visual feedback (B).
+Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback on the appliance.}
+\label{fig:interaction}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=\columnwidth]{figures/stepbystep_multi.png}
+\caption{When multiple clients are within range, they would all have red LEDs illuminated for feedback (A).
+Users call up the disambiguation where all responding clients lights up blue LED while the current hovered one blinks (B).
+Swiping on the touchpad to traverse among responding devices (C).}
+\label{fig:interaction_multi}
+\end{figure}
+
+{\bf Look:} Users select a target device by looking in its general direction.
+Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
+Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}B).
+This enables {\em scanning} the environment with one's gaze to see which devices can be controlled.
+
+{\bf Initiate:} Users confirm their desire to connect to a device by tapping on the Glass touchpad (Figure~\ref{fig:interaction}C).
+The next section on disambiguation deals with cases in which multiple devices received valid IR signals.
+At this point, all further communication switches over to the 802.15.4 wireless network so that line of sight to the target is no longer needed.
+
+{\bf Control:} Glass displays a user interface for parameters of the chosen device.
+Upon connected, the current status of the appliance is retreived by Glass and shown on the UI.
+The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:control_lamp}); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:control_tv}).
+This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the saggital plane (up and down).
+Control commands are sent over XBee radios.
+
+{\bf Disengagement:} Users stay connected to the last selected device up to a timeout period. During that period, users can disengage through down swipes.
 
@@
 
-\subsection{Interaction Flow}
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width=\columnwidth]{figures/stepbystep.png}
-\caption{Targeting interaction: when users turn towards a controllable appliance (A$\rightarrow$B), the appliance shows immediate visual feedback (B).
-Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback on the appliance.}
-\label{fig:interaction}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width=\columnwidth]{figures/stepbystep_multi.png}
-\caption{When multiple clients are within range, they would all have red LEDs illuminated for feedback (A).
-Users call up the disambiguation where all responding clients lights up blue LED while the current hovered one blinks (B).
-Swiping the touchpad is to traverse among responding devices (C).}
-\label{fig:interaction_multi}
-\end{figure}
-
-{\bf Look:} Users select a target device by looking in its general direction.
-Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
-Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}B).
-This enables {\em scanning} the environment with one's gaze to see which devices can be controlled.
-
-{\bf Initiate:} Users confirm their desire to connect to a device by tapping on the Glass touchpad (Figure~\ref{fig:interaction}C).
-The next section on disambiguation deals with cases in which multiple devices received valid IR signals.
-At this point, all further communication switches over to the 802.15.4 wireless network so that line of sight to the target is no longer needed.
-
-{\bf Control:} Glass displays a user interface for parameters of the chosen device.
-The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:control_lamp}); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:control_tv}).
-This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the saggital plane (up and down).
-Control commands are sent over XBee radios.
-
-{\bf Disengagement:} Users stay connected to the last selected device up to a timeout period. During that period, users can disengage through down swipes.
-
 \subsection{Disambiguation}

### 2013-09-18 doc/sigchi14/conclusion.tex
more refinements
refine conclusion

@@
 We introduced a novel method for selecting and controlling smart appliances in physical spaces through infrared targeting based on head orientation.
+Through our solution, we attemp to address the naming and scaling challenges faced by handheld mobile devices.
+The design takes advantage of the fact that visual attention can express intention.
+Along with the visual feedback provided by the target appliances, it helps users remain their focus in the physical world.
 While we present a prototype approach that requires that the user carry additional hardware, all parts can readily be miniaturized and integrated into future head-worn hardware.

### 2013-09-18 doc/sigchi14/targeting.tex
update charts

@@
 \includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
-\caption{Task completion times.} \ben{label mean, change xtick}
+\caption{Boxplot of task completion times for the comparison between {\em infrared} condition and {\em list} condition (A), between IR multiple responses cases and IR single response cases (B).
+The centers of boxes are median values, while white dashed lines are mean values; they are also annotated on the top part.}
 \label{fig:selection-times}
@@
 When multiple devices are in range, the time increases to 9.16 seconds, indicating 2.76 seconds required to disambiguate.
-Though it takes significantly longer (p=0.012 using t-test) in the {\em multiple} case, these cases made up only 10\% of total infrared trials.
+Though it takes significantly longer ($t(19)=-2.7827, p=0.012$ using t-test) in the {\em multiple} case, these cases made up only 10\% of total infrared trials.
 %, even we intentionally placed a few devices nearer to each other. % 19 / (19+168 )
@@
 \includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
-\caption{Average time taken vs. the order of nodes on the list  } \bjoern{Add horizontal lines at $t_IRsingle$ and $t_IRmultiple$.}
+\caption{Time taken to select a device vs. the order in the list.} % \bjoern{Add horizontal lines at $t_{IRsingle}$ and $t_{IRmultiple}$.}
 \label{fig:time-vs-list-order}

### 2013-09-18 data/script/visualize.R

@@ -57,8 +57,15 @@ df <- df[!(grepl("valkerie", df$participant) | grepl("Micheal", df$participant))
 df.IR.single <- df[grepl("IR", df$mode) & grepl("single", df$correct_type), ]
 df.IR.mul <- df[grepl("IR", df$mode) & grepl("mul", df$correct_type), ]
 
+# function for mean labels
+mean.n <- function(x){
+  return(c(y = 3, label = round(median(x),2))) 
+  # experiment with the multiplier to find the perfect position
+}
 
 #### draw the list mode, time vs. position
+## notice this is no longer reproducible for the graph we use in the paper
+## due to my stupid usage of git
 pdf("../../doc/sigchi14/figures/R_List_by_Target.pdf", width=7, height=4)
 df.List <- df[grepl("List", df$mode), ]
 # mapping from id to name
@@ -66,9 +73,12 @@ name <- "GOWFJSYRAM"
 df.List$target <- sapply(df.List$target, function(x) substr(name, x, x))
 label.x <- sapply(names(table(df.List$target)), function(x) regexpr(x, name))
 box <- ggplot(df.List, aes(factor(target), correct_time))
-box <- box + geom_boxplot(lwd=0.5, fill = '#9FB987') + ylim(2.5,15) + xlab("order in list") + ylab("time (seconds)") # + scale_x_discrete(labels=seq(0, 9))
-box <- box + geom_smooth(lwd=1.5, linetype=2, colour = "#477843", aes(group = 1), method="lm", se = F)
-#box <- box + geom_line(data.frame(x = c(0,9), y = c(6.740350,10.85737)), aes(x, y), colour = "red")
+box <- box + geom_boxplot(lwd=0.5, fill = '#9FB987', alpha=0.3) + ylim(2.5,15) + xlab("order in list") + ylab("time (seconds)") + scale_x_discrete(labels=seq(0, 9))
+box <- box + geom_hline(lwd=0.5, fill='black', yintercept=9.16) +
+    geom_hline(lwd=0.5, fill='black', yintercept=6.40) +
+    geom_abline(intercept = 5.04, slope = 0.6, lwd=0.5) +
+##    stat_summary(fun.y = "median", geom = "text", label="---", size= 6, color= "blue") +
[…]

### 2013-09-18 doc/sigchi14/targeting.tex
add error rate

@@
 
+Error rates were low in both conditions (1.1\% in {\em infrared} condition, and 2.9\% in {\em list} condition), precluding us from running a more detailed analysis.
+
 \begin{figure}[t]
@@
 \end{figure}
-
-Error rates were low in both conditions (\bjoern{enter number of user errors in each condition}), precluding us from running a more detailed analysis.
 

### 2013-09-18 doc/sigchi14/targeting.tex
fixing graphs and text descriptions

@@
 \includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
-\caption{Time taken to select a device vs. the order in the list.} % \bjoern{Add horizontal lines at $t_{IRsingle}$ and $t_{IRmultiple}$.}
+\caption{Time taken to select a device vs.~the order in the list.
+The dotted line is a linear fit between the mean times and device orders.
+We also annotate two horizontal lines in {\em infrared} condition to compare different techniques.} % \bjoern{Add horizontal lines at $t_{IRsingle}$ and $t_{IRmultiple}$.}
 \label{fig:time-vs-list-order}
@@
 For each device, $t_{listnav}$ depends on their relative position in the list.
-Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device as a function of its list position - the trend line enables extrapolation to estimate at what number of devices different interaction techniques will outperform {\em list} condition.
-We see a linear trend of the median values, and the fit returns $R^2=0.952$, indicating a strong linear prediction.
+Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices different interaction techniques will outperform {\em list} condition.
+
 % \bjoern{A linear trend is visible, and the regression returns with coefficient of determination ($R^2=0.032$).

### 2013-09-18 data/script/visualize.R

@@ -46,6 +46,48 @@ multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
   }
 }
 
+## Summarizes data.
+## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
+##   data: a data frame.
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   groupvars: a vector containing names of columns that contain grouping variables
+##   na.rm: a boolean that indicates whether to ignore NA's
+##   conf.interval: the percent range of the confidence interval (default is 95%)
+summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE, conf.interval=.95, .drop=TRUE) {
+    require(plyr)
+
+    # New version of length which can handle NA's: if na.rm==T, don't count them
+    length2 <- function (x, na.rm=FALSE) {
+        if (na.rm) sum(!is.na(x))
+        else       length(x)
+    }
+
+    # This is does the summary; it's not easy to understand...
+    datac <- ddply(data, groupvars, .drop=.drop,
+                   .fun= function(xx, col, na.rm) {
+                           c( N    = length2(xx[,col], na.rm=na.rm),
+                              mean = mean   (xx[,col], na.rm=na.rm),
+                              sd   = sd     (xx[,col], na.rm=na.rm)
+                              )
+                          },
+                    measurevar,
+                    na.rm
+             )
+
+    # Rename the "mean" column    
+    datac <- rename(datac, c("mean"=measurevar))
+
+    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
+
+    # Confidence interval multiplier for standard error
+    # Calculate t-statistic for confidence interval: 
+    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
+    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
+    datac$ci <- datac$se * ciMult
+
+    return(datac)
+}
+
 
[…]

### 2013-09-18 doc/sigchi14/introduction.tex
figure cpation

@@
 \includegraphics[width=1.0\columnwidth]{figures/teaser.jpg}
-\caption{Using an augmented head-worn device (1), users can control smart home appliances (2) using head orientation targeting.
-A near-eye display (3) then shows an appliance control UI, which users navigate through multitouch gestures.}
+\caption{Using an augmented head-worn device (1), users can control smart home appliances (2) with head orientation targeting.
+A near-eye display then shows an appliance control UI (3), which users navigate through multitouch gestures.}
 \label{fig:teaser}

### 2013-09-18 doc/sigchi14/introduction.tex
take out 'see'

@@
 If multiple devices fall within communication range, a disambiguation technique that combines on-screen information as well as visual feedback on the target devices lets users select their desired target.
-Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).
+Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}).
 

### 2013-09-18 doc/sigchi14/interaction.tex
figure caption

@@
 \includegraphics[width=\columnwidth]{figures/stepbystep.png}
-\caption{Targeting interaction: when users turn towards a controllable appliance (A$\rightarrow$B), the appliance shows immediate visual feedback (B).
-Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback on the appliance.}
+\caption{Targeting interaction: when users turn towards a controllable appliance (A$\rightarrow$B), the appliance shows immediate visual feedback (red LED) (B).
+Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback (blue LED) on the appliance.}
 \label{fig:interaction}

### 2013-09-18 doc/sigchi14/relatedwork.tex
one disadvantage for laser pointer
fixed url type

@@
 These controllers tend to use wide-angle infrared LEDs.
-Universal remote controls are available as dedicated devices (e.g, Crestron\footnote{\url{http://www.crestron.com}}) or applications for smart phones (e.g., Belkin WeMo\footnote{http://www.belkin.com/us/wemo}).
+Universal remote controls are available as dedicated devices (e.g, Crestron\footnote{\url{http://www.crestron.com}}) or applications for smart phones (e.g., Belkin WeMo\footnote{\url{http://www.belkin.com/us/wemo}}).
 They do not offer spatial selection of target devices, forcing users to browse through lists of pre-configured devices instead.
@@
 The laser dot provides immediate visual feedback to the user what is being selected (though it does not indicate whether the pointed-at object can indeed be controlled).
+Furthermore, laser pointer becomes obstrusive when there are other people in the space.
 Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}.

### 2013-09-18 doc/sigchi14/targeting.tex
changes to targeting.tex

@@
 \caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment.
-Targets were called out by number; for the list condition, participants had to match numbers to letters.}
+Targets were called out by number; for the list mode condition, participants need to match numbers to letters.}
 \label{fig:targeting-study-layout}
@@
 In our within-subjects design, participants performed 15 target acquisition tasks each with two interaction styles.
-In the {\em infrared} condition, participants used our IR targeting approach; in the {\em list} condition, participants had to look up a device's letter code on the printed paper next to the device and then select that letter code from a list displayed on their Glass device.
+In the {\em infrared} mode condition, participants used our IR targeting approach; in the {\em list} mode condition, participants had to look up a device's letter code on the printed paper next to the device and then select that letter code from a list displayed on their Glass device.
 The list was navigated with swipe motions on the Glass touchpad.
@@
 Participants then had to find the corresponding device (by looking for its printed code).
-In the {\em infrared} condition, participants then selected and acquired the target by aiming the infrared beam at the target, and confirmed their selection with a touch pad tap.
+In the {\em infrared} mode, participants then selected and acquired the target by aiming the IR beam at the target, and confirmed their selection with a touch pad tap.
 If more than one target was within range, participants had to either use the disambiguation dialog or reposition themselves.
-In the {\em list} condition, participants had to read the letter next to the number and then select that letter by browsing a linear list shown in their Glass display.
+In the {\em list} mode, participants had to read the letter next to the number and then select that letter by browsing a linear list shown in their Glass display.
 While the list was alphabetized, letter arrangement in the room was not.
@@
 4 wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
-Half of them performed {\em infrared} condition first and the other half did {\em list} condition first.
+Half of them performed {\em infrared} mode first and the other half did {\em list} mode first.
 
 \subsection{Measures}
-The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; {\bf error rate} and {\bf user preference}: which interface users preferred for the task after completing the study.
+The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; and {\bf user preference}: which interface users preferred for the task after completing the study.
 
@@
 In both conditions, participants first have to locate the target announced by the experimenter through visual search ($t_{locate}$).
-In the {\em infrared} condition, participants may then directly confirm their selection if only a single target was selected ($t_{tap}$).
+In the {\em infrared} mode, participants may then directly confirm their selection if only a single target was selected ($t_{tap}$).
 However, if they don't immediately receive feedback that their target was selected, or if multiple targets were selected, users either have change their position or head orientation ($t_{reorient}$) or they have to step through the on-screen disambiguation dialog ($t_{disambiguate}$).
-In the {\em list} condition, participants must scroll through the list to find the desired target identifier ($t_{listnav}$).
+In the {\em list} mode, participants must scroll through the list to find the desired target identifier ($t_{listnav}$).
 Thus, {\em infrared} will show a performance benefit if $(t_{reorient}+t_{disambiguate})<t_{listnav}$.
@@
 \includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
-\caption{Boxplot of task completion times for the comparison between {\em infrared} condition and {\em list} condition (A), between IR multiple responses cases and IR single response cases (B).
+\caption{Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), between IR multiple responses cases and IR single response cases (B).
 The centers of boxes are median values, while white dashed lines are mean values; they are also annotated on the top part.}
@@
 
-To further understand the performance gain in {\em infrared} condition, especially the factor $t_{disambiguate}$, we compare selection times when multiple devices are targeted (and disambiguation is required) to single-device selection times (Figure~\ref{fig:selection-times}B).
+To further understand the performance gain in {\em infrared} mode, especially the factor $t_{disambiguate}$, we compare selection times when multiple devices are targeted (and disambiguation is required) to single-device selection times (Figure~\ref{fig:selection-times}B).
 When there is a single device, $t_{disambiguate}$ is 0 and it takes 6.40 seconds (on average) to complete the connection.
@@
 \includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
-\caption{Time taken to select a device vs.~the order in the list.
+\caption{Times taken to select a device vs.~the order in the list.
 The dotted line is a linear fit between the mean times and device orders.
-We also annotate two horizontal lines in {\em infrared} condition to compare different techniques.} % \bjoern{Add horizontal lines at $t_{IRsingle}$ and $t_{IRmultiple}$.}
+Two horizontal lines of mean target acquisition times in {\em infrared} mode are also annotated for comparison.}
+ % \bjoern{Add horizontal lines at $t_{IRsingle}$ and $t_{IRmultiple}$.}
 \label{fig:time-vs-list-order}
@@
 For each device, $t_{listnav}$ depends on their relative position in the list.
-Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices different interaction techniques will outperform {\em list} condition.
+Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices the {\em infrared} mode interaction techniques will outperform {\em list} mode\footnote{The higher mean value at $order=1$ is caused by one outlier when the participant failed multiple times to connect to the right target}.
+From the figure, we can see that once the target's order has increased to be larger than 6, the average $t_{listnav}$ would be larger than $t_{reorient} + t_{disambiguate}$.
 
@@
 
-Error rates were low in both conditions (1.1\% in {\em infrared} condition, and 2.9\% in {\em list} condition), precluding us from running a more detailed analysis.
+Participants' selection errors do occur in both conditions.
+However, error rates were low (1.1\% in {\em infrared} mode, and 2.9\% in {\em list} mode, repspectively).
+This precludes us from running a more detailed analysis.
 

### 2013-09-18 doc/sigchi14/interaction.tex
smaller pngs so latex doesn't complain
changing 'device' to 'appliance, update ui control figures, unuse subcaption package

@@
 
-{\bf Offer efficient ways to disambiguate orientation input:} It may not always be possible to identify a unique target device based on a user's attention and orientation.
+{\bf Offer efficient ways to disambiguate orientation input:} It may not always be possible to identify a unique target appliance based on a user's attention and orientation.
 Offer ways to supplement orientation-based interaction with screen-based interaction to provide disambiguation information.
@@
 Users call up the disambiguation where all responding clients lights up blue LED while the current hovered one blinks (B).
-Swiping on the touchpad to traverse among responding devices (C).}
+Swiping on the touchpad to traverse among responding appliances (C).}
 \label{fig:interaction_multi}
@@
 
-{\bf Look:} Users select a target device by looking in its general direction.
+{\bf Look:} Users select a target appliance by looking in its general direction.
 Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
 Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}B).
-This enables {\em scanning} the environment with one's gaze to see which devices can be controlled.
+This enables {\em scanning} the environment with one's gaze to see which appliances can be controlled.
 
-{\bf Initiate:} Users confirm their desire to connect to a device by tapping on the Glass touchpad (Figure~\ref{fig:interaction}C).
-The next section on disambiguation deals with cases in which multiple devices received valid IR signals.
+{\bf Initiate:} Users confirm their desire to connect to an appliance by tapping on the Glass touchpad.
+After connected, target appliance toggles on blue LED as visual feedback (Figure~\ref{fig:interaction}C).
+The next section on disambiguation deals with cases in which multiple appliances received valid IR signals.
 At this point, all further communication switches over to the 802.15.4 wireless network so that line of sight to the target is no longer needed.
 
-{\bf Control:} Glass displays a user interface for parameters of the chosen device.
+{\bf Control:} Glass displays a user interface for parameters of the chosen appliance.
 Upon connected, the current status of the appliance is retreived by Glass and shown on the UI.
-The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:control_lamp}); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:control_tv}).
+The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:ui_controls}(A)); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:ui_controls}(B)).
 This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the saggital plane (up and down).
@@
 
-{\bf Disengagement:} Users stay connected to the last selected device up to a timeout period. During that period, users can disengage through down swipes.
+{\bf Disengagement:} Users stay connected to the last selected appliance up to a timeout period. During that period, users can disengage through down swipes.
 
-\begin{figure}
+\begin{figure}[t!]
 \centering
-\begin{subfigure}{.5\columnwidth}
-  \centering
-  \includegraphics[width=0.9\columnwidth]{figures/screenshot_LAMP_CONTROL.png}
-  \caption{Lamp control}
-  \label{fig:control_lamp}
-\end{subfigure}%
-\begin{subfigure}{.5\columnwidth}
-  \centering
-  \includegraphics[width=0.9\columnwidth]{figures/screenshot_TV_CONTROLS.png}
-  \caption{Video player control}
-  \label{fig:control_tv}
-\end{subfigure}
+\includegraphics[width=\columnwidth]{figures/ui_controls_caption.jpg}
 \caption{Screenshots of appliance controls}
-\label{fig:test}
+\label{fig:ui_controls}
 \end{figure}
@@
 To disambiguate, users can either move to adjust their head position, or, alternatively, call up a disambiguation dialog on the Glass display.
-The dialog presents a list filtered to only those devices that are within IR range, whlie devices also lights up blue LED as visual cues (Figure~\ref{fig:interaction_multi}B).
+The dialog presents a list filtered to only those appliances that are within IR range, whlie appliances also use blue LED as visual cues: all responding appliances light up LEDs while the current hovered one blinks (Figure~\ref{fig:interaction_multi}B).
 Users navigate the list using the touchpad (Figure~\ref{fig:interaction_multi}C), and then continue their interaction as described above.

### 2013-09-18 doc/sigchi14/introduction.tex
changing 'device' to 'appliance, update ui control figures, unuse subcaption package

@@
 We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose.
-Users first look in the direction of the device they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume).
-If multiple devices fall within communication range, a disambiguation technique that combines on-screen information as well as visual feedback on the target devices lets users select their desired target.
-Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}).
+Users first look in the direction of the appliance they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume).
+If multiple appliances fall within communication range, a disambiguation technique that combines on-screen information as well as visual feedback on the target appliances lets users select their desired target.
+Once acquired, an appliance specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}).
 
-Our hardware relies on infrared (IR) communication between Glass and target devices to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.
+Our hardware relies on infrared (IR) communication between Glass and target appliances to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.
 Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio.
-Target devices similarly have IR receivers and radios.
-This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target device while issuing control commands.
+Target appliances similarly have IR receivers and radios.
+This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target appliances while issuing control commands.
 

### 2013-09-18 doc/sigchi14/introduction.tex
typoes and fixes from Claire's email

--- 2013-09-18 doc/sigchi14/introduction.tex
@@
 
-To address these challenges, research has introduced techniques of augmenting mobile devices, e.g., with laser pointers, to enable direct aiming at target devices~\cite{beigl_point_1999,patel_2-way_2003}.
+To address these challenges, research has introduced techniques of augmenting mobile devices with accessories like laser pointers to enable direct aiming at target devices~\cite{beigl_point_1999,patel_2-way_2003}.
 While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at the device to-be-controlled.
@@
 We first report measurements of range and beam characteristics of our controller.
-We then conduct a study with $15$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
+We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
 We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.

### 2013-09-18 doc/sigchi14/relatedwork.tex

@@
 Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing --- the user looks at the target device to initiate interaction.
-Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications, we therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.
+Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications.
+We therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.
 
@@
 \subsection{Augmented Reality Interfaces}
-Agumented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
+Augmented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
 Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world, though our techniques could be extended to such configurations.

### 2013-09-18 doc/sigchi14/targeting.tex

@@
 Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs (Figure~\ref{fig:target}).
-An yellow LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple clients are within IR range.
-Next to each client device, a paper sheet shows a number and letter combination, which is used for uniquely identifying the device.
+An yellow LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple targets are within IR range.
+Next to each target, a paper sheet shows a number and letter combination, which is used for uniquely identifying the device.
 The numbers are the primary identifiers, ordered from left to right in the room.

### 2013-09-18 doc/sigchi14/targeting.tex
add a new sentense

@@
 Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices the {\em infrared} mode interaction techniques will outperform {\em list} mode\footnote{The higher mean value at $order=1$ is caused by one outlier when the participant failed multiple times to connect to the right target}.
-From the figure, we can see that once the target's order has increased to be larger than 6, the average $t_{listnav}$ would be larger than $t_{reorient} + t_{disambiguate}$.
+From the figure, we can see that once the target's order has increased to be larger than 6, the average $t_{listnav}$ for that target would be larger than $t_{reorient} + t_{disambiguate}$.
+We expect that, when the number of targets keeps increasing, there would be larger time reduction in {\em infrared} mode.
 

### 2013-09-18 doc/sigchi14/targeting.tex
update footnote

@@
 For each device, $t_{listnav}$ depends on their relative position in the list.
-Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices the {\em infrared} mode interaction techniques will outperform {\em list} mode\footnote{The higher mean value at $order=1$ is caused by one outlier when the participant failed multiple times to connect to the right target}.
+Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices the {\em infrared} mode interaction techniques will outperform {\em list} mode\footnote{The higher mean value at $order=1$ is caused by one outlier when the participant tried multiple times in {\em list} mode to connect to the right target.}.
 From the figure, we can see that once the target's order has increased to be larger than 6, the average $t_{listnav}$ for that target would be larger than $t_{reorient} + t_{disambiguate}$.

### 2013-09-18 doc/sigchi14/conclusion.tex
refinments

--- 2013-09-18 doc/sigchi14/conclusion.tex
@@
 We introduced a novel method for selecting and controlling smart appliances in physical spaces through infrared targeting based on head orientation.
+Through our solution, we attempt to address the naming and scaling challenges faced by handheld mobile devices.
+The design takes advantage of the fact that visual attention can express intention.
+The visual feedback provided by the target appliances helps users keep their focus in the physical world.
 While we present a prototype approach that requires that the user carry additional hardware, all parts can readily be miniaturized and integrated into future head-worn hardware.
@@
 A target acquisition study showed that the technique is efficient; a home control scenario showed promise but also limitations when trying to control complex appliances.
-As our environment continues to be populated by a swarm of sensing and actuation devices, methods to interrogate and control our smart environments may become increasingly important.
+As our environment continues to be populated by a swarm of sensing and actuation devices, methods to interrogate and control our smart environments will become increasingly important.
 %\section{Acknowledgments}

### 2013-09-18 doc/sigchi14/interaction.tex
refinments

@@
 \section{Initiating Interaction through Attention}
-This section describes the design goals of our systems and their realization in particular interaction techniques.
+This section describes the design goals of our system and their realization in particular interaction techniques.
 
@@
 
-{\bf Offer flexible orientation after initiating interaction:} After initiating attention through head orientation, enable the user to reorient their head or body position during the remaining interaction to prevent neck strain.
+{\bf Offer flexible orientation after initiating interaction:} After initiating interaction through head orientation, enable the user to reorient their head or body position during the remaining interaction to prevent neck strain.
 
@@
 \centering
-\includegraphics[width=\columnwidth]{figures/stepbystep.png}
+\includegraphics[width=\columnwidth]{figures/stepbystep_small.png}
 \caption{Targeting interaction: when users turn towards a controllable appliance (A$\rightarrow$B), the appliance shows immediate visual feedback (red LED) (B).
@@
 \centering
-\includegraphics[width=\columnwidth]{figures/stepbystep_multi.png}
-\caption{When multiple clients are within range, they would all have red LEDs illuminated for feedback (A).
-Users call up the disambiguation where all responding clients lights up blue LED while the current hovered one blinks (B).
-Swiping on the touchpad to traverse among responding appliances (C).}
+\includegraphics[width=\columnwidth]{figures/stepbystep_multi_small.png}
+\caption{When multiple appliances are within range, they all have red LEDs illuminated for feedback (A).
+When users initiate connections, all target appliances toggle on blue LEDs while the currently selected one blinks (B).
+Swiping on the touchpad traverses among responding appliances (C).}
 \label{fig:interaction_multi}
@@
 {\bf Initiate:} Users confirm their desire to connect to an appliance by tapping on the Glass touchpad.
-After connected, target appliance toggles on blue LED as visual feedback (Figure~\ref{fig:interaction}C).
+After they are connected, the target appliance toggles on a blue LED as visual feedback (Figure~\ref{fig:interaction}C).
 The next section on disambiguation deals with cases in which multiple appliances received valid IR signals.
@@
 {\bf Control:} Glass displays a user interface for parameters of the chosen appliance.
-Upon connected, the current status of the appliance is retreived by Glass and shown on the UI.
+Upon connection, the current status of the appliance is retreived by Glass and shown on the UI.
 The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:ui_controls}(A)); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:ui_controls}(B)).
@@
 \includegraphics[width=\columnwidth]{figures/ui_controls_caption.jpg}
-\caption{Screenshots of appliance controls}
+\caption{Two screenshots of UI controls for lamp and video player.}
 \label{fig:ui_controls}
@@
 To disambiguate, users can either move to adjust their head position, or, alternatively, call up a disambiguation dialog on the Glass display.
-The dialog presents a list filtered to only those appliances that are within IR range, whlie appliances also use blue LED as visual cues: all responding appliances light up LEDs while the current hovered one blinks (Figure~\ref{fig:interaction_multi}B).
+The dialog presents a list filtered to only those appliances that are within IR range, while appliances also use blue LED as visual cues: all responding appliances light up LEDs while the currently selected one blinks (Figure~\ref{fig:interaction_multi}B).
 Users navigate the list using the touchpad (Figure~\ref{fig:interaction_multi}C), and then continue their interaction as described above.

### 2013-09-18 doc/sigchi14/targeting.tex
refinments

@@
 \includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
-\caption{Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), between IR multiple responses cases and IR single response cases (B).
-The centers of boxes are median values, while white dashed lines are mean values; they are also annotated on the top part.}
+\caption{Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), and between IR multiple responses cases and IR single response cases (B).
+The centers of boxes are median values, while white dashed lines are mean values.}
 \label{fig:selection-times}
@@
 \includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
-\caption{Times taken to select a device vs.~the order in the list.
+\caption{Times taken to select a device vs.~its order in the list.
 The dotted line is a linear fit between the mean times and device orders.
@@
 
-A few perceived weaknesses of IR mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target.
+A few perceived weaknesses of infrared mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target.
 One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}.

### 2013-09-18 doc/sigchi14/device.tex
refinments
typos, adding a example

@@
 This permits wearers to move their head freely after connecting to an appliance.
-In the research prototype, users have to carry an additional microcontroller board that marshalls messages between Glass' Bluetooth radio an IR/802.15.4, but our custom hardware could also be integrated into the wearable device.}
+In the research prototype, users have to carry an additional microcontroller board that marshals messages between Glass' Bluetooth radio an IR/802.15.4, but our custom hardware could also be integrated into the wearable device.}
 \label{fig:architecture}
@@
 \includegraphics[width=1.0\columnwidth]{figures/glass-ir-coverage}
-\caption{Different IR configurations suggest that usable beam widths of 2 to 4 feet and distances up to 16 feet }
+\caption{Different IR configurations suggest usable beam widths of 2 to 4 feet and distances up to 16 feet }
 \label{fig:coverage}

### 2013-09-18 doc/sigchi14/interaction.tex
typos, adding a example

@@
 The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:ui_controls}(A)); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:ui_controls}(B)).
-This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the saggital plane (up and down).
+This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the sagittal plane (up and down).
 Control commands are sent over XBee radios.

### 2013-09-20 doc/sigchi14/sigchi14.tex
bjoern's pass at overall narrative
more text and figures
various improvements
add ui control screenshots
more refinements
update title
copyright section
changing 'device' to 'appliance, update ui control figures, unuse subcaption package
added names

@@
 
+\toappear{Submitted to CHI2014. Do not cite or distribute without written permission from the authors.}
+
 % Arabic page numbers for submission. 
@@
 \usepackage{gensymb}
+\usepackage{afterpage}
 % see http://tex.stackexchange.com/questions/46055/typesetting-with-inch-symbols-and-sizes-in-inches
@@
 \newcommand {\claire}[1]{{\color{blue}\bf{CT: #1}\normalfont}}
-
+\newcommand {\studyquote}[1]{\em ``#1''\normalfont}
 %comment out the above and uncomment these for final submit
@@
 
-\title{Selecting and Interacting with Smart Devices Through a Head-Mounted Display with Infrared Targeting}
+\title{A Context Menu for the Real World: Controlling Physical Appliances Through Head-Worn Infrared Targeting}
 
@@
 \author{
-  \alignauthor Anonymous for submission
+\alignauthor Yu-Hsiang Chen$^{\dagger}$, Ben Zhang$^{\dagger}$, Claire Tuna$^{\dagger}$, Yang Li$^{\ddagger}$, Edward Lee$^{\dagger}$, Bj\"orn Hartmann$^{\dagger}$ \\
+\affaddr{$\dagger$: UC Berkeley EECS \& CITRIS Invention Lab \hspace{0.25in} $\ddagger$: Google Research}\\
+\email{sean.yhc@gmail.com, benzh@eecs.berkeley.edy, clairetuna@gmail.com, yangli@acm.org, eal@eecs.berkeley.edu, bjoern@eecs.berkeley.edu}
+}
+%\author{
+% \alignauthor Anonymous for submission
 %  \alignauthor 3rd Author Name\\
@@
 %    \affaddr{Optional phone number}
-}
 
@@
 \begin{abstract}
-Insert abstract
+\input{abstract}
 \end{abstract}
@@
 \input{relatedwork}
+\input{interaction}
 \input{device}
@@
 \input{control}
+\input{discussion}
 \input{conclusion}
+\section{Acknowledgments}
+This work was partially supported by a Sloan Foundation Fellowship.
 

### 2013-10-23 doc/sigchi14/relatedwork.tex
typos
acks, related work

@@
 \section{Related Work}
+\bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
 Relevant prior work exists in the areas of remote control of physical appliances, evaluations of pointing in physical space and augmented reality applications. We discuss each in turn.
@@
 The laser dot provides immediate visual feedback to the user what is being selected (though it does not indicate whether the pointed-at object can indeed be controlled).
-Furthermore, laser pointer becomes obstrusive when there are other people in the space.
+Furthermore, laser pointer becomes obtrusive when there are other people in the space.
 Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}.

### 2013-10-23 doc/sigchi14/sigchi14.tex
acks, related work

@@
 \section{Acknowledgments}
-This work was partially supported by a Sloan Foundation Fellowship.
+This work was supported in part by the TerraSwarm Research Center, one of six centers supported by the STARnet phase of the Focus Center Research Program (FCRP) a Semiconductor Research Corporation program sponsored by MARCO and DARPA.
+Additional support was provided by a Sloan Foundation Fellowship and a Google Research Award.
 

### 2013-12-10 doc/sigchi14/sigchi14.tex
wrong email address for Ben got fixed

@@
 %% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
-% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is 	granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
+%% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
 Copyrights for components of this work owned by others than ACM must be honored.
@@
 To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee.
-Request permissions from permissions@acm.org.
-\\
+Request permissions from permissions@acm.org.}
 % {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
@@
 
-\toappear{Submitted to CHI2014. Do not cite or distribute without written permission from the authors.}
+\toappear{}
+%Submitted to CHI2014. Do not cite or distribute without written permission from the authors.}
 
@@
 \affaddr{$\dagger$: UC Berkeley EECS \& CITRIS Invention Lab \hspace{0.25in} $\ddagger$: Google Research}\\
-\email{sean.yhc@gmail.com, benzh@eecs.berkeley.edy, clairetuna@gmail.com, yangli@acm.org, eal@eecs.berkeley.edu, bjoern@eecs.berkeley.edu}
+\email{sean.yhc@gmail.com, benzh@eecs.berkeley.edu, clairetuna@gmail.com, yangli@acm.org, eal@eecs.berkeley.edu, bjoern@eecs.berkeley.edu}
 }

### 2014-02-24 doc/uist14/control.tex
create folder uist14 and copy content from sigchi14

--- 2013-09-18 doc/sigchi14/control.tex
@@
 For this study, we only elicited subjective data in the form of Likert data and open-ended responses.
-\begin{figure}[b]
+\begin{figure}[t]
 \centering
@@
 All participants successfully completed the list of tasks.
-Participants stated it was easy to interact with the appliances, but there were distinct differences between appliances.
+They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous study procedure.
+Participants saw benefits of the device for families --- \studyquote{ It might also be useful for people who need to take care of small children that they can complete all the tasks while keeping an eye on their children at the same time}, though settings that require more movement than watching a movie at home (e.g., cooking) may be more appropriate scenarios for wearing the device.
+
+\subsubsection{Simple and Complex Controls}
+Participants rated the ease of control of particular appliances differently.
 Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for the more complex movie player (see Figure~\ref{fig:smarthome-likert}).
 Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
-The screen size and 1D input put a limit on the complexity of interfaces that can be presented.
+The screen size and (largely) 1D input put a limit on the complexity of interfaces that can be presented.
+As one participant remarked: \studyquote{[The media player] does not seem to be more efficient than a tablet device.} The difficulty can also partly be ascribed to our interaction design, which required one finger swipes to switch between parameters and two finger gestures for adjusting parameters --- it was hard for users to exert fine control over two-finger swipes.
+In addition, users did not always remember these mappings as they are not yet part of a standard gesture vocabulary.
 
-\subsubsection{Visual overlap between Near-Eye Display and Target Appliance}
-When a participant looks at a target, the near-eye display may occlude or overlap the target appliance.
+\subsubsection{Eliminating Steps}
+Participants liked the efficiency of our design but also suggested further simplification by eliminating the explicit connection step, they wanted to immediately connect to any appliance that receives the infrared signal --- \studyquote{I intuitively want the screen to automatically appear when the IR detects the appliance rather than having to tap to connect.} Such a design would increase the efficiency of interaction, but at a power tradeoff, as the wireless radio will have to send and receive data each time the user looks at a device --- whether on purpose or inadvertently.
+We leave the study of battery life implications of interaction design choices to future work.
+
+\subsubsection{Feedback On Screen or In The World}
+We found out that the near-eye display may occlude or overlap the target appliance when a participant looks at a target.
 This may make it difficult to read either the on-screen display or see information displayed on the target device.
-As one participant remarked \studyquote{This was especially annoying with the TV because there were two screens overlapping each other.} While it is possible to look away once a device has been acquired to better see the Glass display, this tension is likely to remain for users who want to rely on feedback from the devices themselves instead of the Glass display.
+As one participant remarked \studyquote{This was especially annoying with the TV because there were two screens overlapping each other.} While it is possible to look away once a device has been acquired to better see the Glass display, a tension remains whether users should rely on feedback from the appliances themselves or on the near-eye display.
 

### 2014-02-24 doc/uist14/discussion.tex
create folder uist14 and copy content from sigchi14

\section{Discussion}
Our study procedures demonstrated that users can successfully select and control smart appliances with head-worn infrared targeting, and that this technique outperforms list selection on the Google Glass wearable device. In this section, we revisit some of the results and observations and discuss their larger significance and potential paths for future work.

\paragraph{Meaningful Results?}
While the performance increase in our targeting study is statistically significant, readers may wonder whether it is truly meaningful. We believe it is, for two reasons: first, our technique avoids the problems of {\em naming} and {\em scoping} inherent in any interface that uses representations of objects (e.g., a list of identifiers) rather than the objects themselves. We argue that this disintermediation of interaction leads to a cognitively simpler design. Second, our existing study only showed results for a modest number of targets. Our technique should have a wider margin as the number of targets increases. Of course, the number of targets one can realistically expect may vary across application domains.

\paragraph{Hands-Free Operation}
While head movement is not as precise as hand positioning (e.g., in Patel's mobile laser pointing~\cite{patel_2-way_2003}), one key benefit of our target acquisition step is that it does not require the user's hands. This raises the question if the rest of the interaction (disambiguation and device control) could also be achieved in a hands-free fashion. Voice-command control is an obvious candidate, though such approaches have not found widespread adoption because of social acceptability and other factors.

\paragraph{Hardware Limitations}
In our prototype, the IR emitter is fixed in a single position. Due to different sizes and shapes of the users’ heads, the emitter may not line up exactly with their head orientation. An adjustable emitter (paired with a suitable calibration routing) would improve the performance of our design. Some of our users also mentioned that they had to move closer to some targets to successfully select them. A stronger emitter could overcome these problems, but care has to be taken to avoid possible reflection problems where IR light bounces off a wall and hits an unintended target behind the user.

More importantly though, the main limitation of our design is that extra hardware for infrared communication is needed for the head-mounted device and each controllable appliance. One potential approach to sidestep this requirement would be to combine the growing availability of high-resolution indoor maps with live data from the point-of-view camera on the device to determine what a user is looking at without any infrared data exchange.

\paragraph{The Midas Look}
Our participants suggested eliminating explicit initiation of a connection by the user. However, one of the design guidelines for near-eye displays is to avoid pushing information to the display without an initial request from the user --- flashing device information on screen each time a user moves their head would surely be counterproductive. Future work should investigate how to intelligently decide when and how to initiate interaction for the user.

\paragraph{Where is the Target?}
One open design question of our approach is where infrared receivers should be placed. For a light, one might put a received on the light itself, or on the light switch, to cater to existing expectations. For volume control, the infrared receiver might be located on a speaker or on the amplifier. A thorough study of user preferences would be interesting; though we also point out that our architecture could easily support multiple receivers that end up controlling the same appliance.

%Our prototype does not yet have a general way of communicating device descriptions and capabilities - this means our prototype cannot yet control new, unknown devices - they first have to be modeled. We may extend the PUC work~\cite{nichols} to address this issue.

%bjoern: these below are details about what happened in the experiment; they are less relevant to the research contributions of the paper.
%\item During the experiment, a lot of wireless messages were transferred for communication and logging purposes. Some messages got lost or delayed and caused the users had to redo that particular action. Infrared mode requires a lot of more handshake communications than the list mode. This means that in a more robust infrastructure, the performance of infrared mode can improve more than that of the list mode.

%\item In list mode, users’ eyes focuses more on the display to traverse between the list while in infrared mode, users’ eyes remain in the physical world more.

%\item The list mode could also be improved by making the swiping gestures more responsive, e.g. swiping speed affects traversing speed. % too detailed - partially covered later

%\item In our experiment, after users locates the nodes by ID, they can instantly figure out their names by the label provided. In realistic environments, users have to recall the names of the appliances, which would take a lot more time. Furthermore, finding names such as “Living room corner stand lamp” would be more difficult than an alphabet. On the other hand, infrared mode will most likely to be unaffected since it does not require name-recalling or text matching. %too detailed

### 2014-02-24 doc/uist14/introduction.tex
create folder uist14 and copy content from sigchi14

--- 2013-09-18 doc/sigchi14/introduction.tex
@@
 
-Today, commercial solutions use handheld mobile devices as {\em universal remote controls} to control such appliances.
-Commonly, users first browse a list of all available devices, and then call up a device-specific user interface.
+Today, commercial solutions (such as Belkin WeMo) use handheld mobile devices as {\em universal remote controls} to control such appliances.
+In these solutions, users first browse a list of all available devices and then call up a device-specific user interface.
 This method faces two challenges: {\em naming} and {\em scoping}.

### 2014-02-24 doc/uist14/submitted-for-review/submitted.tex
create folder uist14 and copy content from sigchi14

\documentclass{sigchi}

\toappear{Submitted to CHI2014. Do not cite or distribute.}

\pagenumbering{arabic}

\usepackage{balance}  
\usepackage{graphics} 
\usepackage{times}    
\usepackage{url}      
\usepackage{gensymb}
\usepackage{afterpage}
\usepackage{mathpazo,amsmath}
\def\inch#1{#1''}
\def\ft#1{#1'\thinspace}

\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={SIGCHI Conference Proceedings Format},
pdfauthor={LaTeX},
pdfkeywords={SIGCHI, proceedings, archival format},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

\newcommand\tabhead[1]{\small\textbf{#1}}

\newcommand {\bjoern}[1]{{\color{red}\bf{BH: #1}\normalfont}}
\newcommand {\ben}[1]{{\color{blue}\bf{BZ: #1}\normalfont}}
\newcommand {\sean}[1]{{\color{blue}\bf{SC: #1}\normalfont}}
\newcommand {\claire}[1]{{\color{blue}\bf{CT: #1}\normalfont}}
\newcommand {\studyquote}[1]{\em ``#1''\normalfont}

\begin{document}

\title{A Context Menu for the Real World: Controlling Physical Appliances Through Head-Worn Infrared Targeting}

\numberofauthors{1}
\author{
 \alignauthor Anonymous for submission
}

\maketitle

\begin{abstract}
 % abstract
We introduce a novel method for selecting and controlling smart appliances in physical spaces through a head-worn computing device with near-eye display and wireless communication. We augment a commercial wearable computing device, Google Glass, with a narrow-beam IR emitter for this purpose. This configuration yields a usable beam width of 2 to 4 feet (60 to 120cm) for targeting at room scale. We describe a disambiguation technique if infrared targeting hits multiple targets simultaneously. A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device. We also report qualitative data from using our device to control multiple appliances in a smart home scenario.

\end{abstract}

\keywords{
	smart devices; universal remote control; wearable computing; glass
}

\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}

 % introduction
\section{Introduction}
Increasingly, devices and services in our built environment are networked and can be controlled remotely. The proliferation of smart, controllable devices such as intelligent lighting, AV equipment, HVAC systems, or kitchen appliances raises the question of how to best interact with them.

Today, commercial solutions (such as Belkin WeMo) use handheld mobile devices as {\em universal remote controls} to control such appliances. In these solutions, users first browse a list of all available devices and then call up a device-specific user interface. This method faces two challenges: {\em naming} and {\em scoping}. Assigning clear names is non-trivial. In shared spaces, the person trying to control the device might not be the one that named it - e.g., while an office building manager may know what ``Light 4 in area E'' corresponds to, an occupant may not. Second, without a method of scoping selection to automatically filter non-relevant devices, paging though long lists of names or navigating hierarchies becomes potentially more cumbersome than the physical action the ``convenient'' software solution was meant to replace.

To address these challenges, research has introduced techniques of augmenting mobile devices with accessories like laser pointers to enable direct aiming at target devices~\cite{beigl_point_1999,patel_2-way_2003}. While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at the device to-be-controlled.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/teaser.jpg}
\caption{Using an augmented head-worn device (1), users can control smart home appliances (2) with head orientation targeting. A near-eye display then shows an appliance control UI (3), which users navigate through multitouch gestures.}
\label{fig:teaser}
\end{figure}

In this paper, we introduce a novel method for selecting and controlling smart appliances in physical spaces through the use of a head-worn computing device with near-eye display and wireless communication. We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose. Users first look in the direction of the appliance they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume).  If multiple appliances fall within communication range, a disambiguation technique that combines on-screen information as well as visual feedback on the target appliances lets users select their desired target. Once acquired, an appliance specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}).

Our hardware relies on infrared (IR) communication between Glass and target appliances to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.  Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio. Target appliances similarly have IR receivers and radios. This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target appliances while issuing control commands.

While prior work has tended to focus on proofs-of-concept, we also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control. We first report measurements of range and beam characteristics of our controller. We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface. We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller. We also report qualitative results from participants who use our system for home automation tasks.

 % relatedwork
\section{Related Work}
Relevant prior work exists in the areas of remote control of physical appliances, evaluations of pointing in physical space and augmented reality applications. We discuss each in turn.

\subsection{Remote Control of Physical Appliances}
Standard infrared remote controls for televisions and AV equipment are only meant to control a single device. These controllers tend to use wide-angle infrared LEDs. Universal remote controls are available as dedicated devices (e.g, Crestron\footnote{\url{http://www.crestron.com}}) or applications for smart phones (e.g., Belkin WeMo\footnote{\url{http://www.belkin.com/us/wemo}}). They do not offer spatial selection of target devices, forcing users to browse through lists of pre-configured devices instead. Rukzio found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.

Several approaches to spatial selection with handheld devices exist to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012} or to exchange information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. Key design decisions are the method by which a target device is selected; and the method by which it is then later controlled or configured.

In several techniques, users select objects of interest with laser pointers. The laser dot provides immediate visual feedback to the user what is being selected (though it does not indicate whether the pointed-at object can indeed be controlled). Furthermore, laser pointer becomes obtrusive when there are other people in the space. Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}. Patel extends this technique by modulating the laser light to communicate the controllers' identity~\cite{patel_2-way_2003} to initiate radio communication. These proofs-of-concept do not include thorough evaluations. Kemp et al. use a laser pointer to indicate to robots which item to pick up in a room~\cite{kemp_point-and-click_2008}.

The XWand~\cite{wilson_xwand:_2003} determines its absolute position and orientation and uses a virtual room model to select target devices. Position is determined through two ceiling-mounted cameras; orientation is determined using a built-in IMU. Users can employ physical gestures or utter speech commands to control selected devices. This technique requires room instrumentation and an up-to-date virtual model of device locations. The Tricorder~\cite{lifton_tricorder:_2007} uses IMU orientation coupled with room-localization based on received signal strength indicators (RSSI) to estimate what a user is pointing at.

Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Printed tags like QR codes can also be affixed to devices and read by cameras. Common tagging systems are optimized to be read from a close distance, though it is possible to redesign codes that can be read further away (by encoding less information)~\cite{cross_low-cost_2012}.

Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing --- the user looks at the target device to initiate interaction. Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications. We therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.

\subsection{Evaluation of room-scale selection}
Pausch et al.'s early investigation of head-mounted displays compared head-tracking to handheld orientation control for a target acquisition task in a virtual reality room shown on a head-mounted display~\cite{pausch_user_1993}. They found a clear performance benefit for head-tracking.

On the other hand, Card et al. experimentally determined that the bandwidth of neck muscles is much lower than that of arm, wrist or finger muscle groups~\cite{card_morphological_1991}, which limits the performance of any head orientation-based interaction scheme. However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques.  Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy. Our work only uses head orientation for the initial selection step; since we believe a user's attention is often drawn to the objects they intend to interact with.

Myers et al compared different methods of interacting with displays at a distance~\cite{myers_interacting_2002} and quantified selection time and jitter or position error when using remote handheld pointing. Various techniques outperformed laser pointers.

Our work is complementary as it provides concrete performance data on using head orientation to select targets in a physical environment.

\subsection{Augmented Reality Interfaces}
Augmented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world, though our techniques could be extended to such configurations.

 % interaction
\section{Initiating Interaction through Attention}
This section describes the design goals of our system and their realization in particular interaction techniques.

\subsection{Design Goals}
Our work is motivated by the following design goals that leverage opportunities of head-worn computing, but also acknowledge potential challenges:

{\bf Leverage visual attention:} Take advantage of the fact that visual attention can express intention - initiate interaction based on where a user is already looking.

{\bf Provide immediate feedback about selection targets in the environment:} While a near-eye display can push information to the user, users don't always want to control an object simply because they are looking at it (a problem known in gaze-based interaction as the Midas touch). A calmer~\cite{weiser_coming_1997} approach is to locate visual feedback about selection targets in the environment, to prevent distraction and interruption. Such feedback should be delivered instantaneously, while users look around a room.

{\bf Offer flexible orientation after initiating interaction:} After initiating interaction through head orientation, enable the user to reorient their head or body position during the remaining interaction to prevent neck strain.

{\bf Offer efficient ways to disambiguate orientation input:} It may not always be possible to identify a unique target appliance based on a user's attention and orientation. Offer ways to supplement orientation-based interaction with screen-based interaction to provide disambiguation information.

These design goals find their expression in the following interaction model.

\subsection{Interaction Flow}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/stepbystep_small.png}
\caption{Targeting interaction: when users turn towards a controllable appliance (A$\rightarrow$B), the appliance shows immediate visual feedback (red LED) (B). Users confirm that they wish to connect to this appliance with a tap (C) which triggers connection feedback (blue LED) on the appliance.}
\label{fig:interaction}
\end{figure}

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/stepbystep_multi_small.png}
\caption{When multiple appliances are within range, they all have red LEDs illuminated for feedback (A). When users initiate connections, all target appliances toggle on blue LEDs while the currently selected one blinks (B). Swiping on the touchpad traverses among responding appliances (C).}
\label{fig:interaction_multi}
\end{figure}

{\bf Look:} Users select a target appliance by looking in its general direction.
Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}. Target appliances have IR receivers and offer immediate visual feedback by toggling a red LED whenever a valid id is received (Figure~\ref{fig:interaction}B). This enables {\em scanning} the environment with one's gaze to see which appliances can be controlled.

{\bf Initiate:} Users confirm their desire to connect to an appliance by tapping on the Glass touchpad. After they are connected, the target appliance toggles on a blue LED as visual feedback (Figure~\ref{fig:interaction}C). The next section on disambiguation deals with cases in which multiple appliances received valid IR signals. At this point, all further communication switches over to the 802.15.4 wireless network so that line of sight to the target is no longer needed.

{\bf Control:} Glass displays a user interface for parameters of the chosen appliance. Upon connection, the current status of the appliance is retreived by Glass and shown on the UI. The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:ui_controls}(A)); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:ui_controls}(B)). This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the sagittal plane (up and down). 
Control commands are sent over XBee radios.

{\bf Disengagement:} Users stay connected to the last selected appliance up to a timeout period. During that period, users can disengage through down swipes.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/ui_controls_caption.jpg}
\caption{Two screenshots of UI controls for lamp and video player.}
\label{fig:ui_controls}
\end{figure}

\subsection{Disambiguation}
Head orientation only indicates a general area of visual interest. It does not necessarily match gaze orientation as extra-ocular muscles can move the eyes. The IR beam of our device also has a certain spread (see next section). In an environment dense with potential targets, multiple targets could be within range. Users can tell when multiple feedback LEDs in the environment illuminate (Figure~\ref{fig:interaction_multi}A). To disambiguate, users can either move to adjust their head position, or, alternatively, call up a disambiguation dialog on the Glass display. The dialog presents a list filtered to only those appliances that are within IR range, while appliances also use blue LED as visual cues: all responding appliances light up LEDs while the currently selected one blinks (Figure~\ref{fig:interaction_multi}B). Users navigate the list using the touchpad (Figure~\ref{fig:interaction_multi}C), and then continue their interaction as described above.

 % device

\section{Hardware Device}
\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/glass-with-ir}
\caption{Our augmented Glass prototype has a frame-mounted infrared emitter.}
\label{fig:glass}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/architecture}
\caption{In our system architecture, selection is initiated through infrared but confirmed over 802.15.4. This permits wearers to move their head freely after connecting to an appliance. In the research prototype, users have to carry an additional microcontroller board that marshals messages between Glass' Bluetooth radio an IR/802.15.4, but our custom hardware could also be integrated into the wearable device.}
\label{fig:architecture}
\end{figure}

\subsection{Prototype Implementation}
Our prototype consists of a Google Glass Explorer Edition head-worn computing device, augmented with an infrared emitter that is mounted on the frame, pointing out in the direction of the wearer's view (Figure~\ref{fig:glass}). The IR emitter LED is mounted in an opaque hollow tube, that restricts the outgoing angle of illumination.

In our prototype, Glass communicates over Bluetooth to an additional microcontroller board the user has to wear (Atmel ATMega256). This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame (Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency. We selected XBee 802.15.4 radios to avoid Bluetooth  wake up latencies but we do not claim optimality for our design decisions. Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change. In particular, one could substitute WiFi modules or design an all-Bluetooth network.

\subsection{Device Characterization}
We determined the usable range and accuracy empirically with one IR emitter and two IR receivers. The IR emitter constantly sent out an id signal. The receivers that correctly received the signal turn their LED on for 300 ms.

We placed all three devices at the same height with clear line of sight. The IR emitter is first places 2 feet away from the receivers. The receivers were moved sideways apart from each other until they could no longer receive stable signals. We then recorded the distance of the two receivers for the calculation of coverage angles. The steps are repeated for IR emitters in different distances (as shown in Table~\ref{table:measurements}). We then repeated measurements with the emitter  placed at various depths in the tube (see Figure~\ref{fig:coverage}).

In summary, our measurements suggest that IR communication can be targeted to an area about 2--4' in diameter, up to 16' in front of the user. These values are a reasonable match for selecting appliances in a room-size environment. A wider beam would lead to an increased chance of multiple appliances receiving IR signals simultaneously. A narrower beam will make targeting more challenging, given the precision constraints of human head movement.

\afterpage{
\begin{figure}[t!]
\centering
\includegraphics[width=1.0\columnwidth]{figures/glass-ir-coverage}
\caption{Different IR configurations suggest usable beam widths of 2 to 4 feet and distances up to 16 feet }
\label{fig:coverage}
\end{figure}

\begin{table}[h!]
    \begin{tabular}{l|lllll}
    distance/ depth & \ft2       & \ft4       & \ft8   & \ft{12}  & \ft{16}  \\ \hline
    \inch{0}                     & $74\degree$ & $78\degree$ & N/A  & N/A  & N/A  \\
  \inch{0.5}                   & $60\degree$ & $48\degree$     & $28\degree$ & $22\degree$ & $16\degree$ \\
    \inch{1.0}                     & $46\degree$     & $36\degree$     & $26\degree$ & $18\degree$ & $10\degree$ \\
    \inch{1.5}                   & $36\degree$     & $32\degree$     & $18\degree$ & $14\degree$ & $6\degree$  \\
    \end{tabular}
    \caption{Measured IR coverage angles $\Theta$ at different target distances and different depths of IR emitter inside shielding tube.}
    \label{table:measurements}
\end{table}
}

 % targeting
\section{Physical Target Acquisition Study}
To understand the accuracy and performance of head-orientation-based  selection through our device, we carried out a comparative target acquisition study, where participants had to connect to wireless nodes distributed in a room with our technique, and with an alternate list selection approach.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment. Targets were called out by number; for the list mode condition, participants need to match numbers to letters.}
\label{fig:targeting-study-layout}
\end{figure}

\subsection{Apparatus}
In an indoor environment, 10 wireless nodes are spread across a room at various heights and distances (Figure~\ref{fig:targeting-study-layout}). The nodes are stand-ins for potential smart appliances and have all relevant functionality for targeting and wireless communication, but do not control any actual appliances. Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs (Figure~\ref{fig:target}). An yellow LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple targets are within IR range. Next to each target, a paper sheet shows a number and letter combination, which is used for uniquely identifying the device. The numbers are the primary identifiers, ordered from left to right in the room. This ordering makes it easy to locate them, which simulates looking towards an appliance with a well-known location in a room, and minimizes visual search time.

\begin{figure}[b]
\centering
\includegraphics[width=0.8\columnwidth]{figures/study-node.pdf}
\caption{An example node from the targeting study --- we constructed 10 such nodes - each mounted in a box.}
\label{fig:target}
\end{figure}

\subsection{Methodology}
In our within-subjects design, participants performed 15 target acquisition tasks each with two interaction styles. In the {\em infrared} mode condition, participants used our IR targeting approach; in the {\em list} mode condition, participants had to look up a device's letter code on the printed paper next to the device and then select that letter code from a list displayed on their Glass device. The list was navigated with swipe motions on the Glass touchpad. For each task, participants started at a fixed position in the room. The experimenter called out a number and simultaneously started a timer. Participants then had to find the corresponding device (by looking for its printed code). In the {\em infrared} mode, participants then selected and acquired the target by aiming the IR beam at the target, and confirmed their selection with a touch pad tap. If more than one target was within range, participants had to either use the disambiguation dialog or reposition themselves. In the {\em list} mode, participants had to read the letter next to the number and then select that letter by browsing a linear list shown in their Glass display. While the list was alphabetized, letter arrangement in the room was not. This design required participants to find the target in the room before starting a list navigation to keep visual search times similar in each condition.

Afterwards, participants completed a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.

\subsection{Participants}
We recruited 14 participants from our institution. 13 had never used Glass before. 4 wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle. Half of them performed {\em infrared} mode first and the other half did {\em list} mode first.

\subsection{Measures}
The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; and {\bf user preference}: which interface users preferred for the task after completing the study.

\subsection{Results}
\subsubsection{Performance data}
The time to complete each task can be broken down into the following pieces:

$t_{infrared}=t_{locate}[+t_{reorient}][+t_{disambiguate}]+t_{tap}$

$t_{list}=t_{locate}+t_{listnav}+t_{tap}$

In both conditions, participants first have to locate the target announced by the experimenter through visual search ($t_{locate}$). In the {\em infrared} mode, participants may then directly confirm their selection if only a single target was selected ($t_{tap}$). However, if they don't immediately receive feedback that their target was selected, or if multiple targets were selected, users either have change their position or head orientation ($t_{reorient}$) or they have to step through the on-screen disambiguation dialog ($t_{disambiguate}$). In the {\em list} mode, participants must scroll through the list to find the desired target identifier ($t_{listnav}$).
Thus, {\em infrared} will show a performance benefit if $(t_{reorient}+t_{disambiguate})<t_{listnav}$. This depends on the number of total devices in the environment (increasing $t_{listnav}$), and their density (which will increase $t_{disambiguate}$).

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
\caption{Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), and between IR multiple responses cases and IR single response cases (B). The centers of boxes are median values, while white dashed lines are mean values.}
\label{fig:selection-times}
\end{figure}

We first show results for 10 targets and then discuss extrapolations of these results. Average target acquisition time $t_{infrared}$ was 6.67 seconds, while $t_{list}$ was 8.86 seconds (Figure~\ref{fig:selection-times}A). This difference is significant (Student's t-test, $t(279)=-3.81, p=0.00017$).

To further understand the performance gain in {\em infrared} mode, especially the factor $t_{disambiguate}$, we compare selection times when multiple devices are targeted (and disambiguation is required) to single-device selection times (Figure~\ref{fig:selection-times}B). When there is a single device, $t_{disambiguate}$ is 0 and it takes 6.40 seconds (on average) to complete the connection. When multiple devices are in range, the time increases to 9.16 seconds, indicating 2.76 seconds required to disambiguate. Though it takes significantly longer ($t(19)=-2.7827, p=0.012$ using t-test) in the {\em multiple} case, these cases made up only 10\% of total infrared trials.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
\caption{Times taken to select a device vs.~its order in the list. The dotted line is a linear fit between the mean times and device orders. Two horizontal lines of mean target acquisition times in {\em infrared} mode are also annotated for comparison.}
\label{fig:time-vs-list-order}
\end{figure}

For each device, $t_{listnav}$ depends on their relative position in the list. Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices the {\em infrared} mode interaction techniques will outperform {\em list} mode\footnote{The higher mean value at $order=1$ is caused by one outlier when the participant tried multiple times in {\em list} mode to connect to the right target.}. From the figure, we can see that once the target's order has increased to be larger than 6, the average $t_{listnav}$ for that target would be larger than $t_{reorient} + t_{disambiguate}$. We expect that, when the number of targets keeps increasing, there would be larger time reduction in {\em infrared} mode.

Participants' selection errors do occur in both conditions. However, error rates were low (1.1\% in {\em infrared} mode, and 2.9\% in {\em list} mode, respectively). This precludes us from running a more detailed analysis.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/target-likert.pdf}
\caption{Likert scale ratings for ease of use of aspects of the targeting task. Error bars show standard error. }
\label{fig:target-likert}
\end{figure}

\subsubsection{Preference}

Eleven of 14 users preferred infrared mode over list mode (three preferred list, one was undecided). While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (see Figure~\ref{fig:target-likert}). As self-report data can easily skew positive as participants try to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.

List mode had certain advantages: It was judged to be more accurate and predictable as there was always exactly one device selected in the list (\studyquote{With the list you never have to worry about accidentally picking up two targets}). Also, it did not require a clear line of sight to the target device so participants did not have to move from their starting position (\studyquote{The shortcoming of the IR mode was that you had to be a certain distance away in order for it to detect the appliance}).

On the other hand, list mode was judged to be more \studyquote{annoying} and tedious. The temple-based touchpad for selection was difficult to use for a participant with long hair: \studyquote{List mode was physically difficult for me to navigate, since my long hair wasn't tied back and it kept interfering with my swiping.} Another participant also commented on the ergonomic challenge of touchpad use on Glass: \studyquote{The strength of the IR mode was that I didn't have to use my fingers as much to control. If the items were spaced relatively far apart, it was easy to select a specific appliance.}

One noted benefit of infrared mode was a feeling that it was \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen. One subject called it \studyquote{natural to interact with things just by looking at them}. Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}.

A few perceived weaknesses of infrared mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target. One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}. Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}. Users had to learn the usable angle of the IR emitter before they became successful at controlling the devices: \studyquote{I had to compensate by tilting my head up a little bit.}

 % control
\section{Smart Home Control Scenario}
To understand how our device could be used to interact with smart appliances, we also asked all study participants to work through a concrete scenario. The main goal was to obtain qualitative feedback on the usability and utility of our device with more realistic tasks.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/smarthome-scenario.jpg}
\caption{In the smart home scenario, participants completed a series of appliance control tasks in a simulated living room.}
\label{fig:smarthome}
\end{figure}

\subsection{Methodology}
We recreated a living room environment that had three controllable appliances: a fan, a lamp, and one laptop functioning as a video player (see Figure~\ref{fig:smarthome}). The fan and lamp had binary controls: they could be switched on or off. The laptop had multiple parameterized functions: participants could start, pause, fast forward, rewind, and adjust volume.

We then asked users to work through the following script for controlling the room for watching a movie in the evening:
{\small
\begin{enumerate}
\item Turn off the lights as you want to watch the movie in a darkened room.
\item You feel a little hot in the room, so you turn on the fan.
\item You connect to the Smart TV and start playing the movie. 
\item The volume seems too soft to hear over the fan- turn it up a bit. 
\item After a while, you want to take a break to get a snack. Pause the movie. 
\item When you come back, you've forgotten what was said last - rewind by ~30 seconds and restart the movie. 
\item When the credits roll, you stop the movie and turn the lights back on.
\item After awhile, you turn off the fan and leave the room.
\end{enumerate}
}

For this study, we only elicited subjective data in the form of Likert data and open-ended responses.
\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/scenario-likert.pdf}
\caption{Likert scale ratings for ease of use of aspects of the smart home control scenario. Error bars show standard error.}
\label{fig:smarthome-likert}
\end{figure}

\subsection{Results}
All participants successfully completed the list of tasks. They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous study procedure. Participants saw benefits of the device for families --- \studyquote{ It might also be useful for people who need to take care of small children that they can complete all the tasks while keeping an eye on their children at the same time}, though settings that require more movement than watching a movie at home (e.g., cooking) may be more appropriate scenarios for wearing the device.

\subsubsection{Simple and Complex Controls}
Participants rated the ease of control of particular appliances differently. Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for the more complex movie player (see Figure~\ref{fig:smarthome-likert}). Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}. The screen size and (largely) 1D input put a limit on the complexity of interfaces that can be presented. As one participant remarked: \studyquote{[The media player] does not seem to be more efficient than a tablet device.} The difficulty can also partly be ascribed to our interaction design, which required one finger swipes to switch between parameters and two finger gestures for adjusting parameters --- it was hard for users to exert fine control over two-finger swipes. In addition, users did not always remember these mappings as they are not yet part of a standard gesture vocabulary.

\subsubsection{Eliminating Steps}
Participants liked the efficiency of our design but also suggested further simplification by eliminating the explicit connection step, they wanted to immediately connect to any appliance that receives the infrared signal --- \studyquote{I intuitively want the screen to automatically appear when the IR detects the appliance rather than having to tap to connect.} Such a design would increase the efficiency of interaction, but at a power tradeoff, as the wireless radio will have to send and receive data each time the user looks at a device --- whether on purpose or inadvertently. We leave the study of battery life implications of interaction design choices to future work.

\subsubsection{Feedback On Screen or In The World}
We found out that the near-eye display may occlude or overlap the target appliance when a participant looks at a target. This may make it difficult to read either the on-screen display or see information displayed on the target device.  As one participant remarked \studyquote{This was especially annoying with the TV because there were two screens overlapping each other.} While it is possible to look away once a device has been acquired to better see the Glass display, a tension remains whether users should rely on feedback from the appliances themselves or on the near-eye display.

 % discussion
\section{Discussion}
Our study procedures demonstrated that users can successfully select and control smart appliances with head-worn infrared targeting, and that this technique outperforms list selection on the Google Glass wearable device. In this section, we revisit some of the results and observations and discuss their larger significance and potential paths for future work.

\paragraph{Meaningful Results?}
While the performance increase in our targeting study is statistically significant, readers may wonder whether it is truly meaningful. We believe it is, for two reasons: first, our technique avoids the problems of {\em naming} and {\em scoping} inherent in any interface that uses representations of objects (e.g., a list of identifiers) rather than the objects themselves. We argue that this disintermediation of interaction leads to a cognitively simpler design. Second, our existing study only showed results for a modest number of targets. Our technique should have a wider margin as the number of targets increases. Of course, the number of targets one can realistically expect may vary across application domains.

\paragraph{Hands-Free Operation}
While head movement is not as precise as hand positioning (e.g., in Patel's mobile laser pointing~\cite{patel_2-way_2003}), one key benefit of our target acquisition step is that it does not require the user's hands. This raises the question if the rest of the interaction (disambiguation and device control) could also be achieved in a hands-free fashion. Voice-command control is an obvious candidate, though such approaches have not found widespread adoption because of social acceptability and other factors.

\paragraph{Hardware Limitations}
In our prototype, the IR emitter is fixed in a single position. Due to different sizes and shapes of the users’ heads, the emitter may not line up exactly with their head orientation. An adjustable emitter (paired with a suitable calibration routing) would improve the performance of our design. Some of our users also mentioned that they had to move closer to some targets to successfully select them. A stronger emitter could overcome these problems, but care has to be taken to avoid possible reflection problems where IR light bounces off a wall and hits an unintended target behind the user.

More importantly though, the main limitation of our design is that extra hardware for infrared communication is needed for the head-mounted device and each controllable appliance. One potential approach to sidestep this requirement would be to combine the growing availability of high-resolution indoor maps with live data from the point-of-view camera on the device to determine what a user is looking at without any infrared data exchange.

\paragraph{The Midas Look}
Our participants suggested eliminating explicit initiation of a connection by the user. However, one of the design guidelines for near-eye displays is to avoid pushing information to the display without an initial request from the user --- flashing device information on screen each time a user moves their head would surely be counterproductive. Future work should investigate how to intelligently decide when and how to initiate interaction for the user.

\paragraph{Where is the Target?}
One open design question of our approach is where infrared receivers should be placed. For a light, one might put a received on the light itself, or on the light switch, to cater to existing expectations. For volume control, the infrared receiver might be located on a speaker or on the amplifier. A thorough study of user preferences would be interesting; though we also point out that our architecture could easily support multiple receivers that end up controlling the same appliance.

 % conclusion
\section{Conclusion}
We introduced a novel method for selecting and controlling smart appliances in physical spaces through infrared targeting based on head orientation. Through our solution, we attempt to address the naming and scaling challenges faced by handheld mobile devices. The design takes advantage of the fact that visual attention can express intention. The visual feedback provided by the target appliances helps users keep their focus in the physical world. While we present a prototype approach that requires that the user carry additional hardware, all parts can readily be miniaturized and integrated into future head-worn hardware. We also introduced a disambiguation technique in case head orientation is not sufficient to determine a unique target. We characterized our devices performance, arguing that it is matched well to the amount of head movement people can control without strain. A target acquisition study showed that the technique is efficient; a home control scenario showed promise but also limitations when trying to control complex appliances. As our environment continues to be populated by a swarm of sensing and actuation devices, methods to interrogate and control our smart environments will become increasingly important.

\balance

\bibliographystyle{acm-sigchi}
\begin{thebibliography}{10}

\bibitem{azuma_recent_2001}
Azuma, R., Baillot, Y., Behringer, R., Feiner, S., Julier, S., and {MacIntyre},
  B.
\newblock Recent advances in augmented reality.
\newblock {\em {IEEE} Comput. Graph. Appl. 21}, 6 (Nov. 2001), 34–47.

\bibitem{beigl_point_1999}
Beigl, M.
\newblock Point \& click-interaction in smart environments.
\newblock In {\em Handheld and Ubiquitous Computing}, H.-W. Gellersen, Ed.,
  no.~1707 in Lecture Notes in Computer Science. Springer Berlin Heidelberg,
  Jan. 1999, 311--313.

\bibitem{card_morphological_1991}
Card, S.~K., Mackinlay, J.~D., and Robertson, G.~G.
\newblock A morphological analysis of the design space of input devices.
\newblock {\em {ACM} Trans. Inf. Syst. 9}, 2 (Apr. 1991), 99–122.

\bibitem{costanza_sensortune:_2010}
Costanza, E., Panchard, J., Zufferey, G., Nembrini, J., Freudiger, J., Huang,
  J., and Hubaux, J.-P.
\newblock {SensorTune:} a mobile auditory interface for {DIY} wireless sensor
  networks.
\newblock In {\em Proceedings of the {SIGCHI} Conference on Human Factors in
  Computing Systems}, {CHI} '10, {ACM} (New York, {NY}, {USA}, 2010),
  2317–2326.

\bibitem{cross_low-cost_2012}
Cross, A., Cutrell, E., and Thies, W.
\newblock Low-cost audience polling using computer vision.
\newblock In {\em Proceedings of the 25th annual {ACM} symposium on User
  interface software and technology}, {UIST} '12, {ACM} (New York, {NY}, {USA},
  2012), 45–54.

\bibitem{kemp_point-and-click_2008}
Kemp, C.~C., Anderson, C.~D., Nguyen, H., Trevor, A.~J., and Xu, Z.
\newblock A point-and-click interface for the real world: laser designation of
  objects for mobile manipulation.
\newblock In {\em Proceedings of the 3rd {ACM/IEEE} international conference on
  Human robot interaction}, {HRI} '08, {ACM} (New York, {NY}, {USA}, 2008),
  241–248.

\bibitem{lifton_tricorder:_2007}
Lifton, J., Mittal, M., Lapinski, M., and Paradiso, J.~A.
\newblock Tricorder: A mobile sensor network browser.
\newblock In {\em Proceedings of the {ACM} {CHI} 2007 Conference-Mobile Spatial
  Interaction Workshop} (2007).

\bibitem{mittal_ubicorder:_2011}
Mittal, M., and Paradiso, J.
\newblock Ubicorder: A mobile device for situated interactions with sensor
  networks.
\newblock {\em Sensors Journal, {IEEE} 11}, 3 (2011), 818--828.

\bibitem{myers_interacting_2002}
Myers, B.~A., Bhatnagar, R., Nichols, J., Peck, C.~H., Kong, D., Miller, R.,
  and Long, A.~C.
\newblock Interacting at a distance: measuring the performance of laser
  pointers and other devices.
\newblock In {\em Proceedings of the {SIGCHI} Conference on Human Factors in
  Computing Systems}, {CHI} '02, {ACM} (New York, {NY}, {USA}, 2002), 33–40.

\bibitem{pausch_user_1993}
Pausch, R., Shackelford, M.~A., and Proffitt, D.
\newblock A user study comparing head-mounted and stationary displays.
\newblock In {\em Virtual Reality, 1993. Proceedings., {IEEE} 1993 Symposium on
  Research Frontiers in} (1993), 41–45.

\bibitem{raskar_rfig_2004}
Raskar, R., Beardsley, P., van Baar, J., Wang, Y., Dietz, P., Lee, J., Leigh,
  D., and Willwacher, T.
\newblock {RFIG} lamps: interacting with a self-describing world via
  photosensing wireless tags and projectors.
\newblock In {\em {ACM} {SIGGRAPH} 2004 Papers}, {SIGGRAPH} '04, {ACM} (New
  York, {NY}, {USA}, 2004), 406–415.

\bibitem{rukzio_experimental_2006}
Rukzio, E., Leichtenstern, K., Callaghan, V., Holleis, P., Schmidt, A., and
  Chin, J.
\newblock An experimental comparison of physical mobile interaction techniques:
  Touching, pointing and scanning.
\newblock In {\em {UbiComp} 2006: Ubiquitous Computing}. Springer, 2006,
  87–104.

\bibitem{schmidt_picontrol:_2012}
Schmidt, D., Molyneaux, D., and Cao, X.
\newblock {PICOntrol:} using a handheld projector for direct control of
  physical devices through visible light.
\newblock In {\em Proceedings of the 25th annual {ACM} symposium on User
  interface software and technology}, {UIST} '12, {ACM} (New York, {NY}, {USA},
  2012), 379–388.

\bibitem{patel_2-way_2003}
Shwetak N.~Patel, G. D.~A.
\newblock A 2-way laser-assisted selection scheme for handhelds in a physical
  environment.
\newblock 200--207.

\bibitem{weiser_coming_1997}
Weiser, M., and Brown, J.~S.
\newblock The coming age of calm technolgy.
\newblock In {\em Beyond calculation}, P.~J. Denning and R.~M. Metcalfe, Eds.
  Copernicus, New York, {NY}, {USA}, 1997, 75–85.

\bibitem{willis_sidebyside:_2011}
Willis, K.~D., Poupyrev, I., Hudson, S.~E., and Mahler, M.
\newblock {SideBySide:} ad-hoc multi-user interaction with handheld projectors.
\newblock In {\em Proceedings of the 24th annual {ACM} symposium on User
  interface software and technology}, {UIST} '11, {ACM} (New York, {NY}, {USA},
  2011), 431–440.

\bibitem{wilson_xwand:_2003}
Wilson, A., and Shafer, S.
\newblock {XWand:} {UI} for intelligent spaces.
\newblock In {\em Proceedings of the {SIGCHI} Conference on Human Factors in
  Computing Systems}, {CHI} '03, {ACM} (New York, {NY}, {USA}, 2003),
  545–552.

\end{thebibliography}

\end{document}

### 2014-03-23 test/TSL267_test/serialToOSC.py
added the test file and readme for TSL267 sensor testing

@@ -1,0 +1,77 @@
+# note, select is not guaranteed to be readily used in Windows
+# it's Linux functionality
+
+import serial, glob, argparse, Queue
+import sys, select
+
+parser = argparse.ArgumentParser(description='Pyserial Monitor')
+parser.add_argument('--baud', type=int, action='store', default=9600, help='Specify the baud rate')
+parser.add_argument('--timeout', type=float, action='store', default=1, help='Timeout parameter for serial connection')
+parser.add_argument('--serial', action='store', default=None, help='serial port')
+parser.add_argument('--osc', default=False, action='store_true', help='send readings to OSC for visualization')
+
+# we know for mac it will show as /dev/usb.tty*, so list all of them and ask user to choose
+availables = glob.glob('/dev/tty.*')
+
+arguments = parser.parse_args()
+try:
+  baud = arguments.baud
+  timeout = arguments.timeout
+  serial_port = arguments.serial
+  is_osc_on = arguments.osc
+  print "  baud rate:", baud
+  print "  serial timeout:", timeout
+  print "  serial port:", serial_port
+except Exception as e:
+	print e
+	exit(1)
+
+if serial_port:
+  selected = serial_port
+else:
+  print "All available ports:" 
+  index = 1
+  for port in availables:
+    print '  ', index, port
+    index += 1
+
+  try:
+    num = int(raw_input('Enter the number:'))
+    if num > 0 and num <= len(availables):
+      selected = availables[num-1]
+  except Exception as e:
+    print "invalid input!"
+    exit(1)
+
+try:
+  ser = serial.Serial(selected, baud, timeout=timeout)
+except Exception as e:
+  print "failed to connect the serial port", e
+
+if is_osc_on:
+  import OSC
+  def osc_send(addr, data):
+    msg = OSC.OSCMessage() #  we reuse the same variable msg used above overwriting it
+    msg.setAddress(addr)
+    msg.append(data)
+    client.send(msg) # now we dont need to tell the client the address anymore
+
+  sys.stderr.write( "Starting up OSC...\n")
+  client = OSC.OSCClient()
+  client.connect( ('127.0.0.1', 8000) ) # note that the argument is a tupple and not two arguments
+
+while True:
+  # read line without blocking
+  while sys.stdin in select.select([sys.stdin], [], [], 0)[0]:
+    line = sys.stdin.readline()
+    print '[Console]: ', line,
[…]

### 2014-03-25 test/TSL267_test/measurement.py
changed TSL267 test to include TWO_SENSOR configuration; added python code for measuremnt

@@ -1,0 +1,75 @@
+# note, select is not guaranteed to be readily used in Windows
+# it's Linux functionality
+
+import serial, glob, argparse, Queue
+import sys, select
+
+parser = argparse.ArgumentParser(description='Pyserial Monitor')
+parser.add_argument('--baud', type=int, action='store', default=9600, help='Specify the baud rate')
+parser.add_argument('--timeout', type=float, action='store', default=1, help='Timeout parameter for serial connection')
+parser.add_argument('--serial', action='store', default=None, help='serial port')
+parser.add_argument('--osc', default=False, action='store_true', help='send readings to OSC for visualization')
+
+# we know for mac it will show as /dev/usb.tty*, so list all of them and ask user to choose
+availables = glob.glob('/dev/tty.*')
+
+arguments = parser.parse_args()
+try:
+  baud = arguments.baud
+  timeout = arguments.timeout
+  serial_port = arguments.serial
+  is_osc_on = arguments.osc
+  print "  baud rate:", baud
+  print "  serial timeout:", timeout
+  print "  serial port:", serial_port
+except Exception as e:
+	print e
+	exit(1)
+
+if serial_port:
+  selected = serial_port
+else:
+  print "All available ports:" 
+  index = 1
+  for port in availables:
+    print '  ', index, port
+    index += 1
+
+  try:
+    num = int(raw_input('Enter the number:'))
+    if num > 0 and num <= len(availables):
+      selected = availables[num-1]
+  except Exception as e:
+    print "invalid input!"
+    exit(1)
+
+try:
+  ser = serial.Serial(selected, baud, timeout=timeout)
+except Exception as e:
+  print "failed to connect the serial port", e
+
+if is_osc_on:
+  import OSC
+  def osc_send(addr, data):
+    msg = OSC.OSCMessage() #  we reuse the same variable msg used above overwriting it
+    msg.setAddress(addr)
+    msg.append(data)
+    client.send(msg) # now we dont need to tell the client the address anymore
+
+  sys.stderr.write( "Starting up OSC...\n")
+  client = OSC.OSCClient()
+  client.connect( ('127.0.0.1', 8000) ) # note that the argument is a tupple and not two arguments
+
+while True:
+  # read line without blocking
+  while sys.stdin in select.select([sys.stdin], [], [], 0)[0]:
+    line = sys.stdin.readline()
+    print '[Console]: ', line,
+    ser.write(line)
+  
[…]

### 2014-03-27 server/serial_monitor.py
now support timestamping in messages

@@ -3,6 +3,7 @@
 
 import serial, glob, argparse, Queue
 import sys, select
+from datetime import datetime
 
 parser = argparse.ArgumentParser(description='Pyserial Monitor')
 parser.add_argument('--baud', type=int, action='store', default=9600, help='Specify the baud rate')
@@ -65,11 +66,13 @@ while True:
   # read line without blocking
   while sys.stdin in select.select([sys.stdin], [], [], 0)[0]:
     line = sys.stdin.readline()
-    print '[Console]: ', line,
+    ts = datetime.now().strftime('%Y-%m-%d %H:%M:%S %f')
+    print '[', ts, ', Console]: ', line,
     ser.write(line)
 
   while ser.inWaiting() > 0:
-    print '[Serial]: ',
+    ts = datetime.now().strftime('%Y-%m-%d %H:%M:%S %f')
+    print '[', ts, ', Serial]: ',
     line = ser.readline()
     sys.stdout.write(line)
     if len(line.strip()) > 0 and is_osc_on:

### 2014-03-30 doc/uist14/abstract.tex
new uist2014 intro

\bjoern{I prefer to write the abstract {\em first}, not last, as a blueprint for the rest of the paper.}

Emerging head-worn computing devices can enable hands-free interactions in physical spaces. 
%
We introduce and characterize a method for selecting and interacting with physical targets at a distance using head orientation. We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. The fundamental imprecision of head orientation requires a large selection beam \bjoern{term?} for successful target acquisition. However, a large beam area can lead to ambiguous selections. We introduce a two level disambiguation strategy: first, networked receivers use IR intensity sensors to estimate which target is the intended selection; second, users can fall back to a manual disambiguation technique using their near-eye display. The technique does not require an a priori map of target devices. However, a spatial data structure can be inferred to overcome sensor saturation problems in environments hostile to IR communication (e.g., in sunlight).
%
A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device. Intensity-based target disambiguation reduces selection time and error rate in dense target environments. We apply this technique to an application for controlling smart appliances and report qualitative usage data from a smart home scenario.

### 2014-03-30 doc/uist14/intro.tex
new uist2014 intro

\section{Introduction}

Increasingly, objects in our built environment are equipped with computation and communication, making them potential targets for user interaction. Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for advertising or context-aware computing \bjoern{second part needs to be stronger}. Both uses rely on {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.

Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}. While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed (which also limits use to explicit scenarios); that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at targets in the world.

Emerging head-worn computing devices have the potential to overcome some of these limitations: because they are worn, they do not require retrieval; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to overlay information. \bjoern{needs a bridge from head-worn device to using head orientation as a feature}.

In this paper, we investigate how to leverage head-orientation for target selection in physical spaces. We characterize this interaction modality because it can readily be implemented with small hardware changes to emerging wearable devices. We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose. We use infrared (IR) communication between Glass and target appliances to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.  
%Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio. Target appliances similarly have IR receivers and radios.

Users look in the direction of a target they wish to acquire to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume), then tap to confirm selection.  
%
Head orientation is a useful but imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus. Therefore, it requires area selection \bjoern{reference} combined with disambiguation techniques.
In practice, our IR beam's selection area has a diameter of 60-120cm at distances up to 5m.
If multiple targets fall within communication range, targets compare received IR light intensity readings to disambiguate. Users can override this default selection with an  on-screen target list.

%This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target appliances while issuing control commands.

While prior work has tended to focus on proofs-of-concept, we also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control. We first report measurements of range and beam characteristics of our controller. We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface. We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller. 
We then quantify the additional benefit of using IR-intensity disambiguation \bjoern{and find XYZ.}
%
\bjoern{Say something about the inferred map if we get around to it}.

We then demonstrate how the technique can be applied to a "universal remote control" scenario. In our implementation, user look at a target appliance and tap to load an appliance-specific control UI on the head-mounted display, which enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}) \bjoern{we may need a different figure 1 with the changed focus of the paper}. We also report qualitative results from users performing  home automation tasks with this system.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/teaser.jpg}
\caption{Using an augmented head-worn device (1), users can control smart home appliances (2) with head orientation targeting. A near-eye display then shows an appliance control UI (3), which users navigate through multitouch gestures.}
\label{fig:teaser}
\end{figure}

### 2014-03-30 doc/uist14/abstract.tex
smaller fixes, added area cusor refs

@@
 We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.
-The fundamental imprecision of head orientation requires a large selection beam \bjoern{term?} for successful target acquisition.
-However, a large beam area can lead to ambiguous selections.
+The fundamental imprecision of head orientation requires area selection techniques, which can lead to ambiguous selections.
 We introduce a two level disambiguation strategy: first, networked receivers use IR intensity sensors to estimate which target is the intended selection; second, users can fall back to a manual disambiguation technique using their near-eye display.
@@
 However, a spatial data structure can be inferred to overcome sensor saturation problems in environments hostile to IR communication (e.g., in sunlight).
+\bjoern{speculation, if we get to it}
 %
@@
 Intensity-based target disambiguation reduces selection time and error rate in dense target environments.
-We apply this technique to an application for controlling smart appliances and report qualitative usage data from a smart home scenario.
+\bjoern{rewrite once we have results}.
+We implement this technique in an application for controlling smart appliances and report qualitative user impressions.

### 2014-03-30 doc/uist14/relatedwork.tex
smaller fixes, added area cusor refs

\section{Related Work}
\bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
\bjoern{Need a discussion of area cursors~\cite{kabbash1995prince,worden1997making}.}

### 2014-04-05 doc/uist14/applications.tex
outlined the uist paper

\section{Applications}
\label{sec:applications}

In this section we describe four possible applications that can be built on top of the head orientation-based selection. We focus our discussion on the ``universal remote control'' system.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-05 doc/uist14/conclusion.tex
outlined the uist paper

--- 2013-05-05 doc/report/conclusion.tex
@@
 
-In this project, we explored the novel way of interacting with physical device by expressing users' attention.
-The prototype we have built demonstrate the possibility of identifying user's attention and control various devices.
-
-We will leave X, Y, Z to future work.
+In this paper we have presented our system and solutions for the head orientation selection.
 
@@
 %%% mode: latex
-%%% TeX-master: "main"
+%%% TeX-master: "uist14"
 %%% End:

### 2014-04-05 doc/uist14/disambiguation.tex
outlined the uist paper

\section{Disambiguation Techniques}
\label{sec:disamb-techn}

This section discusses the three different disambiguation techniques we have proposed.

\subsection{Using IR Intensity}
\label{sec:using-ir-intensity}

\subsection{Using Glass Sensors}
\label{sec:using-glass-sensors}

\subsection{Manual Disambiguation}
\label{sec:manu-disamb}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-05 doc/uist14/discussion.tex
outlined the uist paper

@@
 \section{Discussion}
-Our study procedures demonstrated that users can successfully select and control smart appliances with head-worn infrared targeting, and that this technique outperforms list selection on the Google Glass wearable device.
-In this section, we revisit some of the results and observations and discuss their larger significance and potential paths for future work.
+\label{sec:discussion}
 
-\paragraph{Meaningful Results?}
-While the performance increase in our targeting study is statistically significant, readers may wonder whether it is truly meaningful.
-We believe it is, for two reasons: first, our technique avoids the problems of {\em naming} and {\em scoping} inherent in any interface that uses representations of objects (e.g., a list of identifiers) rather than the objects themselves.
-We argue that this disintermediation of interaction leads to a cognitively simpler design.
-Second, our existing study only showed results for a modest number of targets.
-Our technique should have a wider margin as the number of targets increases.
-Of course, the number of targets one can realistically expect may vary across application domains.
-
-\paragraph{Hands-Free Operation}
-While head movement is not as precise as hand positioning (e.g., in Patel's mobile laser pointing~\cite{patel_2-way_2003}), one key benefit of our target acquisition step is that it does not require the user's hands.
-This raises the question if the rest of the interaction (disambiguation and device control) could also be achieved in a hands-free fashion.
-Voice-command control is an obvious candidate, though such approaches have not found widespread adoption because of social acceptability and other factors.
-
-\paragraph{Hardware Limitations}
-In our prototype, the IR emitter is fixed in a single position.
-Due to different sizes and shapes of the users’ heads, the emitter may not line up exactly with their head orientation.
-An adjustable emitter (paired with a suitable calibration routing) would improve the performance of our design.
-Some of our users also mentioned that they had to move closer to some targets to successfully select them.
-A stronger emitter could overcome these problems, but care has to be taken to avoid possible reflection problems where IR light bounces off a wall and hits an unintended target behind the user.
-
-More importantly though, the main limitation of our design is that extra hardware for infrared communication is needed for the head-mounted device and each controllable appliance.
-One potential approach to sidestep this requirement would be to combine the growing availability of high-resolution indoor maps with live data from the point-of-view camera on the device to determine what a user is looking at without any infrared data exchange.
-
-\paragraph{The Midas Look}
-Our participants suggested eliminating explicit initiation of a connection by the user.
-However, one of the design guidelines for near-eye displays is to avoid pushing information to the display without an initial request from the user --- flashing device information on screen each time a user moves their head would surely be counterproductive.
-Future work should investigate how to intelligently decide when and how to initiate interaction for the user.
-
-\paragraph{Where is the Target?}
-One open design question of our approach is where infrared receivers should be placed.
-For a light, one might put a received on the light itself, or on the light switch, to cater to existing expectations.
-For volume control, the infrared receiver might be located on a speaker or on the amplifier.
-A thorough study of user preferences would be interesting; though we also point out that our architecture could easily support multiple receivers that end up controlling the same appliance.
-
-%Our prototype does not yet have a general way of communicating device descriptions and capabilities - this means our prototype cannot yet control new, unknown devices - they first have to be modeled.
-We may extend the PUC work~\cite{nichols} to address this issue.
-
-%bjoern: these below are details about what happened in the experiment; they are less relevant to the research contributions of the paper.
-%\item During the experiment, a lot of wireless messages were transferred for communication and logging purposes.
-Some messages got lost or delayed and caused the users had to redo that particular action.
-Infrared mode requires a lot of more handshake communications than the list mode.
-This means that in a more robust infrastructure, the performance of infrared mode can improve more than that of the list mode.
-
-%\item In list mode, users’ eyes focuses more on the display to traverse between the list while in infrared mode, users’ eyes remain in the physical world more.
-
-%\item The list mode could also be improved by making the swiping gestures more responsive, e.g. swiping speed affects traversing speed. % too detailed - partially covered later
-
-%\item In our experiment, after users locates the nodes by ID, they can instantly figure out their names by the label provided.
-In realistic environments, users have to recall the names of the appliances, which would take a lot more time.
-Furthermore, finding names such as “Living room corner stand lamp” would be more difficult than an alphabet.
-On the other hand, infrared mode will most likely to be unaffected since it does not require name-recalling or text matching.
-%too detailed
+We discuss a few issues with the system.
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "uist14"
+%%% End:

### 2014-04-05 doc/uist14/intro.tex
outlined the uist paper

@@
 
+%% from the swarm vision to the necessity of selection
 Increasingly, objects in our built environment are equipped with computation and communication, making them potential targets for user interaction.
-Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for advertising or context-aware computing \bjoern{second part needs to be stronger}.
+Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for targeted-advertisement or context-aware computing \bjoern{second part needs to be stronger}.
 Both uses rely on {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.
 
+%% previous approaches are limited
 Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}.
@@
 
+%% introducing head-worn computing and head orientation
 Emerging head-worn computing devices have the potential to overcome some of these limitations: because they are worn, they do not require retrieval; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to overlay information.
-\bjoern{needs a bridge from head-worn device to using head orientation as a feature}.
+Also such computing devices are well aligned with the user's line of sight, which is a strong indication of the user's interaction targets \ben{citation needed here}.
 
+%% head orientation limit
+Though natural, head orientation is an imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus. 
+% should we argue gaze tracking doesn't help much but only increases the complexity?
+Therefore, it requires area selection \bjoern{reference} combined with disambiguation techniques.
+
+%% the inaccuracy of head orientation
 In this paper, we investigate how to leverage head-orientation for target selection in physical spaces.
-We characterize this interaction modality because it can readily be implemented with small hardware changes to emerging wearable devices.
-We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose.
-We use infrared (IR) communication between Glass and target appliances to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.
-%Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio. Target appliances similarly have IR receivers and radios.
+We propose a simple low-cost solution that can be readily implemented with small hardware changes to emerging wearable devices.
+Specifically, we augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose.
+We use infrared (IR) communication between Glass and target appliances to capture the user's head orientation.
+Because IR signal usually covers a cone shape of area (a diameter of 60-120cm and distance up to 6m), it's suitable for the area selection modality.
+To achieve the disambiguation when multiple targets have received IR signals, we propose three different ways of disambiguation.
+First, the targets compare the IR signal intensity and we propose some intuitive heuristic to infer the user's intention in disambiguation.
+Second, we incorporate the sensors on Google Glass to track the user's head movement and combine this with IR intensity changes.
+Last but not least, we always leave the option of manually changing the current targets to the user such but limit the scope of potential targets to the ones that have received IR signals.
+%% three ways
 
-Users look in the direction of a target they wish to acquire to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume), then tap to confirm selection.
-%
-Head orientation is a useful but imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus.
-Therefore, it requires area selection \bjoern{reference} combined with disambiguation techniques.
-In practice, our IR beam's selection area has a diameter of 60-120cm at distances up to 5m.
-If multiple targets fall within communication range, targets compare received IR light intensity readings to disambiguate. Users can override this default selection with an  on-screen target list.
-
-%This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target appliances while issuing control commands.
-
-While prior work has tended to focus on proofs-of-concept, we also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control.
+While prior work has tended to focus on proofs-of-concept for head-orientation based interaction, we also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control.
 We first report measurements of range and beam characteristics of our controller.
@@
 We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.
-We then quantify the additional benefit of using IR-intensity disambiguation \bjoern{and find XYZ.}
-%
-\bjoern{Say something about the inferred map if we get around to it}.
+We then quantify the additional benefit of using IR-intensity disambiguation \bjoern{and find XYZ.}. \bjoern{Say something about the inferred map if we get around to it}.
 
-We then demonstrate how the technique can be applied to a "universal remote control" scenario.
-In our implementation, user look at a target appliance and tap to load an appliance-specific control UI on the head-mounted display, which enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}) \bjoern{we may need a different figure 1 with the changed focus of the paper}.
-We also report qualitative results from users performing  home automation tasks with this system.
+Such head orientation-based selection enables a wide range of context-aware applications.
+Examples include smart home remote control, break reminder monitor starer, museum attention tracking, indoor positioning, etc.
+In Figure\,\ref{fig:teaser}, it's a demonstration of the ``universal remote control'' scenario.
+The user can easily select the smart appliances by simply looking at it's general direction and confirm such selection with either voice command or by tapping the Glass input pad.
+Then an appliance-specific control UI will be shown on the head-mounted display.
+For this application, we have asked 14 participants to try the system and we report the qualitative results from them performing home automation tasks.
 
@@
 \end{figure}
+
+%% Contribution
+In this paper, by presenting our work on head orientation-based beam cursor,  we make the following contributions:
+\begin{itemize}
+\item We design and implement a low-cost, easy-to-use head orientation system that enables interaction with physical environment.
+\item To solve the inherent inaccuracy of head orientation, we have proposed three different ways to achieve disambiguation and we evaluate them based on real-world user study.
+\item We also demonstrate that several useful applications can be easily built with the head orientation system.
+\end{itemize}
+
+\ben{Having problem fixing the reference}.
+In the remainder of this paper, we will first describe the related works in physical selection and targeting.
+Given that we focus our scope on head orientation, in Section~\ref{sec:background} we will briefly review human's neck ergonomics as the background for head orientation.
+In Section~\ref{sec:syst-design-prot}, we then present our system designed for the study of head orientation based selection.
+The prototype is also what we have used for building the example applications.
+The disambiguation techniques are discussed in Section~\ref{sec:disamb-techn}, which is followed by the study and evaluation in Section~\ref{sec:evaluation}.
+To further show the usefulness of having such head orientation-based selection, we describe the enabled applications in Section~\ref{sec:applications}, with detailed implementation about the ``universal remote control'' application.
+The discussion and conclusion are in Section~\ref{sec:discussion} and Section~\ref{sec:conclusion} respectively.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "uist14"
+%%% End:

### 2014-04-05 doc/uist14/system.tex
outlined the uist paper

\section{System Design and Prototyping}
\label{sec:syst-design-prot}

\subsection{Design Considerations}
\label{sec:design-cons}

\subsection{IR as beam cursor}
\label{sec:ir-as-beam}

\subsection{Head-up Display}
\label{sec:head-up-display}

\subsection{Wireless Communication}
\label{sec:wirel-comm}

\subsection{3D Printed Adjustable Holder}
\label{sec:3d-print-adjust}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-05 doc/uist14/system.tex
added system design considerations

@@
 \label{sec:syst-design-prot}
+Given the background study on using head orientation for targeting, we have designed and prototyped a low-cost system to study how it can be used in a real-world scenario.
+In this section, we will first briefly describe our design considerations and then for each considerations we propose commercial off-the-shelf (COTS) solutions that enable our prototyping.
 
@@
 \label{sec:design-cons}
+
+\ben{need to refine this by a better wording. Also we need justification for each!}
+
+{\bf Area selection:} The head muscle's fundamental limitation imposes us to choose an area selection technique.
+However, there exists such trade-off between the effective size of the beam cursor and the necessity of doing disambiguation.
+The proper technique should be one that could do both.
+
+% the two visual feedback seems a bit random...
+{\bf Leverage visual feedback:} We consider two types of visual feedback.
+The first type comes from the environment and enables instantaneous visual cue for users to know his head-orientation.
+The second further helps any type of interaction by embedding rich content information in the near-eye display.
+% A calmer~\cite{weiser_coming_1997} approach is to locate visual feedback about selection targets in the environment, to prevent distraction and interruption.
+Such feedback should be delivered instantaneously, while users look around a room.
+
+{\bf Flexible communication:} Since the system enables interaction with the physical environment, it's important to provide communication between each end-node.
+With the advance in wireless communication, we can scatter the nodes in an indoor-environment for both testing and deployment.
+
+{\bf Adjustable angle:} We learned from the study about human's head orientation pointing that there is a certain bias and variance per person.
+To fit each person's line of sight, the system should enable adjustability for calibration.
+
+These design considerations find their expression in our selection of techniques in the following few subsections.
 

### 2014-04-05 doc/uist14/disambiguation.tex
adding system and disambiguation, the texts are badly written now, just outlining

@@
 
-This section discusses the three different disambiguation techniques we have proposed.
+For area selection technique, when multiple targets are within the range, the disambiguation is necessary.
+We propose three different ways of disambiguation solutions in this section.
+Initially our system leaves the disambiguation to the user \ben{cite our paper}.
+However, our studies showed that manual disambiguation is a time-consuming task, so we only consult manual action when the system cannot figure out itself.
+The automatic disambiguation are based on the observation that users tend to adjust their head orientation a bit when they found the target is wrong.
+So we achieve a more fine-grained targeting by exploiting the IR intensity and the change of IR readings, as well as the sensors on Google Glass.
 
@@
 
+From the IR datasheet\footnote{https://www.sparkfun.com/datasheets/Components/LED/YSL-R531FR2C-F1.pdf}, for a directional IR emitter, the illuminance at the receiver falls off as the receiver is positioned at an angle from the emitter.
+Therefore, given multiple targets that are equidistant from the user, the one straight most directly in the line of sight will read the highest intensity values.
+However, illuminance could also drop off as $\frac{1}{d^2}$ where $d$ is the distance between emitter and receiver.
+So a closer targets will read higher intensity values.
+This introduces problems for two targets which are not equidistant from the user, since the closer one will always read a higher value.
+To overcome this problem, we utilize the fact that the user might move his head a little bit and the changes in the intensity readings will give us such information.
+We put a higher weight on those with a higher intensity reading gains.
+
 \subsection{Using Glass Sensors}
 \label{sec:using-glass-sensors}
+The second disambiguation technique is to use the sensors on Google Glass.
+As we discussed before, Google Glass not only provides us with the head-up display, but also many other sensors.
+For head orientation, we uses the rotation vector sensor provided by Android API.
+Such orientation provides the relative movement of user's head.
+However, without the absolute map, we cannot easily map the readings into targets.
+In this case, the combination of IR directions provide an anchor direction and the Glass application calibrates its estimation of orientation.
+From a successful targeting event, we construct the adjacent map during the learning phase and use the model to predict user's targeting if we detect multiple IR readings and also user's head movement.
 
@@
 \label{sec:manu-disamb}
+
+For cases where the ambient IR noise is too high, and the environment is too dense, or when our disambiguation fails, we always enable the manual disambiguation option to the users.
+This is achieved in a way similar to mouse click.
+When the user single click, we only display the current highlighted targets, when the user click again (a double clicke event), we confirm the selection.
+Therefore, before the second click even happens, the user is always free to change the target by swiping forward or backward.
 

### 2014-04-05 doc/uist14/system.tex
adding system and disambiguation, the texts are badly written now, just outlining

@@
 \label{sec:syst-design-prot}
+
+\ben{This section is still far from my expectation, will refine a lot later! Now I am just laying down ideas.}
+
 Given the background study on using head orientation for targeting, we have designed and prototyped a low-cost system to study how it can be used in a real-world scenario.
@@
 {\bf Adjustable angle:} We learned from the study about human's head orientation pointing that there is a certain bias and variance per person.
-To fit each person's line of sight, the system should enable adjustability for calibration.
+To fit each person's line of sight, the system should enable adjust-ability for calibration.
 
@@
 \label{sec:ir-as-beam}
+\ben{will expand on each.}
+We choose IR as the proper hardware for the area selection beam for a couple of reasons.
+First, IR has been used in many home appliances for remote control.
+Both the emitter and receivers are fairly cheap to manufacture and integrated into devices which can be put in physical environment.
+Second, the signal characteristics of IR gives us a cone-shape coverage, which is ideal for area selection.
+While the intensity of IR signal can be easily read from a receiver, it also provides potentials for disambiguation when multiple devices fall in the reception area.
+Third, IR could be immune to most interference.
+The ambient could be annoying, but with a combination of IR signal validation, it's still safe.
+For multipath and reflection exist, the intensity reading can rule out them easily.
 
-\subsection{Head-up Display}
+Therefore, we can add IR emitter to the user and IR receiver in the environment to detect both IR signals and also the intensity of IR.
+For the visual feedback from the environment, we designed the system such that when they receive the IR signal and the intensity reading is the largest amongst the nodes that have received an valid signal, it lights up a LED for the beam cursor indication.
+
+\subsection{Head-up Display with interaction capability}
 \label{sec:head-up-display}
+As we argued for a rich interface, the head-up display (HUD) is an ideal solution to overlay content on top of the physical environment.
+We have considered a few COTS solutions (including a custom hardware solution, Goggle, Google Glass).
+And Google Glass has become our choice because it's desigend for everyday use, capable for a rich set of gesture and voice commands, and the head-up display is easy to use and program.
 
@@
 \label{sec:wirel-comm}
+The IR can be only used for expressing the user's targeting intention, we need another data communication technique that is more reliable and flexible.
+This is not only necessary for our targeting experiement to transmit received IR signals and the IR intensity readings, but also useful for application development that can exchange arbitrary application-specific messages.
+We have considered WiFi, Bluetooth Low Energy (BLE) and 802.15.4 (ZigBee) radio.
+A detailed comparison is out of the scope of this paper, but we have chosen to use 802.15.4 radio, since it is designed for personal area network (PAN) communication, and there are both industrial activity to promote it \footnote{XBee: http://www.digi.com/xbee/} and academia research \cite{watteyne2012openwsn} to make it lower power and more suitable for IoT.
 
@@
 \label{sec:3d-print-adjust}
+Since we have chosen to use IR for targeting, the holder where IR emitter resides can be designed for adjustment. We have designed a ball-joint model and 3D printed it. See Figure 1.
+
+\subsection{Overall System}
+\label{sec:overall-system}
+So far, we have described our choice separately, in this subsection we focus on our system architecture that stitches the components together. 
+Figure 2 is our overall architecture photo. Because we need to take command from Google Glass, we have used another microcontroller that speaks Bluetooth to communicate with the Google Glass.
+
+%% show the figure here
 

### 2014-04-06 doc/uist14/applications.tex
half way of today's writing task, out for dinner

@@
 
-In this section we describe four possible applications that can be built on top of the head orientation-based selection. We focus our discussion on the ``universal remote control'' system.
+In this section we describe four possible applications that can be built on top of the head orientation-based selection.
+We focus our discussion on {\em GlassRemote} -- a universal remote control application.
+
+\subsection{GlassRemote}
+\label{sec:univ-remote-contr}
+
+We extended the head-orientation targeting and existing hardware platform to achieve a universal remote control functionality in smart home scenario (we have a prior study archived in \cite{Chen:EECS-2013-200}).
+The motivation is to solve two challenges that exist in current list-based solutions -- {\em naming} and {\em scoping}.
+First, assigning clear names is non-trivial.
+Especially in shared spaces, the person trying to control the device might not be the one that named it - e.g., while an office building manager may know what ``Light 4 in area E'' corresponds to, an occupant may not.
+Second, without a method of scoping selection to automatically filter non-relevant devices, paging though long lists of names or navigating hierarchies becomes potentially more cumbersome than the physical action that the ``convenient'' software solution was meant to replace.
+Our head-orientation based interaction elegently solves these two problems.
+
+To assess the usability of such applications when the novel interaction scheme is added, we have created a few smart appliances (including a fan, a lamp and a TV) and asked 14 participants to work through a concrete smart home scenario.
+The fan and lamp are relatively dumb and only support turning-on and turning-off.
+The smart TV provides control capability of switching the channels and adjust the volumn, as well as pause/play.
+For the smart home scenario, the participant has been asked to perform a series of actions that swtich the targets and control them.
+We then report some of the results to strength our motivation for the head-orientation interaction.
+
+Durign the study, all participants have successfully completed the list of tasks.
+They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous study procedure.
+Participants saw benefits of the device for families --- \studyquote{It might also be useful for people who need to take care of small children that they can complete all the tasks while keeping an eye on their children at the same time}, though settings that require more movement than watching a movie at home (e.g., cooking) may be more appropriate scenarios for wearing the device.
+
+%% I feel like we should mention a bit about the interaction complexity. But that's a bit out of the scope of this paper.
+%% Participants rated the ease of control of particular appliances differently.
+Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for the more complex movie player (see Figure~\ref{fig:smarthome-likert}).
+Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
+The screen size and (largely) 1D input put a limit on the complexity of interfaces that can be presented.
+As one participant remarked: \studyquote{[The media player] does not seem to be more efficient than a tablet device.} The difficulty can also partly be ascribed to our interaction design, which required one finger swipes to switch between parameters and two finger gestures for adjusting parameters --- it was hard for users to exert fine control over two-finger swipes.
+In addition, users did not always remember these mappings as they are not yet part of a standard gesture vocabulary.
+
+\subsection{Attention Tracking}
+\label{sec:attention-tracking}
+
+The second tier of applications enabled by our system is an implicit use of users' head orientation.
+One sample application is to obtain a personal statistics.
+Nowadays, we are surrounded by all sorts of screens, and many people's job requires them to stare at the monitors for the whole day.
+% citation needed, take a look at this one: E-Readers and Visual Fatigue, http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3873942/
+Knowing how much time you have spent in front of the screen is definitely useful, and the application can be built to remind the user to take a break whenever a certain time threshold is reached.
+
+%% We might be lacking evidence, but let me just write down the draft first.
+Our system enables a flexible and easy-to-use tracking of users' attention without knowing the map.
+Therefore, it's suitable to be deployed in large museums (even when the exihibits positions are adjusted constantly).
+Then such attention tracking will facillitate many social science studies.
+Previous solutions are either using RFID \cite{Hsi:2005:REV:1081992.1082021} or video taping % this is a cv of Adam who has worked on video analysis http://www.exploratorium.edu/vre/cvs/ak_cv.pdf.
+However, head orientation is a better approximation of attention rather than proximity.
+By placing our tracking devices in the environment, we can easily get a more accurate profiling of users intention.
+
+\subsection{Indoor Positioning}
+\label{sec:indoor-positioning}
+
+There is one category of indoor positioning which is based on angle of arrival.
+Traditional approaches on indoor positioning are based on fingerprinting or triangulation, which is known to be bad due to multi-path effect for indoor environment.
+With our system, using the adjacent mapping calibration we can figure out the position of the targets in a room, and then based on a new sample, we can estimate a user's position with some simple math.
 

### 2014-04-06 doc/uist14/evaluation.tex
outlined the uist paper
half way of today's writing task, out for dinner

--- 2013-05-03 doc/report/design.tex
@@
-\section{Design}
-\label{sec:design}
+\section{Evaluation}
+\label{sec:evaluation}
 
-Talks about the system design space.
-What might be available to achieve such a function.
+In this section, we evaluate our targeting system by benchmarking its performance with an user study.
+We first describe our design of the physical targeting study, followed by the results in different scenarios.
+
+\subsection{Apparatus}
+We deployed 10 wireless nodes in an indoor environment at various distance and density.
+Then the participant is asked to stand in a fixed position in the room and look down before the targeting request is made.
+We randomly generate a target and highlight it using a yellow LED, and then measure the amount of the time for a user to achieve a targeting transaction.
+
+\subsection{Methodology}
+\label{sec:methodology}
+
+We break down the time for a complete target acquisition to the following pieces:
+$t_{total}=t_{locate}+t_{reorient}+t_{disambiguate}+t_{tap}$
+And during the study, we video-taped the participants' behavior and measure each pieces of the time.
+
+In our previuos studies, we have made a comparison of using IR and Glass List UI.
+There we reach an conclusion that once the targets numbers have exceeded a certain value (6 in a single room), then list selection (using Google Glass List UI) will be worse than IR targeting.
+However, the average results there comes from cases where disambiguation is not needed and needed.
+And disambiguation is a time-consuming part.
+So in this paper we focus more study on the disambiguation techniques.
+
+\subsection{IR intensity based}
+
+\subsection{IR intensity together with Google Glass}
+
+\subsection{Manual disambiguation}
 

### 2014-04-06 doc/uist14/system.tex
refining system section

--- 2014-04-05 doc/uist14/system.tex
@@
 \label{sec:syst-design-prot}
+
+This section describes the design and implementation of our prototype system.
+We start by summarizing the design considerations in this head-orientation targeting, and for each consideration, we make the choice of commercial off-the-shelf (COTS) components.
+Afterwards, we conclude this section by a depiction of the overall system architecture that is used for both our study and future application development.
 
@@
 
+To effectively enables the physical targeting based on user's head movement, and build such a platform for applications, we have come up with the following design considerations and the justification for each.
+
+{\bf Directional area selection:} To be aligned with the user's head orientation, the sensing modality should be directional.
+Merely direction is not enough, from Section~\ref{sec:background}, the head muscle's fundamental limitation and Fitts' law imposes us to choose an area selection technique.
+However,  with a larger coverage area, there will be a higher chance for disambiguation.
+We should find the proper technique that achieves area selection while still providing ways for a more fine-grained resolution.
+
+% the two visual feedback seems a bit random...
+{\bf Leverage visual feedback:} Since the user cannot see his own head orientation, there should be some effective ways to provide feedback to the users.
+We consider two types of visual feedbacks.
+The first comes from the environment which enables instantaneous visual cue while the user is looking around.
+The second should remain unchanged even if the users' head orientation has changed; in such a way, further interactions are enabled without imposing the burden of staring.
+But such a display shouldn't distract the user's current line-of-sight attention -- which means a near-eye display is better than using a hand-held device.
+
+% A calmer~\cite{weiser_coming_1997} approach is to locate visual feedback about selection targets in the environment, to prevent distraction and interruption.
+Such feedback should be delivered instantaneously, while users look around a room.
+
+{\bf Flexible communication:} The system is for the interaction with the environment, therefore the communication among devices should be flexible.
+Especially there would be a consensus problem if the end-nodes are not agreeing on what the user is looking at.
+The communication would be necessary for resolving the inconsistency in such a distributed environment.
+Also such communication capability can make the system as a platform for other applications.
+
+{\bf Adjustable angle:} From our pilot study, we have learned that human's head orientation pointing is biased for each person.
+In order to align area selection technique, there should be ways to adjust or calibrate the covered area.
+
+These design considerations find their expression in our selection of techniques in the following few subsections.
+
 \subsection{IR as beam cursor}
@@
 
-\subsection{Head-up Display}
+For the directional area selection technique, we have chosen to use infrared (IR) for a couple of reasons.
+
+First, the signal characteristics of IR makes it suitable for directional area selection.
+There are a wide range of viewing angles that we can choose (DigiKey offers IR emitters from $4^\circ$ to $240^\circ$, however, most are around $20^\circ$ of viewing angle).
+This leads to a cone shape-like area coverage and can be well-aligned with humna's line of sight.
+
+Second, for the disambiguation consideration, since the infrared illuminance falls off with respect to both the offset angle from the center and the distance to the emitter, even though the signal might cover multiple targets, using the intensity from the receivers can help disambiguation.
+% for the 4 degree to 240 degree, I used information from http://www.digikey.com/product-search/en/optoelectronics/infrared-uv-visible-emitters/524328?k=IR
+
+In addition, IR has been used for remote control for many years. The emitters and receivers are fairly cheap to manufacture; and it's easy for integration with existing devices. 
+What's more, IR is immune to most interference sources.
+The ambient IR might influence the IR receivers' intensity reading, however with simple IR signal validation, the receiver is still safe.
+There might be issues for signal reflection, either bouncing from the ground to the target or bouncing back to the opposite direction.
+However, such visual feedback is not aligned with the user's attention so no further action from the user might be taken; or the genuine target could easily beat the false positive since the one with reflection tends to travel a longer distance, making the signal strength smaller.
+
+Therefore, we can attach an IR emitter to capture the head orientation and install some IR receivers in the environment for both signals detection and intensity reading.
+For our prototype, we have found some COTS components for IR.
+The IR emitter is the one from OSRAM Opto Semiconductors Inc with manufacturer part number SFH 4545.
+From the datasheet, it has a view angle of $10^\circ$.
+Part of the reason we have chosen it is because the radiant intensity can quite high if we provide it with sufficient current flow (550mW/sr @ 100mA).
+We can then easily adjust it using a resistor to satisfy different demands.
+% transmitter webpage: http://www.digikey.com/product-search/en?x=0&y=0&lang=en&site=us&KeyWords=475-2919-ND
+For the IR receiver, we use 38.0kHz IR Receiver Modules from Vishay Semiconductor Opto Division (manufacturer part number TSOP38238). 
+% receiver webpage http://www.digikey.com/product-detail/en/TSOP38238/751-1227-ND/1681362
+In order to read the IR intensity, we integrates another IR light-to-voltage converter from AMS-TAOS USA Inc (manufacturer part number TSL267-LF) which is pretty sensitive to IR irradiance. 
+% http://www.digikey.com/product-detail/en/TSL267-LF/TSL267-LF-ND/3095052
+
+We use Arduino platform for the prototyping, and there is an easy-to-use IR library\footnote{\url{https://github.com/shirriff/Arduino-IRremote}} where we can customize the IR signal for both transmission and reception.
+To read the IR intensity, an 10-bit ADC is attached to the output of TSL267-LF which gives a reading from 0 to 1023.
+
+\subsection{Environment LEDs and Head-up display}
 \label{sec:head-up-display}
+As we have argued for the visual feedback design considerations, the environment should provide instantaneous visual cue to respond to user's head orientation.
+This can be easily achieved with 1-bit LEDs on each devices that has an IR receiver.
+For the visual feedback that is persistent even when the user's head orientation has changed, we have considered a few COTS solutions that support the head-up display, and ended up with Google Glass because it's easy to program (it supports standard Android programming) and the device is also light-weight for everyday use.
+At the time of writing, we have Google Glass runs on Android 4.0.4 (software version: XE12).
 

### 2014-04-06 doc/uist14/system.tex
still refining system, and added placeholder

@@
 The communication would be necessary for resolving the inconsistency in such a distributed environment.
-Also such communication capability can make the system as a platform for other applications.
+Also providing a flexible communication capability can make the system as a platform for other applications.
 
@@
 For the visual feedback that is persistent even when the user's head orientation has changed, we have considered a few COTS solutions that support the head-up display, and ended up with Google Glass because it's easy to program (it supports standard Android programming) and the device is also light-weight for everyday use.
-At the time of writing, we have Google Glass runs on Android 4.0.4 (software version: XE12).
+At the time of writing, we have Google Glass running Android 4.0.4 (software version: XE12).
 
@@
 \label{sec:wirel-comm}
-The IR can be only used for expressing the user's targeting intention, we need another data communication technique that is more reliable and flexible.
-This is not only necessary for our targeting experiement to transmit received IR signals and the IR intensity readings, but also useful for application development that can exchange arbitrary application-specific messages.
-We have considered WiFi, Bluetooth Low Energy (BLE) and 802.15.4 (ZigBee) radio.
-A detailed comparison is out of the scope of this paper, but we have chosen to use 802.15.4 radio, since it is designed for personal area network (PAN) communication, and there are both industrial activity to promote it \footnote{XBee: http://www.digi.com/xbee/} and academia research \cite{watteyne2012openwsn} to make it lower power and more suitable for IoT.
+
+There are many choices of wireless communication techniques that we could consider to accompany IR's directional targeting.
+WiFi, Bluetooth Low Energy (BLE) and 802.15.4 (ZigBee) radio are all candidates for such communications.
+A detailed comparison is out of the scope of this paper.
+For our prototype, we have chosen to use XBee ZB\footnote{XBee: http://www.digi.com/xbee/}(based on ZigBee protocol and 802.15.4 radio) for the mesh network because of the ease to program and integration to Arduino platform.
+The COTS adapter provides an easy-to-use interface where we could use either hardware serial pins or the software serial library to send and receive messages.
+
+As we mentioned that the communication capability has to be included to resolve the consensus view of head orientation, we use the star topology -- the node which controls IR emission acts as the PAN coordinator.
+It collects the IR received signal strength (RSS) from each targets and runs a local algorithm to determine which one is the current target.
+
+Using such an architecture for communication also makes the integration with Google Glass easier.
+For now, Google Glass is capable of WiFi and Bluetooth; so we have added a Bluetooth modem to establish connections with Google Glass.
+We use BlueSMiRF Silver (WRL-12577) for our prototype.
 
@@
 \label{sec:3d-print-adjust}
-Since we have chosen to use IR for targeting, the holder where IR emitter resides can be designed for adjustment. We have designed a ball-joint model and 3D printed it. See Figure 1.
+To compensate for the bias of each person's line of sight with the IR emitter, we have designed the holder that can be easily attached onto Google Glass while we can adjust the beam angle flexibly.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.9\columnwidth]{figures/placeholder.pdf}
+\caption{we should replace this placeholder with Claire's design and printed version.}
+\label{fig:3d-holder}
+\end{figure}
 

### 2014-04-06 doc/uist14/system.tex
added new texts for system

@@
 There are a wide range of viewing angles that we can choose (DigiKey offers IR emitters from $4^\circ$ to $240^\circ$, however, most are around $20^\circ$ of viewing angle).
-This leads to a cone shape-like area coverage and can be well-aligned with humna's line of sight.
+This leads to a cone shape-like area coverage and can be well-aligned with human's line of sight.
 
@@
 The IR emitter is the one from OSRAM Opto Semiconductors Inc with manufacturer part number SFH 4545.
-From the datasheet, it has a view angle of $10^\circ$.
+From the data-sheet, it has a view angle of $10^\circ$.
 Part of the reason we have chosen it is because the radiant intensity can quite high if we provide it with sufficient current flow (550mW/sr @ 100mA).
@@
 Using such an architecture for communication also makes the integration with Google Glass easier.
-For now, Google Glass is capable of WiFi and Bluetooth; so we have added a Bluetooth modem to establish connections with Google Glass.
+For now, Google Glass is only capable of WiFi and Bluetooth; so we have to add a Bluetooth modem to establish connections with Google Glass\footnote{This is a compromise now, and we are planning to migrate the entire system to BLE once Google upgrade the Glass software to Android 4.4+.
+The prototype now suffice our demand to run the head orientation study.}.
 We use BlueSMiRF Silver (WRL-12577) for our prototype.
@@
 \includegraphics[width=0.9\columnwidth]{figures/placeholder.pdf}
-\caption{we should replace this placeholder with Claire's design and printed version.}
+\caption{we should replace this place-holder with Claire's design and printed version.}
 \label{fig:3d-holder}
@@
 \label{sec:overall-system}
-So far, we have described our choice separately, in this subsection we focus on our system architecture that stitches the components together. 
-Figure 2 is our overall architecture photo. Because we need to take command from Google Glass, we have used another microcontroller that speaks Bluetooth to communicate with the Google Glass.
+So far we have been discussing separate components that fit our design considerations. To stitch them into a single system, we use the following architecture shown in Figure~\ref{fig:architecture}.
 
-%% show the figure here
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.9\columnwidth]{figures/architecture.pdf}
+\caption{In our system architecture, targeting is captured by IR signal but the confirmation and coordination is over 802.15.4. Google Glass communicates with the system through Bluetooth.}
+\label{fig:architecture}
+\end{figure}
+
+\ben{this paragraph needs refinement. Maybe a good idea is to use one single sentence to summarize the disambiguation and say we will talk about it in the next section.}
+In our system, the microcontroller at the Google Glass side (we will call it the master) takes charge of bridging the end nodes and the Glass.
+It constantly transmits IR signal no specific connection with any node has been established.
+When the node receives such signal, they validate the packet and send back IR intensity readings.
+The master may simply use the IR intensity changes to determine which is the intended targets, or pass the information to Google Glasss where the Glass combines both motion sensor readings with the IR signal intensity to disambiguate.
+We will leave the algorithm description in the next section.
+Once a decision has been made, such information will be sent back to the node and the visual feedback is provided to the user.
+The user is free to adjust during the targeting process, our algorithm will dynamically adjust to the user's head movement.
+Once the user confirms the targeting, he could easily tap the touchpad.
+Single tap only enables highlighting, targeting confirmation is done with a second tap.
+This is the last resort for the user to adjust targets.
+Since the disambiguation matters to the overall performance, we will focus on the discussion in the next section.
 

### 2014-04-07 doc/uist14/evaluation.tex
trying to modify evaluation, not clear to me now

@@
 
+\ben{still thinking how to proceed.}
 In this section, we evaluate our targeting system by benchmarking its performance with an user study.
@@
 And during the study, we video-taped the participants' behavior and measure each pieces of the time.
+The experiment setup is similar to a Fitts' Law target acquisition task.
+We highlight one of the targets and ask the user to make the selection.
 
@@
 \subsection{IR intensity based}
+We evaluate how much gain we can get by performing an automatated disambiguation if the IR intensity reading is available. We compare against the case where no IR intensity is available.
 

### 2014-04-10 server/serial_monitor.py
change serial monitor for osc commands

@@ -77,8 +77,6 @@ while True:
     sys.stdout.write(line)
     if len(line.strip()) > 0 and is_osc_on:
       try:
-        osc_send(line[0:2], int(line[3:])/1024.0)
+        osc_send(line[0], int(line[2:])/1024.0)
       except ValueError:
         pass
-
-        

### 2014-04-10 doc/uist14/abstract.tex
pass over abstract, intro, some work on related work

@@
-\bjoern{I prefer to write the abstract {\em first}, not last, as a blueprint for the rest of the paper.}
-
-Emerging head-worn computing devices can enable hands-free interactions in physical spaces. 
+Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
 %
@@
 The technique does not require an a priori map of target devices.
-However, a spatial data structure can be inferred to overcome sensor saturation problems in environments hostile to IR communication (e.g., in sunlight).
-\bjoern{speculation, if we get to it}
+However, a spatial data structure can be inferred to aid ordering of items during manual disambiguation.
 %

### 2014-04-10 doc/uist14/relatedwork.tex
pass over abstract, intro, some work on related work

@@
-\section{Related Work}
+%!TEX root = uist14.tex
+\section{Background and Related Work}
+\subsection{Head and Eye Input}
+Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assitive input technology for cursor control of desktop applications~\cite{}.
+Head joysticks have been evaluated using Fitts' law and \bjoern{what did they find?} However,
+Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse)~\cite{Card:1991:MAD:123078.128726}.
+
+Contrast: we're not working on GUI tasks, but on identifyign and selecting ubicomp objects in physical spaces.
+
+Gaze control of existing graphical user intefaces~\cite{kumar2007eyepoint}.
+There are also wearable gaze trackers~\cite{SMI} -- but while these can provide a concrete point in space where a user is looking, turning this information into a selection information requires either a fixed absolute map of each space with known location and orientation of the user, or...
+Our system works through point-to-point IR communication and does not require any a priori knowledge of where objects are located.
+
+Our work is closer in spirit to Selker's headworn system~\cite{Selker:2001:EGE:634067.634176}.
+
+\subsection{Area Cursors}
+A central hypothesis of this paper is that the area selection paradigm is well matched to head orientation input.
+\bjoern{Need a discussion of area cursors~\cite{kabbash1995prince,worden1997making,Findlater-uist2010}.}
+
+\subsection{Pointing in physical spaces}
+\bjoern{take this from prior CHI submission.}
+
 \bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
-\bjoern{Need a discussion of area cursors~\cite{kabbash1995prince,worden1997making}.}

### 2014-04-11 server/fitts.py
fitts tasks

@@ -1,0 +1,92 @@
+## this is the automatic script to perform Fitts' selection
+import serial, glob, argparse, Queue
+import sys, select, random, threading
+from datetime import datetime
+from random import shuffle
+
+## standard helper for this script
+parser = argparse.ArgumentParser(description='Fitts\' Selection script')
+parser.add_argument('--baud', type=int, action='store', default=9600, help='Specify the baud rate')
+parser.add_argument('--timeout', type=float, action='store', default=1, help='Timeout parameter for serial connection')
+parser.add_argument('--serial', action='store', default=None, help='serial port')
+
+availables = glob.glob('/dev/tty.*')
+
+arguments = parser.parse_args()
+try:
+  baud = arguments.baud
+  timeout = arguments.timeout
+  serial_port = arguments.serial
+  print "  baud rate:", baud
+  print "  serial timeout:", timeout
+  print "  serial port:", serial_port
+except Exception as e:
+  print e
+  exit(1)
+
+if serial_port:
+  selected = serial_port
+else:
+  print "All available ports:" 
+  index = 1
+  for port in availables:
+    print '  ', index, port
+    index += 1
+
+  try:
+    num = int(raw_input('Enter the number:'))
+    if num > 0 and num <= len(availables):
+      selected = availables[num-1]
+  except Exception as e:
+    print "invalid input!"
+    exit(1)
+
+try:
+  ser = serial.Serial(selected, baud, timeout=timeout)
+except Exception as e:
+  print "failed to connect the serial port", e
+
+target_list = [3, 4, 5, 6, 7, 8, 11, 12, 14, 17]
+## target_list = [3, 12, 4]
+random.seed( datetime.now().second )
+
+shuffle(target_list)
+print >> sys.stderr, target_list
+
+for id in target_list[0:5]:
+  ser.write("L" + "%02d" % id)
+  goal = "C" + "%02d" % id
+  print >> sys.stderr, goal
+  start_time = datetime.now()
+  print '[', start_time.strftime('%Y-%m-%d %H:%M:%S %f'), ', Command]: ', goal
+  seconds = 0
+  line = ""
+  while goal not in line:
+    if ( ( datetime.now() - start_time).seconds > 20 ):
+      line = "D"
+      ser.write(line)
+      break
+    
+    if ( ( datetime.now() - start_time).seconds  == seconds + 1 ):
[…]

### 2014-04-11 doc/uist14/intro.tex

--- 2014-03-30 doc/uist14/intro.tex
@@
+%!TEX root = uist14.tex
 \section{Introduction}
 
+%% from the swarm vision to the necessity of selection
 Increasingly, objects in our built environment are equipped with computation and communication, making them potential targets for user interaction. 
-Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for advertising or context-aware computing \bjoern{second part needs to be stronger}.
-Both uses rely on {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.
+%Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for targeted-advertisement or context-aware computing \bjoern{second part needs to be stronger}.
+To initiate such interaction requires {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.
 
+%% previous approaches are limited
 Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}.
-While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed (which also limits use to explicit scenarios); that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at targets in the world.
-
-Emerging head-worn computing devices have the potential to overcome some of these limitations: because they are worn, they do not require retrieval; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to overlay information.
-\bjoern{needs a bridge from head-worn device to using head orientation as a feature}.
-
-In this paper, we investigate how to leverage head-orientation for target selection in physical spaces.
-We characterize this interaction modality because it can readily be implemented with small hardware changes to emerging wearable devices.
-We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose.
-We use infrared (IR) communication between Glass and target appliances to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.
-%Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio. Target appliances similarly have IR receivers and radios.
-
-Users look in the direction of a target they wish to acquire to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume), then tap to confirm selection.
-%
-Head orientation is a useful but imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus.
-Therefore, it requires area selection \bjoern{reference} combined with disambiguation techniques.
-In practice, our IR beam's selection area has a diameter of 60-120cm at distances up to 5m.
-If multiple targets fall within communication range, targets compare received IR light intensity readings to disambiguate. Users can override this default selection with an  on-screen target list.
-
-%This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target appliances while issuing control commands.
-
-While prior work has tended to focus on proofs-of-concept, we also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control.
-We first report measurements of range and beam characteristics of our controller.
-We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
-We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.
-We then quantify the additional benefit of using IR-intensity disambiguation \bjoern{and find XYZ.}
-%
-\bjoern{Say something about the inferred map if we get around to it}.
-
-We then demonstrate how the technique can be applied to a "universal remote control" scenario.
-In our implementation, user look at a target appliance and tap to load an appliance-specific control UI on the head-mounted display, which enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}) \bjoern{we may need a different figure 1 with the changed focus of the paper}.
-We also report qualitative results from users performing  home automation tasks with this system.
+While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed; that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen or keyboard); and that the user's visual attention is now split between looking down at a screen and out at targets in the world.
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/teaser.jpg}
-\caption{Using an augmented head-worn device (1), users can control smart home appliances (2) with head orientation targeting.
+\caption{An example application of head-orientation targeting: Using an augmented head-worn device (1), users can control smart home appliances (2) by looking at them.
 A near-eye display then shows an appliance control UI (3), which users navigate through multitouch gestures.}
@@
 \end{figure}
+
+%% introducing head-worn computing and head orientation
+Emerging head-worn computing devices have the potential to overcome some of these limitations: because they are worn, they do not require retrieval; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information adjacent to physical objects in the wearer's field of view.
+We thus investigate the research question how such computing devices may be used for selection of physical targets in space.
+
+Because they are head-worn, such devices can naturally exploit the user's head orientation, an important, but also imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus.
+While gaze tracking could provide further information about a user's focus, when operating in uncontrolled environments translating gaze information into identification and selection of an object is an unsolved problem.
+We therefore study how to leverage head-orientation alone for target selection in physical spaces.
+
+The imprecision of human head movement suggests adapting area cursor techniques known from assistive devices~\cite{kabbash1995prince,worden1997making,Findlater-uist2010}.
+Such techniques employ a two-step selection process: {\em coarse} selection of an area of interest, followed by {\em refinement} to select a target within that region.
+
+We propose an area-selection technique that can be readily implemented with small hardware changes to emerging wearable devices.
+We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable (IR) communication between Glass and target appliances.
+The cone shape of light emitted by an IR LED (a diameter of 60-120cm and distance up to 6m) provides the {\em coarse} selection area.
+To {\em refine} selection when multiple targets have received IR signals, we introduce three different disambiguation techniques.
+First, targets compare received IR signal intensity to make a default selection among signal recipients.
+Second, we use IR intensity changes from head movement to switch between signal recipients when a user explicitly enters a refinement quasi-mode.
+Third, we always leave the option of manually changing the current targets through explicit list navigation.
+%% three ways
+
+We also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control.
+We first report measurements of range and beam characteristics of our controller.
+We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
+We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.
+We then quantify the additional benefit of using IR-intensity disambiguation \bjoern{and find XYZ.}. \bjoern{Say something about the inferred map if we get around to it}.
+
+%---
+We also demonstrate and example application of our technique for a remote control of smart appliances: a user looks at the appliance she wishes to control and confirms selection by tapping - an appliance-specific user interface is then shown on the user's near-eye display.
+
+%Orientation-based selection enables a wide range of context-aware applications.
+Examples include smart home remote control, break reminder monitor starer, museum attention tracking, indoor positioning, etc.
+In Figure\,\ref{fig:teaser}, it's a demonstration of the ``universal remote control'' scenario.
+The user can easily select the smart appliances by simply looking at it's general direction and confirm such selection with either voice command or by tapping the Glass input pad.
+Then an appliance-specific control UI will be shown on the head-mounted display.
+For this application, we have asked 14 participants to try the system and we report the qualitative results from them performing home automation tasks.
+
+%% Contribution
+In summary, this paper makes the following contributions:
+\begin{itemize}
+\item We design and implement a novel head-orientation based selection technique for physical targets based on IR communication.
+We introduce disambiguation techniques to address the inherent imprecision of head orientation.
+\item We present evaluations that compare head orientation targeting to list selection and quantify the benefits of automatic disambiguation.
+\item We demonstrate a home appliance remote control application built on top of our selection technique.
+\end{itemize}
+
+%\ben{Having problem fixing the reference}.\bjoern{skip this - we want to be brief.}
+%In the remainder of this paper, we will first describe the related works in physical selection and targeting.
+Given that we focus our scope on head orientation, in Section~\ref{sec:background} we will briefly review human's neck ergonomics as the background for head orientation.
+In Section~\ref{sec:syst-design-prot}, we then present our system designed for the study of head orientation based selection.
+The prototype is also what we have used for building the example applications.
+The disambiguation techniques are discussed in Section~\ref{sec:disamb-techn}, which is followed by the study and evaluation in Section~\ref{sec:evaluation}.
+To further show the usefulness of having such head orientation-based selection, we describe the enabled applications in Section~\ref{sec:applications}, with detailed implementation about the ``universal remote control'' application.
+The discussion and conclusion are in Section~\ref{sec:discussion} and Section~\ref{sec:conclusion} respectively.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "uist14"
+%%% End:

### 2014-04-12 doc/uist14/intro.tex
minor changes

@@
 Because they are head-worn, such devices can naturally exploit the user's head orientation, an important, but also imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus.
-While gaze tracking could provide further information about a user's focus, when operating in uncontrolled environments translating gaze information into identification and selection of an object is an unsolved problem.
+While gaze tracking could provide further information about a user's focus, when operating in unconstrained environments, translating gaze information into identification and selection of an object is an open problem.
 We therefore study how to leverage head-orientation alone for target selection in physical spaces.

### 2014-04-12 doc/uist14/relatedwork.tex
minor changes

--- 2014-03-30 doc/uist14/relatedwork.tex
@@
-\section{Related Work}
+%!TEX root = uist14.tex
+\section{Background and Related Work}
+Our approach is related to head and eye-controlled interfaces, selection techniques, and prior work on hardware devices for pointing in physical spaces.
+\subsection{Head and Eye Input}
+Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assitive input technology for cursor control of desktop applications~\cite{radwin1990method}.
+However,
+Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}.
+This prior work often focused on used head orientation for control of graphical interfaces; in contrast, we apply this modality to selection in physical spaces, where users can also move around.
+
+Researchers have also developed methods for gaze control of existing graphical user intefaces~\cite{kumar2007eyepoint}.
+While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires either a map of each space with known location and orientation of the user.
+Our system works through point-to-point IR communication and does not require any a priori knowledge of where objects are located.
+An alternative would be to equip all objects of interest in the environment with cameras that watch the user~\cite{smith2013gaze}, though such an approach would be computationally expensive and may not work at greater distances.
+%Our work is closer in spirit to Selker's headworn system~\cite{Selker:2001:EGE:634067.634176}.
+
+\subsection{Area Cursors}
+A central hypothesis of this paper is that the area selection paradigm is well matched to head orientation input. In area selection cursors in graphical user interfaces, users do X and then Y.
+\bjoern{Need a discussion of area cursors~\cite{kabbash1995prince,worden1997making,Findlater-uist2010}.}
+
+\subsection{Pointing in physical spaces}
+\bjoern{take this from prior CHI submission.}
+
 \bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
-\bjoern{Need a discussion of area cursors~\cite{kabbash1995prince,worden1997making}.}

### 2014-04-12 doc/uist14/intro.tex
finished pass over intro and related work; folded background into rel. work

@@
 
-The imprecision of human head movement suggests adapting area cursor techniques known from assistive devices~\cite{kabbash1995prince,worden1997making,Findlater-uist2010}.
+The imprecision of human head movement suggests adapting area cursor techniques known from assistive devices~\cite{kabbash1995prince,worden1997making,findlater2010enhanced}.
 Such techniques employ a two-step selection process: {\em coarse} selection of an area of interest, followed by {\em refinement} to select a target within that region.

### 2014-04-12 doc/uist14/relatedwork.tex

--- 2014-03-30 doc/uist14/relatedwork.tex
@@
-\section{Related Work}
+%!TEX root = uist14.tex
+\section{Background and Related Work}
+Our approach is related to head and eye-controlled interfaces, selection techniques, and prior work on hardware devices for pointing in physical spaces.
+\subsection{Head and Gaze Input}
+Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assitive input technology for cursor control of desktop applications~\cite{radwin1990method}.
+However, human neck muscles have a much lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
+%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}.
+Prior work often focused on used head orientation for control of graphical interfaces; in contrast, we apply this modality to selection in physical spaces.
+
+Gaze can also be used to control graphical user intefaces~\cite{kumar2007eyepoint}.
+While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map of each space with known target locations.
+Our system works through point-to-point IR communication and does not require any a priori map.
+Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}.
+Such an approach can enable similar benefits as our approach, but is computationally more expensive than our single-pixel sensor solution and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.
+%Our work is closer in spirit to Selker's headworn system~\cite{Selker:2001:EGE:634067.634176}.
+
+\subsection{Area Cursors}
+A central hypothesis of this paper is that the area selection paradigm is well matched to head orientation input.
+In GUI area cursors, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}.
+Area cursers are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}.
+We argue that head orientation pointing has similar challenges (limited pointing performance and accuracy).
+In all area cursors, the bigger cursor activation area can lead to multiple targets being selected and disambiguation is needed.
+We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, users move so the activation area intersects with the target object (and possibly other, unintended targets).
+In the {\em refinement} phase, they adjust so only the intended target will be selected.
+Many disambiguation techniques are possible for refinement.
+We contribute a novel physical method based on received light IR intensity.
+
+\subsection{Pointing in physical spaces}
+Rukzio studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.
+
+Several approaches to spatial selection with handheld devices exist to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or to exchange information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}.
+In some techniques, users select objects of interest with laser pointers.
+The laser dot provides immediate visual feedback to the user what is being selected; however, their small target area make them poorly matched to head orientation input.
+Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.
+
+Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
+Our solution is similar in spirit but relies on simple, single low-cost IR emitters and detectors.
+%However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.
+
 \bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
-\bjoern{Need a discussion of area cursors~\cite{kabbash1995prince,worden1997making}.}

### 2014-04-12 doc/uist14/system.tex
finished pass over intro and related work; folded background into rel. work

@@
+%!TEX root = uist14.tex
+
+\bjoern{The design and prototyping sections are very thorough, but too long.
+If we want to submit a shorter paper we need a more concise description.
+We generally don't need to describe the sequence of steps we took to prototype something, just the final solution and the motivation for each decision.
+If I wanted to describe the whole system on one page, i'd probably structure it as follows:}
+\section{Interaction technique description}
+One paragraph about how the technique works from the user's perspective.
+\section{Implementation: IR-based Head-Orientation}
+\subsection{Head-worn emitter}
+Our selection technique adds an IR emitter to a wearable device frame in line with the wearer's saggital (font-to-back) plane (see Figure).
+IR is a suitable technology because it is directional.
+The angle of emittance leads to a cone-shaped area coverage.
+Many different emitter angles are commercially manufactured; we use $x^circ$, which yields a selection area of diameter Y at distance Z (part number XXXX).
+We attach the IR emitter to the frame with a repositionable 3D-printed mount that can account for individual differences in head shape (users adjust the mount so the emitter points directly forward, level to the ground.)
+
+\subsection{Target receivers}
+Each selectable target is equipped with IR receivers.
+We are interested in sensing two things: 1) IR communication and 2) IR intensity.
+Common IR receivers have automatic gain control built in and hide received signal strength.
+We therefore add an IR intensity sensor.
+In a commercial application, a single sensor that both decodes and reports RSSI could be designed.
+\subsubsection{Visual feedback}
+...
+\subsection{Wireless communication}
+(describe that we only use IR initially, then jump to RF-based communication.)
+
 \section{System Design and Prototyping}

### 2014-04-12 data/201404.Studies/analyze_fitts.py
update data analaysis -- script and data cleaning

@@ -1,0 +1,92 @@
+#!/usr/bin/env python
+## This file processes the fitts law data
+
+from os import listdir
+from os.path import isfile, join
+from datetime import datetime
+import re, argparse
+
+parser = argparse.ArgumentParser(description='Fitts study analysis script')
+parser.add_argument('-d', action='store', dest='dir_name', default=".", help='specify the directory location')
+parser.add_argument('-a', action='store', dest='action', default="", help='actions to do: check')
+
+arguments = parser.parse_args()
+directory = arguments.dir_name
+action = arguments.action
+
+
+class bcolors:
+  HEADER = '\033[95m'
+  OKBLUE = '\033[94m'
+  OKGREEN = '\033[92m'
+  WARNING = '\033[93m'
+  FAIL = '\033[91m'
+  ENDC = '\033[0m'
+
+
+
+def check_files(directory):
+  ''' this function checks if the file logs in a specific directory have the right format'''
+  filenames = map(lambda x: directory + '/' + str(x), range(1, 60))
+  for f in filenames:
+    with open(f) as f_handler:
+      for line in f_handler:
+        if line[0] != '[' and line[0] != ' ':
+          print "file:", f, " formatting is wrong"
+          break
+
+def process_file(filename, tag):
+  target_id, target_time, search_time, acquire_time = None, None, None, None
+  with open(filename) as f:
+    for line in f:
+      if not target_id:
+        m = re.search(r'C(\d{2})', line)
+        if m:
+          ## get target id
+          target_id = m.group()[1:3]
+          target_time = datetime.strptime(line[2:28], '%Y-%m-%d %H:%M:%S %f')
+
+      if target_id and not search_time:
+        m = re.search(target_id + '[uid]', line)
+        if m:
+          ## first time see the target being hovered
+          tn = datetime.strptime(line[2:28], '%Y-%m-%d %H:%M:%S %f')
+          search_time = (tn - target_time).total_seconds() - 1
+
+      if target_id and search_time and not acquire_time:
+        m = re.search('C' + target_id, line)
+        if m:
+          ## acquiring the target
+          tn = datetime.strptime(line[2:28], '%Y-%m-%d %H:%M:%S %f')
+          acquire_time = (tn - target_time).total_seconds() - 1
+
+  if target_id == None:
+    print bcolors.WARNING + " target not found" + bcolors.ENDC
+    
+  if not search_time:
+    search_time = "NA"
+  if not acquire_time:
[…]

### 2014-04-12 data/201404.Studies/data_summary.R

@@ -1,0 +1,4 @@
+d <- read.table("aggregate.txt")
+
+summary(d[d$V5=="intensity", 3:4])
+summary(d[d$V5=="name", 3:4])

### 2014-04-14 doc/uist14/abstract.tex
structural layout as discussed on sunday - iterative development

@@
+%!TEX root = uist14.tex
 Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
@@
 We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.
-The fundamental imprecision of head orientation requires area selection techniques, which can lead to ambiguous selections.
-We introduce a two level disambiguation strategy: first, networked receivers use IR intensity sensors to estimate which target is the intended selection; second, users can fall back to a manual disambiguation technique using their near-eye display.
-The technique does not require an a priori map of target devices.
-However, a spatial data structure can be inferred to aid ordering of items during manual disambiguation.
-%
 A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device.
-Intensity-based target disambiguation reduces selection time and error rate in dense target environments.
-\bjoern{rewrite once we have results}.
-We implement this technique in an application for controlling smart appliances and report qualitative user impressions.
+The fundamental imprecision of head orientation requires disambiguation techniques when the intended target is unclear.
+To improve on the slow manual disambiguation of our first design, we introduce a second design in which receivers use IR intensity sensors to estimate which target is the intended selection.
+A second study demonstrates that using IR intensity successfully reduces ambiguity and enhances selection speed.
+Finally, we describe how a spatial data structure can be inferred to enable a natural mapping of head direction during the disambiguation step and we report first use data.
+We demonstrate an application of our selection technique for controlling smart appliances and report qualitative user impressions.

### 2014-04-14 doc/uist14/applications.tex
structural layout as discussed on sunday - iterative development

@@
-\section{Applications}
+%!TEX root = uist14.tex
+\section{Application Example: Universal Remote Control}
 \label{sec:applications}
 
-In this section we describe four possible applications that can be built on top of the head orientation-based selection.
+%In this section we describe four possible applications that can be built on top of the head orientation-based selection.
 We focus our discussion on {\em GlassRemote} -- a universal remote control application.
 
-\subsection{GlassRemote}
-\label{sec:univ-remote-contr}
+%\subsection{GlassRemote}
+%\label{sec:univ-remote-contr}
 
-We extended the head-orientation targeting and existing hardware platform to achieve a universal remote control functionality in smart home scenario (we have a prior study archived in \cite{Chen:EECS-2013-200}).
+We extended the head-orientation targeting and existing hardware platform to achieve a universal remote control functionality in smart home scenario.
 The motivation is to solve two challenges that exist in current list-based solutions -- {\em naming} and {\em scoping}.
@@
 Our head-orientation based interaction elegently solves these two problems.
+
+\bjoern{you need to tell us more about how the interaction works - namely, that you pull the user interface for an appliance onto the near-eye display. This may not be obvious.}
 
@@
 
-\subsection{Attention Tracking}
-\label{sec:attention-tracking}
+\bjoern{since length will be an issue, let's just focus on glass remote and skip the others?}
+Additional applications such as attention tracking or indoor position and orientation estimation of a wearer based on our technique are conceivable; we leave them to future work.
 
-The second tier of applications enabled by our system is an implicit use of users' head orientation.
+%\subsection{Attention Tracking}
+%\label{sec:attention-tracking}
+
+%The second tier of applications enabled by our system is an implicit use of users' head orientation.
 One sample application is to obtain a personal statistics.
@@
 % citation needed, take a look at this one: E-Readers and Visual Fatigue, http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3873942/
-Knowing how much time you have spent in front of the screen is definitely useful, and the application can be built to remind the user to take a break whenever a certain time threshold is reached.
+%Knowing how much time you have spent in front of the screen is definitely useful, and the application can be built to remind the user to take a break whenever a certain time threshold is reached.
 
 %% We might be lacking evidence, but let me just write down the draft first.
-Our system enables a flexible and easy-to-use tracking of users' attention without knowing the map.
+%Our system enables a flexible and easy-to-use tracking of users' attention without knowing the map.
 Therefore, it's suitable to be deployed in large museums (even when the exihibits positions are adjusted constantly).
@@
 Previous solutions are either using RFID \cite{Hsi:2005:REV:1081992.1082021} or video taping % this is a cv of Adam who has worked on video analysis http://www.exploratorium.edu/vre/cvs/ak_cv.pdf.
-However, head orientation is a better approximation of attention rather than proximity.
+%However, head orientation is a better approximation of attention rather than proximity.
 By placing our tracking devices in the environment, we can easily get a more accurate profiling of users intention.
 
-\subsection{Indoor Positioning}
-\label{sec:indoor-positioning}
+%\subsection{Indoor Positioning}
+%\label{sec:indoor-positioning}
 
-There is one category of indoor positioning which is based on angle of arrival.
+%There is one category of indoor positioning which is based on angle of arrival.
 Traditional approaches on indoor positioning are based on fingerprinting or triangulation, which is known to be bad due to multi-path effect for indoor environment.

### 2014-04-14 doc/uist14/intro.tex
structural layout as discussed on sunday - iterative development

@@
 Because they are head-worn, such devices can naturally exploit the user's head orientation, an important, but also imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus.
-While gaze tracking could provide further information about a user's focus, when operating in unconstrained environments, translating gaze information into identification and selection of an object is an open problem.
+While gaze tracking could provide further information about a user's focus, wearable gaze tracking cannot by itself reliably determine the identity of objects in focus.
+Equipping all devices with stationary gaze trackers is possible~\cite{vertegaal2005media} but much costlier than our approach.
 We therefore study how to leverage head-orientation alone for target selection in physical spaces.
@@
 
-We propose an area-selection technique that can be readily implemented with small hardware changes to emerging wearable devices.
+In this paper, we describe the iterative development and evaluation of an area-selection technique that can be readily implemented with small hardware changes to emerging wearable devices.
 We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable (IR) communication between Glass and target appliances.
 The cone shape of light emitted by an IR LED (a diameter of 60-120cm and distance up to 6m) provides the {\em coarse} selection area.
-To {\em refine} selection when multiple targets have received IR signals, we introduce three different disambiguation techniques.
-First, targets compare received IR signal intensity to make a default selection among signal recipients.
-Second, we use IR intensity changes from head movement to switch between signal recipients when a user explicitly enters a refinement quasi-mode.
-Third, we always leave the option of manually changing the current targets through explicit list navigation.
-%% three ways
-
-We also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control.
-We first report measurements of range and beam characteristics of our controller.
+To {\em refine} selection when multiple targets have received IR signals, we first introduce a manual technique where users select a candidate device using a graphical interface on their wearable device.
 We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
 We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.
-We then quantify the additional benefit of using IR-intensity disambiguation \bjoern{and find XYZ.}. \bjoern{Say something about the inferred map if we get around to it}.
 
+Noting that the refinement UI significantly slowed interaction times in this study, we contribute an improved refinement technique in which target objects compare received IR signal intensity.
+This intensity is both used to make a default selection, and to order the fallback manual UI list\bjoern{rephrase}.
+A second study with $n$ participants shows that IR-intensity disambiguation successfully reduces list navigation for refinement and speeds up selection when compared to the naive technique from the first study.
+
+A shortcoming discovered in the second study is that participants dislike switching between focusing on targets in the world and graphics on their near-eye display during the selection process.
+We solve this problem by inferring a spatial data structure of target locations using Glass' orientation sensors and allow small head tilt motions to move selections during the refinement stage (e.g., nod down to select next target down, turn right to select next target on the right).
+%% three ways
 %---
@@
 %% Contribution
-In summary, this paper makes the following contributions:
-\begin{itemize}
-\item We design and implement a novel head-orientation based selection technique for physical targets based on IR communication.
+%In summary, this paper makes the following contributions:
+%\begin{itemize}
+%\item We design and implement a novel head-orientation based selection technique for physical targets based on IR communication.
 We introduce disambiguation techniques to address the inherent imprecision of head orientation.
-\item We present evaluations that compare head orientation targeting to list selection and quantify the benefits of automatic disambiguation.
-\item We demonstrate a home appliance remote control application built on top of our selection technique.
-\end{itemize}
+%\item We present evaluations that compare head orientation targeting to list selection and quantify the benefits of automatic disambiguation.
+%\item We demonstrate a home appliance remote control application built on top of our selection technique.
+%\end{itemize}
 

### 2014-04-14 doc/uist14/relatedwork.tex
structural layout as discussed on sunday - iterative development

@@
 \section{Background and Related Work}
-Our approach is related to head and eye-controlled interfaces, selection techniques, and prior work on hardware devices for pointing in physical spaces.
+Our approach is related to head and eye-controlled interfaces, area cursors, and prior work on hardware devices for pointing in physical spaces.
 \subsection{Head and Gaze Input}
@@
 In GUI area cursors, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}.
-Area cursers are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}.
+Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}.
 We argue that head orientation pointing has similar challenges (limited pointing performance and accuracy).

### 2014-04-14 doc/uist14/study1.tex
structural layout as discussed on sunday - iterative development

--- 2013-09-18 doc/sigchi14/targeting.tex
@@
-\section{Physical Target Acquisition Study}
+%!TEX root = uist14.tex
+\section{Study 1: Physical Target Acquisition Study}
 To understand the accuracy and performance of head-orientation-based selection through our device, we carried out a comparative target acquisition study, where participants had to connect to wireless nodes distributed in a room with our technique, and with an alternate list selection approach.
+
+\subsection{Apparatus}
+In an indoor environment, 10 wireless nodes are spread across a room (Figure~\ref{fig:targeting-study-layout}).
+Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs %(Figure~\ref{fig:target}).
+%An yellow LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple targets are within IR range.
+Next to each target, a paper sheet shows a number and letter combination, which is used for uniquely identifying the device.
+The numbers are the primary identifiers, ordered from left to right in the room.
 
@@
 \end{figure}
-
-\subsection{Apparatus}
-In an indoor environment, 10 wireless nodes are spread across a room at various heights and distances (Figure~\ref{fig:targeting-study-layout}).
-The nodes are stand-ins for potential smart appliances and have all relevant functionality for targeting and wireless communication, but do not control any actual appliances.
-Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs (Figure~\ref{fig:target}).
-An yellow LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple targets are within IR range.
-Next to each target, a paper sheet shows a number and letter combination, which is used for uniquely identifying the device.
-The numbers are the primary identifiers, ordered from left to right in the room.
-This ordering makes it easy to locate them, which simulates looking towards an appliance with a well-known location in a room, and minimizes visual search time.
-
-\begin{figure}[b]
-\centering
-\includegraphics[width=0.8\columnwidth]{figures/study-node.pdf}
-\caption{An example node from the targeting study --- we constructed 10 such nodes - each mounted in a box.}
-\label{fig:target}
-\end{figure}
+%\begin{figure}[b]
+%\centering
+%\includegraphics[width=0.8\columnwidth]{figures/study-node.pdf}
+%\caption{An example node from the targeting study --- we constructed 10 such nodes - each mounted in a box.}
+%\label{fig:target}
+%\end{figure}
 
@@
 Participants' selection errors do occur in both conditions.
-However, error rates were low (1.1\% in {\em infrared} mode, and 2.9\% in {\em list} mode, repspectively).
+However, error rates were low (1.1\% in {\em infrared} mode, and 2.9\% in {\em list} mode, respectively).
 This precludes us from running a more detailed analysis.
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/target-likert.pdf}
-\caption{Likert scale ratings for ease of use of aspects of the targeting task. Error bars show standard error. }
-\label{fig:target-likert}
-\end{figure}
+%\subsubsection{Preference}
 
-\subsubsection{Preference}
-
-Eleven of 14 users preferred infrared mode over list mode (three preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (see Figure~\ref{fig:target-likert}).
+%Eleven of 14 users preferred infrared mode over list mode (three preferred list, one was undecided).
+While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome.
 As self-report data can easily skew positive as participants try to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.
 
-List mode had certain advantages: It was judged to be more accurate and predictable as there was always exactly one device selected in the list (\studyquote{With the list you never have to worry about accidentally picking up two targets}).
+%List mode had certain advantages: It was judged to be more accurate and predictable as there was always exactly one device selected in the list (\studyquote{With the list you never have to worry about accidentally picking up two targets}).
 Also, it did not require a clear line of sight to the target device so participants did not have to move from their starting position (\studyquote{The shortcoming of the IR mode was that you had to be a certain distance away in order for it to detect the appliance}).
 
-On the other hand, list mode was judged to be more \studyquote{annoying} and tedious.
+%On the other hand, list mode was judged to be more \studyquote{annoying} and tedious.
 The temple-based touchpad for selection was difficult to use for a participant with long hair: \studyquote{List mode was physically difficult for me to navigate, since my long hair wasn't tied back and it kept interfering with my swiping.} Another participant also commented on the ergonomic challenge of touchpad use on Glass: \studyquote{The strength of the IR mode was that I didn't have to use my fingers as much to control.
@@
 
-One noted benefit of infrared mode was a feeling that it was \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
+%One noted benefit of infrared mode was a feeling that it was \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
 One subject called it \studyquote{natural to interact with things just by looking at them}.
@@
 
-A few perceived weaknesses of infrared mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target.
+%A few perceived weaknesses of infrared mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target.
 One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}.
@@
 Users had to learn the usable angle of the IR emitter before they became successful at controlling the devices: \studyquote{I had to compensate by tilting my head up a little bit.}
-
-%IR mode has semantic meaning (like I said).
-I'm likely to want visual feedback from any distant device I'm trying to control, and since I'm already looking at it, this is very easy to get.
-List mode was physically difficult for me to navigate, since my long hair wasn't tied back and it kept interfering with my swiping.
-Another comment about list mode was that swiping faster didn't seem to navigate the list faster, which would have been convenient for accessing items like W and Y very far from the beginning of the list.
-This could also be partially overcome by using a circular linked list (where the list wraps from Y to A) instead of a one-way list.
-%With the list you never have to worry about accidentally picking up two targets, but since you always have to swipe through a (longer) list that's not much of an advantage.
-%List mode felt like I could be more confident in what client I was choosing because I couldn't select more than one client.
-The IR mode was a little more natural because it involved less clicking and navigating, but it felt easier to select the wrong client which made the list mode more comfortable.
-%"The shortcoming of the List mode was that the list didn't wrap around. So for example if I wanted to go to Y, I couldn't just go left from A—I had to scroll through all the letters in between.
-
-%The strength of the list mode was that it's pretty accurate, so as long as I swiped the correct amount, I could activate and power off the appliances.
-I also didn't have to move at all, although perhaps that is a shortcoming because otherwise we might just all turn into Wall-E people.
-
-%The strength of the IR mode was that I didn't have to use my fingers as much to control. If the items were spaced relatively far apart, it was easy to select a specific appliance.
-
-%The shortcoming of the IR mode was that you had to be a certain distance away in order for it to detect the appliance. "
-%The List Mode is easier to choose the desired item when the item list is short.
-It became a bit frustrated when I had to swipe many times in order to get to the targeted item.
-The IR mode works more intuitive in this case, but need a longer time to learn and get used to it.
-%list mode is easier to select what you want, but may take a while to get to since you can only swipe through one at a time; IR is quicker and more intuitive, but unstable
-%"list mode is annoying. i find myself counting how many swipes i have to do.
-
-%IR mode is fun and pretty convenient"

### 2014-04-14 doc/uist14/system1.tex
structural layout as discussed on sunday - iterative development

--- 2014-02-24 doc/uist14/submitted-for-review/submitted.tex
@@
-\documentclass{sigchi}
-
-\toappear{Submitted to CHI2014. Do not cite or distribute.}
-
-\pagenumbering{arabic}
-
-\usepackage{balance}  
-\usepackage{graphics} 
-\usepackage{times}    
-\usepackage{url}      
-\usepackage{gensymb}
-\usepackage{afterpage}
-\usepackage{mathpazo,amsmath}
-\def\inch#1{#1''}
-\def\ft#1{#1'\thinspace}
-
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
-\makeatother
-\urlstyle{leo}
-
-\def\pprw{8.5in}
-\def\pprh{11in}
-\special{papersize=\pprw,\pprh}
-\setlength{\paperwidth}{\pprw}
-\setlength{\paperheight}{\pprh}
-\setlength{\pdfpagewidth}{\pprw}
-\setlength{\pdfpageheight}{\pprh}
-
-\usepackage[pdftex]{hyperref}
-\hypersetup{
-pdftitle={SIGCHI Conference Proceedings Format},
-pdfauthor={LaTeX},
-pdfkeywords={SIGCHI, proceedings, archival format},
-bookmarksnumbered,
-pdfstartview={FitH},
-colorlinks,
-citecolor=black,
-filecolor=black,
-linkcolor=black,
-urlcolor=black,
-breaklinks=true,
-}
-
-\newcommand\tabhead[1]{\small\textbf{#1}}
-
-\newcommand {\bjoern}[1]{{\color{red}\bf{BH: #1}\normalfont}}
-\newcommand {\ben}[1]{{\color{blue}\bf{BZ: #1}\normalfont}}
-\newcommand {\sean}[1]{{\color{blue}\bf{SC: #1}\normalfont}}
-\newcommand {\claire}[1]{{\color{blue}\bf{CT: #1}\normalfont}}
-\newcommand {\studyquote}[1]{\em ``#1''\normalfont}
-
-\begin{document}
-
-\title{A Context Menu for the Real World: Controlling Physical Appliances Through Head-Worn Infrared Targeting}
-
-\numberofauthors{1}
-\author{
- \alignauthor Anonymous for submission
-}
-
-\maketitle
-
-\begin{abstract}
- % abstract
-We introduce a novel method for selecting and controlling smart appliances in physical spaces through a head-worn computing device with near-eye display and wireless communication.
-We augment a commercial wearable computing device, Google Glass, with a narrow-beam IR emitter for this purpose.
-This configuration yields a usable beam width of 2 to 4 feet (60 to 120cm) for targeting at room scale.
-We describe a disambiguation technique if infrared targeting hits multiple targets simultaneously.
-A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device.
-We also report qualitative data from using our device to control multiple appliances in a smart home scenario.
-
-\end{abstract}
-
-\keywords{
-	smart devices; universal remote control; wearable computing; glass
-}
-
-\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}
-
- % introduction
-\section{Introduction}
-Increasingly, devices and services in our built environment are networked and can be controlled remotely.
-The proliferation of smart, controllable devices such as intelligent lighting, AV equipment, HVAC systems, or kitchen appliances raises the question of how to best interact with them.
-
-Today, commercial solutions (such as Belkin WeMo) use handheld mobile devices as {\em universal remote controls} to control such appliances.
-In these solutions, users first browse a list of all available devices and then call up a device-specific user interface.
-This method faces two challenges: {\em naming} and {\em scoping}.
-Assigning clear names is non-trivial.
-In shared spaces, the person trying to control the device might not be the one that named it - e.g., while an office building manager may know what ``Light 4 in area E'' corresponds to, an occupant may not.
-Second, without a method of scoping selection to automatically filter non-relevant devices, paging though long lists of names or navigating hierarchies becomes potentially more cumbersome than the physical action the ``convenient'' software solution was meant to replace.
-
-To address these challenges, research has introduced techniques of augmenting mobile devices with accessories like laser pointers to enable direct aiming at target devices~\cite{beigl_point_1999,patel_2-way_2003}.
-While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at the device to-be-controlled.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/teaser.jpg}
-\caption{Using an augmented head-worn device (1), users can control smart home appliances (2) with head orientation targeting.
-A near-eye display then shows an appliance control UI (3), which users navigate through multitouch gestures.}
-\label{fig:teaser}
-\end{figure}
-
-In this paper, we introduce a novel method for selecting and controlling smart appliances in physical spaces through the use of a head-worn computing device with near-eye display and wireless communication.
-We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose.
-Users first look in the direction of the appliance they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume).
-If multiple appliances fall within communication range, a disambiguation technique that combines on-screen information as well as visual feedback on the target appliances lets users select their desired target.
-Once acquired, an appliance specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (Figure~\ref{fig:teaser}).
-
-Our hardware relies on infrared (IR) communication between Glass and target appliances to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.
-Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio.
-Target appliances similarly have IR receivers and radios.
-This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target appliances while issuing control commands.
-
-While prior work has tended to focus on proofs-of-concept, we also contribute empirical data on the system performance, usability, and user experience of head-orientation targeting and device control.
-We first report measurements of range and beam characteristics of our controller.
-We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
-We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.
-We also report qualitative results from participants who use our system for home automation tasks.
-
- % relatedwork
-\section{Related Work}
-Relevant prior work exists in the areas of remote control of physical appliances, evaluations of pointing in physical space and augmented reality applications. We discuss each in turn.
-
-\subsection{Remote Control of Physical Appliances}
-Standard infrared remote controls for televisions and AV equipment are only meant to control a single device.
-These controllers tend to use wide-angle infrared LEDs.
-Universal remote controls are available as dedicated devices (e.g, Crestron\footnote{\url{http://www.crestron.com}}) or applications for smart phones (e.g., Belkin WeMo\footnote{\url{http://www.belkin.com/us/wemo}}).
-They do not offer spatial selection of target devices, forcing users to browse through lists of pre-configured devices instead.
-Rukzio found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.
-
-Several approaches to spatial selection with handheld devices exist to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012} or to exchange information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}.
-Key design decisions are the method by which a target device is selected; and the method by which it is then later controlled or configured.
-
-In several techniques, users select objects of interest with laser pointers.
-The laser dot provides immediate visual feedback to the user what is being selected (though it does not indicate whether the pointed-at object can indeed be controlled).
-Furthermore, laser pointer becomes obtrusive when there are other people in the space.
-Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}.
-Patel extends this technique by modulating the laser light to communicate the controllers' identity~\cite{patel_2-way_2003} to initiate radio communication.
-These proofs-of-concept do not include thorough evaluations.
-Kemp et al.
-use a laser pointer to indicate to robots which item to pick up in a room~\cite{kemp_point-and-click_2008}.
-
-The XWand~\cite{wilson_xwand:_2003} determines its absolute position and orientation and uses a virtual room model to select target devices.
-Position is determined through two ceiling-mounted cameras; orientation is determined using a built-in IMU.
-Users can employ physical gestures or utter speech commands to control selected devices.
-This technique requires room instrumentation and an up-to-date virtual model of device locations.
-The Tricorder~\cite{lifton_tricorder:_2007} uses IMU orientation coupled with room-localization based on received signal strength indicators (RSSI) to estimate what a user is pointing at.
-
-Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
-Printed tags like QR codes can also be affixed to devices and read by cameras.
-Common tagging systems are optimized to be read from a close distance, though it is possible to redesign codes that can be read further away (by encoding less information)~\cite{cross_low-cost_2012}.
-
-Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing --- the user looks at the target device to initiate interaction.
-Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications.
-We therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.
-
-\subsection{Evaluation of room-scale selection}
-Pausch et al.'s early investigation of head-mounted displays compared head-tracking to handheld orientation control for a target acquisition task in a virtual reality room shown on a head-mounted display~\cite{pausch_user_1993}.
-They found a clear performance benefit for head-tracking.
-
-On the other hand, Card et al.
-experimentally determined that the bandwidth of neck muscles is much lower than that of arm, wrist or finger muscle groups~\cite{card_morphological_1991}, which limits the performance of any head orientation-based interaction scheme.
-However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques.
-Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy.
-Our work only uses head orientation for the initial selection step; since we believe a user's attention is often drawn to the objects they intend to interact with.
-
-Myers et al compared different methods of interacting with displays at a distance~\cite{myers_interacting_2002} and quantified selection time and jitter or position error when using remote handheld pointing.
-Various techniques outperformed laser pointers.
-
-Our work is complementary as it provides concrete performance data on using head orientation to select targets in a physical environment.
-
-\subsection{Augmented Reality Interfaces}
-Augmented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
-Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world, though our techniques could be extended to such configurations.
-
- % interaction
-\section{Initiating Interaction through Attention}
-This section describes the design goals of our system and their realization in particular interaction techniques.
-
-\subsection{Design Goals}
-Our work is motivated by the following design goals that leverage opportunities of head-worn computing, but also acknowledge potential challenges:
-
-{\bf Leverage visual attention:} Take advantage of the fact that visual attention can express intention - initiate interaction based on where a user is already looking.
-
-{\bf Provide immediate feedback about selection targets in the environment:} While a near-eye display can push information to the user, users don't always want to control an object simply because they are looking at it (a problem known in gaze-based interaction as the Midas touch).
-A calmer~\cite{weiser_coming_1997} approach is to locate visual feedback about selection targets in the environment, to prevent distraction and interruption.
-Such feedback should be delivered instantaneously, while users look around a room.
-
-{\bf Offer flexible orientation after initiating interaction:} After initiating interaction through head orientation, enable the user to reorient their head or body position during the remaining interaction to prevent neck strain.
-
-{\bf Offer efficient ways to disambiguate orientation input:} It may not always be possible to identify a unique target appliance based on a user's attention and orientation.
-Offer ways to supplement orientation-based interaction with screen-based interaction to provide disambiguation information.
-
-These design goals find their expression in the following interaction model.
+%!TEX root = uist14.tex
+\section{Iteration 1: IR-based Head Orientation with List Refinement}
 
 \subsection{Interaction Flow}
+\bjoern{I pulled this from the earlier paper}
 
@@
 \label{fig:interaction}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
+\vspace{0.1in}
 \includegraphics[width=\columnwidth]{figures/stepbystep_multi_small.png}
@@
 
-{\bf Look:} Users select a target appliance by looking in its general direction.
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/selectflow01.pdf}
+\caption{Head orientation selection flowchart for our first design.}
+\label{fig:selectflow01}
+\end{figure}
+
+{\bf Look:} Users looking in the general direction of the intended target.
 Glass periodically sends a device id through its IR emitter analogous to Patel's approach~\cite{patel_2-way_2003}.
@@
 
-{\bf Control:} Glass displays a user interface for parameters of the chosen appliance.
-Upon connection, the current status of the appliance is retreived by Glass and shown on the UI.
-The interface is controlled with the temple-mounted touchpad through the following gesture set: tapping toggles discrete parameters (such as power for a lamp as Figure~\ref{fig:ui_controls}(A)); single finger swipe changes between available parameters; double finger swipe adjusts continuous parameters (such as volume for a video player as Figure~\ref{fig:ui_controls}(B)).
-This scheme was chosen because the touchpad is only comfortably operable in the coronal plane (front to back) but not in the sagittal plane (up and down).
-Control commands are sent over XBee radios.
-
-{\bf Disengagement:} Users stay connected to the last selected appliance up to a timeout period. During that period, users can disengage through down swipes.
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width=\columnwidth]{figures/ui_controls_caption.jpg}
-\caption{Two screenshots of UI controls for lamp and video player.}
-\label{fig:ui_controls}
-\end{figure}
-
-\subsection{Disambiguation}
+{\bf Refinement:}
 Head orientation only indicates a general area of visual interest.
 It does not necessarily match gaze orientation as extra-ocular muscles can move the eyes.
-The IR beam of our device also has a certain spread (see next section).
+The IR beam of our device also has a certain spread .
 In an environment dense with potential targets, multiple targets could be within range.
@@
 
- % device
-
-\section{Hardware Device}
+\subsection{Implementation}
 \begin{figure}[t]
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/glass-with-ir}
+\includegraphics[width=0.7\columnwidth]{figures/tube_from_top.jpg}
 \caption{Our augmented Glass prototype has a frame-mounted infrared emitter.}
@@
 
-\subsection{Prototype Implementation}
-Our prototype consists of a Google Glass Explorer Edition head-worn computing device, augmented with an infrared emitter that is mounted on the frame, pointing out in the direction of the wearer's view (Figure~\ref{fig:glass}).
-The IR emitter LED is mounted in an opaque hollow tube, that restricts the outgoing angle of illumination.
+Our prototype consists of a Google Glass Explorer Edition device, augmented with an infrared emitter that is mounted on the frame, pointing out in the direction of the wearer's view (Figure~\ref{fig:glass}).
+The IR emitter LED is mounted in a \bjoern{3D printed housing which can be adjusted so the emitter can point straight forward for different head shapes}.
+
+Our measurements suggest that IR communication can be targeted to an area about 60-120cm in diameter, up to 5m in front of the user.
+These values are a reasonable match for selecting targets in a room-size environment.
+A wider beam would lead to an increased chance of multiple appliances receiving IR signals simultaneously.
+A narrower beam will make targeting more challenging, given the precision constraints of human head movement.
 
@@
 This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame (Figure~\ref{fig:architecture}).
-This architecture was mostly chosen for reasons of expediency.
-We selected XBee 802.15.4 radios to avoid Bluetooth  wake up latencies but we do not claim optimality for our design decisions.
+This architecture was chosen for reasons of expediency and we do not claim optimality for our design decisions.
 Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
@@
 The receivers that correctly received the signal turn their LED on for 300 ms.
-
-We placed all three devices at the same height with clear line of sight.
-The IR emitter is first places 2 feet away from the receivers.
-The receivers were moved sideways apart from each other until they could no longer receive stable signals.
-We then recorded the distance of the two receivers for the calculation of coverage angles.
-The steps are repeated for IR emitters in different distances (as shown in Table~\ref{table:measurements}).
-We then repeated measurements with the emitter  placed at various depths in the tube (see Figure~\ref{fig:coverage}).
-
-In summary, our measurements suggest that IR communication can be targeted to an area about 2--4' in diameter, up to 16' in front of the user.
+Our measurements suggest that IR communication can be targeted to an area about 2--4' in diameter, up to 16' in front of the user.
 These values are a reasonable match for selecting appliances in a room-size environment.
@@
 A narrower beam will make targeting more challenging, given the precision constraints of human head movement.
-
-\afterpage{
-\begin{figure}[t!]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/glass-ir-coverage}
-\caption{Different IR configurations suggest usable beam widths of 2 to 4 feet and distances up to 16 feet }
-\label{fig:coverage}
-\end{figure}
-
-\begin{table}[h!]
-    \begin{tabular}{l|lllll}
-    distance/ depth & \ft2       & \ft4       & \ft8   & \ft{12}  & \ft{16}  \\ \hline
-    \inch{0}                     & $74\degree$ & $78\degree$ & N/A  & N/A  & N/A  \\
-  \inch{0.5}                   & $60\degree$ & $48\degree$     & $28\degree$ & $22\degree$ & $16\degree$ \\
-    \inch{1.0}                     & $46\degree$     & $36\degree$     & $26\degree$ & $18\degree$ & $10\degree$ \\
-    \inch{1.5}                   & $36\degree$     & $32\degree$     & $18\degree$ & $14\degree$ & $6\degree$  \\
-    \end{tabular}
-    \caption{Measured IR coverage angles $\Theta$ at different target distances and different depths of IR emitter inside shielding tube.}
-    \label{table:measurements}
-\end{table}
-}
-
- % targeting
-\section{Physical Target Acquisition Study}
-To understand the accuracy and performance of head-orientation-based  selection through our device, we carried out a comparative target acquisition study, where participants had to connect to wireless nodes distributed in a room with our technique, and with an alternate list selection approach.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
-\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment.
-Targets were called out by number; for the list mode condition, participants need to match numbers to letters.}
-\label{fig:targeting-study-layout}
-\end{figure}
-
-\subsection{Apparatus}
-In an indoor environment, 10 wireless nodes are spread across a room at various heights and distances (Figure~\ref{fig:targeting-study-layout}).
-The nodes are stand-ins for potential smart appliances and have all relevant functionality for targeting and wireless communication, but do not control any actual appliances.
-Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs (Figure~\ref{fig:target}).
-An yellow LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple targets are within IR range.
-Next to each target, a paper sheet shows a number and letter combination, which is used for uniquely identifying the device.
-The numbers are the primary identifiers, ordered from left to right in the room.
-This ordering makes it easy to locate them, which simulates looking towards an appliance with a well-known location in a room, and minimizes visual search time.
-
-\begin{figure}[b]
-\centering
-\includegraphics[width=0.8\columnwidth]{figures/study-node.pdf}
-\caption{An example node from the targeting study --- we constructed 10 such nodes - each mounted in a box.}
-\label{fig:target}
-\end{figure}
-
-\subsection{Methodology}
-In our within-subjects design, participants performed 15 target acquisition tasks each with two interaction styles.
-In the {\em infrared} mode condition, participants used our IR targeting approach; in the {\em list} mode condition, participants had to look up a device's letter code on the printed paper next to the device and then select that letter code from a list displayed on their Glass device.
-The list was navigated with swipe motions on the Glass touchpad.
-For each task, participants started at a fixed position in the room.
-The experimenter called out a number and simultaneously started a timer.
-Participants then had to find the corresponding device (by looking for its printed code).
-In the {\em infrared} mode, participants then selected and acquired the target by aiming the IR beam at the target, and confirmed their selection with a touch pad tap.
-If more than one target was within range, participants had to either use the disambiguation dialog or reposition themselves.
-In the {\em list} mode, participants had to read the letter next to the number and then select that letter by browsing a linear list shown in their Glass display.
-While the list was alphabetized, letter arrangement in the room was not.
-This design required participants to find the target in the room before starting a list navigation to keep visual search times similar in each condition.
-
-Afterwards, participants completed a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
-
-\subsection{Participants}
-We recruited 14 participants from our institution.
-13 had never used Glass before.
-4 wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
-Half of them performed {\em infrared} mode first and the other half did {\em list} mode first.
-
-\subsection{Measures}
-The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; and {\bf user preference}: which interface users preferred for the task after completing the study.
-
-\subsection{Results}
-\subsubsection{Performance data}
-The time to complete each task can be broken down into the following pieces:
-
-$t_{infrared}=t_{locate}[+t_{reorient}][+t_{disambiguate}]+t_{tap}$
-
-$t_{list}=t_{locate}+t_{listnav}+t_{tap}$
-
-In both conditions, participants first have to locate the target announced by the experimenter through visual search ($t_{locate}$).
-In the {\em infrared} mode, participants may then directly confirm their selection if only a single target was selected ($t_{tap}$).
-However, if they don't immediately receive feedback that their target was selected, or if multiple targets were selected, users either have change their position or head orientation ($t_{reorient}$) or they have to step through the on-screen disambiguation dialog ($t_{disambiguate}$).
-In the {\em list} mode, participants must scroll through the list to find the desired target identifier ($t_{listnav}$).
-Thus, {\em infrared} will show a performance benefit if $(t_{reorient}+t_{disambiguate})<t_{listnav}$.
-This depends on the number of total devices in the environment (increasing $t_{listnav}$), and their density (which will increase $t_{disambiguate}$).
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
-\caption{Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), and between IR multiple responses cases and IR single response cases (B).
-The centers of boxes are median values, while white dashed lines are mean values.}
-\label{fig:selection-times}
-\end{figure}
-
-We first show results for 10 targets and then discuss extrapolations of these results.
-Average target acquisition time $t_{infrared}$ was 6.67 seconds, while $t_{list}$ was 8.86 seconds (Figure~\ref{fig:selection-times}A).
-This difference is significant (Student's t-test, $t(279)=-3.81, p=0.00017$).
-
-To further understand the performance gain in {\em infrared} mode, especially the factor $t_{disambiguate}$, we compare selection times when multiple devices are targeted (and disambiguation is required) to single-device selection times (Figure~\ref{fig:selection-times}B).
-When there is a single device, $t_{disambiguate}$ is 0 and it takes 6.40 seconds (on average) to complete the connection.
-When multiple devices are in range, the time increases to 9.16 seconds, indicating 2.76 seconds required to disambiguate.
-Though it takes significantly longer ($t(19)=-2.7827, p=0.012$ using t-test) in the {\em multiple} case, these cases made up only 10\% of total infrared trials.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
-\caption{Times taken to select a device vs.~its order in the list.
-The dotted line is a linear fit between the mean times and device orders.
-Two horizontal lines of mean target acquisition times in {\em infrared} mode are also annotated for comparison.}
-\label{fig:time-vs-list-order}
-\end{figure}
-
-For each device, $t_{listnav}$ depends on their relative position in the list.
-Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices the {\em infrared} mode interaction techniques will outperform {\em list} mode\footnote{The higher mean value at $order=1$ is caused by one outlier when the participant tried multiple times in {\em list} mode to connect to the right target.}.
-From the figure, we can see that once the target's order has increased to be larger than 6, the average $t_{listnav}$ for that target would be larger than $t_{reorient} + t_{disambiguate}$.
-We expect that, when the number of targets keeps increasing, there would be larger time reduction in {\em infrared} mode.
-
-Participants' selection errors do occur in both conditions.
-However, error rates were low (1.1\% in {\em infrared} mode, and 2.9\% in {\em list} mode, respectively).
-This precludes us from running a more detailed analysis.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/target-likert.pdf}
-\caption{Likert scale ratings for ease of use of aspects of the targeting task. Error bars show standard error. }
-\label{fig:target-likert}
-\end{figure}
-
-\subsubsection{Preference}
-
-Eleven of 14 users preferred infrared mode over list mode (three preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, list navigation was also perceived to be cumbersome (see Figure~\ref{fig:target-likert}).
-As self-report data can easily skew positive as participants try to please experimenters, we also asked participants to elucidate why they preferred one interface over the other.
-
-List mode had certain advantages: It was judged to be more accurate and predictable as there was always exactly one device selected in the list (\studyquote{With the list you never have to worry about accidentally picking up two targets}).
-Also, it did not require a clear line of sight to the target device so participants did not have to move from their starting position (\studyquote{The shortcoming of the IR mode was that you had to be a certain distance away in order for it to detect the appliance}).
-
-On the other hand, list mode was judged to be more \studyquote{annoying} and tedious.
-The temple-based touchpad for selection was difficult to use for a participant with long hair: \studyquote{List mode was physically difficult for me to navigate, since my long hair wasn't tied back and it kept interfering with my swiping.} Another participant also commented on the ergonomic challenge of touchpad use on Glass: \studyquote{The strength of the IR mode was that I didn't have to use my fingers as much to control.
-If the items were spaced relatively far apart, it was easy to select a specific appliance.}
-
-One noted benefit of infrared mode was a feeling that it was \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
-One subject called it \studyquote{natural to interact with things just by looking at them}.
-Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}.
-
-A few perceived weaknesses of infrared mode were the necessity to move the head in order to control a device and the imperfect mapping of gaze to target.
-One participant said that it was \studyquote{awkward to be aiming your head at things, tweaking back and forth to get it right}.
-Another noted that observing the head movement didn't capture the site of her attention, because \studyquote {eye movement is an important part of how people look around}.
-Users had to learn the usable angle of the IR emitter before they became successful at controlling the devices: \studyquote{I had to compensate by tilting my head up a little bit.}
-
- % control
-\section{Smart Home Control Scenario}
-To understand how our device could be used to interact with smart appliances, we also asked all study participants to work through a concrete scenario.
-The main goal was to obtain qualitative feedback on the usability and utility of our device with more realistic tasks.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/smarthome-scenario.jpg}
-\caption{In the smart home scenario, participants completed a series of appliance control tasks in a simulated living room.}
-\label{fig:smarthome}
-\end{figure}
-
-\subsection{Methodology}
-We recreated a living room environment that had three controllable appliances: a fan, a lamp, and one laptop functioning as a video player (see Figure~\ref{fig:smarthome}).
-The fan and lamp had binary controls: they could be switched on or off.
-The laptop had multiple parameterized functions: participants could start, pause, fast forward, rewind, and adjust volume.
-
-We then asked users to work through the following script for controlling the room for watching a movie in the evening:
-{\small
-\begin{enumerate}
-\item Turn off the lights as you want to watch the movie in a darkened room.
-\item You feel a little hot in the room, so you turn on the fan.
-\item You connect to the Smart TV and start playing the movie. 
-\item The volume seems too soft to hear over the fan- turn it up a bit. 
-\item After a while, you want to take a break to get a snack. Pause the movie. 
-\item When you come back, you've forgotten what was said last - rewind by ~30 seconds and restart the movie. 
-\item When the credits roll, you stop the movie and turn the lights back on.
-\item After awhile, you turn off the fan and leave the room.
-\end{enumerate}
-}
-
-For this study, we only elicited subjective data in the form of Likert data and open-ended responses.
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/scenario-likert.pdf}
-\caption{Likert scale ratings for ease of use of aspects of the smart home control scenario. Error bars show standard error.}
-\label{fig:smarthome-likert}
-\end{figure}
-
-\subsection{Results}
-All participants successfully completed the list of tasks.
-They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous study procedure.
-Participants saw benefits of the device for families --- \studyquote{ It might also be useful for people who need to take care of small children that they can complete all the tasks while keeping an eye on their children at the same time}, though settings that require more movement than watching a movie at home (e.g., cooking) may be more appropriate scenarios for wearing the device.
-
-\subsubsection{Simple and Complex Controls}
-Participants rated the ease of control of particular appliances differently.
-Ease of use ratings were higher for the lamp and fan which had simple, discrete on/off actions, and lower for the more complex movie player (see Figure~\ref{fig:smarthome-likert}).
-Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
-The screen size and (largely) 1D input put a limit on the complexity of interfaces that can be presented.
-As one participant remarked: \studyquote{[The media player] does not seem to be more efficient than a tablet device.} The difficulty can also partly be ascribed to our interaction design, which required one finger swipes to switch between parameters and two finger gestures for adjusting parameters --- it was hard for users to exert fine control over two-finger swipes.
-In addition, users did not always remember these mappings as they are not yet part of a standard gesture vocabulary.
-
-\subsubsection{Eliminating Steps}
-Participants liked the efficiency of our design but also suggested further simplification by eliminating the explicit connection step, they wanted to immediately connect to any appliance that receives the infrared signal --- \studyquote{I intuitively want the screen to automatically appear when the IR detects the appliance rather than having to tap to connect.} Such a design would increase the efficiency of interaction, but at a power tradeoff, as the wireless radio will have to send and receive data each time the user looks at a device --- whether on purpose or inadvertently.
-We leave the study of battery life implications of interaction design choices to future work.
-
-\subsubsection{Feedback On Screen or In The World}
-We found out that the near-eye display may occlude or overlap the target appliance when a participant looks at a target.
-This may make it difficult to read either the on-screen display or see information displayed on the target device.
-As one participant remarked \studyquote{This was especially annoying with the TV because there were two screens overlapping each other.} While it is possible to look away once a device has been acquired to better see the Glass display, a tension remains whether users should rely on feedback from the appliances themselves or on the near-eye display.
-
- % discussion
-\section{Discussion}
-Our study procedures demonstrated that users can successfully select and control smart appliances with head-worn infrared targeting, and that this technique outperforms list selection on the Google Glass wearable device.
-In this section, we revisit some of the results and observations and discuss their larger significance and potential paths for future work.
-
-\paragraph{Meaningful Results?}
-While the performance increase in our targeting study is statistically significant, readers may wonder whether it is truly meaningful.
-We believe it is, for two reasons: first, our technique avoids the problems of {\em naming} and {\em scoping} inherent in any interface that uses representations of objects (e.g., a list of identifiers) rather than the objects themselves.
-We argue that this disintermediation of interaction leads to a cognitively simpler design.
-Second, our existing study only showed results for a modest number of targets.
-Our technique should have a wider margin as the number of targets increases.
-Of course, the number of targets one can realistically expect may vary across application domains.
-
-\paragraph{Hands-Free Operation}
-While head movement is not as precise as hand positioning (e.g., in Patel's mobile laser pointing~\cite{patel_2-way_2003}), one key benefit of our target acquisition step is that it does not require the user's hands.
-This raises the question if the rest of the interaction (disambiguation and device control) could also be achieved in a hands-free fashion.
-Voice-command control is an obvious candidate, though such approaches have not found widespread adoption because of social acceptability and other factors.
-
-\paragraph{Hardware Limitations}
-In our prototype, the IR emitter is fixed in a single position.
-Due to different sizes and shapes of the users’ heads, the emitter may not line up exactly with their head orientation.
-An adjustable emitter (paired with a suitable calibration routing) would improve the performance of our design.
-Some of our users also mentioned that they had to move closer to some targets to successfully select them.
-A stronger emitter could overcome these problems, but care has to be taken to avoid possible reflection problems where IR light bounces off a wall and hits an unintended target behind the user.
-
-More importantly though, the main limitation of our design is that extra hardware for infrared communication is needed for the head-mounted device and each controllable appliance.
-One potential approach to sidestep this requirement would be to combine the growing availability of high-resolution indoor maps with live data from the point-of-view camera on the device to determine what a user is looking at without any infrared data exchange.
-
-\paragraph{The Midas Look}
-Our participants suggested eliminating explicit initiation of a connection by the user.
-However, one of the design guidelines for near-eye displays is to avoid pushing information to the display without an initial request from the user --- flashing device information on screen each time a user moves their head would surely be counterproductive.
-Future work should investigate how to intelligently decide when and how to initiate interaction for the user.
-
-\paragraph{Where is the Target?}
-One open design question of our approach is where infrared receivers should be placed.
-For a light, one might put a received on the light itself, or on the light switch, to cater to existing expectations.
-For volume control, the infrared receiver might be located on a speaker or on the amplifier.
-A thorough study of user preferences would be interesting; though we also point out that our architecture could easily support multiple receivers that end up controlling the same appliance.
-
- % conclusion
-\section{Conclusion}
-We introduced a novel method for selecting and controlling smart appliances in physical spaces through infrared targeting based on head orientation.
-Through our solution, we attempt to address the naming and scaling challenges faced by handheld mobile devices.
-The design takes advantage of the fact that visual attention can express intention.
-The visual feedback provided by the target appliances helps users keep their focus in the physical world.
-While we present a prototype approach that requires that the user carry additional hardware, all parts can readily be miniaturized and integrated into future head-worn hardware.
-We also introduced a disambiguation technique in case head orientation is not sufficient to determine a unique target.
-We characterized our devices performance, arguing that it is matched well to the amount of head movement people can control without strain.
-A target acquisition study showed that the technique is efficient; a home control scenario showed promise but also limitations when trying to control complex appliances.
-As our environment continues to be populated by a swarm of sensing and actuation devices, methods to interrogate and control our smart environments will become increasingly important.
-
-\balance
-
-\bibliographystyle{acm-sigchi}
-\begin{thebibliography}{10}
-
-\bibitem{azuma_recent_2001}
-Azuma, R., Baillot, Y., Behringer, R., Feiner, S., Julier, S., and {MacIntyre},
-  B.
-\newblock Recent advances in augmented reality.
-\newblock {\em {IEEE} Comput. Graph. Appl. 21}, 6 (Nov. 2001), 34–47.
-
-\bibitem{beigl_point_1999}
-Beigl, M.
-\newblock Point \& click-interaction in smart environments.
-\newblock In {\em Handheld and Ubiquitous Computing}, H.-W. Gellersen, Ed.,
-  no.~1707 in Lecture Notes in Computer Science. Springer Berlin Heidelberg,
-  Jan. 1999, 311--313.
-
-\bibitem{card_morphological_1991}
-Card, S.~K., Mackinlay, J.~D., and Robertson, G.~G.
-\newblock A morphological analysis of the design space of input devices.
-\newblock {\em {ACM} Trans. Inf. Syst. 9}, 2 (Apr. 1991), 99–122.
-
-\bibitem{costanza_sensortune:_2010}
-Costanza, E., Panchard, J., Zufferey, G., Nembrini, J., Freudiger, J., Huang,
-  J., and Hubaux, J.-P.
-\newblock {SensorTune:} a mobile auditory interface for {DIY} wireless sensor
-  networks.
-\newblock In {\em Proceedings of the {SIGCHI} Conference on Human Factors in
-  Computing Systems}, {CHI} '10, {ACM} (New York, {NY}, {USA}, 2010),
-  2317–2326.
-
-\bibitem{cross_low-cost_2012}
-Cross, A., Cutrell, E., and Thies, W.
-\newblock Low-cost audience polling using computer vision.
-\newblock In {\em Proceedings of the 25th annual {ACM} symposium on User
-  interface software and technology}, {UIST} '12, {ACM} (New York, {NY}, {USA},
-  2012), 45–54.
-
-\bibitem{kemp_point-and-click_2008}
-Kemp, C.~C., Anderson, C.~D., Nguyen, H., Trevor, A.~J., and Xu, Z.
-\newblock A point-and-click interface for the real world: laser designation of
-  objects for mobile manipulation.
-\newblock In {\em Proceedings of the 3rd {ACM/IEEE} international conference on
-  Human robot interaction}, {HRI} '08, {ACM} (New York, {NY}, {USA}, 2008),
-  241–248.
-
-\bibitem{lifton_tricorder:_2007}
-Lifton, J., Mittal, M., Lapinski, M., and Paradiso, J.~A.
-\newblock Tricorder: A mobile sensor network browser.
-\newblock In {\em Proceedings of the {ACM} {CHI} 2007 Conference-Mobile Spatial
-  Interaction Workshop} (2007).
-
-\bibitem{mittal_ubicorder:_2011}
-Mittal, M., and Paradiso, J.
-\newblock Ubicorder: A mobile device for situated interactions with sensor
-  networks.
-\newblock {\em Sensors Journal, {IEEE} 11}, 3 (2011), 818--828.
-
-\bibitem{myers_interacting_2002}
-Myers, B.~A., Bhatnagar, R., Nichols, J., Peck, C.~H., Kong, D., Miller, R.,
-  and Long, A.~C.
-\newblock Interacting at a distance: measuring the performance of laser
-  pointers and other devices.
-\newblock In {\em Proceedings of the {SIGCHI} Conference on Human Factors in
-  Computing Systems}, {CHI} '02, {ACM} (New York, {NY}, {USA}, 2002), 33–40.
-
-\bibitem{pausch_user_1993}
-Pausch, R., Shackelford, M.~A., and Proffitt, D.
-\newblock A user study comparing head-mounted and stationary displays.
-\newblock In {\em Virtual Reality, 1993. Proceedings., {IEEE} 1993 Symposium on
-  Research Frontiers in} (1993), 41–45.
-
-\bibitem{raskar_rfig_2004}
-Raskar, R., Beardsley, P., van Baar, J., Wang, Y., Dietz, P., Lee, J., Leigh,
-  D., and Willwacher, T.
-\newblock {RFIG} lamps: interacting with a self-describing world via
-  photosensing wireless tags and projectors.
-\newblock In {\em {ACM} {SIGGRAPH} 2004 Papers}, {SIGGRAPH} '04, {ACM} (New
-  York, {NY}, {USA}, 2004), 406–415.
-
-\bibitem{rukzio_experimental_2006}
-Rukzio, E., Leichtenstern, K., Callaghan, V., Holleis, P., Schmidt, A., and
-  Chin, J.
-\newblock An experimental comparison of physical mobile interaction techniques:
-  Touching, pointing and scanning.
-\newblock In {\em {UbiComp} 2006: Ubiquitous Computing}. Springer, 2006,
-  87–104.
-
-\bibitem{schmidt_picontrol:_2012}
-Schmidt, D., Molyneaux, D., and Cao, X.
-\newblock {PICOntrol:} using a handheld projector for direct control of
-  physical devices through visible light.
-\newblock In {\em Proceedings of the 25th annual {ACM} symposium on User
-  interface software and technology}, {UIST} '12, {ACM} (New York, {NY}, {USA},
-  2012), 379–388.
-
-\bibitem{patel_2-way_2003}
-Shwetak N.~Patel, G. D.~A.
-\newblock A 2-way laser-assisted selection scheme for handhelds in a physical
-  environment.
-\newblock 200--207.
-
-\bibitem{weiser_coming_1997}
-Weiser, M., and Brown, J.~S.
-\newblock The coming age of calm technolgy.
-\newblock In {\em Beyond calculation}, P.~J. Denning and R.~M. Metcalfe, Eds.
-  Copernicus, New York, {NY}, {USA}, 1997, 75–85.
-
-\bibitem{willis_sidebyside:_2011}
-Willis, K.~D., Poupyrev, I., Hudson, S.~E., and Mahler, M.
-\newblock {SideBySide:} ad-hoc multi-user interaction with handheld projectors.
-\newblock In {\em Proceedings of the 24th annual {ACM} symposium on User
-  interface software and technology}, {UIST} '11, {ACM} (New York, {NY}, {USA},
-  2011), 431–440.
-
-\bibitem{wilson_xwand:_2003}
-Wilson, A., and Shafer, S.
-\newblock {XWand:} {UI} for intelligent spaces.
-\newblock In {\em Proceedings of the {SIGCHI} Conference on Human Factors in
-  Computing Systems}, {CHI} '03, {ACM} (New York, {NY}, {USA}, 2003),
-  545–552.
-
-\end{thebibliography}
-
-\end{document}

### 2014-04-14 doc/uist14/system2.tex
structural layout as discussed on sunday - iterative development

%!TEX root = uist14.tex
\section{Iteration 2: IR Intensity refinement}

\bjoern{The results of the first study show that head orientation targeting can outperform linear list selection, but that disambiguation (or refinement step, need to decide on consistent terminology) is costly.

Our approach to avoid paying this penalty is to provide more information to the system which target is more likely than another to be the intended selection. We do so by collecting received IR signal strength at each target node and comparing measurements of all candidate targets.}

\subsection{Interaction Flow}
\bjoern{
Intensity data is used in two ways (see Figure~\ref{fig:selectflow02}): 1) the difference in intensity between the strongest and second target exceeds a threshold, the target with the strongest received signal is automatically selected; 2) if the difference falls below the threshold, the refinement UI from the previous iteration is displayed, but items are ordered in decreasing intensity. This should maximize the chance of being able to confirm the intended target without additional UI navigation actions}.

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/selectflow02.pdf}
\caption{Head orientation selection flowchart for our second design with IR intensity measurement.}
\label{fig:selectflow02}
\end{figure}

\subsection{Implementation}
\bjoern{we use sensor X and LED y which give useful readings at up to N feet. Intensity falls of as $1/d^2$ with distance and approximately $xyz$ with angle. We empirically set the threshold for automatic selection without refinement at Z.}\bjoern{maybe show some measurements for intensity based on distance and angle as we had discussed a while ago. not essential. }

### 2014-04-14 doc/uist14/system3.tex
structural layout as discussed on sunday - iterative development

%!TEX root = uist14.tex
\section{Iteration 3: Orientation-based refinement}
\bjoern{While taking IR intensity into account further reduced the need for manual refinement and increased performance, the UI navigation scheme is problematic - it is not spatially related in any meaningful way to the layout of targets in a room. A better interaction technique would respect that ordering. For example, when refinement is needed, tilting the head slightly to the right could select the target that's adjacent to the right of the currently selected target.

Such spatial navigation requires knowledge about the layout of targets in the environment. However, one of the strengths of our technique so far is that it does not require any map ahead of time. To enable some spatial navigation, we introduce a final iteration in which we build up a spatial data structure by demonstration (i.e., the user looks around the room) and then leverage that data structure during the refinement step of our interaction.}

\bjoern{This technique is based on the assumption that users will generally select targets in indoor environments where targets are spread around the periphery. These assumptions enable us to use orientation data without knowing the user's absolute position.}

\subsection{Implementation}
\bjoern{give implementation details}

\subsection{Informal User Feedback}
\bjoern{We informally evaluate this technique with N users: ...}

### 2014-04-14 doc/uist14/relatedwork.tex
added discussion of sources 10 and 20

@@
 
-\bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
+\claire{Other targeting systems use IR with handheld pointers ~\cite{swindells_that_2002} and as well as wearable devices such as rings and Bluetooth audio earpieces ~\cite{merrill_augmenting_2007} to connect to smart devices.
+Our system tackles an unresolved issue of such IR-based approaches-- navigating an area dense with potential targets.
+}

### 2014-04-14 doc/uist14/abstract.tex
a significant amount of new content and changes on papers

@@
 %!TEX root = uist14.tex
+
+\ben{This parts needs to be changed based on the current flow of the story.}
 Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
@@
 We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.
-A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device.
-The fundamental imprecision of head orientation requires disambiguation techniques when the intended target is unclear.
-To improve on the slow manual disambiguation of our first design, we introduce a second design in which receivers use IR intensity sensors to estimate which target is the intended selection.
-A second study demonstrates that using IR intensity successfully reduces ambiguity and enhances selection speed.
-Finally, we describe how a spatial data structure can be inferred to enable a natural mapping of head direction during the disambiguation step and we report first use data.
+We present our iterative design process that has shown with IR, it outperforms a pure list selection.
+And with IR intensity, the time can be further reduced.
+If head motion is added, we can infer a spatial data structure of the nodes to enable a natural head-motion based disambiguation.
 We demonstrate an application of our selection technique for controlling smart appliances and report qualitative user impressions.

### 2014-04-14 doc/uist14/applications.tex
a significant amount of new content and changes on papers

@@
 \bjoern{since length will be an issue, let's just focus on glass remote and skip the others?}
+\ben{I still think that by listing a few other applications, it helps identifying the importance of physical targeting. We may do that with a short paragraph instead of a subsection.}
 Additional applications such as attention tracking or indoor position and orientation estimation of a wearer based on our technique are conceivable; we leave them to future work.

### 2014-04-14 doc/uist14/interaction.tex
a significant amount of new content and changes on papers

\section{Interaction Models}

In this section, we first describe how head orientation-based interaction works from the users' point of view (Figure~\ref{fig:intearaction}).

{\bf Scan:} The user first scans the environment to locate the position of the target. During this stage, Glass will constantly sends out IR signal, and the target offers immediate feedback when it receives the IR.

{\bf Coarse Selection:} The user confirms his desire to connect to a target by tapping on the Glass touchpad. At this time, Glass collects the replies from targets about the IR signal reception. If there is only a single target in IR range, then it's considered as a successful selection. However, in a relative dense environment where multiple targets are within range, the user needs to perform a fine selection.

{\bf Fine selection:} When disambiguation is needed, the user needs to make an explicit selection among the targets that within his view range. There are multiple ways to achieve it -- either from Glass touchpad that controls a UI list or using the head motion sensor on Glass. We will leave a thorough and complete discussion later. Once the user confirms the current selection, he could tap to complete the acquisition. Since this part is effectively disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/interactionModel.pdf}
\caption{Four different states during the head-orientation targeting.}
\label{fig:interaction}
\end{figure}

From the flow of interaction, we learned that the overall target acquisition time can be broken down into the following pieces:
\begin{equation}
t_{total}=t_{scan}+t_{coarse}+t_{disambiguation}
\end{equation}

Our goal of this paper is to minimize the overall target selection time. In the following few sections, we will describe our iterative design process during which we have explored the design space.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-14 doc/uist14/intro.tex
a significant amount of new content and changes on papers

@@
 While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed; that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen or keyboard); and that the user's visual attention is now split between looking down at a screen and out at targets in the world.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/teaser.jpg}
-\caption{An example application of head-orientation targeting: Using an augmented head-worn device (1), users can control smart home appliances (2) by looking at them.
-A near-eye display then shows an appliance control UI (3), which users navigate through multitouch gestures.}
-\label{fig:teaser}
-\end{figure}
 
@@
 In this paper, we describe the iterative development and evaluation of an area-selection technique that can be readily implemented with small hardware changes to emerging wearable devices.
-We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable (IR) communication between Glass and target appliances.
+We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infared (IR) communication between Glass and target appliances.
 The cone shape of light emitted by an IR LED (a diameter of 60-120cm and distance up to 6m) provides the {\em coarse} selection area.

### 2014-04-14 doc/uist14/iteration1.tex
a significant amount of new content and changes on papers

--- 2014-04-14 doc/uist14/study1.tex
@@
 %!TEX root = uist14.tex
-\section{Study 1: Physical Target Acquisition Study}
-To understand the accuracy and performance of head-orientation-based selection through our device, we carried out a comparative target acquisition study, where participants had to connect to wireless nodes distributed in a room with our technique, and with an alternate list selection approach.
 
-\subsection{Apparatus}
-In an indoor environment, 10 wireless nodes are spread across a room (Figure~\ref{fig:targeting-study-layout}).
-Each node is an embedded wireless system with a microcontroller, IR receiver, a wireless XBee radio, and three status LEDs %(Figure~\ref{fig:target}).
-%An yellow LED indicates that the device is the target that should be selected in the current trial; a red LED lights up whenever the device receives an IR signal from Glass; a blue LED shows when participants have successfully connected to a device, and is also used for disambiguation when multiple targets are within IR range.
-Next to each target, a paper sheet shows a number and letter combination, which is used for uniquely identifying the device.
-The numbers are the primary identifiers, ordered from left to right in the room.
+\section{Iteration 1: IR vs. List}
+Our system design uses IR as the beam cursor for physical target selection because the area coarse selection will remove the targets outside of users' view range.
+And this could help during the fine selection time because there is a smaller set of targets for fine selection.
+The opposite is to have a complete list of all targets when there is no single coarse selection and the disambiguation is always needed.
+This brought up the first question that we asked ourselves -- {\em how much does using IR help in reducing the selection time}.
+To answer that question and quantify the benefit of having IR for the coarse selection, we carried out our first target acquisition study which compares the {\em IR selection} and {\em list selection}.
+
+We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}).
+Each of them has a number as ID and a letter representing the name.
+There are 14 participants from our institution for this study.
+13 had never used Glass before, therefore we offered a Google Glass crash course before the experiment for them to get familiar with Glass gestures.
+4 of them wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
 
@@
 \end{figure}
-%\begin{figure}[b]
-%\centering
-%\includegraphics[width=0.8\columnwidth]{figures/study-node.pdf}
-%\caption{An example node from the targeting study --- we constructed 10 such nodes - each mounted in a box.}
-%\label{fig:target}
-%\end{figure}
 
-\subsection{Methodology}
-In our within-subjects design, participants performed 15 target acquisition tasks each with two interaction styles.
-In the {\em infrared} mode condition, participants used our IR targeting approach; in the {\em list} mode condition, participants had to look up a device's letter code on the printed paper next to the device and then select that letter code from a list displayed on their Glass device.
-The list was navigated with swipe motions on the Glass touchpad.
-For each task, participants started at a fixed position in the room.
-The experimenter called out a number and simultaneously started a timer.
-Participants then had to find the corresponding device (by looking for its printed code).
-In the {\em infrared} mode, participants then selected and acquired the target by aiming the IR beam at the target, and confirmed their selection with a touch pad tap.
-If more than one target was within range, participants had to either use the disambiguation dialog or reposition themselves.
-In the {\em list} mode, participants had to read the letter next to the number and then select that letter by browsing a linear list shown in their Glass display.
-While the list was alphabetized, letter arrangement in the room was not.
-This design required participants to find the target in the room before starting a list navigation to keep visual search times similar in each condition.
+In the within-subject study, half of them performed {\em IR selection} first and the other half did {\em list selection} first.
+For each selection condition, we conducted 15 target selection by random choosing from all the targets.
+During the study, we measure the {\bf target acquisition time} for each target selection, and afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
 
-Afterwards, participants completed a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
-
-\subsection{Participants}
-We recruited 14 participants from our institution.
-13 had never used Glass before.
-4 wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
-Half of them performed {\em infrared} mode first and the other half did {\em list} mode first.
-
-\subsection{Measures}
-The main measures were {\bf target acquisition time}: the time required to identify, select, and connect to a wireless target device; and {\bf user preference}: which interface users preferred for the task after completing the study.
-
-\subsection{Results}
-\subsubsection{Performance data}
-The time to complete each task can be broken down into the following pieces:
-
-$t_{infrared}=t_{locate}[+t_{reorient}][+t_{disambiguate}]+t_{tap}$
-
-$t_{list}=t_{locate}+t_{listnav}+t_{tap}$
-
-In both conditions, participants first have to locate the target announced by the experimenter through visual search ($t_{locate}$).
-In the {\em infrared} mode, participants may then directly confirm their selection if only a single target was selected ($t_{tap}$).
-However, if they don't immediately receive feedback that their target was selected, or if multiple targets were selected, users either have change their position or head orientation ($t_{reorient}$) or they have to step through the on-screen disambiguation dialog ($t_{disambiguate}$).
-In the {\em list} mode, participants must scroll through the list to find the desired target identifier ($t_{listnav}$).
-Thus, {\em infrared} will show a performance benefit if $(t_{reorient}+t_{disambiguate})<t_{listnav}$.
-This depends on the number of total devices in the environment (increasing $t_{listnav}$), and their density (which will increase $t_{disambiguate}$).
+%% remove this for now
+%% Our measurements suggest that IR communication can be targeted to an area about 60-120cm in diameter, up to 5m in front of the user.
+These values are a reasonable match for selecting targets in a room-size environment.
+A wider beam would lead to an increased chance of multiple appliances receiving IR signals simultaneously.
+A narrower beam will make targeting more challenging, given the precision constraints of human head movement.
 
@@
 The centers of boxes are median values, while white dashed lines are mean values.}
-\label{fig:selection-times}
+\label{fig:selection-time}
 \end{figure}
 
-We first show results for 10 targets and then discuss extrapolations of these results.
-Average target acquisition time $t_{infrared}$ was 6.67 seconds, while $t_{list}$ was 8.86 seconds (Figure~\ref{fig:selection-times}A).
-This difference is significant (Student's t-test, $t(279)=-3.81, p=0.00017$).
-% \bjoern{you need to report the t statistic and degrees of freedom as well - t(df)= n,p=m.}
+Our results indicate that {\em IR selection} is generally better than {\em list selection}.
+From Figure~\ref{fig:selection-time}A, the average target acquisition time for {\em IR selection} is 6.67 seconds, while for {\em list selection} it took 8.86 seconds.
+We run Student's t-test on the data and it shows significant difference ($t(279)=-3.81, p=0.00017$).
 
-To further understand the performance gain in {\em infrared} mode, especially the factor $t_{disambiguate}$, we compare selection times when multiple devices are targeted (and disambiguation is required) to single-device selection times (Figure~\ref{fig:selection-times}B).
-When there is a single device, $t_{disambiguate}$ is 0 and it takes 6.40 seconds (on average) to complete the connection.
-When multiple devices are in range, the time increases to 9.16 seconds, indicating 2.76 seconds required to disambiguate.
-Though it takes significantly longer ($t(19)=-2.7827, p=0.012$ using t-test) in the {\em multiple} case, these cases made up only 10\% of total infrared trials.
-%, even we intentionally placed a few devices nearer to each other. % 19 / (19+168 )
+To further understand if the gain comes from the coarse selection, we splited the data from {\em IR selection} to two parts -- ones that require disambiguation and ones that don't.
+The result is shown in Figure~\ref{fig:selection-time}A.
+It shows that it takes 6.40 seconds (on average) to complete the connection if coarse selection has achieved the acquisition, while when multiple devices are in range, the time increases to 9.16 seconds, indicating additional 2.76 seconds for disambiguation.
+Though it takes significantly longer ($t(19)=-2.7827, p=0.012$ using t-test) in the latter case, they only happen 10\% of total {\em IR selection} trials.
+
+From the above result, we learned that with IR, we can have effectively reduced the overall target acquisition time.
+And the major gain comes from the fact that {\em coarse selection} helps reduce the chances of entering {\em fine selection} stage.
 
@@
 
-For each device, $t_{listnav}$ depends on their relative position in the list.
-Figure~\ref{fig:time-vs-list-order} shows the time it takes to select a device (means and standard deviations) as a function of its list position - the trend line (dotted) enables extrapolation to estimate at what number of devices the {\em infrared} mode interaction techniques will outperform {\em list} mode\footnote{The higher mean value at $order=1$ is caused by one outlier when the participant tried multiple times in {\em list} mode to connect to the right target.}.
-From the figure, we can see that once the target's order has increased to be larger than 6, the average $t_{listnav}$ for that target would be larger than $t_{reorient} + t_{disambiguate}$.
-We expect that, when the number of targets keeps increasing, there would be larger time reduction in {\em infrared} mode.
+To further generalize the results, in Figure~\ref{fig:time-vs-list-order}, we show the acquisition time for each individual target in {\em list selection} condition, ordered by their relative position in the list.
+Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}.
+From this figure, we can see that once the target's order is larger than 6, the average acquistion time will be larger than the case when disambiguation is needed in {\em IR selection}.
+Therefore, we argue that once there are more targets in the physical space, the {\em list selection} is not scalable.
 
-% \bjoern{A linear trend is visible, and the regression returns with coefficient of determination ($R^2=0.032$).
---- This R value is basically no correlation at all!
-try to run a correlation only on the median times.
-}
-
-Participants' selection errors do occur in both conditions.
-However, error rates were low (1.1\% in {\em infrared} mode, and 2.9\% in {\em list} mode, respectively).
-This precludes us from running a more detailed analysis.
-
+\ben{qualitative feedback hasn't been added yet for this part.}
 %\subsubsection{Preference}
@@
 Users had to learn the usable angle of the IR emitter before they became successful at controlling the devices: \studyquote{I had to compensate by tilting my head up a little bit.}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "uist14"
+%%% End:

### 2014-04-14 doc/uist14/iteration2.tex
a significant amount of new content and changes on papers

\section{Iteration 2: Binary IR vs. Valued IR}

With our first study, we have learned that with {\em coarse selection}, we can reduce the overall target selection time because in many cases there will be only one target in range and {\em fine selection} is not needed, and in cases where it's needed, the list length has been effectively reduced. However, in a dense environment, there could be more cases when disambiguation is needed.

From Figure~\ref{fig:selection-time}B, the average results when disambiguation is involved is significantly increase. We challenge ourselves that ``can we do a better job in a dense environment?''

Previously we are only using IR reception as a signal for potential target identification. It's a binary result indicating that whether a target has received IR signal or not. But actually the IR intensity at the receiver side can provide more information than just the binary reception. The illuminance falls when the distance increases as well as when the angle inceases.
We have emprically measured the intensity distribution at the receiver side, and it's shown in Figure~\ref{fig:measurement}.

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/placeholder.pdf}
\caption{placeholder for emperical measurement.}
\label{fig:measurement}
\end{figure}

The intensity information is used in two ways:
\begin{enumerate}
\item When multiple targets have received IR signal and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half, which is frequently used because it indicates a 3dB loss in the signal strength.}. In this case, when Binary IR doens't need disambiguation, the new design still doesn't need it. When the environment turns denser, the new design can filter some peripheral targets out, making the chances of entering disambiguation lower.
\item When disambiguation is still needed, we use the IR intensity as an indication of likelihood to sort the list. This will help us reduce the $t_{disambiguation}$ significantly if the first one appears matches the actual target. Given the imprecision of head orientation, this can hardly be always perfect.
\end{enumerate}

Though apparently with more information, we should be able to reduce the selection time. But the question we seek to answer is to further quantify such gain. We then performed another user study to compare the {\em Binary IR} and {\em Valued IR} designs.

For this study, since we are aiming at a denser environment, though we keep the node number to 10, but the environment setup has changed (see Figure~\ref{fig:study-layout2}). In total, we have XX users for this study \ben{more information about users later.} Each user performs the target acquisition study 30 times for each condition. Again half of them perform {\em Binary IR} first and the other half {\em Valued IR} first.

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/study-layout2.jpg}
\caption{The current environment setup \ben{needs annotation and a better photo.}.}
\label{fig:study-layout2}
\end{figure}

We then summarize the result of this study. 
\begin{enumerate}
\item Among the 150 trials in {\em Binary IR} condition, there are 104 times when disambiguation is needed. In {\em Valued IR} condition, only 65 out of 150 trials need disambiguation. We perform a Chi-square test on this result and the result is $\chi^2(1) = 20.611$, $p < 0.0001$. \ben{not sure about the degree of freedom being 1 here, will check later.} This demonstrates that the first intensity gain is from the reduction of chances when disambiguation is needed.
\item In the cases when disambiguation is needed, the {\em Valued IR} will sort the list based on the intensity reading. Indeed, out of the 65 times when disambiguation is involved, there are 38 (roughly 58\%) are right for {\em Valued IR}. However, for {\em Binary IR}, only 37 out of 104 (~35\%) are right. Again a Chi-square test demonstrates the statistical differences (with $\chi^2(1) = 8.487$, $p = 0.0036$. 
\end{enumerate}

Such gain can be seen from Figure~\ref{fig:study2} where the overall target acquisition time is obtained. The average time is 4.32 seconds and 3.71 second for {\em Binary IR} and {\em Valued IR}, respectively; and a t-test shows that they are also significantly different $t(282)=2.0922$, $p=0.03$.

\begin{figure}[t]
\centering
\includegraphics[width=0.9\columnwidth]{figures/study2_time.pdf}
\caption{The target acquisition of in {\em Valued IR} and {\em Binary IR} condition.}
\label{fig:study2}
\end{figure}

However, one side effect that we have abserved in this approach is that the {\em Valued IR} sometimes eliminates the right target during the filter stage. The rate is not that high (16 out of 150 cases, ~90\%), but this is the price to pay for a quicker average in this smarter system.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-14 doc/uist14/iteration3.tex
a significant amount of new content and changes on papers

\section{Iteration 3: Head motion for disambiguation}
\label{sec:iteration-3:-head}

Though our {\em Valued IR} system is already performing well (with 3.71 seconds on average in a relative dense environment), we are only using head-orientation information during the {\em coarse selection} stage. In the {\em fine selection} part, we rely on the near-eye display. This has one fundamental problem -- we are projecting a high dimension distribution to a 1-dimension list. Therefore, it motivates us to explore using head orientation also during the {\em fine selection} stage.

In our third iteration, we ask the question ``can we build a system purely using head orientation and visual feedback from the environment for target selection?''

The sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of users' head. However, such absolute orientation cannot be directly used in any indoor environment because the user might change his location. However, as long as the targets are spread around the periphery, their relative orientation doesn't change. This key observation inspires us to investigate a hybrid solution with IR and sensors on Glass.

Before we describe how this new technique works. The relative orientation information -- {\em adjacent map} -- has to be learned before we can use it. Such a learning process can either be performed a prior, or through demonstration. \ben{need better elaboration here}.

In this hybrid approach, in additional to use IR to reduce the chance of disambiguation and prepare a better list, the IR information helps us identify a base point. When the user wants to use his head-motion for disambiguation, he will put his finger on the touchpad to enter a quasi-mode. Such a gesture comes from a real-world metaphor that people tend to adjust their glasses for a closer inspection. In such quasi-mode, we monitor the changes of his head orientation, and infer the intended next targets according to some algorithm \ben{haven't described yet}.

\ben{We then need evaluation results here, quantitative and qualitative.}

% %!TEX root = uist14.tex
% \section{Iteration 3: Orientation-based refinement}
% \bjoern{While taking IR intensity into account further reduced the need for manual refinement and increased performance, the UI navigation scheme is problematic - it is not spatially related in any meaningful way to the layout of targets in a room. A better interaction technique would respect that ordering. For example, when refinement is needed, tilting the head slightly to the right could select the target that's adjacent to the right of the currently selected target.

% Such spatial navigation requires knowledge about the layout of targets in the environment. However, one of the strengths of our technique so far is that it does not require any map ahead of time. To enable some spatial navigation, we introduce a final iteration in which we build up a spatial data structure by demonstration (i.e., the user looks around the room) and then leverage that data structure during the refinement step of our interaction.}

% \bjoern{This technique is based on the assumption that users will generally select targets in indoor environments where targets are spread around the periphery. These assumptions enable us to use orientation data without knowing the user's absolute position.}

% \subsection{Implementation}
% \bjoern{give implementation details}

% \subsection{Informal User Feedback}
% \bjoern{We informally evaluate this technique with N users: ...}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "uist14"
%%% End:

### 2014-04-14 doc/uist14/system.tex
added a figure
structural layout as discussed on sunday - iterative development
a significant amount of new content and changes on papers

--- 2014-04-06 doc/uist14/system.tex
@@
-\section{System Design and Prototyping}
-\label{sec:syst-design-prot}
-
-This section describes the design and implementation of our prototype system.
-We start by summarizing the design considerations in this head-orientation targeting, and for each consideration, we make the choice of commercial off-the-shelf (COTS) components.
-Afterwards, we conclude this section by a depiction of the overall system architecture that is used for both our study and future application development.
-
-\subsection{Design Considerations}
-\label{sec:design-cons}
-
-To effectively enables the physical targeting based on user's head movement, and build such a platform for applications, we have come up with the following design considerations and the justification for each.
-
-{\bf Directional area selection:} To be aligned with the user's head orientation, the sensing modality should be directional.
-Merely direction is not enough, from Section~\ref{sec:background}, the head muscle's fundamental limitation and Fitts' law imposes us to choose an area selection technique.
-However,  with a larger coverage area, there will be a higher chance for disambiguation.
-We should find the proper technique that achieves area selection while still providing ways for a more fine-grained resolution.
-
-% the two visual feedback seems a bit random...
-{\bf Leverage visual feedback:} Since the user cannot see his own head orientation, there should be some effective ways to provide feedback to the users.
-We consider two types of visual feedbacks.
-The first comes from the environment which enables instantaneous visual cue while the user is looking around.
-The second should remain unchanged even if the users' head orientation has changed; in such a way, further interactions are enabled without imposing the burden of staring.
-But such a display shouldn't distract the user's current line-of-sight attention -- which means a near-eye display is better than using a hand-held device.
-
-% A calmer~\cite{weiser_coming_1997} approach is to locate visual feedback about selection targets in the environment, to prevent distraction and interruption.
-Such feedback should be delivered instantaneously, while users look around a room.
-
-{\bf Flexible communication:} The system is for the interaction with the environment, therefore the communication among devices should be flexible.
-Especially there would be a consensus problem if the end-nodes are not agreeing on what the user is looking at.
-The communication would be necessary for resolving the inconsistency in such a distributed environment.
-Also providing a flexible communication capability can make the system as a platform for other applications.
-
-{\bf Adjustable angle:} From our pilot study, we have learned that human's head orientation pointing is biased for each person.
-In order to align area selection technique, there should be ways to adjust or calibrate the covered area.
-
-These design considerations find their expression in our selection of techniques in the following few subsections.
-
-\subsection{IR as beam cursor}
-\label{sec:ir-as-beam}
-
-For the directional area selection technique, we have chosen to use infrared (IR) for a couple of reasons.
-
-First, the signal characteristics of IR makes it suitable for directional area selection.
-There are a wide range of viewing angles that we can choose (DigiKey offers IR emitters from $4^\circ$ to $240^\circ$, however, most are around $20^\circ$ of viewing angle).
-This leads to a cone shape-like area coverage and can be well-aligned with humna's line of sight.
-
-Second, for the disambiguation consideration, since the infrared illuminance falls off with respect to both the offset angle from the center and the distance to the emitter, even though the signal might cover multiple targets, using the intensity from the receivers can help disambiguation.
-% for the 4 degree to 240 degree, I used information from http://www.digikey.com/product-search/en/optoelectronics/infrared-uv-visible-emitters/524328?k=IR
-
-In addition, IR has been used for remote control for many years. The emitters and receivers are fairly cheap to manufacture; and it's easy for integration with existing devices. 
-What's more, IR is immune to most interference sources.
-The ambient IR might influence the IR receivers' intensity reading, however with simple IR signal validation, the receiver is still safe.
-There might be issues for signal reflection, either bouncing from the ground to the target or bouncing back to the opposite direction.
-However, such visual feedback is not aligned with the user's attention so no further action from the user might be taken; or the genuine target could easily beat the false positive since the one with reflection tends to travel a longer distance, making the signal strength smaller.
-
-Therefore, we can attach an IR emitter to capture the head orientation and install some IR receivers in the environment for both signals detection and intensity reading.
-For our prototype, we have found some COTS components for IR.
-The IR emitter is the one from OSRAM Opto Semiconductors Inc with manufacturer part number SFH 4545.
-From the datasheet, it has a view angle of $10^\circ$.
-Part of the reason we have chosen it is because the radiant intensity can quite high if we provide it with sufficient current flow (550mW/sr @ 100mA).
-We can then easily adjust it using a resistor to satisfy different demands.
-% transmitter webpage: http://www.digikey.com/product-search/en?x=0&y=0&lang=en&site=us&KeyWords=475-2919-ND
-For the IR receiver, we use 38.0kHz IR Receiver Modules from Vishay Semiconductor Opto Division (manufacturer part number TSOP38238). 
-% receiver webpage http://www.digikey.com/product-detail/en/TSOP38238/751-1227-ND/1681362
-In order to read the IR intensity, we integrates another IR light-to-voltage converter from AMS-TAOS USA Inc (manufacturer part number TSL267-LF) which is pretty sensitive to IR irradiance. 
-% http://www.digikey.com/product-detail/en/TSL267-LF/TSL267-LF-ND/3095052
-
-We use Arduino platform for the prototyping, and there is an easy-to-use IR library\footnote{\url{https://github.com/shirriff/Arduino-IRremote}} where we can customize the IR signal for both transmission and reception.
-To read the IR intensity, an 10-bit ADC is attached to the output of TSL267-LF which gives a reading from 0 to 1023.
-
-\subsection{Environment LEDs and Head-up display}
-\label{sec:head-up-display}
-As we have argued for the visual feedback design considerations, the environment should provide instantaneous visual cue to respond to user's head orientation.
-This can be easily achieved with 1-bit LEDs on each devices that has an IR receiver.
-For the visual feedback that is persistent even when the user's head orientation has changed, we have considered a few COTS solutions that support the head-up display, and ended up with Google Glass because it's easy to program (it supports standard Android programming) and the device is also light-weight for everyday use.
-At the time of writing, we have Google Glass running Android 4.0.4 (software version: XE12).
-
-\subsection{Wireless Communication}
-\label{sec:wirel-comm}
-
-There are many choices of wireless communication techniques that we could consider to accompany IR's directional targeting.
-WiFi, Bluetooth Low Energy (BLE) and 802.15.4 (ZigBee) radio are all candidates for such communications.
-A detailed comparison is out of the scope of this paper.
-For our prototype, we have chosen to use XBee ZB\footnote{XBee: http://www.digi.com/xbee/}(based on ZigBee protocol and 802.15.4 radio) for the mesh network because of the ease to program and integration to Arduino platform.
-The COTS adapter provides an easy-to-use interface where we could use either hardware serial pins or the software serial library to send and receive messages.
-
-As we mentioned that the communication capability has to be included to resolve the consensus view of head orientation, we use the star topology -- the node which controls IR emission acts as the PAN coordinator.
-It collects the IR received signal strength (RSS) from each targets and runs a local algorithm to determine which one is the current target.
-
-Using such an architecture for communication also makes the integration with Google Glass easier.
-For now, Google Glass is capable of WiFi and Bluetooth; so we have added a Bluetooth modem to establish connections with Google Glass.
-We use BlueSMiRF Silver (WRL-12577) for our prototype.
-
-\subsection{3D Printed Adjustable Holder}
-\label{sec:3d-print-adjust}
-To compensate for the bias of each person's line of sight with the IR emitter, we have designed the holder that can be easily attached onto Google Glass while we can adjust the beam angle flexibly.
+\section{System Design}
+We first present our system architecture and then dive into each component to justify our design choices.
 
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figures/placeholder.pdf}
-\caption{we should replace this placeholder with Claire's design and printed version.}
-\label{fig:3d-holder}
+\includegraphics[width=0.9\columnwidth]{figures/architecture.pdf}
+\caption{In our system architecture, targeting is captured by IR signal but the confirmation is over 802.15.4 radio.
+In the research prototype, users have to carry an additional microcontroller board that marshals messages between Google Glass' Bluetooth radio and the IR/802.15.4 interface.
+But our customized hardware can be also integrated into the wearable device.}
+\label{fig:architecture}
 \end{figure}
 
-\subsection{Overall System}
-\label{sec:overall-system}
-So far, we have described our choice separately, in this subsection we focus on our system architecture that stitches the components together. 
-Figure 2 is our overall architecture photo. Because we need to take command from Google Glass, we have used another microcontroller that speaks Bluetooth to communicate with the Google Glass.
+%% IR is a proper choice.
+{\bf Area Beam Cursor:} To achieve area selection, our design focuses on using IR signals to detect the general head orientation in physical space.
+Each selectable target is equipped with IR receivers.
+We choose IR for a couple of reasons:
+\begin{itemize}
+\item The signal characteristics of IR makes it suitable for directional area selection.
+In general, IR covers a viewing angle of $20^\circ$ of viewing angle (though a wider range is avaialble in the market).
+This leads to a cone shape-like area coverage and can be well-aligned with users' line of sight.
+\item IR has been extensively used for remote control. The emitters and receivers are cheap to manufacture and easy to integrate.
+\end{itemize}
 
-%% show the figure here
+%% We have chosen to use Google Glass.
+{\bf Head-worn Computing Device:} We use Google Glass among the available COTS head-worn devices because it's easy to program (Android platform) and also light-weight for everyday use.
+
+%% the following is commented out because it's hard to have it while we describe the iterative design procedure
+%% We are interested in sensing two things: 1) IR communication and 2) IR intensity.
+Common IR receivers have automatic gain control built in and hide received signal strength.
+We therefore add an IR intensity sensor.
+In a commercial application, a single sensor that both decodes and reports RSSI could be designed.
+
+{\bf Wireless communication:} Though there is a wide range of radios available for wireless communication, we made the choice of 802.15.4 and ZigBee protocol because it is specifically for mesh network and the nice integration with Arduino platform.
+However, for now Google Glass only supports WiFi and Bluetooth.
+We have to add a Bluetooth modem to bridge Glass and the rest of the system.
+\footnote{This is a compromise now, and we are planning to migrate the entire system to Bluetooth Low Energy (BLE) once Google upgrades the Glass software to Android 4.4+.
+The prototype now suffice our demand to study head orientation-based target selection.}
+
+{\bf Visual Feedback from the environment:} Each target is equipped with multiple LEDs and these LEDs are used for visual feedback during the target selection process.
+
+\ben{discussion on the choices of each}.
+This architecture was chosen for reasons of expediency and we do not claim optimality for our design decisions.
+Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
+In particular, one could substitute WiFi modules or design an all-Bluetooth network.
 

### 2014-04-14 doc/uist14/iteration2.tex
update interactino tex

@@
 We have emprically measured the intensity distribution at the receiver side, and it's shown in Figure~\ref{fig:measurement}.
+\ben{still working on it, not sure how much value we will get out of it.
+Seems useful to support the argument of the benefit though.}
 
@@
 \caption{placeholder for emperical measurement.}
+\label{fig:measurement}
+\end{figure}
+
+From the measurement above we learned that the angluar difference has a large effect on the intensity readings.
+In order to make use of the intensity, we have designed and 3D printed a repositional IR holder that can be clipped onto Glass.
+Then for different users, they can adjust it to align with their line of sight.
+\ben{need to mention in iteration 1 about the crappy tape}.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.9\columnwidth]{figures/tube_from_top.jpg}
+\caption{The repositional IR holder.}
 \label{fig:measurement}

### 2014-04-14 doc/uist14/intro.tex
refined introduction and added comments

@@
 %% from the swarm vision to the necessity of selection
-Increasingly, objects in our built environment are equipped with computation and communication, making them potential targets for user interaction. 
+The number of smart objects with computation and communication powers have grown rapidly in the past years. Therefore, potential interactive targets in our environment has also increased.
 %Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for targeted-advertisement or context-aware computing \bjoern{second part needs to be stronger}.
-To initiate such interaction requires {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.
+To initiate such interaction requires {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world. \sean{this sentence is unclear.}
 
@@
 Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}.
-While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed; that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen or keyboard); and that the user's visual attention is now split between looking down at a screen and out at targets in the world.
+While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and consciously \sean{or do you mean cautiously?} aimed; that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen or keyboard); and that the user's visual attention is split between looking down at a screen and out at targets in the world.
 
 %% introducing head-worn computing and head orientation
-Emerging head-worn computing devices have the potential to overcome some of these limitations: because they are worn, they do not require retrieval; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information adjacent to physical objects in the wearer's field of view.
-We thus investigate the research question how such computing devices may be used for selection of physical targets in space.
+Emerging head-worn computing devices have the potential to overcome some of these limitations: they do not require retrieval since it is already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information adjacent to physical objects in the wearer's field of view.
+We thus investigate the research question of how such computing devices may be used for selection of physical targets in the space.
 
-Because they are head-worn, such devices can naturally exploit the user's head orientation, an important, but also imprecise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus.
-While gaze tracking could provide further information about a user's focus, wearable gaze tracking cannot by itself reliably determine the identity of objects in focus.
+Head-worn devices can naturally exploit the user's head orientation, an important, but also a less precise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus.
+While gaze tracking could provide further information about a user's focus, wearable gaze tracking cannot by itself reliably determine the identity of objects in focus (e.g.
+it also needs to know what the user is seeing).
 Equipping all devices with stationary gaze trackers is possible~\cite{vertegaal2005media} but much costlier than our approach.
@@
 The imprecision of human head movement suggests adapting area cursor techniques known from assistive devices~\cite{kabbash1995prince,worden1997making,findlater2010enhanced}.
-Such techniques employ a two-step selection process: {\em coarse} selection of an area of interest, followed by {\em refinement} to select a target within that region.
+Such techniques employ a two-step selection process: a {\em coarse} selection for an area of interest, followed by a {\em refinement} to select a target within that area.
 
 In this paper, we describe the iterative development and evaluation of an area-selection technique that can be readily implemented with small hardware changes to emerging wearable devices.
-We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infared (IR) communication between Glass and target appliances.
+We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infrared (IR) communication between Glass and target appliances.
 The cone shape of light emitted by an IR LED (a diameter of 60-120cm and distance up to 6m) provides the {\em coarse} selection area.
@@
 
-Noting that the refinement UI significantly slowed interaction times in this study, we contribute an improved refinement technique in which target objects compare received IR signal intensity.
-This intensity is both used to make a default selection, and to order the fallback manual UI list\bjoern{rephrase}.
-A second study with $n$ participants shows that IR-intensity disambiguation successfully reduces list navigation for refinement and speeds up selection when compared to the naive technique from the first study.
+Learning that the refinement UI significantly slowed acquisition times in this study, we designed an improved refinement technique in which target objects compare received IR signal intensity.
+This intensity value allows the system to automatically selects a target when there is a clear winner.
+If disambiguation is still required, the fall-back manual UI list is shown with the targets sorted by their intensity values.
+A second study with $n$ participants shows that IR-intensity disambiguation successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation, comparing to the naive technique from the first study.
 
-A shortcoming discovered in the second study is that participants dislike switching between focusing on targets in the world and graphics on their near-eye display during the selection process.
-We solve this problem by inferring a spatial data structure of target locations using Glass' orientation sensors and allow small head tilt motions to move selections during the refinement stage (e.g., nod down to select next target down, turn right to select next target on the right).
+One shortcoming discovered in the second study was that participants did not like switching their focuses between the targets in the physical world to the GUI on their near-eye display during the disambiguation process.
+We intend to solve this problem by introducing the third technique - disambiguate via head orientation.
+The system infers a spatial data structure of target locations using Glass orientation sensors.
+When disambiguating, Glass senses user head movements and select targets accordingly similar to a joystick controlled by your neck (e.g., nod down to select the target below current selection, tilt right to select next target on the right).
+\sean{need a summary of 3rd study here?}
+
 %% three ways
 %---
-We also demonstrate and example application of our technique for a remote control of smart appliances: a user looks at the appliance she wishes to control and confirms selection by tapping - an appliance-specific user interface is then shown on the user's near-eye display.
+We also demonstrate an example application of our technique used as a remote control of smart appliances: a user looks at the appliance he wishes to control and confirms selection by tapping.
+An appliance-specific user interface is then shown on the user's near-eye display for further interactions.
 

### 2014-04-14 doc/uist14/interaction.tex
update interactino tex
refined interaction and added comments

@@
 
-In this section, we first describe how head orientation-based interaction works from the users' point of view (Figure~\ref{fig:intearaction}).
+In this section, we describe how head orientation-based interaction works from the users' point of views (Figure~\ref{fig:interaction}).
+\sean{do we want to just use the term Glass or explains that it stands for any head-worn computing devices?}
 
 {\bf Scan:} The user first scans the environment to locate the position of the target.
-During this stage, Glass will constantly sends out IR signal, and the target offers immediate feedback when it receives the IR.
+During this stage, Glass constantly sends out IR signals, and the target offers immediate feedback when it receives the signal.
 
 {\bf Coarse Selection:} The user confirms his desire to connect to a target by tapping on the Glass touchpad.
-At this time, Glass collects the replies from targets about the IR signal reception.
-If there is only a single target in IR range, then it's considered as a successful selection.
+Glass collects the responses from targets regarding their IR signal reception.
+If there is only one single target in IR range, then it's automatically connected and the selection process is completed.
 However, in a relative dense environment where multiple targets are within range, the user needs to perform a fine selection.
@@
 {\bf Fine selection:} When disambiguation is needed, the user needs to make an explicit selection among the targets that within his view range.
-There are multiple ways to achieve it -- either from Glass touchpad that controls a UI list or using the head motion sensor on Glass.
-We will leave a thorough and complete discussion later.
-Once the user confirms the current selection, he could tap to complete the acquisition.
-Since this part is effectively disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage.
+There are multiple ways to achieve it -- either from Glass touchpad that controls a GUI list or using the head motion sensor on Glass.
+We will provide a thorough and complete discussion later in the paper.
+Once the user confirms the current selection, he could tap to complete the fine selection process.
+Since the purpose of this stage is to effectively disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage.
 

### 2014-04-14 doc/uist14/intro.tex
refined interaction and added comments

@@
 %Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for targeted-advertisement or context-aware computing \bjoern{second part needs to be stronger}.
-To initiate such interaction requires {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world. \sean{this sentence is unclear.}
+To initiate such interaction requires {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.
+\sean{this sentence is unclear to me.}
 
@@
 Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}.
-While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and consciously \sean{or do you mean cautiously?} aimed; that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen or keyboard); and that the user's visual attention is split between looking down at a screen and out at targets in the world.
+While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed; that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen or keyboard); and that the user's visual attention is split between looking down at a screen and out at targets in the world.
 

### 2014-04-14 doc/uist14/iteration1.tex
refined iteration1, didn't touch qualitative feedback

@@
 \section{Iteration 1: IR vs. List}
-Our system design uses IR as the beam cursor for physical target selection because the area coarse selection will remove the targets outside of users' view range.
-And this could help during the fine selection time because there is a smaller set of targets for fine selection.
-The opposite is to have a complete list of all targets when there is no single coarse selection and the disambiguation is always needed.
+Our system design uses IR as the beam cursor for physical target acqusition because the area coarse selection filters out the targets outside of the user's view range.
+This help during the fine selection time because there is a smaller set of targets to select from.
+The opposite is to have a complete list of all targets when there is no coarse selection and the disambiguation is always needed.
 This brought up the first question that we asked ourselves -- {\em how much does using IR help in reducing the selection time}.
@@
 In the within-subject study, half of them performed {\em IR selection} first and the other half did {\em list selection} first.
-For each selection condition, we conducted 15 target selection by random choosing from all the targets.
+For each selection condition, we conducted 15 target selection by randomly choosing from all the targets.
 During the study, we measure the {\bf target acquisition time} for each target selection, and afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
@@
 
-To further understand if the gain comes from the coarse selection, we splited the data from {\em IR selection} to two parts -- ones that require disambiguation and ones that don't.
-The result is shown in Figure~\ref{fig:selection-time}A.
+To further understand if the gain comes from the coarse selection, we splited the data from {\em IR selection} into two parts -- ones that require disambiguation and ones that don't.
+The result is shown in Figure~\ref{fig:selection-time}B.
 It shows that it takes 6.40 seconds (on average) to complete the connection if coarse selection has achieved the acquisition, while when multiple devices are in range, the time increases to 9.16 seconds, indicating additional 2.76 seconds for disambiguation.
@@
 From this figure, we can see that once the target's order is larger than 6, the average acquistion time will be larger than the case when disambiguation is needed in {\em IR selection}.
-Therefore, we argue that once there are more targets in the physical space, the {\em list selection} is not scalable.
+Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable.
 

### 2014-04-15 doc/uist14/iteration2.tex
refined iteration2, still needs work

--- 2014-04-14 doc/uist14/iteration2.tex
@@
 \section{Iteration 2: Binary IR vs. Valued IR}
+\sean{the term Binary IR and Valued IR can be refined.}
+\subsection{Problem Statement and Goal}
+With our first study, we learn that using IR to perform {\em coarse selection} can reduce the overall target selection time.
+A major reason for the improvement is because disambiguation is only needed 10\% of the time.
+%because in many cases there will be only one target in range and fine selection is not needed, and in cases where it’s needed, the list length has been effectively reduced. 
+However, if the environment becomes denser, more disambiguation will be needed.
+And learning from Figure~\ref{fig:selection-time}B, we know the average results when disambiguation is involved is significantly increased.
+Hence our next research question becomes ``can we improve selection time in a dense environment?''
 
-With our first study, we have learned that with {\em coarse selection}, we can reduce the overall target selection time because in many cases there will be only one target in range and {\em fine selection} is not needed, and in cases where it's needed, the list length has been effectively reduced.
-However, in a dense environment, there could be more cases when disambiguation is needed.
-
-From Figure~\ref{fig:selection-time}B, the average results when disambiguation is involved is significantly increase. We challenge ourselves that ``can we do a better job in a dense environment?''
-
-Previously we are only using IR reception as a signal for potential target identification.
-It's a binary result indicating that whether a target has received IR signal or not.
-But actually the IR intensity at the receiver side can provide more information than just the binary reception.
-The illuminance falls when the distance increases as well as when the angle inceases.
-We have emprically measured the intensity distribution at the receiver side, and it's shown in Figure~\ref{fig:measurement}.
+\subsection{Technique}
+Previously we only used IR reception as a signal for identifying potential targets.
+It's a binary value indicating whether a target has received IR signal or not.
+To further capture the user's attention, we believe that IR intensity at the receiver side can provide more information.
+The IR intensity falls when the distance between IR emitter and receiver increases as well as when the receiving angle increases.
+We have empirically measured the intensity distribution at the receiver side, and it's shown in Figure~\ref{fig:measurement}.
+\ben{still working on it, not sure how much value we will get out of it.
+Seems useful to support the argument of the benefit though.}
 
@@
 
+From the measurement above we learned that the angular difference has a large effect on the intensity readings.
+In order to utilize the intensity, we have designed and 3D printed a repositional IR holder that can be clipped onto Glass.
+Thus, different users can adjust the emitter to align with their line of sight.
+\ben{need to mention in iteration 1 about the crappy tape}.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.9\columnwidth]{figures/tube_from_top.jpg}
+\caption{The repositional IR holder.}
+\label{fig:measurement}
+\end{figure}
+
 The intensity information is used in two ways:
@@
 \item When multiple targets have received IR signal and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half, which is frequently used because it indicates a 3dB loss in the signal strength.}.
-In this case, when Binary IR doens't need disambiguation, the new design still doesn't need it.
+Therefore, when there is only one target within line of sight, the IR intensity approach has the same behavior as the previous iteration - no disambiguation is needed.
 When the environment turns denser, the new design can filter some peripheral targets out, making the chances of entering disambiguation lower.
-\item When disambiguation is still needed, we use the IR intensity as an indication of likelihood to sort the list.
+\item When disambiguation is still needed, meaning that multiple targets have relatively close intensity values, the system sorts the disambiguation list according to the IR intensity, from strongest to weakest.
 This will help us reduce the $t_{disambiguation}$ significantly if the first one appears matches the actual target.
-Given the imprecision of head orientation, this can hardly be always perfect.
 \end{enumerate}
 
-Though apparently with more information, we should be able to reduce the selection time.
-But the question we seek to answer is to further quantify such gain.
-We then performed another user study to compare the {\em Binary IR} and {\em Valued IR} designs.
-
-For this study, since we are aiming at a denser environment, though we keep the node number to 10, but the environment setup has changed (see Figure~\ref{fig:study-layout2}).
-In total, we have XX users for this study \ben{more information about users later.} Each user performs the target acquisition study 30 times for each condition.
-Again half of them perform {\em Binary IR} first and the other half {\em Valued IR} first.
+We seek to quantify the improvement of such design and performed another user study to compare the {\em Binary IR} and {\em Valued IR} approaches.
+For this study, we aim to discover the performances at a denser environment.
+Thus, we take the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}).
+In total, we have XX users for this study \ben{more information about users later.} Each user performs the target acquisition study 30 times for each approach.
+Similarly, half of them perform {\em Binary IR} first and the other half {\em Valued IR} first.
 
@@
 
-We then summarize the result of this study. 
+\subsection{Evaluation}
+Below is a summary of the study result.
 \begin{enumerate}
@@
 We perform a Chi-square test on this result and the result is $\chi^2(1) = 20.611$, $p < 0.0001$.
-\ben{not sure about the degree of freedom being 1 here, will check later.} This demonstrates that the first intensity gain is from the reduction of chances when disambiguation is needed.
-\item In the cases when disambiguation is needed, the {\em Valued IR} will sort the list based on the intensity reading.
-Indeed, out of the 65 times when disambiguation is involved, there are 38 (roughly 58\%) are right for {\em Valued IR}.
-However, for {\em Binary IR}, only 37 out of 104 (~35\%) are right.
+\ben{not sure about the degree of freedom being 1 here, will check later.} This demonstrates that the new approach successfully reduces the chances where disambiguation is needed.
+\item In the cases when disambiguation is inevitable, the {\em Valued IR} will sort the list based on the intensity reading.
+As a result, out of the 65 times when disambiguation is involved, {\em Valued IR} sorts the desired target as the first one in the list 38 times (roughly 58\%).
+In comparison, for {\em Binary IR}, only 37 out of 104 (~35\%) times can the user find the desired target first in the list.
 Again a Chi-square test demonstrates the statistical differences (with $\chi^2(1) = 8.487$, $p = 0.0036$.
@@
 
-However, one side effect that we have abserved in this approach is that the {\em Valued IR} sometimes eliminates the right target during the filter stage.
-The rate is not that high (16 out of 150 cases, ~90\%), but this is the price to pay for a quicker average in this smarter system.
+However, one side effect that we have observed in this approach is that the {\em Valued IR} sometimes eliminates the desired target during the {\em coarse selection} stage.
+The rate is not high (16 out of 150 cases, ~10\%), but this is the trade-off for the overall faster acquisition time.
 

### 2014-04-15 doc/uist14/iteration1.tex
refining iteration 1, half way done.

@@
 \section{Iteration 1: IR vs. List}
-Our system design uses IR as the beam cursor for physical target acqusition because the area coarse selection filters out the targets outside of the user's view range.
+\subsection{Problem Statement and Goal}
+Our system design uses IR as the beam cursor for physical target acquisition because the area coarse selection filters out the targets outside of the user's view range.
 This help during the fine selection time because there is a smaller set of targets to select from.
@@
 To answer that question and quantify the benefit of having IR for the coarse selection, we carried out our first target acquisition study which compares the {\em IR selection} and {\em list selection}.
+
+\subsection{Technique}
 
@@
 
+\subsection{Evaluation}
 Our results indicate that {\em IR selection} is generally better than {\em list selection}.
@@
 Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable.
+Further more, in real-life scenarios, targets would have longer names than an alphabet letter.
+This would increase the searching and text comparison time and also reduce the number of targets being shown in the UI screen.
 

### 2014-04-15 doc/uist14/iteration3.tex
briefly refined the structure. wait until the study is done.

@@
 \label{sec:iteration-3:-head}
-
-Though our {\em Valued IR} system is already performing well (with 3.71 seconds on average in a relative dense environment), we are only using head-orientation information during the {\em coarse selection} stage.
+\subsection{Problem Statement and Goal}
+Though our {\em Valued IR} system is already performing well (with 3.71 seconds on average in a relative dense environment), head-orientation information is only used during the {\em coarse selection} stage.
 In the {\em fine selection} part, we rely on the near-eye display.
-This has one fundamental problem -- we are projecting a high dimension distribution to a 1-dimension list.
-Therefore, it motivates us to explore using head orientation also during the {\em fine selection} stage.
+This has two problems: First, we are projecting a high dimension distribution to a 1-dimension list.
+Second, this requires the user to switch their focus back and forth between the physical world and the near-eye display.
+Therefore, it motivates us design an approach that also utilize head orientation during the {\em fine selection} stage, and that there can be a more direct mapping between the user's gestures and target traversal.
 
@@
 
+\subsection{Technique}
 The sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of users' head.
@@
 The relative orientation information -- {\em adjacent map} -- has to be learned before we can use it.
-Such a learning process can either be performed a prior, or through demonstration.
+Such a learning process can either be performed in advance, or through demonstration.
 \ben{need better elaboration here}.
@@
 
+\subsection{Evaluation}
 \ben{We then need evaluation results here, quantitative and qualitative.}

### 2014-04-15 doc/uist14/iteration3.tex
Iteration 3: Description of orientation system
* Added \achal macro for comment
* Revised most of iteration 3

--- 2014-04-14 doc/uist14/iteration3.tex
@@
 \label{sec:iteration-3:-head}
-
-Though our {\em Valued IR} system is already performing well (with 3.71 seconds on average in a relative dense environment), we are only using head-orientation information during the {\em coarse selection} stage.
-In the {\em fine selection} part, we rely on the near-eye display.
-This has one fundamental problem -- we are projecting a high dimension distribution to a 1-dimension list.
-Therefore, it motivates us to explore using head orientation also during the {\em fine selection} stage.
+\subsection{Problem Statement and Goal}
+Though our {\em Valued IR} system already performs well (with 3.71 seconds on average in a relative dense environment), head-orientation information is only used during the {\em coarse selection} stage.
+In the {\em fine selection} phase, we rely solely on the near-eye display.
+This has two problems: First, we are projecting a high dimension distribution to a 1-dimension list.
+Second, this requires the user to switch their focus back and forth between the physical world and the display.
+This issue motivated us to design an approach that also utilizes orientation information during the {\em fine selection} stage, so that there would be a more direct mapping between the user's gestures and target selection.
 
@@
 
-The sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of users' head.
-However, such absolute orientation cannot be directly used in any indoor environment because the user might change his location.
+\subsection{Technique}
+The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of users' head.
+However, this absolute orientation cannot be directly used in any indoor environment as the user may change his location \achal{Should we use gender neutral language?}.
 However, as long as the targets are spread around the periphery, their relative orientation doesn't change.
@@
 
-Before we describe how this new technique works.
-The relative orientation information -- {\em adjacent map} -- has to be learned before we can use it.
-Such a learning process can either be performed a prior, or through demonstration.
-\ben{need better elaboration here}.
+There are two phases to this technique: learning the relative orientations and intelligently suggesting targets during disambiguation.
 
-In this hybrid approach, in additional to use IR to reduce the chance of disambiguation and prepare a better list, the IR information helps us identify a base point.
-When the user wants to use his head-motion for disambiguation, he will put his finger on the touchpad to enter a quasi-mode.
-Such a gesture comes from a real-world metaphor that people tend to adjust their glasses for a closer inspection.
-In such quasi-mode, we monitor the changes of his head orientation, and infer the intended next targets according to some algorithm \ben{haven't described yet}.
+In the first stage, the user orients his head towards each device, and the system saves the absolute orientations of each device.
+Once the user has added all devices, these orientations are saved for future use.
+It is still possible to add new devices at any point without running the entire calibration step, as the system will simply store the absolute orientation of the new device.
 
+During the second stage, when the user wants to use his head-motion for disambiguation, he will put his finger on the touchpad to enter a quasi-mode.
+This gesture is inspired by observing users who tend to adjust their glasses for closer inspection.
+
+In this mode, only one device has a solid indicator light on at a time.
+The user can then tilt in the direction of the device he wishes to switch to.
+We calculate the direction of movement using a low-passed history of orientations, and look for devices that lay in that direction.
+Once a new device is selected, the frame of reference changes, and the user can repeat the process.
+Note that in reality, this is a completely smooth operation: it is possible to switch between multiple devices in one continuous motion.
+This is achieved by automatically adjusting the frame of reference to the currently lit device, but also creating a ``sticky" threshold so that the user is not constantly jumping between devices.
+
+\achal{Not sure if there should be a conclusion of some sort here.}
+
+\subsection{Evaluation}
 \ben{We then need evaluation results here, quantitative and qualitative.}

### 2014-04-15 doc/uist14/iteration2.tex
updated the iteration 2 results from today's study

@@
 Thus, we take the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}).
-In total, we have XX users for this study \ben{more information about users later.} Each user performs the target acquisition study 30 times for each approach.
-Similarly, half of them perform {\em Binary IR} first and the other half {\em Valued IR} first.
+In total, we have 10 users for this study, and each user performs 30 target acquisition tasks for each approach.
+Similarly to our first within-subject study, half of them perform {\em Binary IR} first and the other half {\em Valued IR} first.
 
@@
 
+% new study results
+% ceiling(c(nrow(name), nrow(name_multiple), wrong_guess, wrong_percentage * 100))
+% intensity
+%  300 167  74  45
+% name
+%  300 225 145  65
+
+%% first one
+%% 
+%             Outcome 1	Outcome 2	     Total
+% Group 1 	167	133	300
+% Group 2	225	75	300
+% Total	        392	208	600
+%% Chi squared equals 24.755 with 1 degrees of freedom. 
+%%  The two-tailed P value is less than 0.0001
+
+% 	    Outcome 1	Outcome 2	     Total
+% Group 1	74	93	167
+% Group 2	145	80	225
+% Total	        219	173	392
+% Chi squared equals 15.758 with 1 degrees of freedom. 
+%   The two-tailed P value is less than 0.0001
+
 \subsection{Evaluation}
@@
 \begin{enumerate}
-\item Among the 150 trials in {\em Binary IR} condition, there are 104 times when disambiguation is needed.
-In {\em Valued IR} condition, only 65 out of 150 trials need disambiguation.
-We perform a Chi-square test on this result and the result is $\chi^2(1) = 20.611$, $p < 0.0001$.
-\ben{not sure about the degree of freedom being 1 here, will check later.} This demonstrates that the new approach successfully reduces the chances where disambiguation is needed.
+\item Among the 300 trials in {\em Binary IR} condition, there are 225 times when disambiguation is needed.
+In {\em Valued IR} condition, only 167 out of 300 trials need disambiguation.
+We perform a Chi-square test on this result and the result is $\chi^2(1) = 24.755$, $p < 0.0001$.
+This demonstrates that the new approach successfully reduces the chances where disambiguation is needed.
 \item In the cases when disambiguation is inevitable, the {\em Valued IR} will sort the list based on the intensity reading.
-As a result, out of the 65 times when disambiguation is involved, {\em Valued IR} sorts the desired target as the first one in the list 38 times (roughly 58\%).
-In comparison, for {\em Binary IR}, only 37 out of 104 (~35\%) times can the user find the desired target first in the list.
-Again a Chi-square test demonstrates the statistical differences (with $\chi^2(1) = 8.487$, $p = 0.0036$.
+As a result, out of the 167 times when disambiguation is involved, {\em Valued IR} sorts the desired target as the first one in the list 93 times (roughly 55\%).
+In comparison, for {\em Binary IR}, only 80 out of 219 (35\%) times can the user find the desired target first in the list.
+Again a Chi-square test demonstrates the statistical difference (with $\chi^2(1) = 15.758$, $p < 0.0001$.
 \end{enumerate}
@@
 Such gain can be seen from Figure~\ref{fig:study2} where the overall target acquisition time is obtained.
-The average time is 4.32 seconds and 3.71 second for {\em Binary IR} and {\em Valued IR}, respectively; and a t-test shows that they are also significantly different $t(282)=2.0922$, $p=0.03$.
+The average time is 4.31 seconds and 3.64 second for {\em Binary IR} and {\em Valued IR}, respectively; and a t-test shows that they are also significantly different $t(555)=3.2945$, $p=0.001$.
+
+%  t.test(name$complete, intensity$complete)
+
+% 	Welch Two Sample t-test
+
+% data:  name$complete and intensity$complete
+% t = 3.2945, df = 555.272, p-value = 0.001049
+% alternative hypothesis: true difference in means is not equal to 0
+% 95 percent confidence interval:
+%  0.2707875 1.0704747
+% sample estimates:
+% mean of x mean of y 
+%  4.311340  3.640709
 
@@
 However, one side effect that we have observed in this approach is that the {\em Valued IR} sometimes eliminates the desired target during the {\em coarse selection} stage.
-The rate is not high (16 out of 150 cases, ~10\%), but this is the trade-off for the overall faster acquisition time.
+The rate is not high (13 out of 300 cases, 4.3\%), but this is the trade-off for the overall faster acquisition time.
 

### 2014-04-15 doc/uist14/abstract.tex
pass through related work

@@
 
-\ben{This parts needs to be changed based on the current flow of the story.}
+\bjoern{updated: }
 Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
 %
-We introduce and characterize a method for selecting and interacting with physical targets at a distance using head orientation.
+We present the iterative design and evaluation of \systemname -- a head-orientation based selections method for such devices, for interacting with physical targets at a distance.
 We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.
-We present our iterative design process that has shown with IR, it outperforms a pure list selection.
-And with IR intensity, the time can be further reduced.
-If head motion is added, we can infer a spatial data structure of the nodes to enable a natural head-motion based disambiguation.
-We demonstrate an application of our selection technique for controlling smart appliances and report qualitative user impressions.
+Our first design shows that a naive IR implementation can outperform list selection, but has poor performance when refinement between multiple targets is needed.
+A second design uses IR intensity measurement at targets to improve refinement.
+To address a lack of natural mapping of on-screen target lists to spatial target location, our third design infers a spatial data structure of the targets from demonstration to enable a natural head-motion based disambiguation.
+%
+We also demonstrate a universal remote control application of the technique and report qualitative user impressions.

### 2014-04-15 doc/uist14/intro.tex
keep wording consistent
pass through related work

@@
 \section{Introduction}
-
+\bjoern{something happened to the figure caption font - it doesn't match the standard template. Please fix.}
 %% from the swarm vision to the necessity of selection
-The number of smart objects with computation and communication powers have grown rapidly in the past years. Therefore, potential interactive targets in our environment has also increased.
-%Such interactions might be {\em explicit} -- controlling smart appliances such as intelligent lighting, AV equipment, or HVAC systems; or {\em implicit} -- tracking the user's attention to gauge interest e.g., for targeted-advertisement or context-aware computing \bjoern{second part needs to be stronger}.
-To initiate such interaction requires {\em selection} information by which a system keeps track of the user's {\em locus of attention}~\cite{raskin} in the world.
-\sean{this sentence is unclear to me.}
+The number of smart objects in our environment with embedded computation and communication has grown rapidly. These objects are all potential targets for interaction.
+To initiate such interaction requires {\em selection} information about the object at the user's {\em locus of attention}~\cite{raskin}. \sean{this sentence is unclear to me.}
 
@@
 Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}.
-While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed; that the user's hands are required to be free for operation (one to hold the device, one to operate the touch screen or keyboard); and that the user's visual attention is split between looking down at a screen and out at targets in the world.
+While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed; that the user's hands are required to be free for operation; and that the user's visual attention is split between looking down at a screen and out at targets in the world.
 
 %% introducing head-worn computing and head orientation
-Emerging head-worn computing devices have the potential to overcome some of these limitations: they do not require retrieval since it is already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information adjacent to physical objects in the wearer's field of view.
+Emerging head-worn computing devices have the potential to overcome some of these limitations: they do not require retrieval since the devices are already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information adjacent to physical objects in the wearer's field of view.
 We thus investigate the research question of how such computing devices may be used for selection of physical targets in the space.
@@
 The imprecision of human head movement suggests adapting area cursor techniques known from assistive devices~\cite{kabbash1995prince,worden1997making,findlater2010enhanced}.
-Such techniques employ a two-step selection process: a {\em coarse} selection for an area of interest, followed by a {\em refinement} to select a target within that area.
+Such techniques employ a two-step selection process: a {\em coarse} selection of an area of interest, followed by a {\em refinement} to select a target within that area.
 
-In this paper, we describe the iterative development and evaluation of an area-selection technique that can be readily implemented with small hardware changes to emerging wearable devices.
+In this paper, we describe the iterative development and evaluation of \systemnamenospace, an area-selection technique that can be readily implemented with small hardware changes to emerging head-worn devices.
 We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infrared (IR) communication between Glass and target appliances.
-The cone shape of light emitted by an IR LED (a diameter of 60-120cm and distance up to 6m) provides the {\em coarse} selection area.
-To {\em refine} selection when multiple targets have received IR signals, we first introduce a manual technique where users select a candidate device using a graphical interface on their wearable device.
-We then conduct a study with $14$ participants that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface.
-We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller.
+In our {\em Naive IR} technique, the cone of light emitted by an IR LED (a diameter of 60-120cm and distance up to 6m) provides the {\em coarse} selection area.
+To {\em refine} selection when multiple targets have received IR signals, we introduce a manual list selection disambiguation technique.
+A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting.
+We find that target acquisition through head orientation is preferred by users and is faster than list selection, but refinements -- if necessary -- are slow.
 
-Learning that the refinement UI significantly slowed acquisition times in this study, we designed an improved refinement technique in which target objects compare received IR signal intensity.
-This intensity value allows the system to automatically selects a target when there is a clear winner.
-If disambiguation is still required, the fall-back manual UI list is shown with the targets sorted by their intensity values.
-A second study with $n$ participants shows that IR-intensity disambiguation successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation, comparing to the naive technique from the first study.
+In response, we designed an improved refinement technique, {\em Intensity IR}, in which target objects compare received IR signal strength.
+This value allows the system to either automatically select the target with the strongest intensity, or to re-order the refinement interface's list by their intensity values.
+A second study with $10$ participants shows that {\em Intensity IR} disambiguation successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.
 
-One shortcoming discovered in the second study was that participants did not like switching their focuses between the targets in the physical world to the GUI on their near-eye display during the disambiguation process.
-We intend to solve this problem by introducing the third technique - disambiguate via head orientation.
-The system infers a spatial data structure of target locations using Glass orientation sensors.
-When disambiguating, Glass senses user head movements and select targets accordingly similar to a joystick controlled by your neck (e.g., nod down to select the target below current selection, tilt right to select next target on the right).
-\sean{need a summary of 3rd study here?}
+A final design addresses the lack of a natural mapping when users select a target in the refinement step using their device's touchpad - the axes of motion do not map directly to the spatial layout of target devices in a room.
+Our {\em Head-motion Refinement} technique first infers a spatial data structure of target locations using Glass' orientation sensors.
+In a refinement step, users can then perform head movements to change selections to spatially adjacent targets.
+For example, nodding down to select the target below current selection, or tilting right to select next target on the right.
+We present initial feedback on this technique.
 

### 2014-04-15 doc/uist14/relatedwork.tex
pass through related work

--- 2014-04-14 doc/uist14/relatedwork.tex
@@
 %Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}.
-Prior work often focused on used head orientation for control of graphical interfaces; in contrast, we apply this modality to selection in physical spaces.
+Prior work often focused on head orientation for controling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.
 
@@
 In all area cursors, the bigger cursor activation area can lead to multiple targets being selected and disambiguation is needed.
-We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, users move so the activation area intersects with the target object (and possibly other, unintended targets).
+We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets).
 In the {\em refinement} phase, they adjust so only the intended target will be selected.
-Many disambiguation techniques are possible for refinement.
-We contribute a novel physical method based on received light IR intensity.
+Many disambiguation techniques are possible for refinement -- this paper describes the tradeoffs between several of them.
 
@@
 
-\bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
+Other targeting systems use IR with handheld pointers ~\cite{swindells_that_2002} and as well as wearable devices such as rings and Bluetooth audio earpieces ~\cite{merrill_augmenting_2007} to connect to smart devices.
+Our system tackles an unresolved issue of such IR-based approaches-- navigating an area dense with potential targets.

### 2014-04-15 doc/uist14/interaction.tex
bjoern's pass

--- 2014-04-14 doc/uist14/interaction.tex
@@
-\section{Interaction Models}
+%!TEX root = uist14.tex
+\section{IR for head orientation sensing}
+\subsection{Hardware}
+A central hypothesis of this paper is that the area cursor paradigm~\cite{kabbash1995prince} is well matched to head orientation selection.
+Head orientation input is imprecise because it does not capture eye movement and has to rely on a low-bandwidth muscle group.
+Therefore, point selection techniques like laser pointers are not appropriate.
 
-In this section, we first describe how head orientation-based interaction works from the users' point of view (Figure~\ref{fig:intearaction}).
+Infrared emitters are a good technology match since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible.
+LEDs with many different fields of view are commercially available.
+\bjoern{say something about receivers and how point to point connections enable us to leverage head orientation without needing a map.}
+
+Another benefit is the low cost and easy integration of both emitters and detectors - detectors often have support for common IR protocol decoding built in and cost cents, making it economically feasible to add detectors to dozens or hundreds of devices in an environment.
+
+Our augmentation to Google Glass adds a \bjoern{add modelnumber and spec details} LED (see Figure~\ref{fig:glass}.
+In the following description, we refer to Glass because the term is succinct; but  other suitable head-worn computing devices could be substituted.
+
+\bjoern{need to mention that all the target nodes also have to be able to talk to each other and to glass outside the IR channel. We use XBees for this purpose.}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.9\columnwidth]{figures/tube_from_top_annotated.pdf}
+\caption{Google Glass augmented with a repositionable IR holder.}
+\label{fig:glass}
+\end{figure}
+
+\subsection{Interaction Model}
+From the user's perspective, interaction with \systemname proceeds in two stages (Figure~\ref{fig:interaction}):
 
 {\bf Scan:} The user first scans the environment to locate the position of the target.
-During this stage, Glass will constantly sends out IR signal, and the target offers immediate feedback when it receives the IR.
+During this stage, Glass constantly sends out IR signals, and  targets offer immediate feedback when they receive a signal.
+The user confirms his desire to connect to a target by tapping on the Glass touchpad.
+Glass collects the responses from targets that have received IR reception (we use an XBee-based backchannel for this traffic).
+If there is only one single target in IR range, it is automatically selected.
+However, in a dense environment where multiple targets are within range, the user needs to perform a refinement.
 
-{\bf Coarse Selection:} The user confirms his desire to connect to a target by tapping on the Glass touchpad.
-At this time, Glass collects the replies from targets about the IR signal reception.
-If there is only a single target in IR range, then it's considered as a successful selection.
-However, in a relative dense environment where multiple targets are within range, the user needs to perform a fine selection.
-
-{\bf Fine selection:} When disambiguation is needed, the user needs to make an explicit selection among the targets that within his view range.
-There are multiple ways to achieve it -- either from Glass touchpad that controls a UI list or using the head motion sensor on Glass.
-We will leave a thorough and complete discussion later.
-Once the user confirms the current selection, he could tap to complete the acquisition.
-Since this part is effectively disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage.
+{\bf Refine:} When disambiguation is needed, the user needs to make an explicit selection among the targets that within his view range.
+We have designed multiple refinement mechanisms -- all enable the user to select one from a subset of targets.
+The user confirms a current selection a tap.
+Since the purpose of this stage is to disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage.
+Again, a tap confirms a selection.
 
@@
 \centering
-\includegraphics[width=\columnwidth]{figures/interactionModel.pdf}
+\includegraphics[width=\columnwidth]{figures/interactionModel2.pdf}
 \caption{Four different states during the head-orientation targeting.}
@@
 
-From the flow of interaction, we learned that the overall target acquisition time can be broken down into the following pieces:
+The overall target acquisition time thus depends on scan and refine times; the probability that refinement is needed; and the time to commit an action (tap):
 \begin{equation}
-t_{total}=t_{scan}+t_{coarse}+t_{disambiguation}
+t_{total}=t_{scan}+P(refine)*t_{refine}+t_{commit}
 \end{equation}
 
-Our goal of this paper is to minimize the overall target selection time. In the following few sections, we will describe our iterative design process during which we have explored the design space.
+In the following sections, we describe our iterative design and evaluation process to minimize the overall target selection time.
 

### 2014-04-15 doc/uist14/iteration1.tex
bjoern's pass

@@
 
-\section{Iteration 1: IR vs. List}
+\section{Iteration 1: Naive IR}
 \subsection{Problem Statement and Goal}
-Our system design uses IR as the beam cursor for physical target acquisition because the area coarse selection filters out the targets outside of the user's view range.
-This help during the fine selection time because there is a smaller set of targets to select from.
-The opposite is to have a complete list of all targets when there is no coarse selection and the disambiguation is always needed.
-This brought up the first question that we asked ourselves -- {\em how much does using IR help in reducing the selection time}.
-To answer that question and quantify the benefit of having IR for the coarse selection, we carried out our first target acquisition study which compares the {\em IR selection} and {\em list selection}.
+Our initial research question is: {\em Can IR-based targeting reduce the selection time compared to UI list navigation on a head-worn device?}
 
 \subsection{Technique}
+In our first implementation, we use IR for scanning as described in the previous section.
+For the refinement technique, we simply show a list of the subset of targets that have received IR signals on the Glass display and let users swipe to select from that list and tap to confirm.
 
+A natural point of comparison is an interface that does not use any orientation information - it always shows 
+a complete list of all targets.
+We implemented a list view where users swiped forwards and backwards on the Glass touchpad to navigate incrementally through the list.
+To quantify the benefit of using IR for the scanning step, we carried out a target acquisition study which compares the {\em Naive IR selection} and {\em list selection}.
+
+\subsection{Method}
 We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}).
 Each of them has a number as ID and a letter representing the name.
-There are 14 participants from our institution for this study.
-13 had never used Glass before, therefore we offered a Google Glass crash course before the experiment for them to get familiar with Glass gestures.
-4 of them wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
+We recruited 14 participants from our institution for this study.
+13 had never used Glass before, so we offered a Google Glass tutorial before the experiment for them to get familiar with Glass gestures.
+Four participants wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
+In this study, the IR LED was fixed and not repositionable as shown in Figure~\ref{fig:glass}.
 
@@
 
-In the within-subject study, half of them performed {\em IR selection} first and the other half did {\em list selection} first.
+In the within-subject study, half of the participants performed {\em IR selection} first and the other half used {\em list selection} first.
 For each selection condition, we conducted 15 target selection by randomly choosing from all the targets.
-During the study, we measure the {\bf target acquisition time} for each target selection, and afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
+During the study, we measure the {\bf target acquisition time} for each target selection.
+Afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
 
@@
 
-\subsection{Evaluation}
-Our results indicate that {\em IR selection} is generally better than {\em list selection}.
-From Figure~\ref{fig:selection-time}A, the average target acquisition time for {\em IR selection} is 6.67 seconds, while for {\em list selection} it took 8.86 seconds.
-We run Student's t-test on the data and it shows significant difference ($t(279)=-3.81, p=0.00017$).
+\subsection{Results}
+Our results indicate that {\em IR selection} outperforms {\em list selection}.
+The average target acquisition time for {\em IR selection} is 6.67 seconds, while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:selection-time}A).
+Student's t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
 
-To further understand if the gain comes from the coarse selection, we splited the data from {\em IR selection} into two parts -- ones that require disambiguation and ones that don't.
-The result is shown in Figure~\ref{fig:selection-time}B.
-It shows that it takes 6.40 seconds (on average) to complete the connection if coarse selection has achieved the acquisition, while when multiple devices are in range, the time increases to 9.16 seconds, indicating additional 2.76 seconds for disambiguation.
-Though it takes significantly longer ($t(19)=-2.7827, p=0.012$ using t-test) in the latter case, they only happen 10\% of total {\em IR selection} trials.
-
-From the above result, we learned that with IR, we can have effectively reduced the overall target acquisition time.
-And the major gain comes from the fact that {\em coarse selection} helps reduce the chances of entering {\em fine selection} stage.
+To understand how scanning and refinement contribute to total selection time, we split the data from {\em IR selection} into two parts -- trials that required refinement and ones that did not.
+It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:selection-time}B).
+This difference is significant ($t(19)=-2.7827, p=0.012$ using t-test \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}).
+Because many targets were spaced far apart, refinements were only necessary in 10\% of total {\em IR selection} trials in this study.
 
@@
 Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}.
-From this figure, we can see that once the target's order is larger than 6, the average acquistion time will be larger than the case when disambiguation is needed in {\em IR selection}.
-Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable.
+From this figure, we can see that once there are more than 6 targets, the average acquistion time will be larger than IR selection, even if disambiguation is needed.
+%Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable.
 Further more, in real-life scenarios, targets would have longer names than an alphabet letter.
 This would increase the searching and text comparison time and also reduce the number of targets being shown in the UI screen.
+
+In summary, {\em Naive IR} can outperform linear list selection, but the refinement step detracts significantly from the efficiency of the technique.
+% learned that with IR, we can have effectively reduced the overall target acquisition time.
+And the major gain comes from the fact that {\em coarse selection} helps reduce the chances of entering {\em fine selection} stage.
 

### 2014-04-15 doc/uist14/iteration2.tex
bjoern's pass

@@
-\section{Iteration 2: Binary IR vs. Valued IR}
-\sean{the term Binary IR and Valued IR can be refined.}
+%!TEX root = uist14.tex
+
+\section{Iteration 2: Intensity IR}
+
 \subsection{Problem Statement and Goal}
-With our first study, we learn that using IR to perform {\em coarse selection} can reduce the overall target selection time.
+%With our first study, we learn that using IR to perform {\em coarse selection} can reduce the overall target selection time.
 A major reason for the improvement is because disambiguation is only needed 10\% of the time.
 %because in many cases there will be only one target in range and fine selection is not needed, and in cases where it’s needed, the list length has been effectively reduced. 
-However, if the environment becomes denser, more disambiguation will be needed.
-And learning from Figure~\ref{fig:selection-time}B, we know the average results when disambiguation is involved is significantly increased.
-Hence our next research question becomes ``can we improve selection time in a dense environment?''
+Performance of the {\em Naive IR} technique will degrade as target density in an environment increases, as increased density will require refinement steps.
+We therefore ask a follow-on research question: ``How might we improve selection time in a dense environment?''
 
 \subsection{Technique}
-Previously we only used IR reception as a signal for identifying potential targets.
-It's a binary value indicating whether a target has received IR signal or not.
-To further capture the user's attention, we believe that IR intensity at the receiver side can provide more information.
-The IR intensity falls when the distance between IR emitter and receiver increases as well as when the receiving angle increases.
-We have empirically measured the intensity distribution at the receiver side, and it's shown in Figure~\ref{fig:measurement}.
+Previously we only used IR reception as a binary signal for identifying potential targets.
+We hypothesize that IR intensity at the receiver side can provide more information about the likelihood that a user intended to select a particular target.
+Received IR intensity falls off with distance between IR emitter and receiver as well as with the angle between emitter and receiver.
+To measure intensity, we add \bjoern{add model number and characteristics} and replace the IR emitter with \bjoern{modelnumber, characteristics}.
+We have empirically measured the intensity distribution at the receiver for this configuration in Figure~\ref{fig:measurement}.
 \ben{still working on it, not sure how much value we will get out of it.
 Seems useful to support the argument of the benefit though.}
-
+Our measurements confirm that angular difference has a large effect on the intensity readings.
 \begin{figure}[t]
@@
 \includegraphics[width=0.9\columnwidth]{figures/placeholder.pdf}
-\caption{placeholder for emperical measurement.}
-\label{fig:measurement}
-\end{figure}
-
-From the measurement above we learned that the angular difference has a large effect on the intensity readings.
-In order to utilize the intensity, we have designed and 3D printed a repositional IR holder that can be clipped onto Glass.
-Thus, different users can adjust the emitter to align with their line of sight.
-\ben{need to mention in iteration 1 about the crappy tape}.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.9\columnwidth]{figures/tube_from_top.jpg}
-\caption{The repositional IR holder.}
+\caption{placeholder for empirical measurement.}
 \label{fig:measurement}
@@
 \begin{enumerate}
-\item When multiple targets have received IR signal and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half, which is frequently used because it indicates a 3dB loss in the signal strength.}.
+\item When multiple targets have received IR signals and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half of the ADC resolution, which is frequently used because it indicates a 3dB loss in the signal strength.}.
 Therefore, when there is only one target within line of sight, the IR intensity approach has the same behavior as the previous iteration - no disambiguation is needed.
-When the environment turns denser, the new design can filter some peripheral targets out, making the chances of entering disambiguation lower.
-\item When disambiguation is still needed, meaning that multiple targets have relatively close intensity values, the system sorts the disambiguation list according to the IR intensity, from strongest to weakest.
-This will help us reduce the $t_{disambiguation}$ significantly if the first one appears matches the actual target.
+When the environment turns denser, the new design can filter some peripheral targets out, reducing $P(refine)$, the likelihood of entering the refinement stage.
+\item When refinement is still needed, meaning that multiple targets have relatively close intensity values, the system sorts the disambiguation list according to the IR intensity, from strongest to weakest.
+We hypothesize that this will reduce time $t_{refine}$ significantly by minimizing extra navigation steps.
+(Ideally, the first list item will match the actual, intended target.)
 \end{enumerate}
 
-We seek to quantify the improvement of such design and performed another user study to compare the {\em Binary IR} and {\em Valued IR} approaches.
-For this study, we aim to discover the performances at a denser environment.
-Thus, we take the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}).
-In total, we have 10 users for this study, and each user performs 30 target acquisition tasks for each approach.
-Similarly to our first within-subject study, half of them perform {\em Binary IR} first and the other half {\em Valued IR} first.
+\subsection{Method}
+To quantify the improvement this design we performed a second study to compare the {\em Naive IR} and {\em Intensity IR} approaches.
+Because we were interested in discovering performance differences in denser environment, we reposition the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}) \bjoern{give measurements in meters}.
+We recruited 10 participants for this study.
+Each  user performed 30 target acquisition tasks for each approach.
+Similarly to our first within-subject study, half of them perform {\em Naive IR} first and the other half {\em Intesity IR} first.
 
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figures/study-layout2.jpg}
-\caption{The current environment setup \ben{needs annotation and a better photo.}.}
+\includegraphics[width=1.0\columnwidth]{figures/study-layout2.jpg}
+\caption{The current environment setup \ben{needs annotation and a better photo.} \bjoern{yes}.}
 \label{fig:study-layout2}
@@
 
-\subsection{Evaluation}
-Below is a summary of the study result.
-\begin{enumerate}
-\item Among the 300 trials in {\em Binary IR} condition, there are 225 times when disambiguation is needed.
-In {\em Valued IR} condition, only 167 out of 300 trials need disambiguation.
-We perform a Chi-square test on this result and the result is $\chi^2(1) = 24.755$, $p < 0.0001$.
-This demonstrates that the new approach successfully reduces the chances where disambiguation is needed.
-\item In the cases when disambiguation is inevitable, the {\em Valued IR} will sort the list based on the intensity reading.
-As a result, out of the 167 times when disambiguation is involved, {\em Valued IR} sorts the desired target as the first one in the list 93 times (roughly 55\%).
-In comparison, for {\em Binary IR}, only 80 out of 219 (35\%) times can the user find the desired target first in the list.
-Again a Chi-square test demonstrates the statistical difference (with $\chi^2(1) = 15.758$, $p < 0.0001$.
-\end{enumerate}
+\subsection{Results}
+{\em Intensity IR} reduces the number of trials in which refinement dialogs are needed from 225 of 300 in {\em Naive IR} to 167 of 300 trials.
+A Chi-square test shows this difference is significant ($\chi^2(1) = 24.755$, $p < 0.0001$).
+This demonstrates that the new approach successfully reduces the probability that a disambiguation dialog is needed.
+
+In the cases when disambiguation is inevitable, {\em Intensity IR} sorts the list based on the intensity reading, while {\em Naive IR} sorts alphabetically.
+{\em Intensity IR} reduces the fraction of refinement trails in which additional list navigation is necessary (i.e., the first, already selected element is incorrect).
+{\em Intensity IR} sorted the desired target as the first one in the list in 55\% of cases (93 of 167).
+In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 219\bjoern{why isn't this 225 as above?}).
+Again, a Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, $p < 0.0001$.
 
 Such gain can be seen from Figure~\ref{fig:study2} where the overall target acquisition time is obtained.
-The average time is 4.31 seconds and 3.64 second for {\em Binary IR} and {\em Valued IR}, respectively; and a t-test shows that they are also significantly different $t(555)=3.2945$, $p=0.001$.
+\bjoern{avoid passive voice throughout!} The average selection time decreases from 4.31 seconds for {\em Naive IR} to 3.64 second for {\em Intensity IR}.
+This difference is also significant ($t(555)=3.2945$, $p=0.001$.) \bjoern{for consistency's sake, can we show a box plot as in figure 5?}
 
@@
 
-However, one side effect that we have observed in this approach is that the {\em Valued IR} sometimes eliminates the desired target during the {\em coarse selection} stage.
-The rate is not high (13 out of 300 cases, 4.3\%), but this is the trade-off for the overall faster acquisition time.
+One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage.
+The this error rate (13 out of 300 cases, 4.3\%), is higher than the error rate for {\em Naive IR} \bjoern{insert error rate}.
+% but this is the trade-off for the overall faster acquisition time.
 

### 2014-04-15 doc/uist14/iteration3.tex
bjoern's pass

@@
-\section{Iteration 3: Head motion for disambiguation}
+%!TEX root = uist14.tex
+\section{Iteration 3: Head motion refinement}
 \label{sec:iteration-3:-head}
 \subsection{Problem Statement and Goal}
-Though our {\em Valued IR} system already performs well (with 3.71 seconds on average in a relative dense environment), head-orientation information is only used during the {\em coarse selection} stage.
-In the {\em fine selection} phase, we rely solely on the near-eye display.
-This has two problems: First, we are projecting a high dimension distribution to a 1-dimension list.
-Second, this requires the user to switch their focus back and forth between the physical world and the display.
-This issue motivated us to design an approach that also utilizes orientation information during the {\em fine selection} stage, so that there would be a more direct mapping between the user's gestures and target selection.
+Both{\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step.
+This has two problems: First, a spatial layout of target is compressed into a uni-dimensional list.
+List navigation commands thus have poor spatial mapping.
+Second, our techniques require the user to switch their focus back and forth between the physical world and the display.
+This issue motivated us to design an approach that also utilizes orientation information during the {\em refinement} stage, to provide a more direct mapping between the user's actions and target selection.
 
@@
 \subsection{Technique}
-The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of users' head.
-However, this absolute orientation cannot be directly used in any indoor environment as the user may change his location \achal{Should we use gender neutral language?}.
+The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of a user's head.
+However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location \achal{Should we use gender neutral language?}\bjoern{yes.
+just did in the previous sentence}.
 However, as long as the targets are spread around the periphery, their relative orientation doesn't change.
-This key observation inspires us to investigate a hybrid solution with IR and sensors on Glass.
+This key observation inspires us to investigate a hybrid solution with IR and orientation sensors on Glass.
 
-There are two phases to this technique: learning the relative orientations and intelligently suggesting targets during disambiguation.
+There are two phases to this technique: learning the relative orientations or targets in a room and intelligently suggesting targets during disambiguation.
 
@@
 In this mode, only one device has a solid indicator light on at a time.
-The user can then tilt in the direction of the device he wishes to switch to.
-We calculate the direction of movement using a low-passed history of orientations, and look for devices that lay in that direction.
+\bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} The user can then tilt in the direction of the device he wishes to switch to.
+We calculate the direction of movement using a low-pass-filtered history of orientations, and look for devices that lie in that direction.
 Once a new device is selected, the frame of reference changes, and the user can repeat the process.
@@
 This is achieved by automatically adjusting the frame of reference to the currently lit device, but also creating a ``sticky" threshold so that the user is not constantly jumping between devices.
+\bjoern{This is called hysteresis or double thresholding.}
+
+\bjoern{how well will this work if the user shifts position within a room?}
 
@@
 \ben{We then need evaluation results here, quantitative and qualitative.}
+\bjoern{got to here in my pass}
 

### 2014-04-15 doc/uist14/relatedwork.tex
bjoern's pass

@@
 \subsection{Area Cursors}
-A central hypothesis of this paper is that the area selection paradigm is well matched to head orientation input.
 In GUI area cursors, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}.

### 2014-04-15 doc/uist14/applications.tex
Filled some of the contents; updated a few figures; just added content to disucssion and conclusion, but the sentences are broken and may not make any sense

--- 2014-04-14 doc/uist14/applications.tex
@@
 %!TEX root = uist14.tex
-\section{Application Example: Universal Remote Control}
+\section{Applications}
 \label{sec:applications}
 
-%In this section we describe four possible applications that can be built on top of the head orientation-based selection.
-We focus our discussion on {\em GlassRemote} -- a universal remote control application.
+Though such head orientation targeting enables a wide range of context-aware applications, in this paper we focus on our GlassRemote system that we have built using the existing hardware platform to achieve a universal smart appliance control functionality.
 
-%\subsection{GlassRemote}
-%\label{sec:univ-remote-contr}
-
-We extended the head-orientation targeting and existing hardware platform to achieve a universal remote control functionality in smart home scenario.
-The motivation is to solve two challenges that exist in current list-based solutions -- {\em naming} and {\em scoping}.
+The huge benefit using head orientation is to solve two challenges that exist in most commercial list-based solutions -- {\em naming} and {\em scoping}.
 First, assigning clear names is non-trivial.
@@
 
-\bjoern{you need to tell us more about how the interaction works - namely, that you pull the user interface for an appliance onto the near-eye display. This may not be obvious.}
+In order to control a smart appliance, the user performs a physical target selection (scan, refine and confirm) first and then the appliance-specific UI will be shown in the near-eye display.
 
-To assess the usability of such applications when the novel interaction scheme is added, we have created a few smart appliances (including a fan, a lamp and a TV) and asked 14 participants to work through a concrete smart home scenario.
+\begin{figure}[t]
+\centering
+\includegraphics[width=1\columnwidth]{figures/smarthome-scenario.pdf}
+\caption{In the smart home scenario, we have built three smart appliances for a user experience study.
+The control interface supports both simple on/off appliances (lamp or fan) and multi-functional appliances (TV, for example).}
+\label{fig:smart-home}
+\end{figure}
+
+To assess the usability of such applications when the novel interaction scheme is added, we have built a few smart appliances (including a fan, a lamp and a TV, Figure~\ref{fig:smart-home}) and asked 14 participants to work through a concrete smart home scenario.
 The fan and lamp are relatively dumb and only support turning-on and turning-off.
 The smart TV provides control capability of switching the channels and adjust the volumn, as well as pause/play.
-For the smart home scenario, the participant has been asked to perform a series of actions that swtich the targets and control them.
-We then report some of the results to strength our motivation for the head-orientation interaction.
+From this study, we report some of the qualitative results to strength our motivation for the head-orientation interaction.
 
-Durign the study, all participants have successfully completed the list of tasks.
+\ben{Will add more from our survey.}
+During the study, all participants have successfully completed the list of tasks.
 They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous study procedure.
@@
 In addition, users did not always remember these mappings as they are not yet part of a standard gesture vocabulary.
-
-\bjoern{since length will be an issue, let's just focus on glass remote and skip the others?}
-Additional applications such as attention tracking or indoor position and orientation estimation of a wearer based on our technique are conceivable; we leave them to future work.
 

### 2014-04-15 doc/uist14/conclusion.tex
Filled some of the contents; updated a few figures; just added content to disucssion and conclusion, but the sentences are broken and may not make any sense

@@
 
-In this paper we have presented our system and solutions for the head orientation selection.
+In this paper, by presenting our iterative design process in head-orientation based targeting, we have learned that IR alone can help reduce the overall acquisition time by reducing the chances when we need to perform refinement.
+With IR intensity added, the targeting can work better in a relative dense environment.
+However, a more natural while challenging approach is to combine IR with the head motion.
+Through our preliminary user studies on this side, we learned that this is a more intuitative way of performing refinement in comparison to the menu-based selections.
+We leave a more comprehensive technical solution of using motion sensors and its evaluation as the future work.
 

### 2014-04-15 doc/uist14/discussion.tex
a significant amount of new content and changes on papers
Filled some of the contents; updated a few figures; just added content to disucssion and conclusion, but the sentences are broken and may not make any sense

@@
 
-We discuss a few issues with the system.
+\ben{cannot focus on writing this part clearly at 4am... I am just laying down whatever has popedup my heads and plan to refine it tomorrow.}
+
+\subsection{Generalization of our results}
+\label{sec:gener-our-results}
+This paper describes three iterative designs and either of them can be applied to a more broader range.
+First, the scan + refinement model is generic.
+Second, with a combination of signal reception and strength measurement, together with the signal strength model, a system can usually do better.
+Third, multiple types of sensors combined can overcome a lot of issues.
+
+\subsection{Using Google Glass}
+\label{sec:using-google-glass}
+In our current implementation, we have chosen to use Google Glass as the main head-worn computing device.
+This decision was made primarily because Glass has the display, touchpad, and IMU sensors and programming Glass is easy (standard Android).
+But this doesn't exclude other head-worn devices.
+
+Comments about that our system made the assumption that users have to wear Google Glass first doesn't hold.
+The work is more about the design exploration of how head orientation can be used for targeting.
+
+\subsection{Gaze Tracking}
+\label{sec:gaze-tracking}
+The reason I brought up it here is because gaze tracking seems to be mentioned by multiple people and having a discussion here is good to clear the questions.
+
 %%% Local Variables: 

### 2014-04-15 doc/uist14/interaction.tex
Filled some of the contents; updated a few figures; just added content to disucssion and conclusion, but the sentences are broken and may not make any sense

@@
 LEDs with many different fields of view are commercially available.
-\bjoern{say something about receivers and how point to point connections enable us to leverage head orientation without needing a map.}
+Each individual target that is equipped with IR detector can react to the signal.
+Such point to point communications enables us to achieve targeting without needing a map.
+%% \bjoern{say something about receivers and how point to point connections enable us to leverage head orientation without needing a map.}
 
@@
 
-Our augmentation to Google Glass adds a \bjoern{add modelnumber and spec details} LED (see Figure~\ref{fig:glass}.
-In the following description, we refer to Glass because the term is succinct; but  other suitable head-worn computing devices could be substituted.
+Our augmentation to Google Glass adds a 940nm 5mm IR emitter\footnote{In the prototype, we have used the IR emitter SFH 4545 from OSRAM Opto Semiconductors Inc.} (see Figure~\ref{fig:glass}).
+\ben{add description about the microcontroller.} In the following description, we refer to Glass because the term is succinct; but other suitable head-worn computing devices could be substituted.
 
-\bjoern{need to mention that all the target nodes also have to be able to talk to each other and to glass outside the IR channel. We use XBees for this purpose.}
+% The IR emitter is the one from OSRAM Opto Semiconductors Inc with manufacturer part number SFH 4545.
+From the datasheet, it has a view angle of $10^\circ$.
+Part of the reason we have chosen it is because the radiant intensity can quite high if we provide it with sufficient current flow (550mW/sr @ 100mA).
+We can then easily adjust it using a resistor to satisfy different demands.
+% % transmitter webpage: http://www.digikey.com/product-search/en?x=0&y=0&lang=en&site=us&KeyWords=475-2919-ND
+% For the IR receiver, we use 38.0kHz IR Receiver Modules from Vishay Semiconductor Opto Division (manufacturer part number TSOP38238). 
+% % receiver webpage http://www.digikey.com/product-detail/en/TSOP38238/751-1227-ND/1681362
+% In order to read the IR intensity, we integrates another IR light-to-voltage converter from AMS-TAOS USA Inc (manufacturer part number TSL267-LF) which is pretty sensitive to IR irradiance. 
+% % http://www.digikey.com/product-detail/en/TSL267-LF/TSL267-LF-ND/3095052
+
+Though IR signal has enabled us achieve head orientation-based targeting (simplex communication), we envision that more complicated applications could be built on top.
+Therefore, we added 802.15.4 radio to Glass and all the target nodes.
 
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figures/tube_from_top_annotated.pdf}
+\includegraphics[width=0.9\columnwidth]{figures/GlassWithHolder.jpg}
 \caption{Google Glass augmented with a repositionable IR holder.}

### 2014-04-15 doc/uist14/iteration2.tex
Filled some of the contents; updated a few figures; just added content to disucssion and conclusion, but the sentences are broken and may not make any sense

@@
 Received IR intensity falls off with distance between IR emitter and receiver as well as with the angle between emitter and receiver.
-To measure intensity, we add \bjoern{add model number and characteristics} and replace the IR emitter with \bjoern{modelnumber, characteristics}.
+To measure intensity, we add an IR light-to-voltage converter (model number TSL267-LF from AMS-TAOS USA Inc and replace the IR emitter with IR emitter (model number SFH 4545 from OSRAM Opto Semiconductors Inc).
+% transmitter webpage: http://www.digikey.com/product-search/en?x=0&y=0&lang=en&site=us&KeyWords=475-2919-ND
+% http://www.digikey.com/product-detail/en/TSL267-LF/TSL267-LF-ND/3095052
+
 We have empirically measured the intensity distribution at the receiver for this configuration in Figure~\ref{fig:measurement}.
-\ben{still working on it, not sure how much value we will get out of it.
-Seems useful to support the argument of the benefit though.}
-Our measurements confirm that angular difference has a large effect on the intensity readings.
+Our measurements confirm that angular difference has a large effect on the intensity readings (see how the IR intensity distribution changes when the offset from the center increases).
+
 \begin{figure}[t]
 \centering
-\includegraphics[width=0.9\columnwidth]{figures/placeholder.pdf}
-\caption{placeholder for empirical measurement.}
+\includegraphics[width=1\columnwidth]{figures/IRIntensityDistribution.pdf}
+\caption{Empirical measurement of IR intensity at different position.
+We measure the intensity at the plane of every 1 meter away from the IR emitter.
+On that plane, we obtain a sample every 5 centimeters.
+The size and color brightness in this graph represents the intensity readings.}
 \label{fig:measurement}
@@
 {\em Intensity IR} sorted the desired target as the first one in the list in 55\% of cases (93 of 167).
-In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 219\bjoern{why isn't this 225 as above?}).
+In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 225).
 Again, a Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, $p < 0.0001$.
 
-Such gain can be seen from Figure~\ref{fig:study2} where the overall target acquisition time is obtained.
-\bjoern{avoid passive voice throughout!} The average selection time decreases from 4.31 seconds for {\em Naive IR} to 3.64 second for {\em Intensity IR}.
-This difference is also significant ($t(555)=3.2945$, $p=0.001$.) \bjoern{for consistency's sake, can we show a box plot as in figure 5?}
+From Figure~\ref{fig:study2}, we can see that the overall target aquisition time has decreased from 4.31 seconds for {\em Naive IR} to 3.64 second for {\em Intensity IR}.
+This difference is also significant ($t(555)=3.2945$, $p=0.001$.) \bjoern{for consistency's sake, can we show a box plot as in figure 5?} \ben{I think cdf is a better way than boxplot?
+It actually provides more information.
+I may consider using cdf consistently.}
 
@@
 \includegraphics[width=0.9\columnwidth]{figures/study2_time.pdf}
-\caption{The target acquisition of in {\em Valued IR} and {\em Binary IR} condition.}
+\caption{The target acquisition of in {\em Naive IR} and {\em Intensity IR} condition.}
 \label{fig:study2}

### 2014-04-16 doc/uist14/iteration3.tex
iteration 3: gender neutral wording

@@
 
-During the second stage, when the user wants to use his head-motion for disambiguation, he will put his finger on the touchpad to enter a quasi-mode.
+During the second stage, when the user wants to use his head-motion for disambiguation, they will put their finger on the touchpad to enter a quasi-mode.
 This gesture is inspired by observing users who tend to adjust their glasses for closer inspection.
@@
 In this mode, only one device has a solid indicator light on at a time.
-\bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} The user can then tilt in the direction of the device he wishes to switch to.
+\bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} The user can then tilt in the direction of the device they wish to switch to.
 We calculate the direction of movement using a low-pass-filtered history of orientations, and look for devices that lie in that direction.

### 2014-04-16 doc/uist14/iteration3.tex
iteration3: remove comment about gender neutral words

@@
 The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of a user's head.
-However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location \achal{Should we use gender neutral language?}\bjoern{yes.
-just did in the previous sentence}.
+However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location.
 However, as long as the targets are spread around the periphery, their relative orientation doesn't change.

### 2014-04-16 doc/uist14/iteration3.tex
iteration3: comment reply about quasi mode refinement set

@@
 In this mode, only one device has a solid indicator light on at a time.
-\bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} The user can then tilt in the direction of the device they wish to switch to.
+\bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} \achal{There is no longer a refinement set - instead the user can select between all the devices using just orientation once they enter quasi mode.} The user can then tilt in the direction of the device they wish to switch to.
 We calculate the direction of movement using a low-pass-filtered history of orientations, and look for devices that lie in that direction.

### 2014-04-16 doc/uist14/iteration3.tex
iteration 3: further gender neutraility

@@
 
-In the first stage, the user orients his head towards each device, and the system saves the absolute orientations of each device.
+In the first stage, the user orients their head towards each device, and the system saves the absolute orientations of each device.
 Once the user has added all devices, these orientations are saved for future use.
@@
 
-During the second stage, when the user wants to use his head-motion for disambiguation, they will put their finger on the touchpad to enter a quasi-mode.
+During the second stage, when the user wants to use head-motion for disambiguation, they will put their finger on the touchpad to enter a quasi-mode.
 This gesture is inspired by observing users who tend to adjust their glasses for closer inspection.

### 2014-04-16 doc/uist14/iteration3.tex
iteration 3: more detailed description of orientation method
* Changed the structure slightly to discuss user facing first, then
  technical details.
* Discussed technical challenges in more detail
* Explained the "sticky" threshold more clearly (hopefully)
* Responded to some comments

@@
 Both{\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step.
-This has two problems: First, a spatial layout of target is compressed into a uni-dimensional list.
+This has two problems: First, a spatial layout of targets is compressed into a uni-dimensional list.
 List navigation commands thus have poor spatial mapping.
 Second, our techniques require the user to switch their focus back and forth between the physical world and the display.
-This issue motivated us to design an approach that also utilizes orientation information during the {\em refinement} stage, to provide a more direct mapping between the user's actions and target selection.
+These issues motivated us to design an approach that also utilizes orientation information during the {\em refinement} stage, to provide a more direct mapping between the user's actions and target selection.
 
@@
 However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location.
-However, as long as the targets are spread around the periphery, their relative orientation doesn't change.
+However, as long as the targets are spread around the periphery, their relative orientation is stable.
 This key observation inspires us to investigate a hybrid solution with IR and orientation sensors on Glass.
@@
 
-In the first stage, the user orients their head towards each device, and the system saves the absolute orientations of each device.
-Once the user has added all devices, these orientations are saved for future use.
-It is still possible to add new devices at any point without running the entire calibration step, as the system will simply store the absolute orientation of the new device.
-
-During the second stage, when the user wants to use head-motion for disambiguation, they will put their finger on the touchpad to enter a quasi-mode.
+In the first stage, the user orients their head towards each device, and the system learns a map of the devices based on their corresponding sensor readings.
+During the second stage, when the user wishes to use head-motion for disambiguation, they will put their finger on the touchpad to enter a quasi-mode in which their head movement is used to select targets.
+In this mode, only one device has a solid indicator light on at a time.
+\bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} \achal{There is no longer a refinement set - instead the user can select between all the devices using just orientation once they enter quasi mode.} The user can then tilt in the direction of the device they wish to switch to.
 This gesture is inspired by observing users who tend to adjust their glasses for closer inspection.
 
-In this mode, only one device has a solid indicator light on at a time.
-\bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} \achal{There is no longer a refinement set - instead the user can select between all the devices using just orientation once they enter quasi mode.} The user can then tilt in the direction of the device they wish to switch to.
-We calculate the direction of movement using a low-pass-filtered history of orientations, and look for devices that lie in that direction.
-Once a new device is selected, the frame of reference changes, and the user can repeat the process.
-Note that in reality, this is a completely smooth operation: it is possible to switch between multiple devices in one continuous motion.
-This is achieved by automatically adjusting the frame of reference to the currently lit device, but also creating a ``sticky" threshold so that the user is not constantly jumping between devices.
-\bjoern{This is called hysteresis or double thresholding.}
+The technical implementation of this interaction involves a number of subtleties due to the noisiness of our data.
+Absolute orientations between devices change frequently at the slightest user movement, and motion sensor data is inherently noisy.
+To alleviate the second issue, a low-pass filter is applied to the measurements, allowing us to extract the direction of motion from this smoothed history.
+Empirically, we found that we can reliably detect the direction of movement with 100ms of history and a sensor resolution of 10ms.
+
+Orientation changes can be overcome by using a hybrid of absolute and relative orientations augmented by IR information.
+While absolute orientations are unstable, relative orientations do not contain information for essential tasks, such as adding devices to a learned map.
+Our system stores the absolute orientations of each device during the learning stage, but utilizes only the relative information embedded in this data for disambiguation.
+Once the user is in quasi-mode, our system calculates the user's direction of motion as described earlier, and searches through the map for the nearest device in that direction.
+Thus, although the absolute orientation of the currently lit device may be incorrect, the relative information can still be used for disambiguation.
+Finally, the system uses hysteresis to prevent the user from accidentally jumping repeatedly between devices by switching devices only once the user has moved beyond a certain threshold in any direction.
+\achal{I understand hysteresis, but I can't figure out how the word can be used naturally (as there's no way to say a system "has" hysteresis).
+This last sentence may require rewording.}
 
 \bjoern{how well will this work if the user shifts position within a room?}
+\achal{I think the updates should explain this now.}
 

### 2014-04-16 doc/uist14/discussion.tex
added data from our study; updated the script

--- 2014-04-05 doc/uist14/discussion.tex
@@
 
-We discuss a few issues with the system.
+\subsection{What is the design space?}
+\label{sec:what-design-space}
+\ben{may change this to a seperate section.}
+Throughout our three iteration, the goal is minimize the overall target selection time (Eq~\ref{eq:time}).
+
+In our first iteration, by introducing IR to for the {\em scan}, we effectively reduce $P(refine)$ (for a less dense environment) and $t_{refine}$ (reduce the target numbers for the {\em refinement} stage.
+
+During the second iteration, with {\em Intensity IR}, $P(refine)$ is the same as {\em Naive IR} while it has been further descreased in a relative dense environment.
+Also $t_{refine}$ is also reduced because the order of targets in {\em refinement} stage is based on the probability of it being the intended target.
+
+Iteraction 3 attempts to solve the problem of mapping targets in the physical space to a 1-d list data structure, and it's mainly in the {\em refinement} stage.
+Preliminary studies have shown that users prefer this more direct {\em refinement} to the menu-based approach.
+
+\ben{Will add a figure to illustrate the above points.}
+
+\subsection{How dense can the environment be?}
+\label{sec:how-dense-can}
+The motivation of our second iteraction comes from the challenge of a relative dense environment.
+However, the term {\it dense} is not well-defined in physical space.
+Given two targets, there are potentially four different parameters to fully describe the {\it density}: separation in x, separation in y, separation in depth and the overall distance from the user to the targets (or the center between them).
+In this paper, we have focused on the design space of user interaction and a comparison of different techniques.
+In {\em Intensity IR} technique, we have done a preliminary study of the distribution of IR signal to justify the benefit of using intensity for {\em scan} and  {\em refine} stage.
+A more systematic and thorough study should be carried to understand the performanc of \systemnamenospace under different scenarios.
+Though the absolute target acquisition time will change when the environment setup differs, our study results still holds.
+We leave the {\em density} study as future work.
+
+\subsection{Why Google Glass?}
+\label{sec:using-google-glass}
+Google Glass is a well-engineered wearable device that has integrated touchpad, display, and IMU and a couple of wireless radios.
+This is a suitable platform because we need a head-worn device that can get users' input and display related UI at the near-eye space.
+The IMU on Google Glass also helps us explore our third technique of monitoring users' head motion for direct {\em refinement}.
+
+Before we settled down on Google Glass, we have actually attempted to build a similar glasses form-factor device.
+Though we successfully integrated the touchpad, and were considering IMU, the near-eye display was difficult to engineer.
+Among the available commercial off-the-shelf head-up display wearable devices, Google Glass stands out because it's light-weight and designed for everyday use.
+It's also easy to program (standard Android) with little learning curve.
+
+\subsection{Computer vision-based physical targeting}
+\label{sec:comp-visi-based}
+One device-free appraoch of performing physical targeting has been brought up during numerous discussions with people when we are working on this project. We see a few challenges in this approach:
+\begin{itemize}
+\item Achieving high accuracy object detection is still an open problem in computer vision. 
+\item problem 2
+\item problem 3
+\end{itemize}
+
+is to use computer vision to detect \ben{Talk about the challenges in using pure computer vision technique.}
+
 %%% Local Variables: 

### 2014-04-16 data/201309.Studies/script/visualize.R

@@ -1,50 +1,6 @@
 require('ggplot2')
 
-# Multiple plot function
-#
-# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
-# - cols:   Number of columns in layout
-# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
-#
-# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
-# then plot 1 will go in the upper left, 2 will go in the upper right, and
-# 3 will go all the way across the bottom.
-#
-multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  require(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
+## to run this script, you need multiplot
 
 ## Summarizes data.
 ## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
@@ -102,7 +58,7 @@ df.IR.mul <- df[grepl("IR", df$mode) & grepl("mul", df$correct_type), ]
 #### draw the list mode, time vs. position
[…]

### 2014-04-16 data/201404.Studies/analyze_fitts.py

@@ -24,20 +24,72 @@ class bcolors:
   ENDC = '\033[0m'
 
 
-
 def check_files(directory):
   ''' this function checks if the file logs in a specific directory have the right format'''
   filenames = map(lambda x: directory + '/' + str(x), range(1, 60))
+  if not all(map(isfile, filenames)) or not isfile(directory + '/glass.txt'):
+    print bcolors.FAIL + " file missing in " + directory + bcolors.ENDC
+    exit(1)
+    
   for f in filenames:
     with open(f) as f_handler:
       for line in f_handler:
         if line[0] != '[' and line[0] != ' ':
           print "file:", f, " formatting is wrong"
-          break
+          exit(1)
+  
+  with open(directory + '/glass.txt') as f:
+    for line in f:
+      if line[0:10] != "E/GlasSees" and line[0:5] != "-----":
+        print "file:", f, " formatting is wrong"
+        break
 
-def process_file(filename, tag):
-  target_id, target_time, search_time, acquire_time = None, None, None, None
-  with open(filename) as f:
+
+### problem now with time synchronization
+## consider doing it later
+def merge(arduino_filename, glass_filename):
+  logs = []
+  start, finish = None, None
+  with open(arduino_filename) as f:
+    for line in f:
+      m = re.search(r"\[ 2014-04", line)
+      if m:
+        time = datetime.strptime(line[2:28], '%Y-%m-%d %H:%M:%S %f')
+        if not start:
+          start = time
+        finish = time
+        event = line[40:]
+        logs.append((time, event))
+
+  print start, finish
+  with open(glass_filename) as f:
+    for line in f:
+      m = re.search(r'time: (\d{13})', line)
+      if m:
+        time = datetime.fromtimestamp(int(m.group()[6:])/1000.0)
+        if time > start and time < finish:
+          event = line[39:]
+          logs.append((time, event))
+
+  logs.sort(key = lambda x: x[0])
+  return logs
+
+## this works and helps!!
+def convert_glass_logs(glass_filename):
+  with open(glass_filename) as f:
+    for line in f:
+      m = re.search(r'time: (\d{13})', line)
+      if m:
+        time = datetime.fromtimestamp(int(m.group()[6:])/1000.0)
+        event = line[39:]
+        print time, event,
+
+def process_file(directory, filename, tag):
+  target_id, target_time, search_time, acquire_time = None, None, [], None
+  first_time = None
+  tap_time, first_target = None, None
[…]

### 2014-04-16 data/201404.Studies/data_summary.R

@@ -1,4 +1,186 @@
-d <- read.table("aggregate.txt")
+## setwd("~/repos/CS294-84Project/data/201404.Studies")
 
-summary(d[d$V5=="intensity", 3:4])
-summary(d[d$V5=="name", 3:4])
+d <- read.table("aggregate.txt")
+colnames(d) <- c("participant", "trial", "target", "first", "visual_search", "tap_time", "complete", "group", "miss_count")
+## for a quick summary
+intensity <- d[d$group =="intensity", ]
+name <- d[d$group =="name", ]
+
+intensity_time <- ggplot(intensity, aes(x = complete, fill = is.na(first))) +
+  geom_density(binwidth = 0.1, alpha = 0.5) +
+  xlab("distribution of acquisition time") +
+  ggtitle("with IR intensity") + 
+  theme_bw() +
+  scale_fill_discrete(name="w/o disambiguation",
+                      labels=c("disambiguation", "direct"))
+name_time <- ggplot(name, aes(x = complete, fill = is.na(first))) + 
+  geom_density(binwidth = 0.1, alpha = 0.5) +
+  xlab("distribution of acquisition time") +
+  ggtitle("without IR intensity") + 
+  theme_bw() + 
+  scale_fill_discrete(name="w/o disambiguation",
+                      labels=c("disambiguation", "direct"))
+
+pdf("figures/single_vs_multiple.pdf", width = 7, height = 6)
+multiplot(intensity_time, name_time, cols = 1)
+dev.off()
+
+## 41% wrong rate for guessing
+intensity_multiple <- intensity[!is.na(intensity$first), ]
+wrong_guess <- sum(intensity_multiple$first != intensity_multiple$target) 
+wrong_percentage <- sum(intensity_multiple$first != intensity_multiple$target) / length(intensity_multiple$target)
+ceiling(c(nrow(intensity), nrow(intensity_multiple), wrong_guess, wrong_percentage * 100))
+
+## 63% wrong rate for guessing
+name_multiple <- name[!is.na(name$first), ]
+wrong_guess <- sum(name_multiple$first != name_multiple$target) 
+wrong_percentage <- sum(name_multiple$first != name_multiple$target) / length(name_multiple$target)
+ceiling(c(nrow(name), nrow(name_multiple), wrong_guess, wrong_percentage * 100))
+
+## this is for individuals
[…]

### 2014-04-16 doc/uist14/applications.tex
fixes, changes to application

@@
 
-Though such head orientation targeting enables a wide range of context-aware applications, in this paper we focus on our GlassRemote system that we have built using the existing hardware platform to achieve a universal smart appliance control functionality.
+Head orientation targeting can enable a wide range of context-aware applications.
+We implement one particular demonstrative application: a universal smart appliance controller.
+Users select a smart device (e.g., light fixture, TV, or home appliance) with Glass --- upon confirming the selection, an appliance-specific UI is shown on the user's near-eye display, and they can control the application (without having to continually look at it) through their device touchpad.
 
-The huge benefit using head orientation is to solve two challenges that exist in most commercial list-based solutions -- {\em naming} and {\em scoping}.
+We implemented a prototype of this remote controller for three physical devices: a lamp and fan that could be switched on or off; and a smart TV control with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
+We switched discrete appliances using relays, and implemented the smart TV with a laptop connected to a 30\inch{} display.
+User interfaces for each were pre-defined in our application.
+
+We set the devices up in a simulated living room environment and invited 14 users to step through a predefined set of tasks to control the appliances at a distance.
+
+%The huge benefit using head orientation is to solve two challenges that exist in most commercial list-based solutions -- {\em naming} and {\em scoping}.
 First, assigning clear names is non-trivial.
@@
 
-In order to control a smart appliance, the user performs a physical target selection (scan, refine and confirm) first and then the appliance-specific UI will be shown in the near-eye display.
+%In order to control a smart appliance, the user performs a physical target selection (scan, refine and confirm) first and then the appliance-specific UI will be shown in the near-eye display.
 
@@
 
-To assess the usability of such applications when the novel interaction scheme is added, we have built a few smart appliances (including a fan, a lamp and a TV, Figure~\ref{fig:smart-home}) and asked 14 participants to work through a concrete smart home scenario.
+%To assess the usability of such applications when the novel interaction scheme is added, we have built a few smart appliances (including a fan, a lamp and a TV, Figure~\ref{fig:smart-home}) and asked 14 participants to work through a concrete smart home scenario.
 The fan and lamp are relatively dumb and only support turning-on and turning-off.
@@
 \ben{Will add more from our survey.}
-During the study, all participants have successfully completed the list of tasks.
+All participants successfully completed the list of tasks.
 They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous study procedure.
-Participants saw benefits of the device for families --- \studyquote{It might also be useful for people who need to take care of small children that they can complete all the tasks while keeping an eye on their children at the same time}, though settings that require more movement than watching a movie at home (e.g., cooking) may be more appropriate scenarios for wearing the device.
+Participants saw benefits of the device for families --- \studyquote{It might also be useful for people who need to take care of small children that they can complete all the tasks while keeping an eye on their children at the same time}, though settings that require more movement than watching a movie at home (e.g., cooking) may ultimately be more compelling scenarios for wearing the device around the home.
 

### 2014-04-16 doc/uist14/interaction.tex
fixes, changes to application

@@
 
-Though IR signal has enabled us achieve head orientation-based targeting (simplex communication), we envision that more complicated applications could be built on top.
-Therefore, we added 802.15.4 radio to Glass and all the target nodes.
+While we use IR signal to establish simplex point-to-point communication from Glass to a target, we envision that more complicated applications can be built on top of the selection functionality.
+Therefore, we added 802.15.4 radios to Glass and all the target nodes to enable bi-directional communication between all nodes and Glass.
+\bjoern{not entirely correct - we also need node-to-glass communication for the Intensity IR technique.}
 

### 2014-04-16 doc/uist14/iteration3.tex
fixes, changes to application

@@
 \subsection{Problem Statement and Goal}
-Both{\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step.
+Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step.
 This has two problems: First, a spatial layout of targets is compressed into a uni-dimensional list.

### 2014-04-16 doc/uist14/iteration1.tex
pushed my changes of figures

@@
 \includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
-\caption{In the targeting study, participants had to find and select one of 10 targets in a lab environment.
-Targets were called out by number; for the list mode condition, participants need to match numbers to letters.}
+\caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment. \ben{add a way to describe the distance in the figure}.}
 \label{fig:targeting-study-layout}
@@
 
+\subsection{Results}
+
 \begin{figure}[t]
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/R_time_by_Category.pdf}
-\caption{Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), and between IR multiple responses cases and IR single response cases (B).
+\includegraphics[width=1.0\columnwidth]{figures/result_study1a.pdf}
+\caption{\ben{fix this later}.
+Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), and between IR multiple responses cases and IR single response cases (B).
 The centers of boxes are median values, while white dashed lines are mean values.}
-\label{fig:selection-time}
+\label{fig:ir_vs_list}
 \end{figure}
 
-\subsection{Results}
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.09\columnwidth]{figures/result_study1b.pdf}
+\caption{With and without refinement.}
+\label{fig:with_without_refinement}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/result_study1c.pdf}
+\caption{Times taken to select a device vs.~its order in the list.
+The dotted line is a linear fit between the mean times and device orders.
+Two horizontal lines of mean target acquisition times in {\em infrared} mode are also annotated for comparison.}
+\label{fig:time-vs-list-order}
+\end{figure}
+
 Our results indicate that {\em IR selection} outperforms {\em list selection}.
-The average target acquisition time for {\em IR selection} is 6.67 seconds, while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:selection-time}A).
+The average target acquisition time for {\em IR selection} is 6.67 seconds, while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}).
 Student's t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
@@
 To understand how scanning and refinement contribute to total selection time, we split the data from {\em IR selection} into two parts -- trials that required refinement and ones that did not.
-It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:selection-time}B).
-This difference is significant ($t(19)=-2.7827, p=0.012$ using t-test \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}).
+It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:with_without_refinement}).
+This difference is significant ($t(19)=-2.7827, p=0.012$ using t-test.
+%  \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}). 
+%% \ben{Because the times when refinement is needed is only 18 times, so the DF is bounded by the samples.}
 Because many targets were spaced far apart, refinements were only necessary in 10\% of total {\em IR selection} trials in this study.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=1.0\columnwidth]{figures/R_List_by_Target.pdf}
-\caption{Times taken to select a device vs.~its order in the list.
-The dotted line is a linear fit between the mean times and device orders.
-Two horizontal lines of mean target acquisition times in {\em infrared} mode are also annotated for comparison.}
- % \bjoern{Add horizontal lines at $t_{IRsingle}$ and $t_{IRmultiple}$.}
-\label{fig:time-vs-list-order}
-\end{figure}
 

### 2014-04-16 doc/uist14/iteration2.tex
pushed my changes of figures

@@
 \includegraphics[width=1.0\columnwidth]{figures/study-layout2.jpg}
-\caption{The current environment setup \ben{needs annotation and a better photo.} \bjoern{yes}.}
+\caption{The current environment setup \ben{needs annotation and a better photo.} \bjoern{yes}.. blackboard is 2*107+122 cm , 183 cm dimension.}
 \label{fig:study-layout2}
@@
 From Figure~\ref{fig:study2}, we can see that the overall target aquisition time has decreased from 4.31 seconds for {\em Naive IR} to 3.64 second for {\em Intensity IR}.
-This difference is also significant ($t(555)=3.2945$, $p=0.001$.) \bjoern{for consistency's sake, can we show a box plot as in figure 5?} \ben{I think cdf is a better way than boxplot?
+This difference is also significant ($t(555)=3.2945$, $p=0.001$.)
+
+%% resolved: \bjoern{for consistency's sake, can we show a box plot as in figure 5?} \ben{I think cdf is a better way than boxplot?
 It actually provides more information.
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figures/study2_time.pdf}
+\includegraphics[width=1.0\columnwidth]{figures/result_study2.pdf}
 \caption{The target acquisition of in {\em Naive IR} and {\em Intensity IR} condition.}

### 2014-04-16 doc/uist14/interaction.tex
pushed my changes of figures
added comments for intro, interaction, iteration1/2

--- 2014-04-15 doc/uist14/interaction.tex
@@
 
-Our augmentation to Google Glass adds a 940nm 5mm IR emitter\footnote{In the prototype, we have used the IR emitter SFH 4545 from OSRAM Opto Semiconductors Inc.} (see Figure~\ref{fig:glass}).
-\ben{add description about the microcontroller.} In the following description, we refer to Glass because the term is succinct; but other suitable head-worn computing devices could be substituted.
+Our augmentation to Google Glass adds a 940nm 5mm IR emitter (SFH 4545 from OSRAM Opto Semiconductors Inc.).
+The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio.
+(see Figure~\ref{fig:glass}).
+In the following description, we use the succint term ``Glass'' to refer to this augmented hardware.
 
@@
 
-Though IR signal has enabled us achieve head orientation-based targeting (simplex communication), we envision that more complicated applications could be built on top.
-Therefore, we added 802.15.4 radio to Glass and all the target nodes.
+While we use IR signal for targeting, wireless communications are needed to efficiently transmit the reception status of IR signal from targets to Glass.
+We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose.
+
+\ben{no mention of target hardware yet. I have checked in an image of the targets ``targets.JPG'' in case we need.}
 
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figures/GlassWithHolder.jpg}
-\caption{Google Glass augmented with a repositionable IR holder.}
+\includegraphics[width=1\columnwidth]{figures/GlassWithArduino.jpg}
+\caption{Our Glass hardware: Google Glass augmented with a repositionable IR holder, and an additional microcontroller that communicates with Google Glass and controls IR emitter.}
 \label{fig:glass}
@@
 The user confirms his desire to connect to a target by tapping on the Glass touchpad.
-Glass collects the responses from targets that have received IR reception (we use an XBee-based backchannel for this traffic).
+Glass collects the responses from targets that have received IR reception through the backchannel of XBee.
 If there is only one single target in IR range, it is automatically selected.
@@
 t_{total}=t_{scan}+P(refine)*t_{refine}+t_{commit}
+\label{eq:time}
 \end{equation}

### 2014-04-16 doc/uist14/intro.tex
Filled some of the contents; updated a few figures; just added content to disucssion and conclusion, but the sentences are broken and may not make any sense
added comments for intro, interaction, iteration1/2

@@
 \section{Introduction}
-\bjoern{something happened to the figure caption font - it doesn't match the standard template. Please fix.}
+% \bjoern{something happened to the figure caption font - it doesn't match the standard template. Please fix.}
+% \ben{fixed. Problem is because of using caption package}.
+
 %% from the swarm vision to the necessity of selection
-The number of smart objects in our environment with embedded computation and communication has grown rapidly. These objects are all potential targets for interaction.
-To initiate such interaction requires {\em selection} information about the object at the user's {\em locus of attention}~\cite{raskin}. \sean{this sentence is unclear to me.}
+The number of smart objects in our environment with embedded computation and communication has grown rapidly.
+These objects are all potential targets for interaction.
+To initiate such interaction, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but yet not well-explored in the physical spaces.
+
+%% To initiate such interaction requires {\em selection} information about the object at the user's {\em locus of attention}~\cite{raskin}. \sean{this sentence is unclear to me.}
 
@@
 We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infrared (IR) communication between Glass and target appliances.
-In our {\em Naive IR} technique, the cone of light emitted by an IR LED (a diameter of 60-120cm and distance up to 6m) provides the {\em coarse} selection area.
-To {\em refine} selection when multiple targets have received IR signals, we introduce a manual list selection disambiguation technique.
-A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting.
-We find that target acquisition through head orientation is preferred by users and is faster than list selection, but refinements -- if necessary -- are slow.
+In our {\em Naive IR} technique, the cone of light emitted by an IR LED (a diameter of 30-60cm and distance up to 8m) provides the {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser}A, where targets that have received IR signal light up).
+To {\em refine} selection when multiple targets have received IR signals, our solution starts with a naive ordering -- based on their names.
+This is shown in Figure~\ref{fig:teaser}B, where the targets in range are ordered numerically on the near-eye display.
+A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting \ben{the ordered-list is also list selection, needs clarification!} .
+We find that target acquisition through head orientation is preferred by users and is faster than list selection but refinements -- if necessary -- are slow.
 
 In response, we designed an improved refinement technique, {\em Intensity IR}, in which target objects compare received IR signal strength.
-This value allows the system to either automatically select the target with the strongest intensity, or to re-order the refinement interface's list by their intensity values.
-A second study with $10$ participants shows that {\em Intensity IR} disambiguation successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.
+This value allows the system to either eliminate some distracting targets on the peripheral, or to re-order the refinement interface's list by their intensity values.
+For example, in Figure~\ref{fig:teaser}B of {\em Intensity IR} technique, 5 is eliminated first and the list is re-ordered based on the intensity readings.
+A second study with $10$ participants proves that {\em Intensity IR} successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.
 
 A final design addresses the lack of a natural mapping when users select a target in the refinement step using their device's touchpad - the axes of motion do not map directly to the spatial layout of target devices in a room.
-Our {\em Head-motion Refinement} technique first infers a spatial data structure of target locations using Glass' orientation sensors.
-In a refinement step, users can then perform head movements to change selections to spatially adjacent targets.
+Our {\em Head-motion Refinement} technique first learns the relative spatial structure of the targets using Glass' orientation sensors.
+Then in future target acquisition stage, users can then perform head movements to change selections to spatially adjacent targets (Figure~\ref{fig:teaser}C).
 For example, nodding down to select the target below current selection, or tilting right to select next target on the right.
@@
 An appliance-specific user interface is then shown on the user's near-eye display for further interactions.
+
+\ben{I am considering adding back the contribution part, because the paper itself mainly focuses on the iterative design process.
+Without an explicit contribution statement, the audience will have to infer and summarize themselves.}
 

### 2014-04-16 doc/uist14/iteration1.tex
added comments for intro, interaction, iteration1/2

@@
 \subsection{Problem Statement and Goal}
-Our initial research question is: {\em Can IR-based targeting reduce the selection time compared to UI list navigation on a head-worn device?}
+Our initial research question is: {\em Can IR-based targeting reduce the selection time compared to the case where only UI list navigation is used on a head-worn device?}
 
@@
 In our first implementation, we use IR for scanning as described in the previous section.
-For the refinement technique, we simply show a list of the subset of targets that have received IR signals on the Glass display and let users swipe to select from that list and tap to confirm.
+For the refinement technique, we simply show a list of the subset of targets that have received IR signals on the Glass display.
+Since we only know that the targets are within the range of users' head orientation.
+We order them using their ids; and in realistic environment target names can also be used if names can be assigned.
+We then let users swipe to select from that list and tap to confirm the intended target.
 
-A natural point of comparison is an interface that does not use any orientation information - it always shows 
-a complete list of all targets.
+A natural point of comparison is an interface that does not use any head orientation information - it always shows a complete list of all targets.
 We implemented a list view where users swiped forwards and backwards on the Glass touchpad to navigate incrementally through the list.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/targeting-study-layout.pdf}
-\caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment. \ben{add a way to describe the distance in the figure}.}
+\includegraphics[width=1.0\columnwidth]{figures/study-layout1.png}
+\caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment.}
 \label{fig:targeting-study-layout}
@@
 For each selection condition, we conducted 15 target selection by randomly choosing from all the targets.
-During the study, we measure the {\bf target acquisition time} for each target selection.
-Afterwards, participants were asked to complete a survey that elicited answers to Likert-scale questions as well as open-ended answers about their experience.
+During the study, we measured the {\bf target acquisition time} for each target selection.
+Afterwards, participants were asked to complete a survey that is mainly open-ended answers about their experience.
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/result_study1a.pdf}
-\caption{\ben{fix this later}.
-Boxplot of task completion times for the comparison between {\em infrared} mode and {\em list} mode (A), and between IR multiple responses cases and IR single response cases (B).
-The centers of boxes are median values, while white dashed lines are mean values.}
+\caption{Boxplot of target acquisition time for {\em IR selection} and {\em list selectioN} is shown on the left.
+The center of each box is the median value, while we annotate the mean value using white dashed lines.}
 \label{fig:ir_vs_list}
@@
 \includegraphics[width=1.09\columnwidth]{figures/result_study1b.pdf}
-\caption{With and without refinement.}
+\caption{Boxplot and CDF of target acquisition time when refinement is needed or not in {\em IR selection}}.
 \label{fig:with_without_refinement}
@@
 \includegraphics[width=1.0\columnwidth]{figures/result_study1c.pdf}
-\caption{Times taken to select a device vs.~its order in the list.
-The dotted line is a linear fit between the mean times and device orders.
-Two horizontal lines of mean target acquisition times in {\em infrared} mode are also annotated for comparison.}
+\caption{Times taken to select a device vs.~its order in the {\em list selection}.
+The dotted line is a linear fit between the average time and target orders in the list.
+Two horizontal solid lines are the average target acquisition times in {\em Naive IR} when refinement is needed or not.}
 \label{fig:time-vs-list-order}
@@
 Our results indicate that {\em IR selection} outperforms {\em list selection}.
-The average target acquisition time for {\em IR selection} is 6.67 seconds, while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}).
+The average target acquisition time for {\em IR selection} is 6.67 seconds \ben{one concern that I have is this is extremely long, 6 seconds.
+Do we need to explain more?
+One possible reason here is that we ask the user not to look at targets until we called out, so the visual search time is around 3 seconds.
+Should we say more or leaving 6 seconds there is fine?}, while {\em list selection} took 8.86 {\em list selection also requires visual search, I think...} seconds (see Figure~\ref{fig:ir_vs_list}).
 Student's t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
@@
 
-In summary, {\em Naive IR} can outperform linear list selection, but the refinement step detracts significantly from the efficiency of the technique.
+As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).
+While both interfaces were judged similarly on overall ease of connecting, list selection was perceived to be cumbersome.
+One user noted benefit of {\em IR selection} being \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
+One subject called it \studyquote{natural to interact with things just by looking at them}.
+Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquistion.
+One subject pointed out that \studyquote{{\em IR selection} doesn't requires you to swipe a lot of times like list mode but list selection}.
+
+In summary, {\em Naive IR} can outperform linear list selection, and most users like this head-orientation based targeting.
+But from our quantified study result, the refinement step detracts significantly from the efficiency of the technique.
+
 % learned that with IR, we can have effectively reduced the overall target acquisition time.
@@
 
-\ben{qualitative feedback hasn't been added yet for this part.}
 %\subsubsection{Preference}

### 2014-04-16 doc/uist14/iteration2.tex
added comments for intro, interaction, iteration1/2

@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/study-layout2.jpg}
-\caption{The current environment setup \ben{needs annotation and a better photo.} \bjoern{yes}.. blackboard is 2*107+122 cm , 183 cm dimension.}
+\includegraphics[width=1.0\columnwidth]{figures/study-layout2.png}
+\caption{The environment setup for our second study. In comparison to our first study, we have delibrately increased the target density.}
 \label{fig:study-layout2}
@@
 
-%% resolved: \bjoern{for consistency's sake, can we show a box plot as in figure 5?} \ben{I think cdf is a better way than boxplot?
-It actually provides more information.
-I may consider using cdf consistently.}
-
 %  t.test(name$complete, intensity$complete)
@@
 \includegraphics[width=1.0\columnwidth]{figures/result_study2.pdf}
-\caption{The target acquisition of in {\em Naive IR} and {\em Intensity IR} condition.}
+\caption{Boxplot and CDF of the target acquisition time in {\em Naive IR} and {\em Intensity IR} conditions.}
 \label{fig:study2}
@@
 One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage.
-The this error rate (13 out of 300 cases, 4.3\%), is higher than the error rate for {\em Naive IR} \bjoern{insert error rate}.
+Out of 300 trials, 13 of them (4.3\%) has been accidentally eliminated.
+This is higher than the error rate for {\em Naive IR} (only 5 out of 300, ~1.6\%).
 % but this is the trade-off for the overall faster acquisition time.

### 2014-04-16 doc/uist14/relatedwork.tex
remove space before the reference in related work

@@
 
-Other targeting systems use IR with handheld pointers ~\cite{swindells_that_2002} and as well as wearable devices such as rings and Bluetooth audio earpieces ~\cite{merrill_augmenting_2007} to connect to smart devices.
-Our system tackles an unresolved issue of such IR-based approaches-- navigating an area dense with potential targets.
+Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} and as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices.
+Our system tackles an unresolved issue of such IR-based approaches -- navigating an area dense with potential targets.

### 2014-04-16 doc/uist14/iteration3.tex
added illustration of graph for iteration3

@@
 
-In our third iteration, we ask the question ``can we build a system purely using head orientation and visual feedback from the environment for target selection?''
+In our third iteration, we ask the question ``can we build a system which further minimize the chance of falling back to list based refinement?''
 
@@
 However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location.
-However, as long as the targets are spread around the periphery, their relative orientation is stable.
+However, as long as the targets are spread around the periphery, their relative orientation to each other is stable (see Figure~\ref{fig:third_principle}).
 This key observation inspires us to investigate a hybrid solution with IR and orientation sensors on Glass.
 
-There are two phases to this technique: learning the relative orientations or targets in a room and intelligently suggesting targets during disambiguation.
+\begin{figure}[t]
+\centering
+\includegraphics[width=1\columnwidth]{figures/third_principle.png}
+\caption{Even when the user's absolution position changes, the relative relationship in the orientation space remain invariant.}
+\label{fig:third_principle}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1\columnwidth]{figures/third_technique.pdf}
+\caption{\ben{figure in interactionModel2.key file.} Our third technique learned each target's absolute orientation and construct the orientation map.
+Though we store the absolute orientation value, during the {\em refinement} stage, the prediction is only based on relative changes to a reference point.
+In such a way, the prediction will work regardless of user's position.}
+\label{fig:third_technique}
+\end{figure}
+
+There are two phases to this technique: learning the relative orientations or targets in a room and intelligently suggesting targets during refinement.
 

### 2014-04-16 doc/uist14/discussion.tex
discussion from yang's comments

--- 2014-04-05 doc/uist14/discussion.tex
@@
+%!TEX root = uist14.tex
 \section{Discussion}
 \label{sec:discussion}
+%%%NOTES FROM YANG%%%%
+The rapid development of sensing technologies has created many opportunities for new ways to interact with smart objects.
+In our exploration of the design space of head orientation-based target selection, we carefully selected sensing techniques that are readily available and easy to deploy; and we added complexity to our system only when necessary.
 
-We discuss a few issues with the system.
+\subsection{Interpretation of results}
+
+A primary goal of this paper was to provide an effective anf efficient method for target selections in physical spaces.
+Targeting is a fundamental building block across many interaction tasks - it has a significant impact on user experiences collectively and can provide seamless interaction when designed well.
+
+We formalized a scan+refine model of head orientation-based selection (Equation~\ref{eq:time}).
+We first introduced head orientation as an alternative to list selection and showed that {\em scanning} can outperform list selection.
+Our redesigns then focused on the case where refinement is needed.
+The two ways to reduce refinement time are 1) to reduce $P(refine)$, the probability that a manual refinement is necessary; and 2) to reduce $t_{refine}$, the time required to perform the refinement interaction itself.
+Using IR intensity readings addresses both these terms, as it can be used to both avoid showing refinement dialogs, and to optimize their display when they are needed.
+
+Our final head-orientation technique improves the nature of the mapping between items in the refinement dialog and the layout of targets in space.
+Informal testing suggests that users prefer using this spatial mapping.
+
+%Point 1: Why an effective technique for target acquisition is important--clarify the motivation
+%This is to address a comment we got from the CHI review. We want to make analogy with target acquisition in the digital world. I copied what we had in the rebuttal here: 
+%“Targeting is a fundamental building block across many interaction tasks - it has a significant impact on user experiences collectively and can provide seamless interaction when designed well.”
+
+%Point 2: The overall research methodology
+%The rapid development of sensing technologies has created many opportunities for new ways to interact with smart objects.
+While we explored the design space of head orientation-based target selection, we cautiously selected sensing techniques that are easy to deploy and added complexity to our system only when necessary.
+Our work thoroughly examined the design space via an iterative design process.
+Theorize the design space as you had in Equation 1.
+Ben already had some good points in the draft.
+
+\subsection{Limitations}
+We introduced new refinement operations based on IR intensity and spatial adjacency tracking that improve the performance of head orientation in environments with dense targets.
+These techniques have several limitations.
+
+First, IR intensity measurements only work within the dynamic range of our sensor.
+Additional strong IR sources like direct sunlight may saturate the sensor and make discrimination impossible.
+Second, our adjacency map is built assuming on stable target locations.
+If targets move, the map will have to be recalculated.
+This may be done incrementally during everyday interactions, but we have not yet tackled this challenge.
+We also have not yet evaluated how robust the adjacency maps is when the user changes her location or view angle.
+
+Finally, we acknowledge several limitations of our study design: we have not yet systematically studied target density variation, our study was performed in a lab environment, and only measured first use.
+We have also not investigated error rates in detail.
+Future work should study how the technique applies in realistic settings over longer periods of time.
+%Point 3: Limitation
+%Highlight the novelty of technique 2 & 3 while point out its weakness:
+%- How sunlight will impact the techniques
+%- Construct the adjacency mapping, what if objects move
+%- How robust its adjacency mapping is when the user changes her location or view angle.
+%- Outline programming by demonstration opportunities for future research
+%- Study limitation: density variation, realistic settings, longitudinal studies to understand how users choose between our techniques and traditional list-based solution.
+
+%%%% PRIOR DRAFT FROM BEN %%%
+%\subsection{What is the design space?}
+%\label{sec:what-design-space}
+%\ben{may change this to a seperate section.}
+%Throughout our three iteration, the goal is minimize the overall target selection time (Eq~\ref{eq:time}).
+
+%In our first iteration, by introducing IR to for the {\em scan}, we effectively reduce $P(refine)$ (for a less dense environment) and $t_{refine}$ (reduce the target numbers for the {\em refinement} stage.
+
+%During the second iteration, with {\em Intensity IR}, $P(refine)$ is the same as {\em Naive IR} while it has been further descreased in a relative dense environment.
+Also $t_{refine}$ is also reduced because the order of targets in {\em refinement} stage is based on the probability of it being the intended target.
+
+%Iteraction 3 attempts to solve the problem of mapping targets in the physical space to a 1-d list data structure, and it's mainly in the {\em refinement} stage.
+Preliminary studies have shown that users prefer this more direct {\em refinement} to the menu-based approach.
+
+%\ben{Will add a figure to illustrate the above points.}
+
+%\subsection{How dense can the environment be?}
+%\label{sec:how-dense-can}
+%The motivation of our second iteraction comes from the challenge of a relative dense environment.
+However, the term {\it dense} is not well-defined in physical space.
+Given two targets, there are potentially four different parameters to fully describe the {\it density}: separation in x, separation in y, separation in depth and the overall distance from the user to the targets (or the center between them).
+In this paper, we have focused on the design space of user interaction and a comparison of different techniques.
+In {\em Intensity IR} technique, we have done a preliminary study of the distribution of IR signal to justify the benefit of using intensity for {\em scan} and  {\em refine} stage.
+A more systematic and thorough study should be carried to understand the performanc of \systemnamenospace under different scenarios.
+Though the absolute target acquisition time will change when the environment setup differs, our study results still holds.
+We leave the {\em density} study as future work.
+
+%\subsection{Why Google Glass?}
+%\label{sec:using-google-glass}
+%Google Glass is a well-engineered wearable device that has integrated touchpad, display, and IMU and a couple of wireless radios.
+This is a suitable platform because we need a head-worn device that can get users' input and display related UI at the near-eye space.
+The IMU on Google Glass also helps us explore our third technique of monitoring users' head motion for direct {\em refinement}.
+
+%Before we settled down on Google Glass, we have actually attempted to build a similar glasses form-factor device.
+Though we successfully integrated the touchpad, and were considering IMU, the near-eye display was difficult to engineer.
+Among the available commercial off-the-shelf head-up display wearable devices, Google Glass stands out because it's light-weight and designed for everyday use.
+It's also easy to program (standard Android) with little learning curve.
+
+%\subsection{Computer vision-based physical targeting}
+%\label{sec:comp-visi-based}
+%One device-free appraoch of performing physical targeting has been brought up during numerous discussions with people when we are working on this project. We see a few challenges in this approach:
+%\begin{itemize}
+%\item Achieving high accuracy object detection is still an open problem in computer vision. 
+%\item problem 2
+%\item problem 3
+%\end{itemize}
+
+%is to use computer vision to detect \ben{Talk about the challenges in using pure computer vision technique.}
+
 %%% Local Variables: 

### 2014-04-16 doc/uist14/iteration3.tex
added fixed figure

@@
 However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location.
-However, as long as the targets are spread around the periphery, their relative orientation to each other is stable (see Figure~\ref{fig:third_principle}).
+However, as long as the targets are spread around the periphery, their relative orientation to each other is relatively stable (see Figure~\ref{fig:third_principle}).
 This key observation inspires us to investigate a hybrid solution with IR and orientation sensors on Glass.
@@
 \centering
-\includegraphics[width=1\columnwidth]{figures/third_technique.pdf}
+\includegraphics[width=1\columnwidth]{figures/third_technique3.pdf}
 \caption{\ben{figure in interactionModel2.key file.} Our third technique learned each target's absolute orientation and construct the orientation map.
@@
 
-There are two phases to this technique: learning the relative orientations or targets in a room and intelligently suggesting targets during refinement.
+There are two phases to this technique: learning the relative orientations or targets in a room and intelligently suggesting targets during refinement (see Figure~\ref{fig:third_technique}).
 

### 2014-04-16 doc/uist14/iteration3.tex
Updates to iteration 3 using Ben's feedback
* Shorten the overall section
* Removed unnecessary details from technical section
* Separated user facing from technical facing

@@
 \subsection{Problem Statement and Goal}
-Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step.
-This has two problems: First, a spatial layout of targets is compressed into a uni-dimensional list.
-List navigation commands thus have poor spatial mapping.
-Second, our techniques require the user to switch their focus back and forth between the physical world and the display.
-These issues motivated us to design an approach that also utilizes orientation information during the {\em refinement} stage, to provide a more direct mapping between the user's actions and target selection.
+Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the
+near-eye display for the refinement step. This has two problems: First, a
+spatial layout of targets is compressed into a uni-dimensional list. List
+navigation commands thus have poor spatial mapping. Second, our techniques
+require the user to switch their focus back and forth between the physical
+world and the display. These issues motivated us to design an approach that
+also utilizes orientation information during the {\em refinement} stage, to
+provide a more direct mapping between the user's actions and target selection.
 
-In our third iteration, we ask the question ``can we build a system which further minimize the chance of falling back to list based refinement?''
+\achal{This used to have an explicit question which I removed, as it sounded
+very unnatural.}
+In our third iteration, we build a system which further minimizes the chance of
+falling back to list based refinement.
 
 \subsection{Technique}
-The motion sensors (including gyroscope, magnetometer and accelerometer) on the head-worn device can help us determine the orientation of a user's head.
-However, this absolute orientation cannot be directly used in any indoor environment as the user may change their location.
-However, as long as the targets are spread around the periphery, their relative orientation to each other is relatively stable (see Figure~\ref{fig:third_principle}).
-This key observation inspires us to investigate a hybrid solution with IR and orientation sensors on Glass.
+The motion sensors (including gyroscope, magnetometer and accelerometer) on the
+head-worn device can help us determine the orientation of a user's head.
+Unfortunately, this absolute orientation cannot be directly used in any indoor
+environment as the user may change their location. As long as the targets are
+spread around the periphery, however, their relative orientations are stable
+(see Figure~\ref{fig:third_principle}). This key observation inspires us to
+investigate a hybrid solution with IR and orientation sensors on Glass.
 
@@
 
-There are two phases to this technique: learning the relative orientations or targets in a room and intelligently suggesting targets during refinement (see Figure~\ref{fig:third_technique}).
+There are two phases to this technique: learning the relative orientations or
+targets in a room and intelligently suggesting targets during refinement (see
+Figure~\ref{fig:third_technique}).
 
-In the first stage, the user orients their head towards each device, and the system learns a map of the devices based on their corresponding sensor readings.
-During the second stage, when the user wishes to use head-motion for disambiguation, they will put their finger on the touchpad to enter a quasi-mode in which their head movement is used to select targets.
-In this mode, only one device has a solid indicator light on at a time.
-\bjoern{revisit prior sentence - there's a problem in the current technique in that the user doesn't know which targets are in the refinement set and which aren't.} \achal{There is no longer a refinement set - instead the user can select between all the devices using just orientation once they enter quasi mode.} The user can then tilt in the direction of the device they wish to switch to.
-This gesture is inspired by observing users who tend to adjust their glasses for closer inspection.
+In the first stage, the user orients their head towards each device, and the
+system learns a map of the devices based on their corresponding sensor readings.
+During the second stage, when the user wishes to use head-motion for
+disambiguation, they will put their finger on the touchpad to enter a
+quasi-mode, where only one device has a solid indicator light on at a time. This
+gesture is inspired by observing users who tend to adjust their glasses for
+closer inspection. The user can then tilt in the direction of the device they
+wish to switch to. The system then uses its understanding of the relative
+locations of the devices to create an {\em adjacency map}.
 
-The technical implementation of this interaction involves a number of subtleties due to the noisiness of our data.
-Absolute orientations between devices change frequently at the slightest user movement, and motion sensor data is inherently noisy.
-To alleviate the second issue, a low-pass filter is applied to the measurements, allowing us to extract the direction of motion from this smoothed history.
-Empirically, we found that we can reliably detect the direction of movement with 100ms of history and a sensor resolution of 10ms.
-
-Orientation changes can be overcome by using a hybrid of absolute and relative orientations augmented by IR information.
-While absolute orientations are unstable, relative orientations do not contain information for essential tasks, such as adding devices to a learned map.
-Our system stores the absolute orientations of each device during the learning stage, but utilizes only the relative information embedded in this data for disambiguation.
-Once the user is in quasi-mode, our system calculates the user's direction of motion as described earlier, and searches through the map for the nearest device in that direction.
-Thus, although the absolute orientation of the currently lit device may be incorrect, the relative information can still be used for disambiguation.
-Finally, the system uses hysteresis to prevent the user from accidentally jumping repeatedly between devices by switching devices only once the user has moved beyond a certain threshold in any direction.
-\achal{I understand hysteresis, but I can't figure out how the word can be used naturally (as there's no way to say a system "has" hysteresis).
-This last sentence may require rewording.}
-
-\bjoern{how well will this work if the user shifts position within a room?}
-\achal{I think the updates should explain this now.}
-
-\achal{Not sure if there should be a conclusion of some sort here.}
+The technical implementation of this interaction involves a number of subtleties
+due to the noisiness of our data. As mentioned earlier, absolute orientations
+of devices change frequently at the slightest user movement. Unfortunately,
+while relative orientation information is more stable, it does not contain
+the necessary information for essential tasks, such as adding devices to a
+learned map. To overcome these issues, we implement a hybrid method using
+absolute and relative orientations augmented by IR information. Our system
+stores the absolute orientations of each device during the learning stage, but
+utilizes only the relative information for disambiguation. Once the user is in
+quasi-mode, our system calculates the user's direction of motion using a
+low-passed history of sensor measurements, and searches through the devices for
+the nearest device in that direction. Thus, although the absolute orientations
+may be incorrect, the relative information can still be used for disambiguation.
 

### 2014-04-16 doc/uist14/iteration3.tex
problem statement

@@
 Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the
-near-eye display for the refinement step. This has two problems: First, a
-spatial layout of targets is compressed into a uni-dimensional list. List
-navigation commands thus have poor spatial mapping. Second, our techniques
-require the user to switch their focus back and forth between the physical
-world and the display. These issues motivated us to design an approach that
-also utilizes orientation information during the {\em refinement} stage, to
-provide a more direct mapping between the user's actions and target selection.
+near-eye display for the refinement step.
+The list interface has a few clear shortcomings: navigation actions map poorly to real world results, and the users must switch their focus back and forth between the physical scene and the near-eye display.
+These problems motivated us to design an approach that harnesses the head orientation during the {\em refinement} stage.
 
-\achal{This used to have an explicit question which I removed, as it sounded
-very unnatural.}
 In our third iteration, we build a system which further minimizes the chance of
-falling back to list based refinement.
+falling back to list based {\em refinement}.
 

### 2014-04-16 doc/uist14/iteration3.tex
clarifying technique

@@
 \subsection{Technique}
-The motion sensors (including gyroscope, magnetometer and accelerometer) on the
-head-worn device can help us determine the orientation of a user's head.
-Unfortunately, this absolute orientation cannot be directly used in any indoor
-environment as the user may change their location. As long as the targets are
-spread around the periphery, however, their relative orientations are stable
+Using the motion sensors (including gyroscope, magnetometer and accelerometer) on the
+head-worn device and IR, we can discover the spatial relationships between targets and the head orientation of the user.
+
+Unfortunately, this absolute orientation cannot be applied to all indoor
+environments, since the user’s movements through the space could change the relationships between targets.
+However, with the constraint that the targets are spread around the periphery, their relative orientations are stable
 (see Figure~\ref{fig:third_principle}). This key observation inspires us to
-investigate a hybrid solution with IR and orientation sensors on Glass.
+investigate a hybrid solution using a combination of IR and orientation sensors on Glass.
 

### 2014-04-16 doc/uist14/abstract.tex
Filled some of the contents; updated a few figures; just added content to disucssion and conclusion, but the sentences are broken and may not make any sense
updated a bunch of figures
minor change to abstract
fig1 caption; resolve abstract comment
pass the paper before noon submission

@@
 
-\bjoern{updated: }
-Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
+%% Yang suggests: Interacting with smart objects in the physical space efficiently is a realistic challenge as these objects become ubiquitous.
+In this paper, we contribute HOBS, a set of novel methods for selecting physical objects at a distance using infrared-sensed head orientations.
+We augment a commercial head-worn device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.
+We present the iterative design process of our methods, involving a series of interaction technique and hardware design and user evaluations....
+Emerging head-worn computing devices can enable interactions with smart objects in physical spaces. 
 %
-We present the iterative design and evaluation of \systemname -- a head-orientation based selections method for such devices, for interacting with physical targets at a distance.
+We present the iterative design and evaluation of \systemname -- a head-orientation based selection method for interacting with these devices at a distance.
 We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.
@@
 %
-We also demonstrate a universal remote control application of the technique and report qualitative user impressions.
+Finally, we demonstrate a universal remote control application using HOBS and report qualitative user impressions.

### 2014-04-16 doc/uist14/intro.tex
pass the paper before noon submission

@@
 These objects are all potential targets for interaction.
-To initiate such interaction, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but yet not well-explored in the physical spaces.
+To initiate such an interaction, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but is not yet well-explored in the physical spaces.
 
@@
 %% introducing head-worn computing and head orientation
-Emerging head-worn computing devices have the potential to overcome some of these limitations: they do not require retrieval since the devices are already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information adjacent to physical objects in the wearer's field of view.
-We thus investigate the research question of how such computing devices may be used for selection of physical targets in the space.
+Emerging head-worn computing devices have the potential to overcome some of these limitations: they do not require retrieval since the devices are already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information in the wearer's field of view.
+We thus investigate the research question of how such computing devices may be used for selection of physical targets.
 
-Head-worn devices can naturally exploit the user's head orientation, an important, but also a less precise indicator of the user's locus of attention: it contains the general direction, but not the particular point of focus.
+Head-worn devices can naturally exploit the user's head orientation, an important but less precise indicator of the user's locus of attention.
+It contains the general direction, but not the particular point of focus.
 While gaze tracking could provide further information about a user's focus, wearable gaze tracking cannot by itself reliably determine the identity of objects in focus (e.g.
@@
 We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infrared (IR) communication between Glass and target appliances.
-In our {\em Naive IR} technique, the cone of light emitted by an IR LED (a diameter of 30-60cm and distance up to 8m) provides the {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser}A, where targets that have received IR signal light up).
-To {\em refine} selection when multiple targets have received IR signals, our solution starts with a naive ordering -- based on their names.
-This is shown in Figure~\ref{fig:teaser}B, where the targets in range are ordered numerically on the near-eye display.
+In our {\em Naive IR} technique, the cone of light emitted by an IR LED %(a diameter of 30-60cm and distance up to 8m)%
+ provides the {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser} {\em left}), where targets that have received IR signal light up.
+To {\em refine} selection when multiple targets have received IR signals, the {\em Naive IR} solution starts with an ordering based on their names.
+This is shown in the center of Figure~\ref{fig:teaser}, where targets in range are ordered numerically on the near-eye display.
 A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting \ben{the ordered-list is also list selection, needs clarification!} .
-We find that target acquisition through head orientation is preferred by users and is faster than list selection but refinements -- if necessary -- are slow.
+We find that target acquisition through head orientation is preferred by users and is faster than list selection.
+But when refinements are needed, it will become slow.
 
@@
 Our {\em Head-motion Refinement} technique first learns the relative spatial structure of the targets using Glass' orientation sensors.
-Then in future target acquisition stage, users can then perform head movements to change selections to spatially adjacent targets (Figure~\ref{fig:teaser}C).
+Then in future target acquisition stage, users can perform head movements to change selections to spatially adjacent targets (see the right of Figure~\ref{fig:teaser}).
 For example, nodding down to select the target below current selection, or tilting right to select next target on the right.
-We present initial feedback on this technique.
+We present preliminary feedback from participants on this technique.
 
@@
 An appliance-specific user interface is then shown on the user's near-eye display for further interactions.
-
-\ben{I am considering adding back the contribution part, because the paper itself mainly focuses on the iterative design process.
-Without an explicit contribution statement, the audience will have to infer and summarize themselves.}
 
@@
 
-%% Contribution
-%In summary, this paper makes the following contributions:
-%\begin{itemize}
-%\item We design and implement a novel head-orientation based selection technique for physical targets based on IR communication.
-We introduce disambiguation techniques to address the inherent imprecision of head orientation.
-%\item We present evaluations that compare head orientation targeting to list selection and quantify the benefits of automatic disambiguation.
-%\item We demonstrate a home appliance remote control application built on top of our selection technique.
-%\end{itemize}
-
-%\ben{Having problem fixing the reference}.\bjoern{skip this - we want to be brief.}
-%In the remainder of this paper, we will first describe the related works in physical selection and targeting.
-Given that we focus our scope on head orientation, in Section~\ref{sec:background} we will briefly review human's neck ergonomics as the background for head orientation.
-In Section~\ref{sec:syst-design-prot}, we then present our system designed for the study of head orientation based selection.
-The prototype is also what we have used for building the example applications.
-The disambiguation techniques are discussed in Section~\ref{sec:disamb-techn}, which is followed by the study and evaluation in Section~\ref{sec:evaluation}.
-To further show the usefulness of having such head orientation-based selection, we describe the enabled applications in Section~\ref{sec:applications}, with detailed implementation about the ``universal remote control'' application.
-The discussion and conclusion are in Section~\ref{sec:discussion} and Section~\ref{sec:conclusion} respectively.
+% In summary, this paper makes the following contributions:
+% \begin{itemize}
+% \item We presented our three iteractions of design.
+% \item We present evaluations that compare head orientation targeting to list selection and quantify the benefits of automatic disambiguation.
+% \item We demonstrate a home appliance remote control application built on top of our selection technique.
+% \end{itemize}
 

### 2014-04-16 doc/uist14/iteration1.tex
pass the paper before noon submission

@@
 
-Our results indicate that {\em IR selection} outperforms {\em list selection}.
-The average target acquisition time for {\em IR selection} is 6.67 seconds \ben{one concern that I have is this is extremely long, 6 seconds.
-Do we need to explain more?
-One possible reason here is that we ask the user not to look at targets until we called out, so the visual search time is around 3 seconds.
-Should we say more or leaving 6 seconds there is fine?}, while {\em list selection} took 8.86 {\em list selection also requires visual search, I think...} seconds (see Figure~\ref{fig:ir_vs_list}).
+Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
+while {\em list selection} took 8.86 {\em list selection also requires visual search, I think...} seconds (see Figure~\ref{fig:ir_vs_list}).
 Student's t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
@@
 It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:with_without_refinement}).
-This difference is significant ($t(19)=-2.7827, p=0.012$ using t-test.
+This difference is significant ($t(19)=-2.7827, p=0.012)$ using t-test.
 %  \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}). 

### 2014-04-16 doc/uist14/iteration2.tex
fixed fig 9 (study 2 setup)
pass the paper before noon submission

@@
 To quantify the improvement this design we performed a second study to compare the {\em Naive IR} and {\em Intensity IR} approaches.
-Because we were interested in discovering performance differences in denser environment, we reposition the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}) \bjoern{give measurements in meters}.
+Because we were interested in discovering performance differences in denser environment, we reposition the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}).
 We recruited 10 participants for this study.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/study-layout2.png}
+\includegraphics[width=1.0\columnwidth]{figures/study-layout2.pdf}
 \caption{The environment setup for our second study. In comparison to our first study, we have delibrately increased the target density.}

### 2014-04-16 doc/uist14/iteration3.tex
further edited technique

@@
 \subsection{Technique}
-Using the motion sensors (including gyroscope, magnetometer and accelerometer) on the
-head-worn device and IR, we can discover the spatial relationships between targets and the head orientation of the user.
+
+We introduce a third technique that uses a combination of motion sensors and IR to learn the relative orientations of
+targets in a room and intelligently suggest targets during refinement (see
+Figure~\ref{fig:third_technique}).
 
@@
 However, with the constraint that the targets are spread around the periphery, their relative orientations are stable
-(see Figure~\ref{fig:third_principle}). This key observation inspires us to
-investigate a hybrid solution using a combination of IR and orientation sensors on Glass.
+(see Figure~\ref{fig:third_principle}).
 
@@
 \end{figure}
-
-There are two phases to this technique: learning the relative orientations or
-targets in a room and intelligently suggesting targets during refinement (see
-Figure~\ref{fig:third_technique}).
 

### 2014-04-16 doc/uist14/iteration3.tex

@@
 
-In the first stage, the user orients their head towards each device, and the
-system learns a map of the devices based on their corresponding sensor readings.
-During the second stage, when the user wishes to use head-motion for
-disambiguation, they will put their finger on the touchpad to enter a
-quasi-mode, where only one device has a solid indicator light on at a time. This
-gesture is inspired by observing users who tend to adjust their glasses for
-closer inspection. The user can then tilt in the direction of the device they
-wish to switch to. The system then uses its understanding of the relative
-locations of the devices to create an {\em adjacency map}.
+In the learning stage, the user scans over the targets, and the
+system attains the absolute orientation of each device from IR and motion sensors. From this information it can abstract out their relative positions and build an {\em adjacency map}.
 
-The technical implementation of this interaction involves a number of subtleties
-due to the noisiness of our data. As mentioned earlier, absolute orientations
-of devices change frequently at the slightest user movement. Unfortunately,
-while relative orientation information is more stable, it does not contain
-the necessary information for essential tasks, such as adding devices to a
-learned map. To overcome these issues, we implement a hybrid method using
-absolute and relative orientations augmented by IR information. Our system
-stores the absolute orientations of each device during the learning stage, but
-utilizes only the relative information for disambiguation. Once the user is in
-quasi-mode, our system calculates the user's direction of motion using a
-low-passed history of sensor measurements, and searches through the devices for
-the nearest device in that direction. Thus, although the absolute orientations
-may be incorrect, the relative information can still be used for disambiguation.
+After the map is created, the user can hold down on the touchpad to enter a
+quasi-mode for refinement.
+In this quasi-mode, one device lights up at a time.
+When the user turns his head in the direction of another device, the light switches to that device.
+Therefore, the user can move between devices one at a time with slight head movements.
+We implement this interaction by calculating the user's direction of motion using a low-passed history of sensor measurements and searching through the adjacency map for
+the nearest device in that direction.
 

### 2014-04-16 doc/uist14/iteration3.tex
updated claire's section

@@
-%!TEX root = uist14.tex
-\section{Iteration 3: Head motion refinement}
-\label{sec:iteration-3:-head}
-\subsection{Problem Statement and Goal}
-Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the
-near-eye display for the refinement step.
-The list interface has a few clear shortcomings: navigation actions map poorly to real world results, and the users must switch their focus back and forth between the physical scene and the near-eye display.
-These problems motivated us to design an approach that harnesses the head orientation during the {\em refinement} stage.
-
-In our third iteration, we build a system which further minimizes the chance of
-falling back to list based {\em refinement}.
-
 \subsection{Technique}
@@
 
-Unfortunately, this absolute orientation cannot be applied to all indoor
+This absolute orientation cannot be applied to all indoor
 environments, since the user’s movements through the space could change the relationships between targets.
@@
 \subsection{Evaluation}
-\ben{We then need evaluation results here, quantitative and qualitative.}
-\bjoern{got to here in my pass}
-
-% %!TEX root = uist14.tex
-% \section{Iteration 3: Orientation-based refinement}
-% \bjoern{While taking IR intensity into account further reduced the need for manual refinement and increased performance, the UI navigation scheme is problematic - it is not spatially related in any meaningful way to the layout of targets in a room.
-A better interaction technique would respect that ordering.
-For example, when refinement is needed, tilting the head slightly to the right could select the target that's adjacent to the right of the currently selected target.
-
-% Such spatial navigation requires knowledge about the layout of targets in the environment.
-However, one of the strengths of our technique so far is that it does not require any map ahead of time.
-To enable some spatial navigation, we introduce a final iteration in which we build up a spatial data structure by demonstration (i.e., the user looks around the room) and then leverage that data structure during the refinement step of our interaction.}
-
-% \bjoern{This technique is based on the assumption that users will generally select targets in indoor environments where targets are spread around the periphery.
-These assumptions enable us to use orientation data without knowing the user's absolute position.}
-
-% \subsection{Implementation}
-% \bjoern{give implementation details}
-
-% \subsection{Informal User Feedback}
-% \bjoern{We informally evaluate this technique with N users: ...}
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "uist14"
-%%% End:
+We evaluate the head motion refinement method by holding an informal study and collecting qualitative feedback from a subset of 4 users from the Iteration 2 study.
+In the study, we asked users to try cycling through the targets using the new quasi-mode.
+The users had strong preferences for the new method of refinement.
+On a scale from 1-7, 1 being the least mental effort and 7 being the most mental effort users rated the old technique 4.25 and the new technique 2 on average.
+100\% indicated a preference for head movement to list navigation.
+One user referenced the issue of naming targets that the list necessitates, preferring the experience of ``matching visual cues rather than numbers''.
+Another participant remarked that it ``just made more sense'' and was a ``more natural way for demonstrating intentionality.'' The users prefered the new mapping in relation to the whole environment: ``it leveraged the spatial sense that I already had just by using the system''.
+They were also delighted to avoid list navigation, which they now called ``difficult'' and ``painful''.

### 2014-04-16 doc/uist14/intro.tex
intro recited raskin

@@
 
-Head-worn devices can naturally exploit the user's head orientation, an important but less precise indicator of the user's locus of attention.
+Head-worn devices can naturally exploit the user's head orientation, an important but less precise indicator of the user's {\em locus of attention}~\cite{raskin}.
 It contains the general direction, but not the particular point of focus.

### 2014-04-16 doc/uist14/relatedwork.tex
micro edits to related work

@@
 \section{Background and Related Work}
-Our approach is related to head and eye-controlled interfaces, area cursors, and prior work on hardware devices for pointing in physical spaces.
+Our approach is related to head and eye-controlled interfaces, area cursors and prior work on hardware devices for pointing in physical spaces.
 \subsection{Head and Gaze Input}
@@
 While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map of each space with known target locations.
-Our system works through point-to-point IR communication and does not require any a priori map.
+Our system works through point-to-point IR communication and does not require an {\em a priori} map.
 Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}.
@@
 
-Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} and as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices.
+Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices.
 Our system tackles an unresolved issue of such IR-based approaches -- navigating an area dense with potential targets.

### 2014-04-16 doc/uist14/interaction.tex
pass the paper before noon submission
grammar in interaction

@@
 Each individual target that is equipped with IR detector can react to the signal.
-Such point to point communications enables us to achieve targeting without needing a map.
+Such point to point communication enables us to achieve targeting without needing a map.
 %% \bjoern{say something about receivers and how point to point connections enable us to leverage head orientation without needing a map.}
@@
 (see Figure~\ref{fig:glass}).
-In the following description, we use the succint term ``Glass'' to refer to this augmented hardware.
+In the following description, we use the term ``Glass'' to refer to this augmented hardware.
 
@@
 
-\ben{no mention of target hardware yet. I have checked in an image of the targets ``targets.JPG'' in case we need.}
+%% \ben{no mention of target hardware yet. I have checked in an image of the targets ``targets.JPG'' in case we need.}
 
@@
 
-{\bf Refine:} When disambiguation is needed, the user needs to make an explicit selection among the targets that within his view range.
+{\bf Refine:} When disambiguation is needed, the user must make an explicit selection among the targets within his view range.
 We have designed multiple refinement mechanisms -- all enable the user to select one from a subset of targets.
-The user confirms a current selection a tap.
+The user confirms a current selection with a tap.
 Since the purpose of this stage is to disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage.
-Again, a tap confirms a selection.
+Finally, a tap confirms a decision.
 

### 2014-04-16 doc/uist14/iteration1.tex
grammar checked iteration 1

@@
 In our first implementation, we use IR for scanning as described in the previous section.
-For the refinement technique, we simply show a list of the subset of targets that have received IR signals on the Glass display.
-Since we only know that the targets are within the range of users' head orientation.
-We order them using their ids; and in realistic environment target names can also be used if names can be assigned.
-We then let users swipe to select from that list and tap to confirm the intended target.
+For the refinement stage, we simply show a list of the subset of targets that have received IR signals on the Glass display.
+Users swipe to select from that list and tap to confirm the intended target.
 
 A natural point of comparison is an interface that does not use any head orientation information - it always shows a complete list of all targets.
-We implemented a list view where users swiped forwards and backwards on the Glass touchpad to navigate incrementally through the list.
+We implemented a list view where users swiped forward and backward on the Glass touchpad to navigate incrementally through the list.
 To quantify the benefit of using IR for the scanning step, we carried out a target acquisition study which compares the {\em Naive IR selection} and {\em list selection}.
@@
 In the within-subject study, half of the participants performed {\em IR selection} first and the other half used {\em list selection} first.
-For each selection condition, we conducted 15 target selection by randomly choosing from all the targets.
+For each selection condition, we conducted 15 target selections by randomly choosing from all the targets.
 During the study, we measured the {\bf target acquisition time} for each target selection.
-Afterwards, participants were asked to complete a survey that is mainly open-ended answers about their experience.
+Afterward, participants were asked to complete a survey of mainly open-ended answers about their experience.
 
@@
 Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
-while {\em list selection} took 8.86 {\em list selection also requires visual search, I think...} seconds (see Figure~\ref{fig:ir_vs_list}).
-Student's t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
+while {\em list selection} took 8.86 {\em list} seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
 
@@
 One subject called it \studyquote{natural to interact with things just by looking at them}.
-Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquistion.
-One subject pointed out that \studyquote{{\em IR selection} doesn't requires you to swipe a lot of times like list mode but list selection}.
+Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.
+One subject pointed out that \studyquote{{\em IR selection} doesn't requires you to swipe a lot of times like list mode}.
 
 In summary, {\em Naive IR} can outperform linear list selection, and most users like this head-orientation based targeting.
-But from our quantified study result, the refinement step detracts significantly from the efficiency of the technique.
+But from our quantitative result, the refinement step detracts significantly from the efficiency of the technique.
 

### 2014-04-16 doc/uist14/iteration2.tex
minor changes to some parts
grammar iteration 2

@@
 
-\subsection{Problem Statement and Goal}
 %With our first study, we learn that using IR to perform {\em coarse selection} can reduce the overall target selection time.
@@
 Performance of the {\em Naive IR} technique will degrade as target density in an environment increases, as increased density will require refinement steps.
-We therefore ask a follow-on research question: ``How might we improve selection time in a dense environment?''
+We therefore ask a follow-on research question: {\em``How might we improve selection time in a dense environment?’’}
 
@@
 \subsection{Method}
-To quantify the improvement this design we performed a second study to compare the {\em Naive IR} and {\em Intensity IR} approaches.
-Because we were interested in discovering performance differences in denser environment, we reposition the 10 nodes and set it up in a smaller area (see Figure~\ref{fig:study-layout2}).
+To quantify the improvement in this design we performed a second study to compare the {\em Naive IR} and {\em Intensity IR} approaches.
+Because we were interested in discovering performance differences in denser environment, we reposition the 10 nodes and set them up in a smaller area (see Figure~\ref{fig:study-layout2}).
 We recruited 10 participants for this study.
 Each  user performed 30 target acquisition tasks for each approach.
-Similarly to our first within-subject study, half of them perform {\em Naive IR} first and the other half {\em Intesity IR} first.
+As in our first within-subject study, half of them perform {\em Naive IR} first and the other half {\em Intesity IR} first.
 
@@
 In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 225).
-Again, a Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, $p < 0.0001$.
+A Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, $p < 0.0001$).
 

### 2014-04-16 doc/uist14/iteration3.tex
grammar iter3

@@
 Figure~\ref{fig:third_technique}).
+
+In the learning stage, the user scans over the targets, and the
+system attains the absolute orientation of each device from IR and motion sensors. From this information it can abstract out their relative positions and build an {\em adjacency map}.
 
@@
 
-In the learning stage, the user scans over the targets, and the
-system attains the absolute orientation of each device from IR and motion sensors. From this information it can abstract out their relative positions and build an {\em adjacency map}.
-
 After the map is created, the user can hold down on the touchpad to enter a
@@
 In the study, we asked users to try cycling through the targets using the new quasi-mode.
+
 The users had strong preferences for the new method of refinement.
@@
 100\% indicated a preference for head movement to list navigation.
-One user referenced the issue of naming targets that the list necessitates, preferring the experience of ``matching visual cues rather than numbers''.
-Another participant remarked that it ``just made more sense'' and was a ``more natural way for demonstrating intentionality.'' The users prefered the new mapping in relation to the whole environment: ``it leveraged the spatial sense that I already had just by using the system''.
-They were also delighted to avoid list navigation, which they now called ``difficult'' and ``painful''.
+One user referenced the issue of naming targets that the list necessitates, preferring the experience of {\em ``matching visual cues rather than numbers’’}.
+Another participant remarked that it {\em ``just made more sense’’} and was a {\em``more natural way for demonstrating intentionality.’’} The users preferred the new mapping in relation to the whole environment: {\em ``it leveraged the spatial sense that I already had just by using the system’’}.
+They were also delighted to avoid list navigation, which they now called {\em ``difficult’’} and {\em ``painful’’}.

### 2014-04-16 doc/uist14/abstract.tex
abstract: Valkyrie's changes

@@
 A second design uses IR intensity measurement at targets to improve refinement.
-To address a lack of natural mapping of on-screen target lists to spatial target location, our third design infers a spatial data structure of the targets from demonstration to enable a natural head-motion based disambiguation.
+To address the lack of natural mapping of on-screen target lists to spatial target location, our third design infers a spatial data structure of the targets from demonstration enabling a natural head-motion based disambiguation.
 %

### 2014-04-16 doc/uist14/intro.tex
intro: grammar/minor content

@@
 These objects are all potential targets for interaction.
-To initiate such an interaction, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but is not yet well-explored in the physical spaces.
+To initiate such an interaction, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but is not yet well-explored in physical spaces.
 
@@
 Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}.
-While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and consciously aimed; that the user's hands are required to be free for operation; and that the user's visual attention is split between looking down at a screen and out at targets in the world.
+While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and cautiously aimed; that the user's hands are required to be free for operation; and that the user's visual attention is split between looking down at a screen and out at targets in the world.
 
@@
 Emerging head-worn computing devices have the potential to overcome some of these limitations: they do not require retrieval since the devices are already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information in the wearer's field of view.
-We thus investigate the research question of how such computing devices may be used for selection of physical targets.
+We thus investigate how such computing devices may be used for the selection of physical targets.
 
 Head-worn devices can naturally exploit the user's head orientation, an important but less precise indicator of the user's {\em locus of attention}~\cite{raskin}.
-It contains the general direction, but not the particular point of focus.
-While gaze tracking could provide further information about a user's focus, wearable gaze tracking cannot by itself reliably determine the identity of objects in focus (e.g.
+It suggests the general direction, but not the particular point of focus.
+While gaze tracking could provide further information about a user's focus, wearable gaze tracking cannot by itself reliably determine the identity of objects in focus (i.e.
 it also needs to know what the user is seeing).
@@
 
-The imprecision of human head movement suggests adapting area cursor techniques known from assistive devices~\cite{kabbash1995prince,worden1997making,findlater2010enhanced}.
+The imprecision of human head movement suggests adapting area cursor techniques developed for assistive devices~\cite{kabbash1995prince,worden1997making,findlater2010enhanced}.
 Such techniques employ a two-step selection process: a {\em coarse} selection of an area of interest, followed by a {\em refinement} to select a target within that area.
@@
 In our {\em Naive IR} technique, the cone of light emitted by an IR LED %(a diameter of 30-60cm and distance up to 8m)%
- provides the {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser} {\em left}), where targets that have received IR signal light up.
+ provides the {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser} {\em left}), where targets that have received an IR signal light up.
 To {\em refine} selection when multiple targets have received IR signals, the {\em Naive IR} solution starts with an ordering based on their names.
@@
 A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting \ben{the ordered-list is also list selection, needs clarification!} .
-We find that target acquisition through head orientation is preferred by users and is faster than list selection.
-But when refinements are needed, it will become slow.
+We find that target acquisition through head orientation is preferred by users and is faster than list selection, but refinement is still time consuming.
 
 In response, we designed an improved refinement technique, {\em Intensity IR}, in which target objects compare received IR signal strength.
-This value allows the system to either eliminate some distracting targets on the peripheral, or to re-order the refinement interface's list by their intensity values.
-For example, in Figure~\ref{fig:teaser}B of {\em Intensity IR} technique, 5 is eliminated first and the list is re-ordered based on the intensity readings.
+This value allows the system to eliminate some distracting targets on the peripheral and to re-order the refinement interface's list by their intensity values.
+For example, in Figure~\ref{fig:teaser}B of {\em Intensity IR} technique, device 5 is eliminated first and the list is re-ordered based on the intensity readings.
 A second study with $10$ participants proves that {\em Intensity IR} successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.
@@
 Our {\em Head-motion Refinement} technique first learns the relative spatial structure of the targets using Glass' orientation sensors.
-Then in future target acquisition stage, users can perform head movements to change selections to spatially adjacent targets (see the right of Figure~\ref{fig:teaser}).
-For example, nodding down to select the target below current selection, or tilting right to select next target on the right.
+In a future target acquisition stage, users can perform head movements to change selections to spatially adjacent targets (see the right of Figure~\ref{fig:teaser}).
+For example, nodding down to select the target below current selection, or tilting right to select the next target on the right.
 We present preliminary feedback from participants on this technique.

### 2014-04-16 doc/uist14/applications.tex
minor changes to some parts
finished what Claire has left over

@@
 
-\ben{Will add more from our survey.}
 All participants successfully completed the list of tasks.
 They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous study procedure.
-Participants saw benefits of the device for families --- \studyquote{It might also be useful for people who need to take care of small children that they can complete all the tasks while keeping an eye on their children at the same time}, though settings that require more movement than watching a movie at home (e.g., cooking) may ultimately be more compelling scenarios for wearing the device around the home.
+Participants saw potential benefits of the device for families; one user remarked that he could imagine people using the system \studyquote{while keeping an eye on their children at the same time}.
 

### 2014-04-16 doc/uist14/iteration3.tex
added section name, use studyquote

--- 2014-04-16 doc/uist14/iteration3.tex
@@
+%!TEX root = uist14.tex
+\section{Iteration 3: Head motion refinement}
+\label{sec:iteration-3:-head}
+
+Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step.
+The list interface has a few clear shortcomings: navigation actions map poorly to real world results, and the users must switch their focus back and forth between the physical scene and the near-eye display.
+These problems motivated us to design an approach that harnesses the head orientation during the {\em refinement} stage.
+
+In our third iteration, we build a system which further minimizes the chance of falling back to list based {\em refinement}.
+
 \subsection{Technique}
@@
 100\% indicated a preference for head movement to list navigation.
-One user referenced the issue of naming targets that the list necessitates, preferring the experience of ``matching visual cues rather than numbers''.
-Another participant remarked that it ``just made more sense'' and was a ``more natural way for demonstrating intentionality.'' The users prefered the new mapping in relation to the whole environment: ``it leveraged the spatial sense that I already had just by using the system''.
-They were also delighted to avoid list navigation, which they now called ``difficult'' and ``painful''.
+One user referenced the issue of naming targets that the list necessitates, preferring the experience of \studyquote{matching visual cues rather than numbers}.
+Another participant remarked that it \studyquote{just made more sense} and was a \studyquote{more natural way for demonstrating intentionality}.
+The users preferred the new mapping in relation to the whole environment: \studyquote{it leveraged the spatial sense that I already had just by using the system}.
+They were also delighted to avoid list navigation, which they now called \studyquote{difficult} and \studyquote{painful}.

### 2014-04-16 doc/uist14/interaction.tex
interaction: grammar/minor content

@@
 \section{IR for head orientation sensing}
+
 \subsection{Hardware}
@@
 
-Another benefit is the low cost and easy integration of both emitters and detectors - detectors often have support for common IR protocol decoding built in and cost cents, making it economically feasible to add detectors to dozens or hundreds of devices in an environment.
+Another benefit is the low cost and easy integration of both emitters and detectors - detectors often have support for common IR protocol decoding built in and are inexpensive, making it economically feasible to add detectors to dozens or hundreds of devices in an environment.
 
@@
 If there is only one single target in IR range, it is automatically selected.
-However, in a dense environment where multiple targets are within range, the user needs to perform a refinement.
+However, in a dense environment where multiple targets are within range, the user needs refine his selection.
 
 {\bf Refine:} When disambiguation is needed, the user must make an explicit selection among the targets within his view range.
-We have designed multiple refinement mechanisms -- all enable the user to select one from a subset of targets.
+We have designed multiple refinement mechanisms -- all of which enable the user to select one from a subset of targets.
 The user confirms a current selection with a tap.
@@
 
-The overall target acquisition time thus depends on scan and refine times; the probability that refinement is needed; and the time to commit an action (tap):
+The overall target acquisition time thus depends on scan and refine times, the probability that refinement is needed, and the time to commit an action (tap):
 \begin{equation}

### 2014-04-16 doc/uist14/relatedwork.tex
relatedwork: newlines for readability

@@
 \section{Background and Related Work}
-Our approach is related to head and eye-controlled interfaces, area cursors and prior work on hardware devices for pointing in physical spaces.
+Our approach is related to head- and eye-controlled interfaces, area cursors and prior work on hardware devices for pointing in physical spaces.
+
 \subsection{Head and Gaze Input}
-Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assitive input technology for cursor control of desktop applications~\cite{radwin1990method}.
+Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}.
 However, human neck muscles have a much lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
+
 %Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}.
-Prior work often focused on head orientation for controling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.
 
-Gaze can also be used to control graphical user intefaces~\cite{kumar2007eyepoint}.
+Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.
+
+Gaze can also be used to control graphical user interfaces~\cite{kumar2007eyepoint}.
 While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map of each space with known target locations.
@@
 Such an approach can enable similar benefits as our approach, but is computationally more expensive than our single-pixel sensor solution and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.
+
 %Our work is closer in spirit to Selker's headworn system~\cite{Selker:2001:EGE:634067.634176}.
@@
 In all area cursors, the bigger cursor activation area can lead to multiple targets being selected and disambiguation is needed.
+
 We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets).
 In the {\em refinement} phase, they adjust so only the intended target will be selected.
-Many disambiguation techniques are possible for refinement -- this paper describes the tradeoffs between several of them.
+Many disambiguation techniques are possible for refinement -- this paper describes the trade-offs between several of them.
 
 \subsection{Pointing in physical spaces}
-Rukzio studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.
+Rukzio studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list~\cite{rukzio_experimental_2006}.
 
-Several approaches to spatial selection with handheld devices exist to control appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or to exchange information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}.
+Several approaches to spatial selection with handheld devices exist for controlling appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}.
 In some techniques, users select objects of interest with laser pointers.
-The laser dot provides immediate visual feedback to the user what is being selected; however, their small target area make them poorly matched to head orientation input.
+The laser dot provides immediate visual feedback to the user about what is being selected; however, its small target area makes it poorly matched to head orientation input.
+
 Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.
@@
 Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
-Our solution is similar in spirit but relies on simple, single low-cost IR emitters and detectors.
+Our solution is similar in spirit but relies on simple low-cost IR emitters and detectors.
 %However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.

### 2014-04-16 doc/uist14/iteration1.tex
iteration1: grammar/minor content updates

@@
 \section{Iteration 1: Naive IR}
-\subsection{Problem Statement and Goal}
 Our initial research question is: {\em Can IR-based targeting reduce the selection time compared to the case where only UI list navigation is used on a head-worn device?}
@@
 We recruited 14 participants from our institution for this study.
-13 had never used Glass before, so we offered a Google Glass tutorial before the experiment for them to get familiar with Glass gestures.
-Four participants wore prescription glasses, which may have affected their task performance as wearing glasses beneath Glass makes it more cumbersome to secure the position of Glass and to adjust the screen to the optimal angle.
+13 had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
+Four participants wore prescription glasses, which makes Glass more cumbersome to use and adjust, and may have affected their task performance.
 In this study, the IR LED was fixed and not repositionable as shown in Figure~\ref{fig:glass}.
@@
 During the study, we measured the {\bf target acquisition time} for each target selection.
-Afterward, participants were asked to complete a survey of mainly open-ended answers about their experience.
+Afterward, participants were asked to complete a survey of primarily open-ended answers about their experience.
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/result_study1a.pdf}
-\caption{Boxplot of target acquisition time for {\em IR selection} and {\em list selectioN} is shown on the left.
+\caption{Boxplot of target acquisition time for {\em IR selection} and {\em list selection} is shown on the left.
 The center of each box is the median value, while we annotate the mean value using white dashed lines.}
@@
 It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:with_without_refinement}).
-This difference is significant ($t(19)=-2.7827, p=0.012)$ using t-test.
+This difference is significant $(t(19)=-2.7827, p=0.012)$ using t-test.
 %  \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}). 
@@
 
-In summary, {\em Naive IR} can outperform linear list selection, and most users like this head-orientation based targeting.
-But from our quantitative result, the refinement step detracts significantly from the efficiency of the technique.
+In summary, {\em Naive IR} can outperform linear list selection, and most users prefer this head-orientation based targeting.
+From our quantitative result, however, we foudn that the refinement step detracts significantly from the efficiency of the technique.
 

### 2014-04-16 doc/uist14/iteration2.tex
update iteration 2 quotes, and model

@@
 Performance of the {\em Naive IR} technique will degrade as target density in an environment increases, as increased density will require refinement steps.
-We therefore ask a follow-on research question: {\em``How might we improve selection time in a dense environment?’’}
+We therefore ask a follow-on research question: {\em``How might we improve selection time in a dense environment?''}
 
@@
 Received IR intensity falls off with distance between IR emitter and receiver as well as with the angle between emitter and receiver.
-To measure intensity, we add an IR light-to-voltage converter (model number TSL267-LF from AMS-TAOS USA Inc and replace the IR emitter with IR emitter (model number SFH 4545 from OSRAM Opto Semiconductors Inc).
+To measure intensity, we add an IR light-to-voltage converter TSL267-LF by AMS-TAOS USA Inc.
 % transmitter webpage: http://www.digikey.com/product-search/en?x=0&y=0&lang=en&site=us&KeyWords=475-2919-ND

### 2014-04-16 doc/uist14/iteration2.tex
iteration2: grammar/minor content updates

@@
 %because in many cases there will be only one target in range and fine selection is not needed, and in cases where it’s needed, the list length has been effectively reduced.
+
 Performance of the {\em Naive IR} technique will degrade as target density in an environment increases, as increased density will require refinement steps.
-We therefore ask a follow-on research question: {\em``How might we improve selection time in a dense environment?''}
+We therefore ask a follow-up research question: {\em``How might we improve selection time in a dense environment?''}
 
@@
 We hypothesize that IR intensity at the receiver side can provide more information about the likelihood that a user intended to select a particular target.
-Received IR intensity falls off with distance between IR emitter and receiver as well as with the angle between emitter and receiver.
+Received IR intensity falls off with distance between IR emitter and receiver as well as with the angle between the emitter and receiver.
 To measure intensity, we add an IR light-to-voltage converter TSL267-LF by AMS-TAOS USA Inc.
+
 % transmitter webpage: http://www.digikey.com/product-search/en?x=0&y=0&lang=en&site=us&KeyWords=475-2919-ND
@@
 \caption{Empirical measurement of IR intensity at different position.
-We measure the intensity at the plane of every 1 meter away from the IR emitter.
+We measure the intensity at a plane every meter away from the IR emitter.
 On that plane, we obtain a sample every 5 centimeters.
@@
 \begin{enumerate}
-\item When multiple targets have received IR signals and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half of the ADC resolution, which is frequently used because it indicates a 3dB loss in the signal strength.}.
+\item When multiple targets have received IR signals and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half of the ADC resolution, which is frequently used as it indicates a 3dB loss in the signal strength.}.
 Therefore, when there is only one target within line of sight, the IR intensity approach has the same behavior as the previous iteration - no disambiguation is needed.
@@
 \item When refinement is still needed, meaning that multiple targets have relatively close intensity values, the system sorts the disambiguation list according to the IR intensity, from strongest to weakest.
-We hypothesize that this will reduce time $t_{refine}$ significantly by minimizing extra navigation steps.
-(Ideally, the first list item will match the actual, intended target.)
+We hypothesize that this will reduce $t_{refine}$ significantly by minimizing extra navigation steps, as the first list item will generally match the intended target.
 \end{enumerate}
@@
 
-In the cases when disambiguation is inevitable, {\em Intensity IR} sorts the list based on the intensity reading, while {\em Naive IR} sorts alphabetically.
+In the cases where disambiguation is inevitable, {\em Intensity IR} sorts the list based on the intensity reading, while {\em Naive IR} sorts alphabetically.
 {\em Intensity IR} reduces the fraction of refinement trails in which additional list navigation is necessary (i.e., the first, already selected element is incorrect).

### 2014-04-16 doc/uist14/conclusion.tex
applications, conclusion, discussion, iteration3: grammar/minor content

@@
 With IR intensity added, the targeting can work better in a relative dense environment.
-However, a more natural while challenging approach is to combine IR with the head motion.
-Through our preliminary user studies on this side, we learned that this is a more intuitative way of performing refinement in comparison to the menu-based selections.
+However, a more natural approach is to combine IR with head motion.
+Through our preliminary user studies, we learned that this is a more intuitative way of performing refinement in comparison to the menu-based selections.
 We leave a more comprehensive technical solution of using motion sensors and its evaluation as the future work.

### 2014-04-16 doc/uist14/discussion.tex
applications, conclusion, discussion, iteration3: grammar/minor content

@@
 
-A primary goal of this paper was to provide an effective anf efficient method for target selections in physical spaces.
+A primary goal of this paper was to provide an effective and efficient method for target selections in physical spaces.
 Targeting is a fundamental building block across many interaction tasks - it has a significant impact on user experiences collectively and can provide seamless interaction when designed well.
 
-We formalized a scan+refine model of head orientation-based selection (Equation~\ref{eq:time}).
+We formalized a scan and refine model of head orientation-based selection (Equation~\ref{eq:time}).
 We first introduced head orientation as an alternative to list selection and showed that {\em scanning} can outperform list selection.
@@
 Additional strong IR sources like direct sunlight may saturate the sensor and make discrimination impossible.
-Second, our adjacency map is built assuming on stable target locations.
+Second, our adjacency map is built assuming stable relative target locations.
 If targets move, the map will have to be recalculated.
 This may be done incrementally during everyday interactions, but we have not yet tackled this challenge.
-We also have not yet evaluated how robust the adjacency maps is when the user changes her location or view angle.
+We also have not yet evaluated how robust the adjacency map is when the user changes her location or view angle.
 

### 2014-04-16 doc/uist14/iteration3.tex
applications, conclusion, discussion, iteration3: grammar/minor content

@@
 
-In our third iteration, we build a system which further minimizes the chance of falling back to list based {\em refinement}.
+In our third iteration, we ask the question {\em can we build a system purely using head orientation and visual feedback from the environment for target selection?}
 
@@
 Though we store the absolute orientation value, during the {\em refinement} stage, the prediction is only based on relative changes to a reference point.
-In such a way, the prediction will work regardless of user's position.}
+Thus, the prediction will work regardless of user's position.}
 \label{fig:third_technique}

### 2014-04-16 doc/uist14/interaction.tex
removed a dot
fixed two stage, four states

@@
 Our augmentation to Google Glass adds a 940nm 5mm IR emitter (SFH 4545 from OSRAM Opto Semiconductors Inc.).
-The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio.
-(see Figure~\ref{fig:glass}).
+The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio (see Figure~\ref{fig:glass}).
 In the following description, we use the term ``Glass'' to refer to this augmented hardware.
@@
 \includegraphics[width=\columnwidth]{figures/interactionModel2.pdf}
-\caption{Four different states during the head-orientation targeting.}
+\caption{Two main stages during the interaction with HOBS. For completeness, we have also show the final state of {\bf commit}.} 
 \label{fig:interaction}

### 2014-04-16 doc/uist14/relatedwork.tex
doing vfill and eject also make sure titles are captical cases in related works

@@
 
-\subsection{Pointing in physical spaces}
+\subsection{Pointing in Physical Spaces}
 Rukzio studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list~\cite{rukzio_experimental_2006}.

### 2014-04-16 doc/uist14/intro.tex

@@
 This is shown in the center of Figure~\ref{fig:teaser}, where targets in range are ordered numerically on the near-eye display.
-A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting \ben{the ordered-list is also list selection, needs clarification!} .
+A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting \ben{the ordered-list is also list selection, needs clarification!}.
 We find that target acquisition through head orientation is preferred by users and is faster than list selection, but refinement is still time consuming.

### 2014-04-16 doc/uist14/iteration1.tex
doing vfill and eject also make sure titles are captical cases in related works
ispell chekcing

--- 2014-04-16 doc/uist14/iteration1.tex
@@
 
+Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
+while {\em list selection} took 8.86 {\em list} seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
+
 \begin{figure}[t]
@@
 \caption{Boxplot of target acquisition time for {\em IR selection} and {\em list selection} is shown on the left.
-The center of each box is the median value, while we annotate the mean value using white dashed lines.}
+The center of each box is the median value, while we annotate the mean value using white dashed lines.
+The cumulative distributions are shown on the right.}
 \label{fig:ir_vs_list}
@@
 
-Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
-while {\em list selection} took 8.86 {\em list} seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
-
 To understand how scanning and refinement contribute to total selection time, we split the data from {\em IR selection} into two parts -- trials that required refinement and ones that did not.
@@
 Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}.
-From this figure, we can see that once there are more than 6 targets, the average acquistion time will be larger than IR selection, even if disambiguation is needed.
+From this figure, we can see that once there are more than 6 targets, the average acquisition time will be larger than IR selection, even if disambiguation is needed.
 %Therefore, we argue that once there are more targets in the physical space, the {\em list selection} method is not scalable.
@@
 In summary, {\em Naive IR} can outperform linear list selection, and most users prefer this head-orientation based targeting.
-From our quantitative result, however, we foudn that the refinement step detracts significantly from the efficiency of the technique.
+From our quantitative result, however, we found that the refinement step detracts significantly from the efficiency of the technique.
 

### 2014-04-17 doc/uist14/iteration3.tex
made some final changes before submission

@@
 \includegraphics[width=1\columnwidth]{figures/third_principle.png}
-\caption{Even when the user's absolution position changes, the relative relationship in the orientation space remain invariant.}
+\caption{This illustrates that when the change of user's absolution position doesn't change the relative relationship of physical targets.}
 \label{fig:third_principle}
@@
 \includegraphics[width=1\columnwidth]{figures/third_technique3.pdf}
-\caption{\ben{figure in interactionModel2.key file.} Our third technique learned each target's absolute orientation and construct the orientation map.
-Though we store the absolute orientation value, during the {\em refinement} stage, the prediction is only based on relative changes to a reference point.
-Thus, the prediction will work regardless of user's position.}
+\caption{Our third technique learned each target's absolute orientation and construct the adjacency map.
+During the {\em refinement} stage, the prediction is based on relative changes to a reference point.}
 \label{fig:third_technique}

### 2014-04-17 data/201309.Studies/script/visualize.R
check in my changes to R file and visualizatio figure

@@ -1,3 +1,4 @@
+setwd("~/repos/CS294-84Project/data/201309.Studies/script/")
 require('ggplot2')
 
 ## to run this script, you need multiplot
@@ -223,14 +224,16 @@ dev.off()
 toPlot <- mode.bind.type[grepl("IR", mode.bind.type$mode_type), ]
 for (i in 1:nrow(toPlot)) {
   if (toString(toPlot$mode_type[i]) == "IR single") {
-    toPlot$type[i] <- "without refinement"
+    toPlot$type[i] <- "no refinement"
   }
   else {
     toPlot$type[i] <- "with refinement"
   }
 }
 
-box2 <- ggplot(toPlot, aes(factor(type), correct_time, fill = toPlot$mode_type))
+box2$type <- factor(box2$type, c("without refinement ","with refinement"))
+
+box2 <- ggplot(toPlot, aes(factor(type), correct_time, fill = factor(toPlot$type)))
 box2 <- box2 + geom_boxplot(lwd=0.5) + ylim(2.5,25) + xlab("") +
   scale_fill_manual(values = c("#0072B2", "#56B4E9")) +
   annotate("text", x = 1, y = 25, label = "mean: 9.16", size=3.6) +
@@ -240,9 +243,10 @@ box2 <- box2 + geom_boxplot(lwd=0.5) + ylim(2.5,25) + xlab("") +
   stat_summary(fun.y = "mean", geom = "text", label="---", size= 8, color= "white") +
   theme_bw() +
   theme(legend.position = "none", axis.title.y=element_blank()) +
-  scale_x_discrete(labels=c("with refinement", "without refinement"))
+  scale_x_discrete(labels=c("no refinement", "with refinement"))
 
 cdf2 <- ggplot(toPlot, aes(x = correct_time, colour = type)) +
+  scale_colour_manual(values = c("#0072B2", "#56B4E9")) +
   stat_ecdf() +
   ylab("cumulative distribution function") +
   xlab("acquisition time") +

### 2014-04-17 server/fitts.py
changed the fitts law study script, this version is what we have used for the second iteration study

@@ -1,6 +1,6 @@
 ## this is the automatic script to perform Fitts' selection
 import serial, glob, argparse, Queue
-import sys, select, random, threading
+import sys, select, random, threading, time
 from datetime import datetime
 from random import shuffle
 
@@ -9,6 +9,7 @@ parser = argparse.ArgumentParser(description='Fitts\' Selection script')
 parser.add_argument('--baud', type=int, action='store', default=9600, help='Specify the baud rate')
 parser.add_argument('--timeout', type=float, action='store', default=1, help='Timeout parameter for serial connection')
 parser.add_argument('--serial', action='store', default=None, help='serial port')
+parser.add_argument('--serial2', action='store', default=None, help='serial port 2')
 
 availables = glob.glob('/dev/tty.*')
 
@@ -17,9 +18,11 @@ try:
   baud = arguments.baud
   timeout = arguments.timeout
   serial_port = arguments.serial
+  serial_2 = arguments.serial2
   print "  baud rate:", baud
   print "  serial timeout:", timeout
   print "  serial port:", serial_port
+  print "  serial for Arduino:", serial_2
 except Exception as e:
   print e
   exit(1)
@@ -48,45 +51,46 @@ except Exception as e:
 
 target_list = [3, 4, 5, 6, 7, 8, 11, 12, 14, 17]
 ## target_list = [3, 12, 4]
-random.seed( datetime.now().second )
+random.seed( datetime.now().microsecond )
 
-shuffle(target_list)
-print >> sys.stderr, target_list
+id = random.sample(target_list, 1)[0]
 
-for id in target_list[0:5]:
-  ser.write("L" + "%02d" % id)
-  goal = "C" + "%02d" % id
-  print >> sys.stderr, goal
-  start_time = datetime.now()
-  print '[', start_time.strftime('%Y-%m-%d %H:%M:%S %f'), ', Command]: ', goal
-  seconds = 0
-  line = ""
-  while goal not in line:
-    if ( ( datetime.now() - start_time).seconds > 20 ):
-      line = "D"
-      ser.write(line)
-      break
+goal = "C" + "%02d" % id
+
+print >> sys.stderr, goal
+start_time = datetime.now()
+print '[', start_time.strftime('%Y-%m-%d %H:%M:%S %f'), ', Command]:', goal
+seconds = 0
+line = ""
+
[…]

### 2014-06-23 doc/sui14/abstract.tex
created sui2014 dir

--- 2014-04-16 doc/uist14/abstract.tex
@@
 A second design uses IR intensity measurement at targets to improve refinement.
-To address a lack of natural mapping of on-screen target lists to spatial target location, our third design infers a spatial data structure of the targets from demonstration to enable a natural head-motion based disambiguation.
+To address the lack of natural mapping of on-screen target lists to spatial target location, our third design infers a spatial data structure of the targets enabling a natural head-motion based disambiguation.
 %

### 2014-06-23 doc/sui14/applications.tex
created sui2014 dir

--- 2014-04-16 doc/uist14/applications.tex
@@
 
-We implemented a prototype of this remote controller for three physical devices: a lamp and fan that could be switched on or off; and a smart TV control with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
+We implemented a prototype of this remote controller for three physical devices: a lamp and fan that could be switched on or off, and a smart TV control with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
 We switched discrete appliances using relays, and implemented the smart TV with a laptop connected to a 30\inch{} display.
@@
 \caption{In the smart home scenario, we have built three smart appliances for a user experience study.
-The control interface supports both simple on/off appliances (lamp or fan) and multi-functional appliances (TV, for example).}
+The interface supports both simple on/off appliances (lamp or fan) and multi-functional appliances (TV, for example).}
 \label{fig:smart-home}

### 2014-06-23 doc/sui14/disambiguation.tex
created sui2014 dir

--- 2014-04-05 doc/uist14/disambiguation.tex
@@
 
-For area selection technique, when multiple targets are within the range, the disambiguation is necessary.
-We propose three different ways of disambiguation solutions in this section.
-Initially our system leaves the disambiguation to the user \ben{cite our paper}.
-However, our studies showed that manual disambiguation is a time-consuming task, so we only consult manual action when the system cannot figure out itself.
-The automatic disambiguation are based on the observation that users tend to adjust their head orientation a bit when they found the target is wrong.
-So we achieve a more fine-grained targeting by exploiting the IR intensity and the change of IR readings, as well as the sensors on Google Glass.
+In our previous section we have described how we design and choose the COTS components for our prototype.
+However, for area selection technique, when multiple targets are within the range, the disambiguation is necessary.
+We could just leave this as a manual operation to the user; however, from previous studies \cite{Chen:EECS-2013-200} we have found that this becomes the bottleneck for seamless interaction in many cases.
+Therefore, we consider the problem of simplifying disambiguation in physical targeting contexts by incorporating multiple sources of sensors.
+
+The first and most straightforward way is to use the signal strength of area selection technique.
+For IR, we have added IR light-to-voltage converter to read the absolute IR intensity.
+The second class of sensors are those available on wearable devices.
+Google Glass' motion sensors are the ones that we have used in our study.
+The automatic disambiguation will definitely save time, but leaving the option of manual switching targets will still be useful.
+But this doesn't necessarily mean the naive list browsing technique.
+We consider how the IR and motion sensors can helps us build a smarter list -- adjacent map -- when manual switching is performed.
+We talk about each technique in the following few subsections.
 

### 2014-06-23 doc/sui14/intro.tex
created sui2014 dir

--- 2014-04-16 doc/uist14/intro.tex
@@
 This value allows the system to eliminate some distracting targets on the peripheral and to re-order the refinement interface's list by their intensity values.
-For example, in Figure~\ref{fig:teaser}B of {\em Intensity IR} technique, device 5 is eliminated first and the list is re-ordered based on the intensity readings.
+For example, in Figure~\ref{fig:teaser} of {\em Intensity IR} technique, device 5 is eliminated first and the list is re-ordered based on the intensity readings.
 A second study with $10$ participants proves that {\em Intensity IR} successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.

### 2014-06-23 doc/sui14/sui14.tex
created sui2014 dir

--- 2013-09-20 doc/sigchi14/sigchi14.tex
@@
 Request permissions from permissions@acm.org.
-\\
+\\\textbf{}
 % {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
@@
 %% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
-
-\toappear{Submitted to CHI2014. Do not cite or distribute without written permission from the authors.}
+\toappear{Submitted for review to SUI 2014.}
 
@@
 \usepackage{url}      % llt: nicely formatted URLs
-\usepackage{gensymb}
-\usepackage{afterpage}
-% see http://tex.stackexchange.com/questions/46055/typesetting-with-inch-symbols-and-sizes-in-inches
-\usepackage{mathpazo,amsmath}
-\def\inch#1{#1''}
-\def\ft#1{#1'\thinspace}
-
 % llt: Define a global style for URLs, rather that the default one
@@
 \urlstyle{leo}
+
+\usepackage{gensymb}
+\usepackage{afterpage}
+% see http://tex.stackexchange.com/questions/46055/typesetting-with-inch-symbols-and-sizes-in-inches
+% \usepackage{mathpazo}
+\usepackage{amsmath}
+\def\inch#1{#1''}
+\def\ft#1{#1'\thinspace}
+
+% comment macros are in inputmacros.tex
+\input{inputmacros}
 
@@
 
-%use these commands while writing
-\newcommand {\bjoern}[1]{{\color{red}\bf{BH: #1}\normalfont}}
-\newcommand {\ben}[1]{{\color{blue}\bf{BZ: #1}\normalfont}}
-\newcommand {\sean}[1]{{\color{blue}\bf{SC: #1}\normalfont}}
-\newcommand {\claire}[1]{{\color{blue}\bf{CT: #1}\normalfont}}
-\newcommand {\studyquote}[1]{\em ``#1''\normalfont}
-%comment out the above and uncomment these for final submit
-%\newcommand {\bjoern}[1]{}
-%\newcommand {\ben}[1]{}
-%\newcommand {\sean}[1]{}
-%\newcommand {\claire}[1]{}
-
 % End of preamble. Here it comes the document.
@@
 
-\title{A Context Menu for the Real World: Controlling Physical Appliances Through Head-Worn Infrared Targeting}
+% as we discussed last time, the "Sees" and "Line-of-sight" suggests some computer vision direction 
+% need a different title for sure, but not in a hurry
+\title{\systemnamenospace: Head Orientation Based Selection\\in Physical Spaces}
+%Project GlasSees: Direct Interaction Through Line-of-sight}
 
+\iftoggle{anonymous}{
+\author{
+ \alignauthor Anonymous for submission\\
+    %\affaddr{...}\\
+    %\email{...}\\
+  }
+}{ %else
   \numberofauthors{1}
   \author{
-\alignauthor Yu-Hsiang Chen$^{\dagger}$, Ben Zhang$^{\dagger}$, Claire Tuna$^{\dagger}$, Yang Li$^{\ddagger}$, Edward Lee$^{\dagger}$, Bj\"orn Hartmann$^{\dagger}$ \\
+  \alignauthor  Ben Zhang$^{\dagger}$, Yu-Hsiang Chen$^{\dagger}$, Claire Tuna$^{\dagger}$, Achal Dave$^{\dagger}$,\\Yang Li$^{\ddagger}$, Edward Lee$^{\dagger}$, Bj\"orn Hartmann$^{\dagger}$ \\
   \affaddr{$\dagger$: UC Berkeley EECS \& CITRIS Invention Lab \hspace{0.25in} $\ddagger$: Google Research}\\
-\email{sean.yhc@gmail.com, benzh@eecs.berkeley.edy, clairetuna@gmail.com, yangli@acm.org, eal@eecs.berkeley.edu, bjoern@eecs.berkeley.edu}
+  \email{\{benzh,clairetuna,achal,eal,bjoern\}@berkeley.edu, sean.yhc@gmail.com, yangli@acm.org}
   }
-%\author{
-% \alignauthor Anonymous for submission
-%  \alignauthor 3rd Author Name\\
-%    \affaddr{Affiliation}\\
-%    \affaddr{Address}\\
-%    \email{e-mail address}\\
-%    \affaddr{Optional phone number}
+}
+
+\teaser{
+  \centering
+  \includegraphics[width=\textwidth]{figures/teaser.png}
+  \caption{{\em Left}: Our head orientation-based selection techniques use an IR emitter -- multiple targets may fall within its illumination area.
+{\em Center}: We offer two list-based refinement techniques -- {\em Naive IR} uses alphabetical ordering; {\em Intensity IR} orders targets by IR intensity.
+{\em Right}: Alternatively, users can refine their selection through head orientation refinement in a quasi-mode.}
+  \label{fig:teaser}
+}
+
+%% the third part of the teaser: An example application of head-orientation targeting: Using an augmented head-worn device (1), users can control smart home appliances (2) by looking at them.
+A near-eye display then shows an appliance control UI (3), which users navigate through multitouch gestures.}
 
@@
 \end{abstract}
+\vspace{-0.05in}
+\keywords{
+  Wearable computing; selection; glass; infrared
+}
+\vspace{-0.1in}
+%% refer to http://www.acm.org/about/class/ccs98-html
+\category{H.5.2.}{Information Interfaces and Presentation (e.g. HCI)}{Interaction styles (e.g., commands, menus, forms, direct manipulation).}
 
-\keywords{
-	smart devices; universal remote control; wearable computing; glass
-}
-
-\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}
-
-%$See: \url{http://www.acm.org/about/class/1998/}
-%for more information and the full list of ACM classifiers
-%and descriptors. 
-%\textcolor{red}{Mandatory section to be included in your
-%final version. On the submission page only the classifiers'
-%letter-number combination will need to be entered.}
-
-\input{introduction}
+\input{intro}
 \input{relatedwork}
 \input{interaction}
-\input{device}
-\input{targeting}
-\input{control}
+\input{iteration1}
+\input{iteration2}
+\input{iteration3}
+%\input{disambiguation}
+%\input{evaluation}
+
+%% the rest of the paper seems to be fine; it's mainly the texts
+\input{applications}
 \input{discussion}
 \input{conclusion}
+
+\iftoggle{anonymous}{
+% no acks in anonymous submission
+}{
   \section{Acknowledgments}
-This work was partially supported by a Sloan Foundation Fellowship.
 
+  This work was supported in part by the TerraSwarm Research Center, one of six centers supported by the STARnet phase of the Focus Center Research Program (FCRP) a Semiconductor Research Corporation program sponsored by MARCO and DARPA.
+Additional support was provided by a Sloan Foundation Fellowship and a Google Research Award.
+}
 %figure template:
@@
 %
-\balance
-
-% REFERENCES FORMAT
-% References must be the same font size as other body text.
+%% \balance
 
 \bibliographystyle{acm-sigchi}
-\bibliography{sigchi14}
+\bibliography{sui14}
 \end{document}

### 2014-06-25 doc/sui14/discussion.tex
created sui2014 dir
first pass with rebuttals

--- 2014-04-16 doc/uist14/discussion.tex
@@
 This may be done incrementally during everyday interactions, but we have not yet tackled this challenge.
-We also have not yet evaluated how robust the adjacency map is when the user changes her location or view angle.
 
 Finally, we acknowledge several limitations of our study design: we have not yet systematically studied target density variation, our study was performed in a lab environment, and only measured first use.
-We have also not investigated error rates in detail.
 Future work should study how the technique applies in realistic settings over longer periods of time.
@@
 %- Study limitation: density variation, realistic settings, longitudinal studies to understand how users choose between our techniques and traditional list-based solution.
+
+\subsection{Computer-vision Solution?}
+Reviewers suggest considering computer vision-based techniques with markers in the environment (such as QR codes, or Bokode [Mohan] which extends the working distance to a few meters for visual tags).
+We considered this direction and found several drawbacks, including computational complexity, low accuracy when targets are far, the absence of real-time environment feedback with passive tags, and aesthetic issues with QR codes.
+While we don't claim our choice of IR is optimal, it has important advantages -- it is low-cost, readily available and quite suitable for area selection.
+We can weaken our claim and add explanations of how disambiguation techniques might generalize to other methods of implementing target identification.
 

### 2014-06-25 doc/sui14/interaction.tex
created sui2014 dir
first pass with rebuttals

--- 2014-04-16 doc/uist14/interaction.tex
@@
 \subsection{Hardware}
+
+\ben{a bit more on system description.}
+
 A central hypothesis of this paper is that the area cursor paradigm~\cite{kabbash1995prince} is well matched to head orientation selection.
@@
 \includegraphics[width=\columnwidth]{figures/interactionModel2.pdf}
-\caption{Two main stages during the interaction with HOBS. For completeness, we have also show the final state of {\bf commit}.} 
+\caption{{\em scan} and {\em refine} -- two main stages during the interaction with HOBS. For completeness, we have also added the final state of {\em commit}.} 
 \label{fig:interaction}

### 2014-06-25 doc/sui14/iteration2.tex
first pass with rebuttals

--- 2014-04-16 doc/uist14/iteration2.tex
@@
 Performance of the {\em Naive IR} technique will degrade as target density in an environment increases, as increased density will require refinement steps.
-We therefore ask a follow-up research question: {\em``How might we improve selection time in a dense environment?''}
+We therefore ask a follow-up research question: {\em How might we improve selection time in a dense environment?}
 
@@
 Each  user performed 30 target acquisition tasks for each approach.
-As in our first within-subject study, half of them perform {\em Naive IR} first and the other half {\em Intesity IR} first.
+As in our first within-subject study, half of them perform {\em Naive IR} first and the other half {\em Intensity IR} first.
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/study-layout2.pdf}
-\caption{The environment setup for our second study. In comparison to our first study, we have delibrately increased the target density.}
+\caption{The environment setup for our second study. In comparison to our first study, we have deliberately increased the target density.}
 \label{fig:study-layout2}
@@
 {\em Intensity IR} reduces the fraction of refinement trails in which additional list navigation is necessary (i.e., the first, already selected element is incorrect).
-{\em Intensity IR} sorted the desired target as the first one in the list in 55\% of cases (93 of 167).
+{\em Intensity IR} sorted the desired target as the first one in the list in 55\% of cases (93 of 167) \ben{explicitly saying this is higher than random guess?}.
 In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 225).
@@
 
-From Figure~\ref{fig:study2}, we can see that the overall target aquisition time has decreased from 4.31 seconds for {\em Naive IR} to 3.64 second for {\em Intensity IR}.
+From Figure~\ref{fig:study2}, we can see that the overall target acquisition time has decreased from 4.31 seconds for {\em Naive IR} to 3.64 second for {\em Intensity IR}.
 This difference is also significant ($t(555)=3.2945$, $p=0.001$.)

### 2014-06-25 doc/sui14/relatedwork.tex
first pass with rebuttals

--- 2014-04-16 doc/uist14/relatedwork.tex
@@
 
+\ben{new related works from reviewers}.
+\changes{RFIG Lamps and photosensing wireless tags [Raskar et al] use a projector and data is encoded in each pixel so that tags can localize themselves for target selection.
+Our IR emitter, a relatively low-cost solution, doesn’t have this feature but shares the commonality of covering an area for the ease of selection.
+At the end of their paper, they envision an IR-based system to solve ambient light problems.
+Our work lies in that direction and also contributes evaluations with user studies.
+Progressive Refinement [Bacima] studies several progressive refinement selection modalities, which is close to our two-stage selection; but our contexts are head-orientation as it reflects users' point of interest.
+}
+
 \subsection{Pointing in Physical Spaces}

### 2014-06-25 doc/sui14/intro.tex
minor comments and changes

@@
-%!TEX root = uist14.tex
+%!TEX root = sui14.tex
 \section{Introduction}
-% \bjoern{something happened to the figure caption font - it doesn't match the standard template. Please fix.}
-% \ben{fixed. Problem is because of using caption package}.
+\bjoern{We should see if we can change language in a few places to strengthen the connection to ``Spatial User Interaction", the title of the conference.}
 
@@
 These objects are all potential targets for interaction.
-To initiate such an interaction, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but is not yet well-explored in physical spaces.
-
-%% To initiate such interaction requires {\em selection} information about the object at the user's {\em locus of attention}~\cite{raskin}. \sean{this sentence is unclear to me.}
+To initiate spatial interactions, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but is not yet well-explored in physical spaces.
 
@@
 Emerging head-worn computing devices have the potential to overcome some of these limitations: they do not require retrieval since the devices are already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information in the wearer's field of view.
-We thus investigate how such computing devices may be used for the selection of physical targets.
+We thus investigate how such computing devices may be used for the selection and control of devices in physical spaces.
 
-Head-worn devices can naturally exploit the user's head orientation, an important but less precise indicator of the user's {\em locus of attention}~\cite{raskin}.
+Head-worn devices can naturally exploit the user's head orientation, an important (but imprecise) indicator of the user's {\em locus of attention}~\cite{raskin}.
 It suggests the general direction, but not the particular point of focus.
@@
 We therefore study how to leverage head-orientation alone for target selection in physical spaces.
+\bjoern{I think this needs reworking.
+In essence, reviewers didn't believe us that head orientation was a sensible way to go when compared to more complex gaze tracking or computer vision approaches.}
 

### 2014-06-25 doc/sui14/interaction.tex
more on hardware

@@
-%!TEX root = uist14.tex
+%!TEX root = sui14.tex
 \section{IR for head orientation sensing}
@@
 Therefore, point selection techniques like laser pointers are not appropriate.
-
 Infrared emitters are a good technology match since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible.
@@
 Such point to point communication enables us to achieve targeting without needing a map.
-%% \bjoern{say something about receivers and how point to point connections enable us to leverage head orientation without needing a map.}
 
-Another benefit is the low cost and easy integration of both emitters and detectors - detectors often have support for common IR protocol decoding built in and are inexpensive, making it economically feasible to add detectors to dozens or hundreds of devices in an environment.
+%Another benefit is the low cost and easy integration of both emitters and detectors - detectors often have support for common IR protocol decoding built in and are inexpensive, making it economically feasible to add detectors to dozens or hundreds of devices in an environment.
+\bjoern{added some more detail back in}
+\changes{Our augmentation to Google Glass adds a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545).
+The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio, since Glass does not directly support hardware modification (see Figure~\ref{fig:glass}).
+Target devices use Vishay TSOP38238 IR receivers.
+Data is encoded using standard IR remote protocols at 38.0kHz.
 
-Our augmentation to Google Glass adds a 940nm 5mm IR emitter (SFH 4545 from OSRAM Opto Semiconductors Inc.).
-The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio (see Figure~\ref{fig:glass}).
-In the following description, we use the term ``Glass'' to refer to this augmented hardware.
+IR signals are used intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send IR signal status from targets to back to Glass and for device control itself.
+We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose.
+ (Figure~\ref{fig:architecture}).
+This architecture was mostly chosen for reasons of expediency.
+We selected XBee 802.15.4 radios to avoid Bluetooth  wake up latencies but we do not claim optimality for our design decisions.
+Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
+In particular, one could substitute WiFi modules or design an all-Bluetooth network and forego intermediate layers in our design.}
 
@@
 
-While we use IR signal for targeting, wireless communications are needed to efficiently transmit the reception status of IR signal from targets to Glass.
-We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose.
+%% \ben{no mention of target hardware yet. I have checked in an image of the targets ``targets.JPG'' in case we need.}
 
-%% \ben{no mention of target hardware yet. I have checked in an image of the targets ``targets.JPG'' in case we need.}
+%For the IR receiver, we use 38.0kHz IR Receiver Modules from Vishay Semiconductor Opto Division (manufacturer part number TSOP38238). 
+% receiver webpage http://www.digikey.com/product-detail/en/TSOP38238/751-1227-ND/1681362
+%In order to read the IR intensity, we integrates another IR light-to-voltage converter from AMS-TAOS USA Inc (manufacturer part number TSL267-LF) which is pretty sensitive to IR irradiance. 
+% % http://www.digikey.com/product-detail/en/TSL267-LF/TSL267-LF-ND/3095052
 
@@
 \label{fig:glass}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/architecture}
+\caption{In our system architecture, selection is initiated through infrared but confirmed over 802.15.4.
+This permits wearers to move their head freely after connecting to an appliance.
+In the research prototype, users have to carry an additional microcontroller board that marshals messages between Glass' Bluetooth radio an IR/802.15.4, but our custom hardware could also be integrated into the wearable device.}
+\label{fig:architecture}
 \end{figure}

### 2014-06-25 doc/sui14/relatedwork.tex
folded in a missing reference

--- 2014-04-16 doc/uist14/relatedwork.tex
@@
-%!TEX root = uist14.tex
+%!TEX root = sui14.tex
 \section{Background and Related Work}
@@
 Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}.
-However, human neck muscles have a much lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
+However, human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
 
@@
 \subsection{Area Cursors}
-In GUI area cursors, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}.
+\changes{
+In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}.
+We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy).
 Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}.
-We argue that head orientation pointing has similar challenges (limited pointing performance and accuracy).
-In all area cursors, the bigger cursor activation area can lead to multiple targets being selected and disambiguation is needed.
+Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~
+\cite{bacim2013design}. In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed.
 
@@
 Many disambiguation techniques are possible for refinement -- this paper describes the trade-offs between several of them.
+}
 
@@
 
-Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
+Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
 Our solution is similar in spirit but relies on simple low-cost IR emitters and detectors.
+\bjoern{need a stronger statement here}
 %However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.
@@
 Our system tackles an unresolved issue of such IR-based approaches -- navigating an area dense with potential targets.
+
+%\ben{new related works from reviewers}. \bjoern{folded in}
+%\changes{RFIG Lamps and photosensing wireless tags [Raskar et al] use a projector and data is encoded in each pixel so that tags can localize themselves for target selection.
+Our IR emitter, a relatively low-cost solution, doesn’t have this feature but shares the commonality of covering an area for the ease of selection.
+At the end of their paper, they envision an IR-based system to solve ambient light problems.
+Our work lies in that direction and also contributes evaluations with user studies.
+%Progressive Refinement [Bacima] studies several progressive refinement selection modalities, which is close to our two-stage selection; but our contexts are head-orientation as it reflects users' point of interest.
+%}

### 2014-06-26 doc/sui14/relatedwork.tex
Added info about CV approaches

@@
 
+\subsection{Markers and Vision Methods}
+\achal{Not sure where to put this, feel free to move.}
+\changes{Many alternative solutions for detecting devices in contained spaces
+rely on passive tags on devices, combined with computer vision methods for
+detection. Unfortunately, these methods either impose significant constraints
+on the camera used for detection, or require large, obtrusive tags, and
+inevitably lack feedback from passive devices in the environment. \cite{Bokode},
+for example, relies on visually imperceptible codes, but has significant
+drawbacks: despite using a 12.1 megapixel DSLR camera, detections are no
+longer robust past an angle of 20. Similarly, in our experiments, we found
+that striking a balance between the size of the tag, the prices of the
+device, and the computational cost of detection was unfeasible with current
+cameras and algorithms.}
+
+\ben{Reviewers suggest considering computer vision-based techniques with markers in the environment (such as QR codes, or Bokode [Mohan] which extends the working distance to a few meters for visual tags).
+We considered this direction and found several drawbacks, including computational complexity, low accuracy when targets are far, the absence of real-time environment feedback with passive tags, and aesthetic issues with QR codes.
+While we don't claim our choice of IR is optimal, it has important advantages -- it is low-cost, readily available and quite suitable for area selection.
+We can weaken our claim and add explanations of how disambiguation techniques might generalize to other methods of implementing target identification.}
+
+\achal{Read and discussed the Bokode paper and CV in general. Unsure about what
+to do regarding "weakening our claim."}
+
 %\ben{new related works from reviewers}. \bjoern{folded in}

### 2014-06-26 doc/sui14/interaction.tex
updated the system architecture figure

@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/architecture}
-\caption{In our system architecture, selection is initiated through infrared but confirmed over 802.15.4.
-This permits wearers to move their head freely after connecting to an appliance.
-In the research prototype, users have to carry an additional microcontroller board that marshals messages between Glass' Bluetooth radio an IR/802.15.4, but our custom hardware could also be integrated into the wearable device.}
+\includegraphics[width=1\columnwidth]{figures/architecture_new}
+\caption{Our System architecture diagram. The selection is initiated through infrared but confirmed over 802.15.4.}
 \label{fig:architecture}

### 2014-06-26 doc/sui14/intro.tex
emphasize spatial

@@
 These objects are all potential targets for interaction.
-To initiate spatial interactions, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but is not yet well-explored in physical spaces.
+To initiate {\em spatial interactions}, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but is not yet well-explored in {\em physical spaces}.
 

### 2014-06-26 doc/sui14/relatedwork.tex
commenting on relatedwork“

@@
 \achal{Not sure where to put this, feel free to move.}
+\ben{For CV, I think related work is one option to go; however, the requirement of being related work is that -- we need to reference them.
+There are a few examples ``require large, obtrusive tags, and inevitably lack feedback from passive devices in the environment'' which should be referenced.
+That's why I am thinking of putting this to discussion.}
 \changes{Many alternative solutions for detecting devices in contained spaces

### 2014-06-26 doc/sui14/interaction.tex
rewording interaction section

@@
 %!TEX root = sui14.tex
-\section{IR for head orientation sensing}
+\section{IR for head orientation-based targeting}
+\changes{In this section, we present our hardware platform (IR targeting) and interaction model (head-orientation based) that we will use throughout our iterative designs.}
 
 \subsection{Hardware}
-
-\ben{a bit more on system description.}
-
 A central hypothesis of this paper is that the area cursor paradigm~\cite{kabbash1995prince} is well matched to head orientation selection.
@@
 Therefore, point selection techniques like laser pointers are not appropriate.
-Infrared emitters are a good technology match since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible.
-LEDs with many different fields of view are commercially available.
-Each individual target that is equipped with IR detector can react to the signal.
-Such point to point communication enables us to achieve targeting without needing a map.
 
-%Another benefit is the low cost and easy integration of both emitters and detectors - detectors often have support for common IR protocol decoding built in and are inexpensive, making it economically feasible to add detectors to dozens or hundreds of devices in an environment.
-\bjoern{added some more detail back in}
-\changes{Our augmentation to Google Glass adds a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545).
+Infrared (IR) emitters are a good technology match since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible.
+IR LEDs with many different fields of view are commercially available.
+\changes{We augment Google Glass with a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545).
 The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio, since Glass does not directly support hardware modification (see Figure~\ref{fig:glass}).
 Target devices use Vishay TSOP38238 IR receivers.
-Data is encoded using standard IR remote protocols at 38.0kHz.
+Data is encoded using standard IR remote protocols at 38.0kHz.}
 
-IR signals are used intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send IR signal status from targets to back to Glass and for device control itself.
-We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose.
- (Figure~\ref{fig:architecture}).
-This architecture was mostly chosen for reasons of expediency.
-We selected XBee 802.15.4 radios to avoid Bluetooth  wake up latencies but we do not claim optimality for our design decisions.
-Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change.
+\changes{IR signals are used for intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send IR signal status from targets to back to Glass and for device control itself.
+We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}).
+This architecture was mostly chosen for reasons of expediency and we do not claim optimality for our design decisions.
+Future head-mounted devices could clearly integrate IR emitters; the choice of wireless technology could also change.
 In particular, one could substitute WiFi modules or design an all-Bluetooth network and forego intermediate layers in our design.}

### 2014-06-27 doc/sui14/iteration3.tex
minor changes: grammar in iteration3, line separater in relatedwork
[iteration3]: remove merge conflict

--- 2014-04-17 doc/uist14/iteration3.tex
@@
 
-Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the near-eye display for the refinement step.
-The list interface has a few clear shortcomings: navigation actions map poorly to real world results, and the users must switch their focus back and forth between the physical scene and the near-eye display.
-These problems motivated us to design an approach that harnesses the head orientation during the {\em refinement} stage.
+Both {\em Naive IR} and {\em Intensity IR} rely on list navigation on the
+near-eye display for the refinement step. The list interface has a few clear
+shortcomings: navigation actions map poorly to real world results, and
+users must switch their focus back and forth between the physical scene and the
+near-eye display. These problems motivated us to design an approach that
+harnesses the head orientation during the {\em refinement} stage.
 
-In our third iteration, we ask the question {\em can we build a system purely using head orientation and visual feedback from the environment for target selection?}
+In our third iteration, we ask the question {\em can we build a system purely
+using head orientation and visual feedback from the environment for target
+selection?}
 
@@
 
-We introduce a third technique that uses a combination of motion sensors and IR to learn the relative orientations of
-targets in a room and intelligently suggest targets during refinement (see
-Figure~\ref{fig:third_technique}).
+We introduce a third technique that uses a combination of motion sensors and IR
+to learn the relative orientations of targets in a room and intelligently
+suggest targets during refinement (see Figure~\ref{fig:third_technique}).
 
-In the learning stage, the user scans over the targets, and the
-system attains the absolute orientation of each device from IR and motion sensors. From this information it can abstract out their relative positions and build an {\em adjacency map}.
+In the learning stage, the user scans over the targets, and the system attains
+the absolute orientation of each device from IR and motion sensors. From this
+information it can abstract out their relative positions and build an {\em
+adjacency map}.
 
-This absolute orientation cannot be applied to all indoor
-environments, since the user’s movements through the space could change the relationships between targets.
-However, with the constraint that the targets are spread around the periphery, their relative orientations are stable
-(see Figure~\ref{fig:third_principle}).
+This absolute orientation cannot be applied to all indoor environments, since
+the user’s movements through the space could change the relationships between
+targets. However, with the constraint that the targets are spread around the
+periphery, their relative orientations are stable (see
+Figure~\ref{fig:third_principle}).
 
@@
 \includegraphics[width=1\columnwidth]{figures/third_principle.png}
-\caption{This illustrates that when the change of user's absolution position doesn't change the relative relationship of physical targets.}
+\caption{This illustrates that the change of a user's absolution position
+doesn't change the relative relationship of physical targets.}
 \label{fig:third_principle}
@@
 \centering
-\includegraphics[width=1\columnwidth]{figures/third_technique3.pdf}
-\caption{Our third technique learned each target's absolute orientation and construct the adjacency map.
-During the {\em refinement} stage, the prediction is based on relative changes to a reference point.}
+\includegraphics[width=1\columnwidth]{figures/third_technique.pdf}
+\caption{Our third technique learned each target's absolute orientation and
+construct the adjacency map. During the {\em refinement} stage, the prediction
+is based on relative changes to a reference point.}
 \label{fig:third_technique}
@@
 After the map is created, the user can hold down on the touchpad to enter a
-quasi-mode for refinement.
-In this quasi-mode, one device lights up at a time.
-When the user turns his head in the direction of another device, the light switches to that device.
-Therefore, the user can move between devices one at a time with slight head movements.
-We implement this interaction by calculating the user's direction of motion using a low-passed history of sensor measurements and searching through the adjacency map for
-the nearest device in that direction.
+quasi-mode for refinement. In this quasi-mode, one device lights up at a time.
+When the user turns his head in the direction of another device, the light
+switches to that device. Therefore, the user can move between devices one at a
+time with slight head movements. We implement this interaction by calculating
+the user's direction of motion using a low-passed history of sensor
+measurements and searching through the adjacency map for the nearest device in
+that direction.
 
 \subsection{Evaluation}
-We evaluate the head motion refinement method by holding an informal study and collecting qualitative feedback from a subset of 4 users from the Iteration 2 study.
-In the study, we asked users to try cycling through the targets using the new quasi-mode.
+We evaluate the head motion refinement method by holding an informal study and
+collecting qualitative feedback from a subset of 4 users from the Iteration 2
+study. In the study, we asked users to try cycling through the targets using
+the new quasi-mode.
 
-The users had strong preferences for the new method of refinement.
-On a scale from 1-7, 1 being the least mental effort and 7 being the most mental effort users rated the old technique 4.25 and the new technique 2 on average.
-100\% indicated a preference for head movement to list navigation.
-One user referenced the issue of naming targets that the list necessitates, preferring the experience of \studyquote{matching visual cues rather than numbers}.
-Another participant remarked that it \studyquote{just made more sense} and was a \studyquote{more natural way for demonstrating intentionality}.
-The users preferred the new mapping in relation to the whole environment: \studyquote{it leveraged the spatial sense that I already had just by using the system}.
-They were also delighted to avoid list navigation, which they now called \studyquote{difficult} and \studyquote{painful}.
+The users had strong preferences for the new method of refinement. On a scale
+from 1-7, 1 being the least mental effort and 7 being the most mental effort
+users rated the old technique 4.25 and the new technique 2 on average. 100\%
+indicated a preference for head movement to list navigation. One user
+referenced the issue of naming targets that the list necessitates, preferring
+the experience of \studyquote{matching visual cues rather than numbers}.
+Another participant remarked that it \studyquote{just made more sense} and was
+a \studyquote{more natural way for demonstrating intentionality}. The users
+preferred the new mapping in relation to the whole environment: \studyquote{it
+leveraged the spatial sense that I already had just by using the system}. They
+were also delighted to avoid list navigation, which they now called
+\studyquote{difficult} and \studyquote{painful}.

### 2014-06-27 doc/sui14/applications.tex
first pass with rebuttals
add comment and change proposals

@@
 \label{sec:applications}
+
+\ben{We may need to add some discussion of applications}.
 
@@
 
+\sean{here are some parts I can think of to add to this section, feel free to refine/elaborate/remove.}
+
+\changes{During our study, we learned a few things.
+(1)Our soulution works best on simple applications such as on/off or volume up/down controls.
+For more complicated controls, we suggest to use a hybrid solution in which HOBS completes the target selection process and seamlessly transfer information to user's smartphone and triggers corresponding app.
+(2)Our solution excels at tasks where the user needs to focus primarily on the objects and the information provided by the near-eye display is only supportive, for example, museums or automobile body shops.
+(3)Our soulution also works well for roles whose hands are constantly occupied from using handheld devices, e.g.
+chefs.}
 %% I feel like we should mention a bit about the interaction complexity. But that's a bit out of the scope of this paper.

### 2014-06-27 doc/sui14/intro.tex
add comment and change proposals

@@
 While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and cautiously aimed; that the user's hands are required to be free for operation; and that the user's visual attention is split between looking down at a screen and out at targets in the world.
+
+\sean{Ben mentioned about discussing the limitations of list approach. Below is the paragraph from chi submission for your reference.}
+
+\changes{ Today, commercial solutions (such as Belkin WeMo) use handheld mobile devices as {\em universal remote controls} to control such appliances.
+In these solutions, users first browse a list of all available devices and then call up a device-specific user interface.
+This method faces two challenges: {\em naming} and {\em scoping}.
+Assigning clear names is non-trivial.
+In shared spaces, the person trying to control the device might not be the one that named it - e.g., while an office building manager may know what ``Light 4 in area E'' corresponds to, an occupant may not.
+Second, without a method of scoping selection to automatically filter non-relevant devices, paging though long lists of names or navigating hierarchies becomes potentially more cumbersome than the physical action the ``convenient'' software solution was meant to replace.}
 

### 2014-06-27 doc/sui14/abstract.tex
upper case acronym letters in abstract, minor grammatical fixes

@@
 %
-We present the iterative design and evaluation of \systemname -- a head-orientation based selection method for interacting with these devices at a distance.
+We present the iterative design and evaluation of \systemname -- a Head-Orientation Based Method for interacting with these devices at a distance.
 We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.

### 2014-06-27 doc/sui14/intro.tex
intro shortened

--- 2014-06-25 doc/sui14/intro.tex
@@
 \section{Introduction}
-\bjoern{We should see if we can change language in a few places to strengthen the connection to ``Spatial User Interaction", the title of the conference.}
+%\bjoern{We should see if we can change language in a few places to strengthen the connection to ``Spatial User Interaction", the title of the conference.}
 
@@
 These objects are all potential targets for interaction.
-To initiate spatial interactions, a user needs to first acquire the target object -- a fundamental task in an interaction process that has been extensively studied in the digital world, but is not yet well-explored in physical spaces.
+To initiate {\em spatial interactions}, a user needs to first acquire the target object -- a fundamental task that has been extensively studied in graphical user interfaces, but is not yet well-explored in {\em physical spaces}.
 
+%\sean{Ben mentioned about discussing the limitations of list approach. Below is the paragraph from chi submission for your reference.}
+%
+\changes{Today, commercial solutions for controlling physical devices (such as Belkin WeMo) use smartphones as {\em universal remote controls} -- users first browse a list of all available devices and then call up a device-specific user interface.
+This method faces {\em naming} issues (where is ``Light 4 in Area E''?) and {\em scaling} issues as the number of controlled devices increases.}
+%
 %% previous approaches are limited
-Past research has introduced techniques of augmenting hand-held mobile devices with accessories like laser pointers to enable direct aiming at target devices in space~\cite{beigl_point_1999,patel_2-way_2003}.
-While promising, some drawbacks of using hand-held devices are that the device first has to be retrieved (e.g., from a pocket) and cautiously aimed; that the user's hands are required to be free for operation; and that the user's visual attention is split between looking down at a screen and out at targets in the world.
+Past research has used direct aiming at target devices in space with phones to overcome these problems~\cite{beigl_point_1999,patel_2-way_2003}.
+In such techniques, the aiming device first has to be retrieved (e.g., from a pocket); user's hands have to be free for operation; and the user's visual attention is split between looking down at a screen and out at targets in the world.
 
 %% introducing head-worn computing and head orientation
-Emerging head-worn computing devices have the potential to overcome some of these limitations: they do not require retrieval since the devices are already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information in the wearer's field of view.
+Emerging head-worn computing devices do not require retrieval since the devices are already worn; they may enable hands-free or uni-manual interactions; and they offer near-eye or see-through displays to present information in the wearer's field of view.
 We thus investigate how such computing devices may be used for the selection and control of devices in physical spaces.
@@
 It suggests the general direction, but not the particular point of focus.
-While gaze tracking could provide further information about a user's focus, wearable gaze tracking cannot by itself reliably determine the identity of objects in focus (i.e.
-it also needs to know what the user is seeing).
-Equipping all devices with stationary gaze trackers is possible~\cite{vertegaal2005media} but much costlier than our approach.
-We therefore study how to leverage head-orientation alone for target selection in physical spaces.
-\bjoern{I think this needs reworking.
-In essence, reviewers didn't believe us that head orientation was a sensible way to go when compared to more complex gaze tracking or computer vision approaches.}
-
-The imprecision of human head movement suggests adapting area cursor techniques developed for assistive devices~\cite{kabbash1995prince,worden1997making,findlater2010enhanced}.
+We draw an analogy to assistive area cursors and adapt area cursor techniques~\cite{kabbash1995prince,worden1997making,findlater2010enhanced} for physical selection.
 Such techniques employ a two-step selection process: a {\em coarse} selection of an area of interest, followed by a {\em refinement} to select a target within that area.
@@
 We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} to enable infrared (IR) communication between Glass and target appliances.
-In our {\em Naive IR} technique, the cone of light emitted by an IR LED %(a diameter of 30-60cm and distance up to 8m)%
- provides the {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser} {\em left}), where targets that have received an IR signal light up.
-To {\em refine} selection when multiple targets have received IR signals, the {\em Naive IR} solution starts with an ordering based on their names.
-This is shown in the center of Figure~\ref{fig:teaser}, where targets in range are ordered numerically on the near-eye display.
-A study with $14$ participants compares acquisition times for physical targets in a room for our technique and an alternative list selection interface without any IR targeting \ben{the ordered-list is also list selection, needs clarification!}.
-We find that target acquisition through head orientation is preferred by users and is faster than list selection, but refinement is still time consuming.
+We contribute and evaluate new methods for addressing selection ambiguity in this context.
+In all our techniques, the emitted IR beam %(a diameter of 30-60cm and distance up to 8m)%
+ provides an initial {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser} {\em left}) --- all targets that receive this IR signal can be selected.
+To {\em refine} selection when multiple targets have received IR signals, we describe and evaluate three techniques:
 
-In response, we designed an improved refinement technique, {\em Intensity IR}, in which target objects compare received IR signal strength.
-This value allows the system to eliminate some distracting targets on the peripheral and to re-order the refinement interface's list by their intensity values.
+ Our {\em Naive IR} technique shows an alphabetically ordered disambiguation list on the near-eye display (Figure~\ref{fig:teaser} center).
+A study with $14$ participants finds that target acquisition with naive IR targeting is preferred by users and is faster than pure list selection without IR, but refinement is still time consuming.
+
+Our {\em Intensity IR} techniques improves refinement as target objects compare received IR signal strength (RSS).
+This value allows the system to eliminate some peripheral targets and to re-order the refinement interface's list by their intensity values.
 For example, in Figure~\ref{fig:teaser} of {\em Intensity IR} technique, device 5 is eliminated first and the list is re-ordered based on the intensity readings.
-A second study with $10$ participants proves that {\em Intensity IR} successfully reduces both the chance of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.
+A second study with $10$ participants shows that {\em Intensity IR} successfully reduces both the probability of needing to do refinement as well as the time spent in list navigation when compared to {\em Naive IR}.
 
-A final design addresses the lack of a natural mapping when users select a target in the refinement step using their device's touchpad - the axes of motion do not map directly to the spatial layout of target devices in a room.
-Our {\em Head-motion Refinement} technique first learns the relative spatial structure of the targets using Glass' orientation sensors.
-In a future target acquisition stage, users can perform head movements to change selections to spatially adjacent targets (see the right of Figure~\ref{fig:teaser}).
+Our final {\em Head-motion Refinement} addresses the lack of a natural mapping when users select a target in the refinement step using their device's touchpad --- the axes of motion do not map directly to the spatial layout of target devices in a room.
+We first learn the relative spatial structure of the targets using Glass' orientation sensors.
+Users can then perform head movements to change selections to spatially adjacent targets (see the right of Figure~\ref{fig:teaser}).
 For example, nodding down to select the target below current selection, or tilting right to select the next target on the right.
@@
 %---
-We also demonstrate an example application of our technique used as a remote control of smart appliances: a user looks at the appliance he wishes to control and confirms selection by tapping.
+We also demonstrate an example application of our technique used as a remote control of smart appliances such as lighting and TV sets: a user looks at the appliance he wishes to control and confirms selection by tapping.
 An appliance-specific user interface is then shown on the user's near-eye display for further interactions.
@@
 %%% mode: latex
-%%% TeX-master: "uist14"
+%%% TeX-master: "sui14"
 %%% End:

### 2014-06-27 doc/sui14/relatedwork.tex
minor changes: grammar in iteration3, line separater in relatedwork
Strip trailing spaces
intro shortened
relatedwork: more CV citations + minor updates

@@
 Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}.
-However, human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
-
+However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
+%
 %Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}.
-
+%
 Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.
@@
 \subsection{Pointing in Physical Spaces}
-Rukzio studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list~\cite{rukzio_experimental_2006}.
+Rukzio ~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list.
 
@@
 Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
-Our solution is similar in spirit but relies on simple low-cost IR emitters and detectors.
-\bjoern{need a stronger statement here}
+Our solution is similar in spirit but requires only small, low-cost IR emitters and detectors.
+\bjoern{need a stronger statement here} \achal{Tried to add a small contrast of small/low cost, but don't want to explicitly say projectors are large and unwieldy as there may be some exceptions.}
 %However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.
@@
 \achal{Not sure where to put this, feel free to move.}
+
 \ben{For CV, I think related work is one option to go; however, the requirement of being related work is that -- we need to reference them.
@@
 That's why I am thinking of putting this to discussion.}
+
 \changes{Many alternative solutions for detecting devices in contained spaces
 rely on passive tags on devices, combined with computer vision methods for
-detection. Unfortunately, these methods either impose significant constraints
-on the camera used for detection, or require large, obtrusive tags, and
-inevitably lack feedback from passive devices in the environment. \cite{Bokode},
+detection. Unfortunately, these methods impose significant constraints
+on the camera used for detection~\cite{Bokode}, require large, obtrusive tags, or are
+designed to work specifically at short distances~\cite{CyberCode}; inevitably,
+they also lack feedback from passive devices in the environment. \cite{Bokode},
 for example, relies on visually imperceptible codes, but has significant
-drawbacks: despite using a 12.1 megapixel DSLR camera, detections are no
-longer robust past an angle of 20. Similarly, in our experiments, we found
-that striking a balance between the size of the tag, the prices of the
-device, and the computational cost of detection was unfeasible with current
-cameras and algorithms.}
+drawbacks: despite using a 12.1 megapixel DSLR camera, detections are no longer
+robust past an angle of 20. Similarly, in our experiments, we found that
+striking a balance between the size of the tag, the prices of the device, and
+the computational cost of detection was unfeasible with current cameras and
+algorithms.}
 

### 2014-06-27 doc/sui14/applications.tex
applications.tex: minor grammar / style

@@
 We implement one particular demonstrative application: a universal smart appliance controller.
-Users select a smart device (e.g., light fixture, TV, or home appliance) with Glass --- upon confirming the selection, an appliance-specific UI is shown on the user's near-eye display, and they can control the application (without having to continually look at it) through their device touchpad.
+Users select a smart device (e.g., light fixture, TV, or home appliance) with Glass --- upon confirming the selection, an device-specific UI is shown on the user's near-eye display, and they can control the application (without having to continually look at it) through their device touchpad.
 
 We implemented a prototype of this remote controller for three physical devices: a lamp and fan that could be switched on or off, and a smart TV control with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
-We switched discrete appliances using relays, and implemented the smart TV with a laptop connected to a 30\inch{} display.
+We switched discrete appliances using relays, and implemented controls for a smart TV with a laptop connected to a 30\inch{} display.
 User interfaces for each were pre-defined in our application.

### 2014-06-27 doc/sui14/interaction.tex
interaction.tex: minor grammar, style, spelling

@@
 
-\changes{IR signals are used for intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send IR signal status from targets to back to Glass and for device control itself.
+\changes{IR signals are used for initial line-of-sight targeting; subsequently, we use bi-directional wireless communication to send IR signal status from targets to back to Glass and for device control itself.
 We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}).
-This architecture was mostly chosen for reasons of expediency and we do not claim optimality for our design decisions.
+This architecture was mostly chosen for reasons of expediency, and could be further optimized.
 Future head-mounted devices could clearly integrate IR emitters; the choice of wireless technology could also change.
-In particular, one could substitute WiFi modules or design an all-Bluetooth network and forego intermediate layers in our design.}
+In particular, one could substitute WiFi modules or design an all-Bluetooth network and forego the intermediate layers in our design.}
 
@@
 During this stage, Glass constantly sends out IR signals, and  targets offer immediate feedback when they receive a signal.
-The user confirms his desire to connect to a target by tapping on the Glass touchpad.
+The user confirms their desire to connect to a target by tapping on the Glass touchpad.
 Glass collects the responses from targets that have received IR reception through the backchannel of XBee.
-If there is only one single target in IR range, it is automatically selected.
-However, in a dense environment where multiple targets are within range, the user needs refine his selection.
+If there is only one target in IR range, it is automatically selected.
+However, in a dense environment where multiple targets are within range, the user needs to refine their selection.
 
-{\bf Refine:} When disambiguation is needed, the user must make an explicit selection among the targets within his view range.
+{\bf Refine:} When disambiguation is needed, the user must make an explicit selection among the targets within their view range.
 We have designed multiple refinement mechanisms -- all of which enable the user to select one from a subset of targets.

### 2014-06-27 doc/sui14/iteration1.tex
created sui2014 dir
fixed figure width problem
iteration1.tex: added comment regarding user quotations

--- 2014-04-16 doc/uist14/iteration1.tex
@@
 \caption{Boxplot of target acquisition time for {\em IR selection} and {\em list selection} is shown on the left.
-The center of each box is the median value, while we annotate the mean value using white dashed lines.
-The cumulative distributions are shown on the right.}
+The center is the median value, and the mean value is shown using white dashed lines.
+The cumulative distributions are on the right.}
 \label{fig:ir_vs_list}
@@
 \centering
-\includegraphics[width=1.09\columnwidth]{figures/result_study1b.pdf}
+\includegraphics[width=1.0\columnwidth]{figures/result_study1b.pdf}
 \caption{Boxplot and CDF of target acquisition time when refinement is needed or not in {\em IR selection}}.
@@
 One user noted benefit of {\em IR selection} being \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
-One subject called it \studyquote{natural to interact with things just by looking at them}.
+\achal{This seems rather weak...
+Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}.
 Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.

### 2014-06-27 doc/uist14/submission/rebuttal.txt
edits

We thank the reviewers for their detailed feedback. Below, we answer the
raised questions and clarify some points of confusion.

CONTRIBUTION  LEVEL (AC, R1, R2, R3)
Reviewers find this work’s contribution limited and suggest focusing on
interacting with complex targets. In the past, we built interfaces to
control complex appliances including smart TVs; however, CHI reviewers on
a prior application-centric paper suggested focusing on fundamental
selection techniques instead, resulting in the abbreviated description of
applications in our paper. Some examples can be found in the APPLICATION
section; we are happy to reintroduce more material in our revision.

We focus on the fundamental question of target selection itself, which is
likely to be a problem in a future rich with smart devices. Existing
commercial solutions (primarily list selection) don’t scale well (we
envision a swarm of interactive devices as the contexts) and exhibit
naming problems when multiple identical objects are within the same
space. Our work contributes a sophisticated understanding of methods for
using head-orientation as an indication of user’s point of interest in
physical space.

LOW DISAMBIGUATION SUCCESS RATE (R1)
R1 comments that our technique barely outperforms chance. However, our
success rate of 55% in iteration 2 is much higher than random chance
given that there are at least 4 targets in the user’s field of view in
our setup (in this case, chance would be 25%).

COMPUTER VISION APPROACH (AC, R2, R3)
Reviewers suggest considering computer vision-based techniques with
markers in the environment (such as QR codes, or Bokode [Mohan] which
extends the working distance to a few meters for visual tags). We
considered this direction and found several drawbacks, including
computational complexity, low accuracy when targets are far, the absence
of real-time environment feedback with passive tags, and aesthetic issues
with QR codes. While we don’t claim our choice of IR is optimal, it has
important advantages -- it is low-cost, readily available and quite
suitable for area selection. We can weaken our claim and add explanations
of how disambiguation techniques might generalize to other methods of
implementing target identification.

SYSTEM DESCRIPTION (AC)
The AC commented that “the description of the system is very
lacking”. We provided brief implementation information in order to
focus on the design and evaluation within a shorter paper to match the
magnitude of contribution. This may have been counterproductive. We list
part numbers in the "Hardware" subsection but can describe how IDs are
exchanged over IR and 802.15.4 in more detail.

MISSING RELATED WORK (AC, R3)
We will add the additional references.
RFIG Lamps and photosensing wireless tags [Raskar et al] use a projector
and data is encoded in each pixel so that tags can localize themselves
for target selection. Our IR emitter, a relatively low-cost solution,
doesn't have this feature but shares the commonality of covering an area
for the ease of selection. At the end of their paper, they envision an
IR-based system to solve ambient light problems. Our work lies in that
direction and also contributes evaluations with user studies.
Progressive Refinement [Bacima] studies several progressive refinement
selection modalities, which is close to our two-stage selection; but our
contexts are head-orientation as it reflects users’ point of interest.
Tinmith-metro [Piekarski et al] tracks the users’ hand movements and
uses a HUD for augmented reality, while our work focuses on direct
selection on physical targets.

MISC (AC)
The AC asked why users couldn't swipe in a certain direction on Glass for
disambiguation. Two major factors preclude this interaction: 1) the Glass
physical form factor makes upward swipes awkward to perform; 2) the plane
of the touchpad is orthogonal to the plane of the targets, leading to
ambiguity what lateral swipes mean.

### 2014-06-27 doc/sui14/applications.tex
edits

--- 2014-06-27 doc/sui14/applications.tex
@@
 \sean{here are some parts I can think of to add to this section, feel free to refine/elaborate/remove.} 
-
+\bjoern{1 is problematic - because it sounds like that the head worn device is in the end not worth it and you should just use your phone, even if it takes longer.
+I'm not sure we can confidently say something about museums or body shops - this sounds speculative.
+Needs section needs to be rephrased., e.g.
+as in: Our solution succeeds at X because of Y.
+This suggests that future applications in domains like Z could be especially promising.}
 \changes{During our study, we learned a few things.

### 2014-06-27 doc/sui14/interaction.tex
edits

--- 2014-06-26 doc/sui14/interaction.tex
@@
 \subsection{Hardware}
-A central hypothesis of this paper is that the area cursor paradigm~\cite{kabbash1995prince} is well matched to head orientation selection.
+%A central hypothesis of this paper is that the area cursor paradigm~\cite{kabbash1995prince} is well matched to head orientation selection.
 Head orientation input is imprecise because it does not capture eye movement and has to rely on a low-bandwidth muscle group.
@@
 
-Infrared (IR) emitters are a good technology match since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible.
-IR LEDs with many different fields of view are commercially available.
+We hypothesize that infrared (IR) emitters are a good technology match for head orientation selection, since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible.
+IR LEDs with many different beam angles are commercially available.
 \changes{We augment Google Glass with a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545).
@@
 
-\changes{IR signals are used for intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send IR signal status from targets to back to Glass and for device control itself.
+\changes{IR signals are used for intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass.
 We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}).
@@
 {\bf Scan:} The user first scans the environment to locate the position of the target.
-During this stage, Glass constantly sends out IR signals, and  targets offer immediate feedback when they receive a signal.
+During this stage, Glass constantly sends out IR signals, and  targets offer immediate visual feedback when they receive a signal.
 The user confirms his desire to connect to a target by tapping on the Glass touchpad.

### 2014-06-27 doc/sui14/iteration1.tex
edits

@@
 Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
-while {\em list selection} took 8.86 {\em list} seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
+while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
 
@@
 One user noted benefit of {\em IR selection} being \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
-\achal{This seems rather weak...
-Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}.
+One subject called it \studyquote{natural to interact with things just by looking at them}.
 Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.

### 2014-06-27 doc/sui14/relatedwork.tex
edits

--- 2014-06-26 doc/sui14/relatedwork.tex
@@
 \subsection{Head and Gaze Input}
-Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}.
-However, human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
-
+\changes{Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}.
+However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.}
+%
 %Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}.
-
+%
 Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.
@@
 Gaze can also be used to control graphical user interfaces~\cite{kumar2007eyepoint}.
-While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map of each space with known target locations.
-Our system works through point-to-point IR communication and does not require an {\em a priori} map.
+While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map with known target locations or the use of visually obtrusive markers such as QR codes.
+Our system works through point-to-point IR communication and does not require an {\em a priori} map or markers.
 Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}.
@@
 Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~
-\cite{bacim2013design}. In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed.
+\cite{bacim2013design}.
+In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed.
+This paper describes the trade-offs between several disambiguation approaches.
 
-We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets).
+%We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets).
 In the {\em refinement} phase, they adjust so only the intended target will be selected.
@@
 
-Several approaches to spatial selection with handheld devices exist for controlling appliances~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}.
+Several approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist.
+%and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. 
 In some techniques, users select objects of interest with laser pointers.
 The laser dot provides immediate visual feedback to the user about what is being selected; however, its small target area makes it poorly matched to head orientation input.
+Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.
+%Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices.
+Our system tackles an unresolved issue of prior approaches -- navigating an area dense with potential targets and resolving selection ambiguity.
 
-Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.
-
+\subsection{Vision- and Projection-Based Target Selection}
+\changes{Many alternative solutions for detecting devices in contained spaces
+rely on computer-vision recognition of printed tags on devices. Unfortunately, these methods either impose significant constraints
+on the camera used for detection, or require large, obtrusive tags, and
+inevitably lack feedback from passive markers in the environment. \bjoern{need a better reference than bokode that uses QR codes for initiating interaction.}
 Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
 Our solution is similar in spirit but relies on simple low-cost IR emitters and detectors.
-\bjoern{need a stronger statement here}
+\bjoern{need a stronger statement here}}
 %However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.
 
-Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices.
-Our system tackles an unresolved issue of such IR-based approaches -- navigating an area dense with potential targets.
+\bjoern{I tried to fold in some of the points about QR codes into the text above.}
+%\subsection{Markers and Vision Methods}
+%\achal{Not sure where to put this, feel free to move.}
+%
+%\ben{For CV, I think related work is one option to go; however, the requirement of being related work is that -- we need to reference them.
+There are a few examples ``require large, obtrusive tags, and inevitably lack feedback from passive devices in the environment'' which should be referenced.
+That's why I am thinking of putting this to discussion.}
+%
 
-\subsection{Markers and Vision Methods}
-\achal{Not sure where to put this, feel free to move.}
-\changes{Many alternative solutions for detecting devices in contained spaces
-rely on passive tags on devices, combined with computer vision methods for
-detection. Unfortunately, these methods either impose significant constraints
-on the camera used for detection, or require large, obtrusive tags, and
-inevitably lack feedback from passive devices in the environment. \cite{Bokode},
-for example, relies on visually imperceptible codes, but has significant
-drawbacks: despite using a 12.1 megapixel DSLR camera, detections are no
-longer robust past an angle of 20. Similarly, in our experiments, we found
-that striking a balance between the size of the tag, the prices of the
-device, and the computational cost of detection was unfeasible with current
-cameras and algorithms.}
-
-\ben{Reviewers suggest considering computer vision-based techniques with markers in the environment (such as QR codes, or Bokode [Mohan] which extends the working distance to a few meters for visual tags).
+%\ben{Reviewers suggest considering computer vision-based techniques with markers in the environment (such as %QR codes, or Bokode [Mohan] which extends the working distance to a few meters for visual tags).
 We considered this direction and found several drawbacks, including computational complexity, low accuracy when targets are far, the absence of real-time environment feedback with passive tags, and aesthetic issues with QR codes.
@@
 
-\achal{Read and discussed the Bokode paper and CV in general. Unsure about what
-to do regarding "weakening our claim."}
+%\achal{Read and discussed the Bokode paper and CV in general. Unsure about what
+%to do regarding "weakening our claim."}
 

### 2014-06-27 doc/sui14/iteration2.tex
rewording interaction section

@@
 \item When multiple targets have received IR signals and reported the intensity readings, we discard those whose intensities are significantly lower than the largest reading\footnote{In our current implementation, we empirically set it to be half of the ADC resolution, which is frequently used as it indicates a 3dB loss in the signal strength.}.
-Therefore, when there is only one target within line of sight, the IR intensity approach has the same behavior as the previous iteration - no disambiguation is needed.
+Therefore, when there is only one target within the line of sight, the IR intensity approach has the same behavior as the previous iteration - no disambiguation is needed.
 When the environment turns denser, the new design can filter some peripheral targets out, reducing $P(refine)$, the likelihood of entering the refinement stage.
@@
 {\em Intensity IR} reduces the fraction of refinement trails in which additional list navigation is necessary (i.e., the first, already selected element is incorrect).
-{\em Intensity IR} sorted the desired target as the first one in the list in 55\% of cases (93 of 167) \ben{explicitly saying this is higher than random guess?}.
+{\em Intensity IR} sorted the desired target as the first one in the list in 55\% of cases (93 of 167).
 In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 225).

### 2014-06-27 doc/sui14/relatedwork.tex

@@
 rely on computer-vision recognition of printed tags on devices. Unfortunately, these methods either impose significant constraints
-on the camera used for detection, or require large, obtrusive tags, and
-inevitably lack feedback from passive markers in the environment. \bjoern{need a better reference than bokode that uses QR codes for initiating interaction.}
+on the camera used for detection~\cite{Bokode}, require large, obtrusive tags, or are
+designed to work specifically at short distances~\cite{CyberCode}. Passive markers also cannot show visual feedback in the environment. 
 Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
@@
 \bjoern{need a stronger statement here}}
+
 %However, visible tags at appropriate sizes may be rejected by users because of their negative aesthetic effect on the space.
 
-\bjoern{I tried to fold in some of the points about QR codes into the text above.}
 %\subsection{Markers and Vision Methods}

### 2014-06-27 doc/sui14/abstract.tex
intro shortened
HOBS in the right texts

@@
 %
-We present the iterative design and evaluation of \systemname -- a Head-Orientation Based Method for interacting with these devices at a distance.
+We present the iterative design and evaluation of \systemname -- a Head-Orientation Based Selection technique for interacting with these devices at a distance.
 We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.

### 2014-06-27 doc/sui14/sui14.tex
first pass with rebuttals
emphasize spatial
intro shortened
update caption for teaser

@@
 {\em Center}: We offer two list-based refinement techniques -- {\em Naive IR} uses alphabetical ordering; {\em Intensity IR} orders targets by IR intensity.
-{\em Right}: Alternatively, users can refine their selection through head orientation refinement in a quasi-mode.}
+{\em Right}: Using {\em Head-motion Refinement} technique, users can refine their selection through head orientation refinement in a quasi-mode when they hold the touchpad.}
   \label{fig:teaser}
@@
 \keywords{
-  Wearable computing; selection; glass; infrared
+  Wearable computing; spatial interaction; selection; Glass%; infrared
 }
@@
 %% refer to http://www.acm.org/about/class/ccs98-html
-\category{H.5.2.}{Information Interfaces and Presentation (e.g. HCI)}{Interaction styles (e.g., commands, menus, forms, direct manipulation).}
+\category{H.5.2.}{Information Interfaces and Presentation (e.g. HCI)}{Interaction styles}% (e.g., commands, menus, forms, direct manipulation).}
 

### 2014-06-27 doc/sui14/intro.tex
vfill for intro
fixed typos

@@
 %!TEX root = sui14.tex
+\vfill
 \section{Introduction}
@@
 These objects are all potential targets for interaction.
-To initiate {\em spatial interactions}, a user needs to first acquire the target object -- a fundamental task that has been extensively studied in graphical user interfaces, but is not yet well-explored in {\em physical spaces}.
+To initiate {\em spatial interactions}, a user needs to first acquire the target object -- a fundamental task that has been extensively studied in graphical user interfaces, but not yet well-explored in {\em physical spaces}.
 
@@
 Past research has used direct aiming at target devices in space with phones to overcome these problems~\cite{beigl_point_1999,patel_2-way_2003}.
-In such techniques, the aiming device first has to be retrieved (e.g., from a pocket); user's hands have to be free for operation; and the user's visual attention is split between looking down at a screen and out at targets in the world.
+Such techniques have a few drawbacks: the aiming device first has to be retrieved (e.g., from a pocket); user's hands have to be free for operation; and the user's visual attention is split between looking down at a screen and out at targets in the world.
 
@@
 
-Our {\em Intensity IR} techniques improves refinement as target objects compare received IR signal strength (RSS).
+Our {\em Intensity IR} technique improves refinement as target objects compare IR received signal strength (RSS).
 This value allows the system to eliminate some peripheral targets and to re-order the refinement interface's list by their intensity values.

### 2014-06-27 doc/sui14/relatedwork.tex
vfill and rewording in related works

@@
 
+\vfill
+\ben{make sure no section title appears at the end of any column.}
 \subsection{Head and Gaze Input}
@@
 Gaze can also be used to control graphical user interfaces~\cite{kumar2007eyepoint}.
-While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map with known target locations or the use of visually obtrusive markers such as QR codes.
+While there are wearable gaze trackers~\cite{bulling2009wearable}, turning information about a concrete point in space where a user is looking into a selection requires a map with known target locations.
 Our system works through point-to-point IR communication and does not require an {\em a priori} map or markers.

### 2014-06-27 doc/sui14/applications.tex
refining applications (with limitation for complex devices) and commented 'comments' in disucssion

--- 2014-06-27 doc/sui14/applications.tex
@@
 \label{sec:applications}
-
-\ben{We may need to add some discussion of applications}.
 
@@
 
-\sean{here are some parts I can think of to add to this section, feel free to refine/elaborate/remove.}
+% \bjoern{1 is problematic - because it sounds like that the head worn device is in the end not worth it and you should just use your phone, even if it takes longer.
+I'm not sure we can confidently say something about museums or body shops - this sounds speculative.
+Needs section needs to be rephrased., e.g.
+as in: Our solution succeeds at X because of Y.
+This suggests that future applications in domains like Z could be especially promising.}
 
-\changes{During our study, we learned a few things.
-(1)Our soulution works best on simple applications such as on/off or volume up/down controls.
-For more complicated controls, we suggest to use a hybrid solution in which HOBS completes the target selection process and seamlessly transfer information to user's smartphone and triggers corresponding app.
-(2)Our solution excels at tasks where the user needs to focus primarily on the objects and the information provided by the near-eye display is only supportive, for example, museums or automobile body shops.
-(3)Our soulution also works well for roles whose hands are constantly occupied from using handheld devices, e.g.
-chefs.}
+\changes{Though our solution successfully solves the problem of {\em targets selection}, we haven't addressed the controlling issue when the device becomes complex.
+Users reported that \studyquote{for more complex appliances (TV), interaction was more difficult}.
+This problem can be partially ascribed to the limitations of Glass -- the touchpad can only detect a few gestures and the screen is too small.
+A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on the target selection in this paper.}
+
 %% I feel like we should mention a bit about the interaction complexity. But that's a bit out of the scope of this paper.
@@
 
-%\subsection{Attention Tracking}
-%\label{sec:attention-tracking}
-
-%The second tier of applications enabled by our system is an implicit use of users' head orientation.
-One sample application is to obtain a personal statistics.
-Nowadays, we are surrounded by all sorts of screens, and many people's job requires them to stare at the monitors for the whole day.
-% citation needed, take a look at this one: E-Readers and Visual Fatigue, http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3873942/
-%Knowing how much time you have spent in front of the screen is definitely useful, and the application can be built to remind the user to take a break whenever a certain time threshold is reached.
-
-%% We might be lacking evidence, but let me just write down the draft first.
-%Our system enables a flexible and easy-to-use tracking of users' attention without knowing the map.
-Therefore, it's suitable to be deployed in large museums (even when the exihibits positions are adjusted constantly).
-Then such attention tracking will facillitate many social science studies.
-Previous solutions are either using RFID \cite{Hsi:2005:REV:1081992.1082021} or video taping % this is a cv of Adam who has worked on video analysis http://www.exploratorium.edu/vre/cvs/ak_cv.pdf.
-%However, head orientation is a better approximation of attention rather than proximity.
-By placing our tracking devices in the environment, we can easily get a more accurate profiling of users intention.
-
-%\subsection{Indoor Positioning}
-%\label{sec:indoor-positioning}
-
-%There is one category of indoor positioning which is based on angle of arrival.
-Traditional approaches on indoor positioning are based on fingerprinting or triangulation, which is known to be bad due to multi-path effect for indoor environment.
-With our system, using the adjacent mapping calibration we can figure out the position of the targets in a room, and then based on a new sample, we can estimate a user's position with some simple math.
-
 %%% Local Variables: 

### 2014-06-27 doc/sui14/relatedwork.tex
relatedwork: add one more citation

@@
 \changes{Many alternative solutions for detecting devices in contained spaces
-rely on computer-vision recognition of printed tags on devices. Unfortunately, these methods either impose significant constraints
-on the camera used for detection~\cite{Bokode}, require large, obtrusive tags, or are
-designed to work specifically at short distances~\cite{CyberCode}. Passive markers also cannot show visual feedback in the environment. 
-Handheld projectors can both display a user interface in space and communicate control information optically, e.g., by encoding information temporally (using Gray codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared spectrum in SideBySide~\cite{willis_sidebyside:_2011}).
-Our solution is similar in spirit but relies on simple low-cost IR emitters and detectors.
-\bjoern{need a stronger statement here}}
+rely on computer-vision recognition of printed tags on devices. Unfortunately,
+these methods either impose significant constraints on the camera used for
+detection~\cite{Bokode}, require large or obtrusive tags~\cite{Dataglyphs}, or
+are designed to work specifically at short distances~\cite{CyberCode}. Passive
+markers also cannot show visual feedback in the environment.  Handheld
+projectors can both display a user interface in space and communicate control
+information optically, e.g., by encoding information temporally (using Gray
+codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG
+Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared
+spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Our solution is similar
+in spirit but relies on simple low-cost IR emitters and detectors. \bjoern{need
+a stronger statement here}}
 

### 2014-06-27 doc/sui14/relatedwork.tex
relatedwork: add back wording removed from merge conflict

@@
 spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Our solution is similar
-in spirit but relies on simple low-cost IR emitters and detectors. \bjoern{need
-a stronger statement here}}
+in spirit but requires only small, low-cost IR emitters and detectors.
+\bjoern{need a stronger statement here}} \achal{Tried emphasizing the smaller
+requirements, not sure what else we can say.}
 

### 2014-06-27 doc/sui14/intro.tex
editing for grammar

@@
 %
-\changes{Today, commercial solutions for controlling physical devices (such as Belkin WeMo) use smartphones as {\em universal remote controls} -- users first browse a list of all available devices and then call up a device-specific user interface.
-This method faces {\em naming} issues (where is ``Light 4 in Area E''?) and {\em scaling} issues as the number of controlled devices increases.}
+\changes{Today, companies like Samsung and Whirlpool are making smart appliances with companion applications that use smartphones as {\em universal remote controls}.
+With these applications, the user can select a device from a list in order to control it with a device-specific user interface.
+This method faces {\em naming} issues (i.e.
+"what do we name the lamp on the left?") and {\em scaling} issues as the number of controlled devices increases.
+These solutions also present a necessarily flawed mapping from the positions of the appliances in the rich, 3-dimensional world to their place in a 1D or 2D list presented on the screen.
+}
 %
@@
 Past research has used direct aiming at target devices in space with phones to overcome these problems~\cite{beigl_point_1999,patel_2-way_2003}.
-Such techniques have a few drawbacks: the aiming device first has to be retrieved (e.g., from a pocket); user's hands have to be free for operation; and the user's visual attention is split between looking down at a screen and out at targets in the world.
+Such techniques have a few drawbacks: the aiming device first has to be retrieved; the user's hands have to be free for operation; and the user's visual attention is split between looking down at a screen and out at targets in the world.
 

### 2014-06-27 doc/sui14/iteration1.tex
read for grammar

@@
 13 had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
-Four participants wore prescription glasses, which makes Glass more cumbersome to use and adjust, and may have affected their task performance.
+Four participants wore prescription glasses, which makes Glass more cumbersome to use and adjust and may have affected their task performance.
 In this study, the IR LED was fixed and not repositionable as shown in Figure~\ref{fig:glass}.
@@
 As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, list selection was perceived to be cumbersome.
-One user noted benefit of {\em IR selection} being \studyquote{more direct [than list mode]}, allowing users to focus on the targeted objects instead of the screen.
-One subject called it \studyquote{natural to interact with things just by looking at them}.
+While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and pleasurable.
+One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen.
+\achal{This seems rather weak...
+Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}.
 Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.
-One subject pointed out that \studyquote{{\em IR selection} doesn't requires you to swipe a lot of times like list mode}.
 

### 2014-06-27 doc/sui14/iteration1.tex
marked changes

@@
 As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and pleasurable.
+While both interfaces were judged similarly on overall ease of connecting, \changes{IR selection was perceived to be more direct and pleasurable.
 One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen.
@@
 Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}.
-Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.
+Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.}
 

### 2014-06-27 doc/sui14/applications.tex
worked on grammar and added discussion of complex interactions

@@
 
-We implemented a prototype of this remote controller for three physical devices: a lamp and fan that could be switched on or off, and a smart TV control with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
+We implemented a prototype of this remote controller for three physical devices: a lamp and fan that could be switched on or off and a smart TV control with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
 We switched discrete appliances using relays, and implemented controls for a smart TV with a laptop connected to a 30\inch{} display.
@@
 
-\changes{Though our solution successfully solves the problem of {\em targets selection}, we haven't addressed the controlling issue when the device becomes complex.
-Users reported that \studyquote{for more complex appliances (TV), interaction was more difficult}.
-This problem can be partially ascribed to the limitations of Glass -- the touchpad can only detect a few gestures and the screen is too small.
-A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on the target selection in this paper.}
+\changes{Though our solution successfully solves the problem of {\em target selection}, we have not fully addressed the issue of controlling smart devices.
+One user reported that \studyquote{for more complex appliances (TV), interaction was more difficult}.
+A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on target selection.
+When the device is more complicated, it is possible that the best solution may be a hybrid-- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv?
+I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.}
 

### 2014-06-27 doc/sui14/iteration2.tex
worked on grammar and added discussion of complex interactions

@@
 Therefore, when there is only one target within the line of sight, the IR intensity approach has the same behavior as the previous iteration - no disambiguation is needed.
-When the environment turns denser, the new design can filter some peripheral targets out, reducing $P(refine)$, the likelihood of entering the refinement stage.
+When the environment becomes more populated, the new design can filter some peripheral targets out, reducing $P(refine)$, the likelihood of entering the refinement stage.
 \item When refinement is still needed, meaning that multiple targets have relatively close intensity values, the system sorts the disambiguation list according to the IR intensity, from strongest to weakest.

### 2014-06-27 doc/sui14/applications.tex
add feedback

@@
 When the device is more complicated, it is possible that the best solution may be a hybrid-- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv?
-I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.}
+I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.} \sean{ I wouldn't state that.
+We could be challenged if the reviewers bring up smart fridge, microwave, etc.}
 

### 2014-06-27 doc/sui14/iteration1.tex
add missing '}' that caused error

@@
 Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}.
-Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.}
+Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.}}
 

### 2014-06-27 doc/sui14/relatedwork.tex
editing for grammar
fix for pg 2 whitespace

@@
 %!TEX root = sui14.tex
+\vfill
+%\ben{make sure no section title appears at the end of any column.}
 \section{Background and Related Work}
@@
 
-\vfill
-\ben{make sure no section title appears at the end of any column.}
 \subsection{Head and Gaze Input}
@@
 \cite{bacim2013design}.
-In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed.
+All area and space cursors necessitate disambiguation when there are multiple potential targets and no clear winner.
 This paper describes the trade-offs between several disambiguation approaches.

### 2014-06-27 doc/sui14/applications.tex
rephrase hybrid approach

@@
 A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on target selection.
-When the device is more complicated, it is possible that the best solution may be a hybrid-- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv?
+When the device is more complicated, it is possible that a better solution may be a hybrid -- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv?
 I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.} \sean{ I wouldn't state that.

### 2014-06-28 doc/sui14/iteration1.tex
fix } mismatch
use 0.95 for all figures
fixed } again

--- 2014-06-27 doc/sui14/iteration1.tex
@@
 
+\vfill
 \subsection{Technique}
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/study-layout1.png}
+\includegraphics[width=0.9\columnwidth]{figures/study-layout1.png}
 \caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment.}
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/result_study1a.pdf}
+\includegraphics[width=0.95\columnwidth]{figures/result_study1a.pdf}
 \caption{Boxplot of target acquisition time for {\em IR selection} and {\em list selection} is shown on the left.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/result_study1b.pdf}
+\includegraphics[width=0.95\columnwidth]{figures/result_study1b.pdf}
 \caption{Boxplot and CDF of target acquisition time when refinement is needed or not in {\em IR selection}}.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/result_study1c.pdf}
+\includegraphics[width=0.95\columnwidth]{figures/result_study1c.pdf}
 \caption{Times taken to select a device vs.~its order in the {\em list selection}.
@@
 While both interfaces were judged similarly on overall ease of connecting, \changes{IR selection was perceived to be more direct and pleasurable.
-One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen.
+One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen}.
 \achal{This seems rather weak...
 Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}.
-Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}, and many participants compared the overall time of target acquisition.}
+Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}}.
 

### 2014-06-28 doc/sui14/applications.tex
use 0.95 for all figures
shorten the paper to fit 8 pages

--- 2014-06-27 doc/sui14/applications.tex
@@
 
-We implemented a prototype of this remote controller for three physical devices: a lamp and fan that could be switched on or off and a smart TV control with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
-We switched discrete appliances using relays, and implemented controls for a smart TV with a laptop connected to a 30\inch{} display.
+Our prototype includes three smart devices: a lamp and a fan that could be switched on and off; a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
+The on/off appliances are using relays, and the smart TV is a 30\inch{} display connected to a laptop.
 User interfaces for each were pre-defined in our application.
@@
 \centering
-\includegraphics[width=1\columnwidth]{figures/smarthome-scenario.pdf}
+\includegraphics[width=0.95\columnwidth]{figures/smarthome-scenario.pdf}
 \caption{In the smart home scenario, we have built three smart appliances for a user experience study.
@@
 One user reported that \studyquote{for more complex appliances (TV), interaction was more difficult}.
-A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on target selection.
-When the device is more complicated, it is possible that the best solution may be a hybrid-- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv?
+A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on target selection.} \ben{ Suggest to remove this one: When the device is more complicated, it is possible that a better solution may be a hybrid -- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv?
 I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.} \sean{ I wouldn't state that.

### 2014-06-28 doc/sui14/discussion.tex
edits
refining applications (with limitation for complex devices) and commented 'comments' in disucssion
shorten the paper to fit 8 pages

--- 2014-04-16 doc/uist14/discussion.tex
@@
 The rapid development of sensing technologies has created many opportunities for new ways to interact with smart objects.
-In our exploration of the design space of head orientation-based target selection, we carefully selected sensing techniques that are readily available and easy to deploy; and we added complexity to our system only when necessary.
+In our exploration of the design space of HOBS, we carefully selected sensing techniques that are readily available and easy to deploy; and we added complexity to our system only when necessary.
 
@@
 A primary goal of this paper was to provide an effective and efficient method for target selections in physical spaces.
-Targeting is a fundamental building block across many interaction tasks - it has a significant impact on user experiences collectively and can provide seamless interaction when designed well.
+Targeting is a fundamental building block across many interaction tasks -- it has a significant impact on user experiences collectively and can provide seamless interaction when designed well.
 
-We formalized a scan and refine model of head orientation-based selection (Equation~\ref{eq:time}).
+We formalized a {\em scan} and {\em refine} model of head orientation-based selection (Equation~\ref{eq:time}).
 We first introduced head orientation as an alternative to list selection and showed that {\em scanning} can outperform list selection.
@@
 \subsection{Limitations}
-We introduced new refinement operations based on IR intensity and spatial adjacency tracking that improve the performance of head orientation in environments with dense targets.
-These techniques have several limitations.
 
+Our system faces a few limitations.
 First, IR intensity measurements only work within the dynamic range of our sensor.
@@
 This may be done incrementally during everyday interactions, but we have not yet tackled this challenge.
-We also have not yet evaluated how robust the adjacency map is when the user changes her location or view angle.
 
-Finally, we acknowledge several limitations of our study design: we have not yet systematically studied target density variation, our study was performed in a lab environment, and only measured first use.
-We have also not investigated error rates in detail.
+We also acknowledge several limitations of our study design: we have not yet systematically studied target density variation, our study was performed in a lab environment, and only measured first use.
 Future work should study how the technique applies in realistic settings over longer periods of time.
+
 %Point 3: Limitation
@@
 %- Study limitation: density variation, realistic settings, longitudinal studies to understand how users choose between our techniques and traditional list-based solution.
+
+%% \bjoern{removed CV subsection; handled in related work}
+
+%\subsection{Computer-vision Solution?}
+%Reviewers suggest considering computer vision-based techniques with markers in the environment (such as QR codes, or Bokode [Mohan] which extends the working distance to a few meters for visual tags).
+We considered this direction and found several drawbacks, including computational complexity, low accuracy when targets are far, the absence of real-time environment feedback with passive tags, and aesthetic issues with QR codes.
+While we don't claim our choice of IR is optimal, it has important advantages -- it is low-cost, readily available and quite suitable for area selection.
+We can weaken our claim and add explanations of how disambiguation techniques might generalize to other methods of implementing target identification.
 

### 2014-06-28 doc/sui14/relatedwork.tex
use 0.95 for all figures
shorten the paper to fit 8 pages

--- 2014-06-27 doc/sui14/relatedwork.tex
@@
 %!TEX root = sui14.tex
-\section{Background and Related Work}
-Our approach is related to head- and eye-controlled interfaces, area cursors and prior work on hardware devices for pointing in physical spaces.
-
 \vfill
 \ben{make sure no section title appears at the end of any column.}
+
+\section{Background and Related Work}
+Our approach is related to head- and eye-controlled interfaces, area cursors, pointing in physical spaces and computer vision-based selections.
+
 \subsection{Head and Gaze Input}
@@
 Target objects in the environment can also be equipped with individual cameras that watch the user~\cite{smith2013gaze,vertegaal2005media}.
-Such an approach can enable similar benefits as our approach, but is computationally more expensive than our single-pixel sensor solution and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.
+Such an approach can enable similar benefits as our approach, but is computationally more expensive and may not work at greater distances or angles, because it relies on finding the user's pupils in a camera image.
 
@@
 Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}.
-Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~
-\cite{bacim2013design}.
-In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed.
+Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~\cite{bacim2013design}.
+All area and space cursors necessitate disambiguation when there are multiple targets and no clear winner.
 This paper describes the trade-offs between several disambiguation approaches.
@@
 Rukzio ~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list.
-
-Several approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist.
+Several other approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist.
 %and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}. 
-In some techniques, users select objects of interest with laser pointers.
-The laser dot provides immediate visual feedback to the user about what is being selected; however, its small target area makes it poorly matched to head orientation input.
+In some techniques, users select objects of interest with laser pointers; however, the laser dot's small target area makes it poorly matched to head orientation input.
 Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.

### 2014-06-28 doc/sui14/interaction.tex
use 0.95 for all figures

@@
 \changes{IR signals are used for intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass.
-We have chosen a commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}).
-This architecture was mostly chosen for reasons of expediency and we do not claim optimality for our design decisions.
-Future head-mounted devices could clearly integrate IR emitters; the choice of wireless technology could also change.
-In particular, one could substitute WiFi modules or design an all-Bluetooth network and forego intermediate layers in our design.}
+We have chosen the commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}).
+This architecture was mostly chosen for reasons of expediency and we do not claim optimality for prototyping decisions.
+Future head-mounted devices could clearly integrate IR emitters; other wireless techniques (WiFi, Bluetooth) can also be used.}
 
@@
 \centering
-\includegraphics[width=1\columnwidth]{figures/GlassWithArduino.jpg}
+\includegraphics[width=0.95\columnwidth]{figures/GlassWithArduino.jpg}
 \caption{Our Glass hardware: Google Glass augmented with a repositionable IR holder, and an additional microcontroller that communicates with Google Glass and controls IR emitter.}
@@
 \centering
-\includegraphics[width=1\columnwidth]{figures/architecture_new}
+\includegraphics[width=0.95\columnwidth]{figures/architecture_new}
 \caption{Our System architecture diagram. The selection is initiated through infrared but confirmed over 802.15.4.}
@@
 
-{\bf Refine:} When disambiguation is needed, the user must make an explicit selection among the targets within his view range.
+{\bf Refine:} When disambiguation is needed, the user must make an explicit selection among the targets within their view range.
 We have designed multiple refinement mechanisms -- all of which enable the user to select one from a subset of targets.

### 2014-06-28 doc/sui14/intro.tex
correct punctuation syntax

@@
 
-%\sean{Ben mentioned about discussing the limitations of list approach. Below is the paragraph from chi submission for your reference.}
-%
 \changes{Today, companies like Samsung and Whirlpool are making smart appliances with companion applications that use smartphones as {\em universal remote controls}.
 With these applications, the user can select a device from a list in order to control it with a device-specific user interface.
-This method faces {\em naming} issues (i.e.
-"what do we name the lamp on the left?") and {\em scaling} issues as the number of controlled devices increases.
+However, this method faces {\em naming} issues (i.e.
+``what do we name the lamp on the left?'') and {\em scaling} issues as the number of controlled devices increases.
 These solutions also present a necessarily flawed mapping from the positions of the appliances in the rich, 3-dimensional world to their place in a 1D or 2D list presented on the screen.
@@
 We thus investigate how such computing devices may be used for the selection and control of devices in physical spaces.
-
 Head-worn devices can naturally exploit the user's head orientation, an important (but imprecise) indicator of the user's {\em locus of attention}~\cite{raskin}.
@@
 In all our techniques, the emitted IR beam %(a diameter of 30-60cm and distance up to 8m)%
- provides an initial {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser} {\em left}) --- all targets that receive this IR signal can be selected.
+ provides an initial {\em coarse} selection area (illustrated in Figure~\ref{fig:teaser} {\em left}).
 To {\em refine} selection when multiple targets have received IR signals, we describe and evaluate three techniques:
@@
 
-%% three ways
-%---
 We also demonstrate an example application of our technique used as a remote control of smart appliances such as lighting and TV sets: a user looks at the appliance he wishes to control and confirms selection by tapping.

### 2014-06-28 doc/sui14/iteration3.tex
iteration3.tex: his->their
use 0.95 for all figures

@@
 
-In the learning stage, the user scans over the targets, and the system attains
-the absolute orientation of each device from IR and motion sensors. From this
-information it can abstract out their relative positions and build an {\em
-adjacency map}.
+In the learning stage, as the user scans over targets in their normal use, the
+system attains the absolute orientation of each device from IR and motion
+sensors. From this information it can abstract out their relative positions and
+build an {\em adjacency map}. This process is transparent to the user, and the
+map can be built without an explicit calibration phase.
 
 This absolute orientation cannot be applied to all indoor environments, since
-the user’s movements through the space could change the relationships between
+the user's movements through the space could change the relationships between
 targets. However, with the constraint that the targets are spread around the
@@
 \centering
-\includegraphics[width=1\columnwidth]{figures/third_principle.png}
+\includegraphics[width=0.95\columnwidth]{figures/third_principle.png}
 \caption{This illustrates that the change of a user's absolution position
@@
 \centering
-\includegraphics[width=1\columnwidth]{figures/third_technique.pdf}
+\includegraphics[width=0.95\columnwidth]{figures/third_technique.pdf}
 \caption{Our third technique learned each target's absolute orientation and
@@
 quasi-mode for refinement. In this quasi-mode, one device lights up at a time.
-When the user turns his head in the direction of another device, the light
+When the user turns their head in the direction of another device, the light
 switches to that device. Therefore, the user can move between devices one at a

### 2014-06-28 doc/sui14/applications.tex
rewrote last sentence of applications section, fixed refs

@@
-%!TEX root = uist14.tex
+%!TEX root = sui14.tex
 \section{Applications}
@@
 User interfaces for each were pre-defined in our application.
-
+%
 We set the devices up in a simulated living room environment and invited 14 users to step through a predefined set of tasks to control the appliances at a distance.
@@
 All participants successfully completed the list of tasks.
-They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous study procedure.
+They commented positively on the universal remote control functionality (e.g., \studyquote{I didn't have to search for different remote controllers for different appliances}) and stated it was easy to target and connect to appliances, in line with the findings of the previous studies.
 Participants saw potential benefits of the device for families; one user remarked that he could imagine people using the system \studyquote{while keeping an eye on their children at the same time}.
@@
 This suggests that future applications in domains like Z could be especially promising.}
-
-\changes{Though our solution successfully solves the problem of {\em target selection}, we have not fully addressed the issue of controlling smart devices.
+\changes{While our system enabled users to select and control the example devices successfully, work remains in devising appropriate interfaces for complex devices.
+Users rated ease of use higher for the lamp and fan which had simple, discrete on/off actions, and lower for the TV control, which had more options.
+Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
+We leave addressing this fundamental usability challenge of wearable devices with near-eye displays to future work.}
+%\changes{Though our solution successfully solves the problem of {\em target selection}, we have not fully addressed the issue of controlling smart devices.
 One user reported that \studyquote{for more complex appliances (TV), interaction was more difficult}.

### 2014-06-28 doc/sui14/applications.tex
grammer fixes

@@
 This suggests that future applications in domains like Z could be especially promising.}
-\changes{While our system enabled users to select and control the example devices successfully, work remains in devising appropriate interfaces for complex devices.
+\changes{While our system enables users to select and control the example devices successfully, work remains in devising appropriate interfaces for complex devices.
 Users rated ease of use higher for the lamp and fan which had simple, discrete on/off actions, and lower for the TV control, which had more options.
-Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{Most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
+Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
 We leave addressing this fundamental usability challenge of wearable devices with near-eye displays to future work.}

### 2014-06-28 doc/sui14/conclusion.tex
grammer fixes

--- 2014-04-16 doc/uist14/conclusion.tex
@@
 However, a more natural approach is to combine IR with head motion.
-Through our preliminary user studies, we learned that this is a more intuitative way of performing refinement in comparison to the menu-based selections.
+Through our preliminary user studies, we learned that this is a more intuitive way of performing refinement in comparison to the menu-based selections.
 We leave a more comprehensive technical solution of using motion sensors and its evaluation as the future work.

### 2014-06-28 doc/sui14/discussion.tex
grammer fixes

@@
 
-We also acknowledge several limitations of our study design: we have not yet systematically studied target density variation, our study was performed in a lab environment, and only measured first use.
+We also acknowledge several limitations of our study design: we have not yet systematically studied target density variation; our study was performed in a lab environment; and only measured first use.
 Future work should study how the technique applies in realistic settings over longer periods of time.

### 2014-06-28 doc/sui14/interaction.tex
grammer fixes

@@
 
-\changes{IR signals are used for intial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass.
+\changes{IR signals are used for initial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass.
 We have chosen the commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}).
@@
 We have designed multiple refinement mechanisms -- all of which enable the user to select one from a subset of targets.
-The user confirms a current selection with a tap.
+The user confirms the current selection with a tap.
 Since the purpose of this stage is to disambiguate among potential targets, we will also use {\em disambiguation} to refer to this stage.

### 2014-06-28 doc/sui14/iteration2.tex
use 0.95 for all figures
grammer fixes

@@
 \centering
-\includegraphics[width=1\columnwidth]{figures/IRIntensityDistribution.pdf}
-\caption{Empirical measurement of IR intensity at different position.
+\includegraphics[width=0.95\columnwidth]{figures/IRIntensityDistribution.pdf}
+\caption{Empirical measurement of IR intensity at different positions.
 We measure the intensity at a plane every meter away from the IR emitter.
@@
 To quantify the improvement in this design we performed a second study to compare the {\em Naive IR} and {\em Intensity IR} approaches.
-Because we were interested in discovering performance differences in denser environment, we reposition the 10 nodes and set them up in a smaller area (see Figure~\ref{fig:study-layout2}).
+Because we were interested in discovering performance differences in denser environment, we re-positioned the 10 nodes and set them up in a smaller area (see Figure~\ref{fig:study-layout2}).
 We recruited 10 participants for this study.
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/study-layout2.pdf}
+\includegraphics[width=0.95\columnwidth]{figures/study-layout2.pdf}
 \caption{The environment setup for our second study. In comparison to our first study, we have deliberately increased the target density.}
@@
 From Figure~\ref{fig:study2}, we can see that the overall target acquisition time has decreased from 4.31 seconds for {\em Naive IR} to 3.64 second for {\em Intensity IR}.
-This difference is also significant ($t(555)=3.2945$, $p=0.001$.)
+This difference is also significant ($t(555)=3.2945$, $p=0.001$).
 
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/result_study2.pdf}
+\includegraphics[width=0.95\columnwidth]{figures/result_study2.pdf}
 \caption{Boxplot and CDF of the target acquisition time in {\em Naive IR} and {\em Intensity IR} conditions.}

### 2014-08-06 doc/sui14/applications.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

@@
 This suggests that future applications in domains like Z could be especially promising.}
-\changes{While our system enables users to select and control the example devices successfully, work remains in devising appropriate interfaces for complex devices.
+While our system enables users to select and control the example devices successfully, work remains in devising appropriate interfaces for complex devices.
 Users rated ease of use higher for the lamp and fan which had simple, discrete on/off actions, and lower for the TV control, which had more options.
 Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}.
-We leave addressing this fundamental usability challenge of wearable devices with near-eye displays to future work.}
+We leave addressing this fundamental usability challenge of wearable devices with near-eye displays to future work.
 %\changes{Though our solution successfully solves the problem of {\em target selection}, we have not fully addressed the issue of controlling smart devices.

### 2014-08-06 doc/sui14/interaction.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

@@
 \section{IR for head orientation-based targeting}
-\changes{In this section, we present our hardware platform (IR targeting) and interaction model (head-orientation based) that we will use throughout our iterative designs.}
+In this section, we present our hardware platform (IR targeting) and interaction model (head-orientation based) that we will use throughout our iterative designs.
 
@@
 IR LEDs with many different beam angles are commercially available.
-\changes{We augment Google Glass with a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545).
+We augment Google Glass with a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545).
 The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio, since Glass does not directly support hardware modification (see Figure~\ref{fig:glass}).
 Target devices use Vishay TSOP38238 IR receivers.
-Data is encoded using standard IR remote protocols at 38.0kHz.}
+Data is encoded using standard IR remote protocols at 38.0kHz.
 
-\changes{IR signals are used for initial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass.
+IR signals are used for initial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass.
 We have chosen the commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}).
 This architecture was mostly chosen for reasons of expediency and we do not claim optimality for prototyping decisions.
-Future head-mounted devices could clearly integrate IR emitters; other wireless techniques (WiFi, Bluetooth) can also be used.}
+Future head-mounted devices could clearly integrate IR emitters; other wireless techniques (WiFi, Bluetooth) can also be used.
 

### 2014-08-06 doc/sui14/intro.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

@@
 
-\changes{Today, companies like Samsung and Whirlpool are making smart appliances with companion applications that use smartphones as {\em universal remote controls}.
+Today, companies like Samsung and Whirlpool are making smart appliances with companion applications that use smartphones as {\em universal remote controls}.
 With these applications, the user can select a device from a list in order to control it with a device-specific user interface.
@@
 These solutions also present a necessarily flawed mapping from the positions of the appliances in the rich, 3-dimensional world to their place in a 1D or 2D list presented on the screen.
-}
 %

### 2014-08-06 doc/sui14/iteration1.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

@@
 
-\vfill
 \subsection{Technique}
@@
 As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, \changes{IR selection was perceived to be more direct and pleasurable.
+While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and pleasurable.
 One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen}.
@@
 Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}.
-Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}}.
+Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}.
 

### 2014-08-06 doc/sui14/relatedwork.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

@@
 \vfill
-\ben{make sure no section title appears at the end of any column.}
+%\ben{make sure no section title appears at the end of any column.}
 
@@
 \subsection{Head and Gaze Input}
-\changes{Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}.
-However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.}
+Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}.
+However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
 %
@@
 \subsection{Area Cursors}
-\changes{
+
 In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}.
@@
 Many disambiguation techniques are possible for refinement -- this paper describes the trade-offs between several of them.
-}
 
@@
 \subsection{Vision- and Projection-Based Target Selection}
-\changes{Many alternative solutions for detecting devices in contained spaces
+Many alternative solutions for detecting devices in contained spaces
 rely on computer-vision recognition of printed tags on devices. Unfortunately,
@@
 in spirit but requires only small, low-cost IR emitters and detectors.
-\bjoern{need a stronger statement here}} \achal{Tried emphasizing the smaller
-requirements, not sure what else we can say.}
 

### 2014-08-06 doc/sui14/sui14.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

@@
 %% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
-\toappear{Submitted for review to SUI 2014.}
+\toappear{Permission to make digital or hard copies of part or all of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
+Copyrights for third-party components of this work must be honored.
+For all other uses, contact the owner/author(s).
+Copyright is held by the author/owner(s).\\
+{\emph{SUI'14}}, October 4–5, 2014, Honolulu, HI, USA.\\
+ACM 978-1-4503-2820-3/14/10.\\
+\changes{TODO: DOI string goes here.}}
 
@@
 % Remove this line to eliminate page numbers for the camera ready copy
-\pagenumbering{arabic}
+%\pagenumbering{arabic}
 

### 2014-08-06 doc/sui14/sui14.tex
updated copyright block

@@
 Copyrights for third-party components of this work must be honored.
-For all other uses, contact the owner/author(s).
-Copyright is held by the author/owner(s).\\
-{\emph{SUI'14}}, October 4–5, 2014, Honolulu, HI, USA.\\
+For all other uses, contact the Owner/Author.\\
+\\
+Copyright is held by the owner/author(s).\\
+\\
+{\emph{SUI '14}}, Oct 04-05 2014, Honolulu, HI, USA\\
 ACM 978-1-4503-2820-3/14/10.\\
-\changes{TODO: DOI string goes here.}}
+http://dx.doi.org/10.1145/2659766.2659773}
 

### 2014-08-08 doc/sui14/iteration2.tex
p < 0.0001 to p < 0.001

@@
 {\em Intensity IR} reduces the number of trials in which refinement dialogs are needed from 225 of 300 in {\em Naive IR} to 167 of 300 trials.
-A Chi-square test shows this difference is significant ($\chi^2(1) = 24.755$, $p < 0.0001$).
+A Chi-square test shows this difference is significant ($\chi^2(1) = 24.755$, $p < 0.001$).
 This demonstrates that the new approach successfully reduces the probability that a disambiguation dialog is needed.
@@
 In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 225).
-A Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, $p < 0.0001$).
+A Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, $p < 0.001$).
 

### 2014-08-08 doc/sui14/interaction.tex
Fix grammar issues pointed out by SUI R2

@@
 
-IR signals are used for initial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass.
+IR signals are used for initial line-of-sight targeting; subsequently, we use bidirectional wireless communication to send information such as target IDs and signal strength from targets back to Glass.
 We have chosen the commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}).
@@
 If there is only one single target in IR range, it is automatically selected.
-However, in a dense environment where multiple targets are within range, the user needs refine his selection.
+However, in a dense environment where multiple targets are within range, the user needs to refine his selection.
 

### 2014-08-08 doc/sui14/iteration1.tex
Fix grammar issues pointed out by SUI R2

@@
 During the study, we measured the {\bf target acquisition time} for each target selection.
-Afterward, participants were asked to complete a survey of primarily open-ended answers about their experience.
+Afterward, participants were asked to complete a survey of primarily open-ended questions about their experience.
 
@@
 As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and pleasurable.
+While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and pleasant.
 One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen}.

### 2014-08-08 doc/sui14/relatedwork.tex
Fix grammar issues pointed out by SUI R2

@@
 \subsection{Pointing in Physical Spaces}
-Rukzio ~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list.
+Rukzio et al.
+~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list.
 Several other approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist.

### 2014-08-08 doc/sui14/iteration3.tex
Added distance constraint for iteration3

@@
 targets. However, with the constraint that the targets are spread around the
-periphery, their relative orientations are stable (see
+periphery at similar distances, their relative orientations are stable (see
 Figure~\ref{fig:third_principle}).

### 2014-08-08 doc/sui14/iteration1.tex
Add \changes

@@
 During the study, we measured the {\bf target acquisition time} for each target selection.
-Afterward, participants were asked to complete a survey of primarily open-ended questions about their experience.
+Afterward, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.
 
@@
 As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).
-While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and pleasant.
+While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and \changes{pleasant}.
 One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen}.

### 2014-08-08 doc/sui14/iteration2.tex
Add \changes

@@
 {\em Intensity IR} reduces the number of trials in which refinement dialogs are needed from 225 of 300 in {\em Naive IR} to 167 of 300 trials.
-A Chi-square test shows this difference is significant ($\chi^2(1) = 24.755$, $p < 0.001$).
+A Chi-square test shows this difference is significant ($\chi^2(1) = 24.755$, \changes{$p < 0.001$}).
 This demonstrates that the new approach successfully reduces the probability that a disambiguation dialog is needed.
@@
 In comparison, for {\em Naive IR}, only 35\% of trials sorted the desired target as the first one in the list (80 out of 225).
-A Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, $p < 0.001$).
+A Chi-square test show that this difference is significant (with $\chi^2(1) = 15.758$, \changes{$p < 0.001$}).
 

### 2014-08-08 doc/sui14/iteration3.tex
Add \changes

--- 2014-06-28 doc/sui14/iteration3.tex
@@
 targets. However, with the constraint that the targets are spread around the
-periphery, their relative orientations are stable (see
-Figure~\ref{fig:third_principle}).
+periphery \changes{at similar distances}, their relative orientations are
+stable (see Figure~\ref{fig:third_principle}).
 

### 2014-08-08 doc/sui14/relatedwork.tex
Add \changes

--- 2014-08-06 doc/sui14/relatedwork.tex
@@
 \subsection{Pointing in Physical Spaces}
-Rukzio ~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list.
+Rukzio \changes{et al.} ~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list.
 Several other approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist.

### 2014-08-10 doc/sui14/applications.tex
for remote control application, added info that we are using naive IR

@@
 
-Our prototype includes three smart devices: a lamp and a fan that could be switched on and off; a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
+Our prototype \changes{(based on {\em Naive IR}\footnote{For this particular configuration, the environment is not dense enough, and {\em Naive IR} performs well for head orientation based targeting proof-of-concept.})} includes three smart devices: a lamp and a fan that could be switched on and off; a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
 The on/off appliances are using relays, and the smart TV is a 30\inch{} display connected to a laptop.
 User interfaces for each were pre-defined in our application.
+
 %

### 2014-08-10 data/201404.Studies/data_summary.R
updated figure 9 to use angle rather than distance, note: text not updated yet

@@ -171,16 +171,17 @@ df <- cbind(new, offset, distance)
 df_mirror <- df[df$offset > 0,]
 df_mirror$offset <- -df_mirror$offset
 whole <- rbind(df, df_mirror)
+whole$angle <- atan(whole$offset * 0.01 / whole$distance) / pi * 180
 
-p <- ggplot(whole, aes(x=distance, y=offset, colour = value)) +
+p <- ggplot(whole, aes(x=distance, y=angle, colour = value)) +
   geom_point(aes(size= log(value))) +
   theme_bw() +
   theme(legend.position="none") +
   xlab("distance (m) between IR and the receiver") +
-  ylab("offset (cm) from the center") +
+  ylab("angle (degree) from the center") +
   ggtitle("IR intensity distribution (measured)") +
   scale_fill_brewer() +
   xlim(0, 8.1)
-pdf("figures/IRIntensityDistribution.pdf", width = 10, height = 3)
+pdf("figures/IRIntensityDistribution.pdf", width = 10, height = 3.5)
 p
 dev.off()

### 2014-08-10 doc/sui14/iteration2.tex
added description about our intensity measurement

@@
 We have empirically measured the intensity distribution at the receiver for this configuration in Figure~\ref{fig:measurement}.
-Our measurements confirm that angular difference has a large effect on the intensity readings (see how the IR intensity distribution changes when the offset from the center increases).
+Our measurements confirm that angular difference has a large effect on the intensity readings (see how the IR intensity distribution changes when the \changes{angle has increased}).
 
@@
 \includegraphics[width=0.95\columnwidth]{figures/IRIntensityDistribution.pdf}
-\caption{Empirical measurement of IR intensity at different positions.
+\caption{This figure shows an empirical measurement of IR intensity at different positions.
 We measure the intensity at a plane every meter away from the IR emitter.
-On that plane, we obtain a sample every 5 centimeters.
-The size and color brightness in this graph represents the intensity readings.}
+On that plane, a sample is mesaured every 5 centimeters, \changes{and we stop when the reading is smaller than ambient noise.} \changes{To match with our assumption about the relationship between angles and intensities, the vertical axis is changed to the degree of the angle.} The size and color brightness in this graph represents the intensity readings for visualization.}
 \label{fig:measurement}

### 2014-08-11 doc/sui14/iteration1.tex
fixed another p < 0.001 issue
adding info about recruitment
remove useless newlines

@@
 Each of them has a number as ID and a letter representing the name.
-We recruited 14 participants from our institution for this study.
+We recruited 14 participants from our institution \changes{from email invitations} for this study.
 13 had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
@@
 \subsection{Results}
-
 Our results indicate that {\em IR selection} outperforms {\em list selection}.  The average target acquisition time for {\em IR selection} is 6.67 seconds 
-while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, p=0.00017$).
+while {\em list selection} took 8.86 seconds (see Figure~\ref{fig:ir_vs_list}). A t-test shows a significant difference ($t(279)=-3.81, \changes{p < 0.001}$).
 

### 2014-08-11 doc/sui14/iteration3.tex
refining texts for iteration3, more descriptions are needed

--- 2014-06-28 doc/sui14/iteration3.tex
@@
 
-In our third iteration, we ask the question {\em can we build a system purely
-using head orientation and visual feedback from the environment for target
-selection?}
+In our third iteration, we ask the \changes{research} question --
+{\em can we build a system purely using head orientation and visual
+  feedback from the environment for target selection?}
 
@@
 
-In the learning stage, as the user scans over targets in their normal use, the
-system attains the absolute orientation of each device from IR and motion
-sensors. From this information it can abstract out their relative positions and
-build an {\em adjacency map}. This process is transparent to the user, and the
+In the learning stage, as the user scans over targets in their normal
+use, the system attains the absolute orientation of each device from
+IR and motion sensors. From this information it can abstract out
+\changes{the} relative positions \changes{of targets} and construct an
+{\em adjacency map}. The absolute orientations in the map cannot be
+applied to all indoor environments, since the user's movements through
+the space could change the relationships between targets. However,
+with the constraint that the targets are spread around the periphery
+\changes{at similar distances}, their relative orientations are stable
+(see Figure~\ref{fig:third_principle}). And since we are using IR at
+the same time, the learning process is transparent to the user and the
 map can be built without an explicit calibration phase.
-
-This absolute orientation cannot be applied to all indoor environments, since
-the user's movements through the space could change the relationships between
-targets. However, with the constraint that the targets are spread around the
-periphery, their relative orientations are stable (see
-Figure~\ref{fig:third_principle}).
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.95\columnwidth]{figures/third_principle.png}
-\caption{This illustrates that the change of a user's absolution position
-doesn't change the relative relationship of physical targets.}
-\label{fig:third_principle}
-\end{figure}
 
@@
 
-After the map is created, the user can hold down on the touchpad to enter a
-quasi-mode for refinement. In this quasi-mode, one device lights up at a time.
-When the user turns their head in the direction of another device, the light
-switches to that device. Therefore, the user can move between devices one at a
-time with slight head movements. We implement this interaction by calculating
-the user's direction of motion using a low-passed history of sensor
-measurements and searching through the adjacency map for the nearest device in
-that direction.
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.95\columnwidth]{figures/third_principle.png}
+\caption{This illustrates that the change of a user's absolution position
+doesn't change the relative relationship of physical targets.}
+\label{fig:third_principle}
+\end{figure}
+
+\changes{After the map is created, the user can enter a quasi-mode for
+  refinement by holding down on the touchpad. In this quasi-mode, one
+  device lights up at a time. When the user turns their head in the
+  direction of another device, the light switches to that
+  device. Therefore, the user can move between devices one at a time
+  with slight head movements. This prediction is implemented by
+  calculating the user's direction of motion using a low-passed
+  history of sensor measurements and searching through the adjacency
+  map for the nearest device in that direction.}
 

### 2014-08-11 doc/sui14/iteration2.tex
in iteration 2, we count the error trials in data analysis

@@
 This is higher than the error rate for {\em Naive IR} (only 5 out of 300, ~1.6\%).
-% but this is the trade-off for the overall faster acquisition time.
+\changes{Even though the higher error rate increases the chance of multiple attempts in target selection, from our analysis above which has included the error trials, we see that the overall performance of {\em Intensity IR} is still better than {\em Naive IR}.
+}  % but this is the trade-off for the overall faster acquisition time.
 

### 2014-08-11 doc/sui14/iteration3.tex
task performance for iteration 3

@@
 
-The users had strong preferences for the new method of refinement. On a scale
-from 1-7, 1 being the least mental effort and 7 being the most mental effort
-users rated the old technique 4.25 and the new technique 2 on average. 100\%
-indicated a preference for head movement to list navigation. One user
-referenced the issue of naming targets that the list necessitates, preferring
-the experience of \studyquote{matching visual cues rather than numbers}.
-Another participant remarked that it \studyquote{just made more sense} and was
-a \studyquote{more natural way for demonstrating intentionality}. The users
-preferred the new mapping in relation to the whole environment: \studyquote{it
-leveraged the spatial sense that I already had just by using the system}. They
-were also delighted to avoid list navigation, which they now called
-\studyquote{difficult} and \studyquote{painful}.
+The users had strong preferences for the new method of refinement.
+\changes{Though we lack formal data to evaluate the performance of
+this method, from our observeation during the experiments, we found
+that each trial became much easier than previous iterations. We
+conducted a survey to collect qualitative feedback after the
+experiement.} On a scale from 1-7, 1 being the least mental effort and
+7 being the most mental effort users rated the old technique 4.25 and
+the new technique 2 on average. 100\% indicated a preference for head
+movement to list navigation. One user referenced the issue of naming
+targets that the list necessitates, preferring the experience of
+\studyquote{matching visual cues rather than numbers}.  Another
+participant remarked that it \studyquote{just made more sense} and was
+a \studyquote{more natural way for demonstrating intentionality}. The
+users preferred the new mapping in relation to the whole environment:
+\studyquote{it leveraged the spatial sense that I already had just by
+using the system}. They were also delighted to avoid list navigation,
+which they now called \studyquote{difficult} and \studyquote{painful}.

### 2014-08-13 doc/sui14/iteration1.tex
add more description for i1

@@
 During the study, we measured the {\bf target acquisition time} for each target selection.
-Afterward, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.
+\changes{At the beginning of each acquisition, participants were asked to stand at a fixed position approximately 10-12 feet away from the nodes.
+They would have to look down until the target is given.
+As the start to connect, they may move around if they decided it would help them select.
+After each task, they needed to move back to the starting position.} Afterward, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.
 

### 2014-08-13 doc/sui14/iteration1.tex
add physical description, will update more later

@@
 
+\changes{Most participants walked one or two steps during each task.
+In the survey we asked \studyquote{When you have multiple candidate targets within sight, which method for selecting between did you use?} (multiple choice), among 16 participates, X\% chose {\em adjust head so that only one candidate is signaled}, Y\% chose {\em walk closer to the candidate for a better result}, and only Z\% chose {\em tap (to enter refinement stage) and then choose between list}.
+\sean{will update numbers and add a few sentences here.}}
 As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).

### 2014-08-13 doc/sui14/iteration1.tex
update participant number & demographics. TBC

@@
 Each of them has a number as ID and a letter representing the name.
-We recruited 14 participants from our institution \changes{from email invitations} for this study.
-13 had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
+We recruited \changes{16} participants from our institution \changes{from email invitations} for this study \changes{(8 males, 6 females)} And figure X shows their department distribution.
+\sean{will calculate later.
+shall we use a pie chart?}  \changes{14} had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
 Four participants wore prescription glasses, which makes Glass more cumbersome to use and adjust and may have affected their task performance.
@@
 \sean{will update numbers and add a few sentences here.}}
+
 As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).

### 2014-08-13 doc/sui14/iteration2.tex
comment: suggest add physical movement desc. in i2

@@
 Each  user performed 30 target acquisition tasks for each approach.
-As in our first within-subject study, half of them perform {\em Naive IR} first and the other half {\em Intensity IR} first.
+\sean{Need to add desciption about participants physical movements or restrictions here.
+Ben or Claire?} As in our first within-subject study, half of them perform {\em Naive IR} first and the other half {\em Intensity IR} first.
 
@@
 
-One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage.
+\sean{If possible, add how we observed participants physical movements in this iteration.
+Better if we can compare to i1.} One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage.
 Out of 300 trials, 13 of them (4.3\%) has been accidentally eliminated.

### 2014-08-13 doc/sui14/iteration1.tex
changed grammar in method

@@
 \subsection{Method}
-We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}).
-Each of them has a number as ID and a letter representing the name.
-We recruited \changes{16} participants from our institution \changes{from email invitations} for this study \changes{(8 males, 6 females)} And figure X shows their department distribution.
+We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}), each with a number ID and a letter representing the name.
+For the study, we recruited \changes{16} participants from our institution \changes{(8 males and 6 females)} \changes{using email invitations}.
+Figure X shows their department distribution.
 \sean{will calculate later.

### 2014-08-13 doc/sui14/applications.tex
adding tasks for applications
added some commas

--- 2014-08-06 doc/sui14/applications.tex
@@
 
-Our prototype includes three smart devices: a lamp and a fan that could be switched on and off; a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
+Our prototype, \changes{(based on {\em Naive IR}, \footnote{For this particular configuration, the environment is not dense enough, and {\em Naive IR} performs well for head orientation based targeting proof-of-concept.})} includes three smart devices: a lamp and a fan that could be switched on and off; a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
 The on/off appliances are using relays, and the smart TV is a 30\inch{} display connected to a laptop.
 User interfaces for each were pre-defined in our application.
+
 %
 We set the devices up in a simulated living room environment and invited 14 users to step through a predefined set of tasks to control the appliances at a distance.
+\changes{These tasks include turning off the light, playing movie, turning up the volume, etc., which are some sample tasks that people would normally do at home.}
+
+%% set of predefined tasks:
+% \item Turn off the lights as you want to watch the movie in a darkened room.
+% \item You feel a little hot in the room, so you turn on the fan.
+% \item You connect to the Smart TV and start playing the movie. 
+% \item The volume seems too soft to hear over the fan- turn it up a bit. 
+% \item After a while, you want to take a break to get a snack. Pause the movie. 
+% \item When you come back, you've forgotten what was said last - rewind by ~30 seconds and restart the movie. 
+% \item When the credits roll, you stop the movie and turn the lights back on.
+% \item After awhile, you turn off the fan and leave the room.
 

### 2014-08-13 doc/sui14/iteration2.tex
changed grammar

@@
 We measure the intensity at a plane every meter away from the IR emitter.
-On that plane, a sample is mesaured every 5 centimeters, \changes{and we stop when the reading is smaller than ambient noise.} \changes{To match with our assumption about the relationship between angles and intensities, the vertical axis is changed to the degree of the angle.} The size and color brightness in this graph represents the intensity readings for visualization.}
+On that plane, a sample is taken every 5 centimeters, \changes{and we stop when the reading is smaller than ambient noise.} \changes{To match our assumption about the relationship between angles and intensities, the vertical axis is changed to the degree of the angle.} The size and color brightness represent the intensity of the readings for visualization.}
 \label{fig:measurement}

### 2014-08-13 doc/sui14/iteration2.tex
grammar grammar

@@
 Better if we can compare to i1.} One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage.
-Out of 300 trials, 13 of them (4.3\%) has been accidentally eliminated.
-This is higher than the error rate for {\em Naive IR} (only 5 out of 300, ~1.6\%).
-\changes{Even though the higher error rate increases the chance of multiple attempts in target selection, from our analysis above which has included the error trials, we see that the overall performance of {\em Intensity IR} is still better than {\em Naive IR}.
+Out of 300 trials, \changes{the target is accidentally eliminated in 13 (4.3\%).} This is higher than the error rate for {\em Naive IR} (5 out of 300, ~1.6\%).
+\changes{Even though the higher error rate increases the chance of multiple attempts in target selection, our analysis above illustrates that overall performance of {\em Intensity IR} is still better than {\em Naive IR}.
 }  % but this is the trade-off for the overall faster acquisition time.

### 2014-08-13 doc/sui14/iteration3.tex
fixed spelling errors

@@
 \changes{Though we lack formal data to evaluate the performance of
-this method, from our observeation during the experiments, we found
+this method, from our observation during the experiments, we found
 that each trial became much easier than previous iterations. We
 conducted a survey to collect qualitative feedback after the
-experiement.} On a scale from 1-7, 1 being the least mental effort and
+experiment.} On a scale from 1-7, 1 being the least mental effort and
 7 being the most mental effort users rated the old technique 4.25 and

### 2014-08-13 doc/sui14/iteration3.tex
Added back implementation details for iter3
These existed earlier but were removed at some point...

@@
 \changes{After the map is created, the user can enter a quasi-mode for
-  refinement by holding down on the touchpad. In this quasi-mode, one
-  device lights up at a time. When the user turns their head in the
-  direction of another device, the light switches to that
-  device. Therefore, the user can move between devices one at a time
-  with slight head movements. This prediction is implemented by
-  calculating the user's direction of motion using a low-passed
-  history of sensor measurements and searching through the adjacency
-  map for the nearest device in that direction.}
+refinement by holding down on the touchpad. In this quasi-mode, one device
+lights up at a time. When the user turns his head in the direction of another
+device, the light switches to that device. Therefore, the user can move between
+devices one at a time with slight head movements. This prediction is
+implemented by first calculating the user's direction of motion, and searching
+through the adjacency map for the nearest device in that direction.  To
+calculate the direction, we maintain a circular buffer of the last 10 sensor
+measurements, passed through a low pass filter. The direction is then
+calculated as the difference between the last sample and the first sample.
+Note that we use a hysteresis method to temporarily lock to the current device
+- if the variance of the buffer is below a threshold, we assume no movement has
+occurred and do not select another device.}
 

### 2014-08-13 doc/sui14/iteration1.tex
Grammar changes and comments

@@
 During the study, we measured the {\bf target acquisition time} for each target selection.
-\changes{At the beginning of each acquisition, participants were asked to stand at a fixed position approximately 10-12 feet away from the nodes.
-They would have to look down until the target is given.
-As the start to connect, they may move around if they decided it would help them select.
-After each task, they needed to move back to the starting position.} Afterward, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.
+\changes{At the beginning of each acquisition, participants were asked to stand at a fixed position approximately 10-12 feet away from the nodes, looking down until the target is provided.
+As they attempt to connect, they may move around if they decided it would help them select.
+After each task, they were asked to move back to the starting position.} Afterward, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.
 
@@
 \changes{Most participants walked one or two steps during each task.
-In the survey we asked \studyquote{When you have multiple candidate targets within sight, which method for selecting between did you use?} (multiple choice), among 16 participates, X\% chose {\em adjust head so that only one candidate is signaled}, Y\% chose {\em walk closer to the candidate for a better result}, and only Z\% chose {\em tap (to enter refinement stage) and then choose between list}.
+In the survey we asked \studyquote{When you have multiple candidate targets within sight, which method for selecting between them did you use?} (multiple choice), among 16 participates, X\% chose {\em adjust head so that only one candidate is signaled}, Y\% chose {\em walk closer to the candidate for a better result}, and only Z\% chose {\em tap (to enter refinement stage) and then choose between list}.
 \sean{will update numbers and add a few sentences here.}}

### 2014-08-13 doc/sui14/iteration3.tex
Grammar changes and comments

@@
 conducted a survey to collect qualitative feedback after the
-experiment.} On a scale from 1-7, 1 being the least mental effort and
+experiment.} \achal{I vote changing to "Thoguh we lack... during the experiments,
+we conducted..."} On a scale from 1-7, 1 being the least mental effort and
 7 being the most mental effort users rated the old technique 4.25 and

### 2014-08-14 doc/sui14/iteration1.tex
update demographics

--- 2014-08-13 doc/sui14/iteration1.tex
@@
 For the study, we recruited \changes{16} participants from our institution \changes{(8 males and 6 females)} \changes{using email invitations}.
-Figure X shows their department distribution.
-\sean{will calculate later.
-shall we use a pie chart?}  \changes{14} had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
+\changes{They are composed of undergrad students, graduate students, and university staff.
+Their background ranges from Information(6), Engineering(4), Math and Science(4), Design(1), Others(1)}  \changes{14} had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
 Four participants wore prescription glasses, which makes Glass more cumbersome to use and adjust and may have affected their task performance.

### 2014-08-14 doc/sui14/iteration1.tex
more discussion on physical movements

@@
 
-\changes{Most participants walked one or two steps during each task.
-In the survey we asked \studyquote{When you have multiple candidate targets within sight, which method for selecting between did you use?} (multiple choice), among 16 participates, X\% chose {\em adjust head so that only one candidate is signaled}, Y\% chose {\em walk closer to the candidate for a better result}, and only Z\% chose {\em tap (to enter refinement stage) and then choose between list}.
-\sean{will update numbers and add a few sentences here.}}
+As for the qualitative feedback, 11 out of \changes{16} participants preferred IR selection over list mode (\changes{3} preferred list, \changes{2} was undecided).
+\changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between did you use?} (multiple choices), among 16 participates, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result}.
+Although through our observation, many participants walked one or two steps during some tasks, this suggests that participants viewed their moving as a followed-through action of adjusting their heads instead of an intention to walk closer to the targets.
 
-As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided).
 While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and \changes{pleasant}.

### 2014-08-14 doc/sui14/iteration1.tex

@@
 As for the qualitative feedback, 11 out of \changes{16} participants preferred IR selection over list mode (\changes{3} preferred list, \changes{2} was undecided).
-\changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between did you use?} (multiple choices), among 16 participates, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result}.
+\changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between them did you use?} (multiple choices), among 16 participates, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result}.
 Although through our observation, many participants walked one or two steps during some tasks, this suggests that participants viewed their moving as a followed-through action of adjusting their heads instead of an intention to walk closer to the targets.

### 2014-08-14 doc/sui14/iteration2.tex

@@
 We have empirically measured the intensity distribution at the receiver for this configuration in Figure~\ref{fig:measurement}.
-Our measurements confirm that angular difference has a large effect on the intensity readings (see how the IR intensity distribution changes when the \changes{angle has increased}).
+Our measurements confirm that angular difference has a large effect on the intensity readings (see how the IR intensity distribution changes when the \changes{angle has increased}) \achal{this sounds awkward to me - maybe "Note the changes in IR intensity distribution when the angle increases}.
 
@@
 Better if we can compare to i1.} One side effect that we have observed in this approach is that the {\em Intensity IR} sometimes eliminates the desired target during the {\em scanning} stage.
-Out of 300 trials, \changes{the target is accidentally eliminated in 13 (4.3\%).} This is higher than the error rate for {\em Naive IR} (5 out of 300, ~1.6\%).
+Out of 300 trials, \changes{the target was accidentally eliminated in 13 (4.3\%).} This is higher than the error rate for {\em Naive IR} (5 out of 300, ~1.6\%).
 \changes{Even though the higher error rate increases the chance of multiple attempts in target selection, our analysis above illustrates that overall performance of {\em Intensity IR} is still better than {\em Naive IR}.

### 2014-08-14 doc/sui14/iteration1.tex
bug fix

@@
 As for the qualitative feedback, 11 out of \changes{16} participants preferred IR selection over list mode (\changes{3} preferred list, \changes{2} was undecided).
-\changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between them did you use?} (multiple choices), among 16 participates, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result}.
+\changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between them did you use?} (multiple choices), among 16 participates, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result}}.
 Although through our observation, many participants walked one or two steps during some tasks, this suggests that participants viewed their moving as a followed-through action of adjusting their heads instead of an intention to walk closer to the targets.

### 2014-08-14 doc/sui14/sui14.tex
bug fix

@@
 For all other uses, contact the Owner/Author.\\
-\\
+%\\
 Copyright is held by the owner/author(s).\\
-\\
+%\\
 {\emph{SUI '14}}, Oct 04-05 2014, Honolulu, HI, USA\\

### 2014-08-14 doc/sui14/applications.tex
rewrote sentence about sample tasks
Grammar changes and comments
bjoern's editing pass

@@
 
-Our prototype, \changes{(based on {\em Naive IR}, \footnote{For this particular configuration, the environment is not dense enough, and {\em Naive IR} performs well for head orientation based targeting proof-of-concept.})} includes three smart devices: a lamp and a fan that could be switched on and off; a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
-The on/off appliances are using relays, and the smart TV is a 30\inch{} display connected to a laptop.
+Our prototype, includes three smart devices: a lamp and a fan that could be switched on and off; and a smart TV with playback, volume and navigation controls (see Figure~\ref{fig:smart-home}).
+\changes{The prototype used \em{Naive IR}, as the small number of target devices made selection without disambiguation possible.} The appliances are switched with 120V AC relays.
+The smart TV is a 30\inch{} display connected to a laptop.
 User interfaces for each were pre-defined in our application.
@@
 We set the devices up in a simulated living room environment and invited 14 users to step through a predefined set of tasks to control the appliances at a distance.
-\changes{These tasks include turning off the light, playing movie, turning up the volume, etc., which are some sample tasks that people would normally do at home.}
+\changes{The tasks included turning off a lamp, playing a movie and turning up the volume.}
 

### 2014-08-14 doc/sui14/iteration1.tex
bjoern's editing pass

--- 2014-08-14 doc/sui14/iteration1.tex
@@
-%!TEX root = uist14.tex
+%!TEX root = sui14.tex
 
@@
 
+\newpage
 \subsection{Technique}
@@
 We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}), each with a number ID and a letter representing the name.
-For the study, we recruited \changes{16} participants from our institution \changes{(8 males and 6 females)} \changes{using email invitations}.
-\changes{They are composed of undergrad students, graduate students, and university staff.
-Their background ranges from Information(6), Engineering(4), Math and Science(4), Design(1), Others(1)}  \changes{14} had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
+For the study, \changes{we recruited 16 participants from our institution (8 males and 6 females) by email.} \changes{Participants included undergraduate and graduate students, as well as university staff.
+Their educational backgrounds included Information Science (6), Engineering (4), Math and Science (4), Design (1), Others(1).}  \changes{14} had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.
 Four participants wore prescription glasses, which makes Glass more cumbersome to use and adjust and may have affected their task performance.
@@
 \centering
-\includegraphics[width=0.9\columnwidth]{figures/study-layout1.png}
+\includegraphics[width=1.0\columnwidth]{figures/study-layout1.png}
 \caption{In the targeting study, participants were asked to find and select one of 10 targets in the lab environment.}
@@
 During the study, we measured the {\bf target acquisition time} for each target selection.
-\changes{At the beginning of each acquisition, participants were asked to stand at a fixed position approximately 10-12 feet away from the nodes, looking down until the target is provided.
-As they attempt to connect, they may move around if they decided it would help them select.
-After each task, they were asked to move back to the starting position.} Afterward, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.
+\changes{At the beginning of each acquisition, participants were asked to stand at a fixed position approximately 10-12 feet away from the nodes, looking down until a target is announced.
+Participants were allowed to physically move towards targets if they decided that this would help them with the task.
+After each task, they were asked to move back to the starting position.} Afterwards, participants were asked to complete a survey of primarily open-ended \changes{questions} about their experience.
 
@@
 A narrower beam will make targeting more challenging, given the precision constraints of human head movement.
-
+\newpage %bjoern - hack fix to avoid orphaned heading
 \subsection{Results}
@@
 \includegraphics[width=0.95\columnwidth]{figures/result_study1b.pdf}
-\caption{Boxplot and CDF of target acquisition time when refinement is needed or not in {\em IR selection}}.
+\caption{Boxplot and CDF of target acquisition time when refinement is needed or not in {\em IR selection}.}
 \label{fig:with_without_refinement}
@@
 It takes 6.40 seconds (on average) to complete  a selection without any refinement, but 9.16 seconds with refinement, indicating that an additional 2.76 seconds are needed for disambiguation (see Figure~\ref{fig:with_without_refinement}).
-This difference is significant $(t(19)=-2.7827, p=0.012)$ using t-test.
+This difference is significant $(t(19)=-2.7827, \changes{p<0.05})$.% using t-test.
+%p=0.012
 %  \bjoern{I don't understand why we have 279 DF in the previous test but only 19 here??}). 
@@
 
-To further generalize the results, in Figure~\ref{fig:time-vs-list-order}, we show the acquisition time for each individual target in {\em list selection} condition, ordered by their relative position in the list.
+To further generalize the results, in Figure~\ref{fig:time-vs-list-order}, we show the acquisition time for each individual target in the {\em list selection} condition, ordered by their relative position in the list.
 Since with IR technique, the acquisition time is invariant from each target's order, we use solid lines to represent the average performance of {\em IR selection}.
@@
 
-As for the qualitative feedback, 11 out of \changes{16} participants preferred IR selection over list mode (\changes{3} preferred list, \changes{2} was undecided).
-\changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between them did you use?} (multiple choices), among 16 participates, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result}.
-Although through our observation, many participants walked one or two steps during some tasks, this suggests that participants viewed their moving as a followed-through action of adjusting their heads instead of an intention to walk closer to the targets.
+\changes{In the elicited qualitative feedback, 11 out of 16 participants preferred IR selection over list mode (3 preferred list, 2 were undecided)}.
+\changes{When asked: \studyquote{When you have multiple candidate targets within sight, which method for selecting between them did you use?}, 81\% chose {\em adjust head so that only one candidate is signaled}, 56\% chose {\em tap (to enter refinement stage) and then choose between list}, and 25\% chose {\em walk closer to the candidate for a better result} (multiple choices were allowed)}.
+\changes{While we observed more participants walking one or two steps during some tasks, this suggests that participants viewed their moving as a followed-through action of adjusting their heads instead of an intentional to walk closer to the targets.}
 

### 2014-08-14 doc/sui14/iteration2.tex
bjoern's editing pass

@@
 We have empirically measured the intensity distribution at the receiver for this configuration in Figure~\ref{fig:measurement}.
-Our measurements confirm that angular difference has a large effect on the intensity readings (see how the IR intensity distribution changes when the \changes{angle has increased}) \achal{this sounds awkward to me - maybe "Note the changes in IR intensity distribution when the angle increases}.
+\changes{Our measurements confirm that angular difference has a large effect on the intensity readings, with rapid fall-off at increasing angles.}
+%(see how the IR intensity distribution changes when the \changes{angle has increased}) \achal{this sounds awkward to me - maybe "Note the changes in IR intensity distribution when the angle increases}.
 
@@
 \includegraphics[width=0.95\columnwidth]{figures/IRIntensityDistribution.pdf}
-\caption{This figure shows an empirical measurement of IR intensity at different positions.
-We measure the intensity at a plane every meter away from the IR emitter.
-On that plane, a sample is taken every 5 centimeters, \changes{and we stop when the reading is smaller than ambient noise.} \changes{To match our assumption about the relationship between angles and intensities, the vertical axis is changed to the degree of the angle.} The size and color brightness represent the intensity of the readings for visualization.}
+\caption{\changes{Empirical measurement of IR intensity at different positions.
+We measure the intensity on a horizontal plane at one-meter intervals from the IR emitter.} On that plane, a sample is taken every 5 centimeters, \changes{and we stop when the reading is smaller than ambient noise.} \changes{To match our assumption about the relationship between angles and intensities, the vertical axis is changed to the degree of the angle.} The size and color brightness represent the intensity of the readings for visualization.}
 \label{fig:measurement}
@@
 \centering
-\includegraphics[width=0.95\columnwidth]{figures/study-layout2.pdf}
+\includegraphics[width=1.0\columnwidth]{figures/study-layout2.pdf}
 \caption{The environment setup for our second study. In comparison to our first study, we have deliberately increased the target density.}
@@
 \centering
-\includegraphics[width=0.95\columnwidth]{figures/result_study2.pdf}
+\includegraphics[width=1.0\columnwidth]{figures/result_study2.pdf}
 \caption{Boxplot and CDF of the target acquisition time in {\em Naive IR} and {\em Intensity IR} conditions.}
@@
 }  % but this is the trade-off for the overall faster acquisition time.
-
 %%% Local Variables: 

### 2014-08-14 doc/sui14/iteration3.tex
bjoern's editing pass

--- 2014-08-13 doc/sui14/iteration3.tex
@@
-%!TEX root = uist14.tex
+%!TEX root = sui14.tex
+\newpage
 \section{Iteration 3: Head motion refinement}
@@
 \changes{at similar distances}, their relative orientations are stable
-(see Figure~\ref{fig:third_principle}). And since we are using IR at
-the same time, the learning process is transparent to the user and the
-map can be built without an explicit calibration phase.
+(see Figure~\ref{fig:third_principle}). \changes{This learning process can be transparent to the user and the
+map can be built without an explicit calibration phase.}
 
@@
 \changes{After the map is created, the user can enter a quasi-mode for
-refinement by holding down on the touchpad. In this quasi-mode, one device
+refinement by holding down on the touchpad. In this quasi-mode, a single selected device
 lights up at a time. When the user turns his head in the direction of another
-device, the light switches to that device. Therefore, the user can move between
+device, the selection (and indicator light) switches to that device. Therefore, the user can move between
 devices one at a time with slight head movements. This prediction is
-implemented by first calculating the user's direction of motion, and searching
+implemented by first calculating the user's direction of head motion, and searching
 through the adjacency map for the nearest device in that direction.  To
@@
 calculated as the difference between the last sample and the first sample.
-Note that we use a hysteresis method to temporarily lock to the current device
-- if the variance of the buffer is below a threshold, we assume no movement has
+We use a hysteresis method to avoid spurrious selection changes
+--- if the variance of the buffer is below a threshold, we assume no movement has
 occurred and do not select another device.}
@@
 \subsection{Evaluation}
-We evaluate the head motion refinement method by holding an informal study and
-collecting qualitative feedback from a subset of 4 users from the Iteration 2
-study. In the study, we asked users to try cycling through the targets using
+We evaluated the head motion refinement method through an informal study and
+collected qualitative feedback from a subset of 4 users from the Iteration 2
+study. In this evaluation, we asked users to cycle through multiple targets using
 the new quasi-mode.
@@
 The users had strong preferences for the new method of refinement.
-\changes{Though we lack formal data to evaluate the performance of
-this method, from our observation during the experiments, we found
-that each trial became much easier than previous iterations. We
+\changes{Our observations suggest that each trial became much easier than previous iterations. We
 conducted a survey to collect qualitative feedback after the
-experiment.} On a scale from 1-7, 1 being the least mental effort and
+experiment.} 
+On a scale from 1-7, 1 being the least mental effort and
 7 being the most mental effort users rated the old technique 4.25 and
-the new technique 2 on average. 100\% indicated a preference for head
+the new technique 2 on average. All users indicated a preference for head
 movement to list navigation. One user referenced the issue of naming

### 2014-08-14 doc/sui14/iteration1.tex
update i1 participant gender

@@
 We deployed 10 wireless nodes in an indoor environment (Figure~\ref{fig:targeting-study-layout}), each with a number ID and a letter representing the name.
-For the study, \changes{we recruited 16 participants from our institution (8 males and 6 females) by email.} \changes{Participants included undergraduate and graduate students, as well as university staff.
+For the study, \changes{we recruited 16 participants from our institution (9 males and 7 females) by email.} \changes{Participants included undergraduate and graduate students, as well as university staff.
 Their educational backgrounds included Information Science (6), Engineering (4), Math and Science (4), Design (1), Others(1).}  \changes{14} had never used Google Glass before, so we offered a tutorial before the experiment in order to introduce the device.

### 2014-08-14 doc/sui14/iteration2.tex
updated description of figure 9, this leads to page overflow, so removed \newpage from iteration 3

@@
 \caption{\changes{Empirical measurement of IR intensity at different positions.
-We measure the intensity on a horizontal plane at one-meter intervals from the IR emitter.} On that plane, a sample is taken every 5 centimeters, \changes{and we stop when the reading is smaller than ambient noise.} \changes{To match our assumption about the relationship between angles and intensities, the vertical axis is changed to the degree of the angle.} The size and color brightness represent the intensity of the readings for visualization.}
+We measure the intensity on a horizontal plane at one-meter intervals from the IR emitter.} On that plane, a sample is taken every 5 centimeters, \changes{and we stop when the intensity reading is around the level of ambient noise.
+To match our assumption about the relationship between angles and intensities, the vertical axis shows the relative degree of the angle rather than the distance to the center of the plane.} The size and color brightness represent the intensity of the readings for visualization.}
 \label{fig:measurement}

### 2014-08-14 doc/sui14/sui14.tex
bjoern's editing pass
Capitalize the first letter of key words

@@
 % need a different title for sure, but not in a hurry
-\title{\systemnamenospace: Head Orientation Based Selection\\in Physical Spaces}
+\title{\systemnamenospace: Head Orientation-Based Selection\\in Physical Spaces}
 %Project GlasSees: Direct Interaction Through Line-of-sight}
@@
 \end{abstract}
-\vspace{-0.05in}
+%\vspace{-0.05in}
 \keywords{
-  Wearable computing; spatial interaction; selection; Glass%; infrared
+  Wearable Computing; Spatial Interaction; Selection; Glass%; infrared
 }
-\vspace{-0.1in}
+%\vspace{-0.1in}
 %% refer to http://www.acm.org/about/class/ccs98-html
@@
 %
-%% \balance
+\balance
 

### 2014-08-14 doc/sui14/relatedwork.tex
Minor, remove trailing spaces
added related work of Hipp'09, also tried to use 80-column rules for the texts
added low cost solution argument to hipp
removed useless whitespace

--- 2014-08-06 doc/sui14/relatedwork.tex
@@
 \subsection{Pointing in Physical Spaces}
-Rukzio ~\cite{rukzio_experimental_2006} studied alternative methods for selecting devices in physical spaces and found that users strongly preferred either tapping target appliances with a mobile device or pointing at a distance to browsing a list.
-Several other approaches to spatial selection with handheld devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008} or finger-worn devices~\cite{merrill_augmenting_2007} exist.
+Rukzio \changes{et al.}~\cite{rukzio_experimental_2006} studied alternative
+methods for selecting devices in physical spaces and found that users strongly
+preferred either tapping target appliances with a mobile device or pointing at a
+distance to browsing a list. Several other approaches to spatial selection with
+handheld
+devices~\cite{beigl_point_1999,patel_2-way_2003,wilson_xwand:_2003,schmidt_picontrol:_2012,kemp_point-and-click_2008}
+or finger-worn devices~\cite{merrill_augmenting_2007} exist.
 %and exchanging information with smart infrastructure sensor networks ~\cite{lifton_tricorder:_2007,mittal_ubicorder:_2011,costanza_sensortune:_2010}.
-In some techniques, users select objects of interest with laser pointers; however, the laser dot's small target area makes it poorly matched to head orientation input.
-Other approaches rely on virtual room models in which a user's location is estimated using IMU-based orientation sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our technique does not require a static map ahead of time.
-%Other targeting systems use IR with handheld pointers~\cite{swindells_that_2002} as well as wearable devices such as rings and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to smart devices.
-Our system tackles an unresolved issue of prior approaches -- navigating an area dense with potential targets and resolving selection ambiguity.
+
+In some techniques, users select objects of interest with laser pointers;
+however, the laser dot's small target area makes it poorly matched to head
+orientation input.  Other approaches rely on virtual room models in which a
+user's location is estimated using IMU-based orientation
+sensing~\cite{wilson_xwand:_2003,lifton_tricorder:_2007} -- in contrast, our
+technique does not require a static map ahead of time. \changes{In the FreeMote
+system, \cite{hipp2009universal} an IR camera in a handheld controller interprets readings from IR emitters on target appliances.
+In contrast, we explore a lower cost alternative, using only IR emitters and receivers.
+}
+% Other targeting systems use IR with handheld
+% pointers~\cite{swindells_that_2002} as well as wearable devices such as rings
+% and Bluetooth audio earpieces~\cite{merrill_augmenting_2007} to connect to
+% smart devices.
+Our system tackles an unresolved issue of prior approaches -- navigating an area
+dense with potential targets and resolving selection ambiguity.
 

### 2014-08-14 doc/sui14/abstract.tex
orientation-based rather than head-orientation

--- 2014-06-27 doc/sui14/abstract.tex
@@
 %
-We present the iterative design and evaluation of \systemname -- a Head-Orientation Based Method for interacting with these devices at a distance.
+We present the iterative design and evaluation of \systemname -- a Head Orientation-Based Selection technique for interacting with these devices at a distance.
 We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers.

### 2014-08-14 doc/sui14/conclusion.tex
orientation-based rather than head-orientation

@@
 
-In this paper, by presenting our iterative design process in head-orientation based targeting, we have learned that IR alone can help reduce the overall acquisition time by reducing the chances when we need to perform refinement.
+In this paper, by presenting our iterative design process in head orientation-based targeting, we have learned that IR alone can help reduce the overall acquisition time by reducing the chances when we need to perform refinement.
 With IR intensity added, the targeting can work better in a relative dense environment.

### 2014-08-14 doc/sui14/interaction.tex
bjoern's editing pass
orientation-based rather than head-orientation

@@
 \section{IR for head orientation-based targeting}
-In this section, we present our hardware platform (IR targeting) and interaction model (head-orientation based) that we will use throughout our iterative designs.
+In this section, we present our hardware platform (IR targeting) and interaction model (head orientation-based) that we will use throughout our iterative designs.
 
@@
 
+\newpage %bjoern hack fix
 \subsection{Interaction Model}

### 2014-10-03 doc/sui14/intro.tex
should be 14 participants

@@
  Our {\em Naive IR} technique shows an alphabetically ordered disambiguation list on the near-eye display (Figure~\ref{fig:teaser} center).
-A study with $14$ participants finds that target acquisition with naive IR targeting is preferred by users and is faster than pure list selection without IR, but refinement is still time consuming.
+A study with $16$ participants finds that target acquisition with naive IR targeting is preferred by users and is faster than pure list selection without IR, but refinement is still time consuming.
 

### 2014-10-05 glass/slider.h
cleaned the code a bit

@@ -1,3 +1,6 @@
+// This code serves as the driver to use softpot slider
+// After we use Google Glass, this code is obselete now.
+
 #ifndef slider_h
 #define slider_h
 
@@ -68,12 +71,12 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
 
       } else {
         //within duration but still unpressed -> keep waiting
-        
+
         return gHOVER;
       }
     }
 
-    
+
   }
 
   if(checkDTap) {
@@ -103,7 +106,7 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
       release_time = millis();
       DEBUG_PRINT("pressed duration: ");
       DEBUG_PRINT(release_time - start_time);
-      
+
       DEBUG_PRINTLN();
       if(release_time - start_time < HOLD_THRESHOLD) {
         //a tap -> check for double tap
@@ -137,7 +140,7 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
           // return gTAP;
         }
 
-        
+
       } else {
         //relase from hover (long hold)
 
@@ -156,32 +159,32 @@ gesture_t sliderEvent(int* sDelta, int* sVal) {
         // return gRELEASE;
       }
 
-      
-    } 
+
+    }
     else {
       // has been idle
       sliderState = gsIDLE;
       return gNONE;
     }
-  } 
+  }
   else {  // slider is being pressed
-    
+
     if(sliderState == gsPRESS) {
       // continuously pressing
       *sDelta = softpotReading - *sVal;
       if(abs(*sDelta) > SOFTPOT_DELTA_THRESHOLD){ //avoid noise
         *sVal = softpotReading;
-        
+
         return gHOVERCHANGE;
       } else {
-        
-        return gHOVER;    
+
+        return gHOVER;
       }
-      
-    } 
+
+    }
     else {
       // first time pressed
-      
+
       sliderState = gsPRESS;
       start_time = millis();
       *sVal = softpotReading;
